### 2022-10-18 doc/paper.tex
Add files from previous project

\documentclass[twocolumn,10pt,cleanfoot]{asme2ej}

\usepackage{graphicx} %% for loading jpg figures
\usepackage{bm}
\usepackage{nicefrac}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{parskip}
\usepackage{listings}
\usepackage{tablefootnote}
\usepackage{float}
\usepackage{xcolor}
\usepackage{hyperref}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,
    breaklines=true,
    captionpos=b,
    keepspaces=true,
    numbers=left,
    numbersep=5pt,
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2
}

\lstset{breaklines=true, style=mystyle}
%% The class has several options
%  onecolumn/twocolumn - format for one or two columns per page
%  10pt/11pt/12pt - use 10, 11, or 12 point font
%  oneside/twoside - format for oneside/twosided printing
%  final/draft - format for final/draft copy
%  cleanfoot - take out copyright info in footer leave page number
%  cleanhead - take out the conference banner on the title page
%  titlepage/notitlepage - put in titlepage or leave out titlepage
%  
%% The default is oneside, onecolumn, 10pt, final

\title{Linear regression for approximating real and synthetic data}

%%% first author
\author{Jonatan H. Hanssen
    \affiliation{
	Bachelor Student, Robotics and Intelligent Systems\\
	Department of Informatics\\
	The faculty of Mathmatics and Natural Sciences\\
    Email: jonatahh@ifi.uio.no
    }
}

\author{Eric E. Reber
    \affiliation{
	Bachelor Student, Robotics and Intelligent Systems\\
	Department of Informatics\\
	The faculty of Mathmatics and Natural Sciences\\
    Email: ericer@ifi.uio.no
    }
}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Abstract}
We used linear regression to create a fit of a synthetic dataset generated by the Franke function, and of a real dataset of elevation data from a terrain in Norway. More specifically, we used Ordinary Least Squares (OLS), Ridge and Lasso regression to explore how the different methods perform on different datasets. On the synthetic dataset, we found that Ridge regression allowed us to increase our model complexity and achieve lower MSE scores than OLS, but that Lasso struggled due to computational constraints. On the real dataset, we found that regularization had little impact, as we did not reach model complexities in which overfitting occured.
% Comparing the Mean Squared Error (MSE) for the different methods, we found that Ridge and Lasso outperform OLS for small datasets and high polynomial degrees, but that OLS, given the optimal polynomial degree for the dataset, yields the lowest MSE. Overall we found that the different forms of linear regression fit the synthetic data well. However, when looking at the real datasets, we found that all forms of linear regression performed relatively poorly. We concluded that linear regression perhaps is not the best way of predicting terrain data with rapid and irregular changes in elevation.

% \tableofcontents
% \newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{nomenclature}
% \entry{A}{You may include nomenclature here.}
% \entry{$\alpha$}{There are two arguments for each entry of the nomemclature environment, the symbol and the definition.}
% \end{nomenclature}
%
% The primary text heading is  boldface and flushed left with the left margin.  The spacing between the  text and the heading is two line spaces.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

One of the core challenges in machine learning is model selection, because we cannot make a priori assessments about how well the model will fit our dataset. A multitude of factors determine the efficacy of our model, the machine learning method we use, the hyperparameters of said method, and the model complexity we select. An uninformed decision in model selection could result in an under- or overfitting to the data, yielding poor results. Thus, it is important that we rigorously test our tuning such that these factors coincide at an intersection, providing the minimal error in unity.

In this paper, we will provide a detailed inspection of the Ordinary Least Squares (OLS), Ridge and Lasso linear regression methods, their hyperparameters and varying model complexities with the aim of making an informed decision about model selection when fitting to different datasets. The datasets consist of synthetic data created by the Franke function, and a real dataset of terrain in Norway. We inform our decisions by plotting the mean squared error (MSE), our cost function, over different polynomial degrees describing model complexity, over different resampling techniques and over different values of lambda, which is the hyperparameter responsible for a suitable regularization. We will perform a bias-variance decomposition for the different linear regression methods to check for under- and overfitting and plot our implementations of the methods up against popular implementations such as scikit learn to ensure that the behaviour of our models is expected.

First, we will go more into depth about the methods and the theory behind them, along with a brief description of our implementation. We will further continue to the results, alongside a critical discussion of their implications. Finally, we will give a conclusion on the optimal model selection for the different datasets.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Method}

\subsection{Theory}

This sections covers the theoretic background which underlies the methods used in this paper. We use the following shorthands for our mathematical notation.

\subsubsection{Linear regression}

To describe how linear regression works, let us first describe the problem it aims to solve. Assume we have a dataset, which contains a series of inputs $\{x_i\}$, and corresponding target values $\{y_i\}$. Our assumption is that there is some relationship $y_i = f(x_i) + \epsilon_i$ where $\epsilon_i \sim N(0,\sigma^2)$, and we wish to create a model which approximates this relationship. More specifically, we wish to find a vector $\beta$ which gives us the "best" approximation $\bm{\tilde{y}}$ of $\bm{y}$, by using the parameters in the following equation:

\begin{equation}
X \bm{\beta} = \bm{\tilde{y}}
\label{first}
\end{equation}

$X$ is called the design matrix, and is an $n$ by $p$ matrix created by our input data and our choice of basis functions, where $n$ is the amount of different input states we have and $p$ is the number of features we use for our fit. For our problem, $n$ is the number of distinct $x$ values times the number of distinct $y$ values, and $p$ is based on the polynomial degree we use. The basis functions are the summands of a two variable polynomial of degree $N$. The element $x_{ij}$ is calculated by evaluating the basis function corresponding the coloumn $j$ for the input data corresponding to the row $i$. As we can see from Eq.~\ref{first}, each row of this matrix is scalar multiplied by the vector $\beta$, giving us a linear combination of the values of $X_{i*}$ for each $y_i$ in the vector $\bm{\tilde{y}}$. The vector $\bm{\beta}$ is called the parameter vector, and finding the optimal values $\hat{\bm{\beta}}$ is the problem we wish to solve. But for this, we need to define what an "optimal" prediction $\bm{\tilde{y}}$ is.

\subsubsection{The cost function}

To define the optimal prediction, we use a cost function, and the choice of cost function greatly influences what optimal parameters we get.\footnote{It is here important to keep in mind that the cost function used when fitting the model (training) rarely is the same as the one used for testing our prediction} An intuitive cost function is the mean squared error (MSE). This function simply computes the average distance between our prediction and the target value, squared:

\begin{equation}
	C_{mse}(\tilde{\bm{y}}) = \sum_{i=0}^{n-1}(y_i-\tilde{y_i})^2
\end{equation}

We may now plug in the definition of $\bm{\tilde{y}}$.

\begin{equation}
	C_{mse}(\bm{\beta}) = \sum_{i=0}^{n-1}(y_i-X_{i*}\bm{\beta})^2
\end{equation}

We see that we have a function of our parameters $\bm{\beta}$, which describe how much our prediction deviates from our target values. Our problem now is to optimize this function, that is, find the values of $\beta$ which gives the MSE the lowest value. Luckily for us, it can be proven that the second derivative of this function is always positive \cite[106]{linreg}, which means that we can find the global minimum by solving the equation

\begin{gather}
\frac{\delta}{\delta \beta} \sum_{i=0}^{n-1}(y_i-X_{i*}\bm{\beta})^2 = 0 \\
 - \frac{2}{n} X^T(\bm{y} - X\bm{\beta}) = 0
\end{gather}

Solving this for $\beta$ gives us the so called normal equation, an analytical solution\footnote{This solution requires that $X^TX$ is non-singular, which may not be the case. This is solved by using the singular value decomposition of the design matrix, which we will not go into here} for the optimal $\hat{\beta}$.

% \begin{equation}
\begin{gather}
- \frac{2}{n} X^T(\bm{y} - X\bm{\beta}) = 0 \\
X^T(\bm{y} - X\bm{\beta}) = 0 \\
X^T\bm{y} - X^TX\bm{\beta} = 0 \\
X^T\bm{y} = X^TX\bm{\beta} \\
\bm{\hat{\beta}} = (X^TX)^{-1}X^T\bm{y}
\end{gather}

The process we have just described, with MSE as the cost function, is what is known as Ordinary Least Squares (OLS) regression. Using different cost functions when fitting the model leads to different equations for $\bm{\hat{\beta}}$. This leads us to the two next forms of linear regression we use in this paper. But before this, a...

\subsubsection{Brief statistical intermission}

We must remember that our target values are defined as $y_i = f + \epsilon_i$, that is, they are distorted by a stochastic noise $\epsilon_i \sim N(0,\sigma^2)$. We should therefore take a brief intermission to prove some of the statistical qualities of the method we have just described, so as to reassure ourselves that we are justified in applying linear regression to noisy data. We will start by showing that

\begin{equation}
E[y_i] = \sum_j X_{ij}\beta_j
\end{equation}

The proof is as follows:

$\bm{y}$ is defined as the $f(\bm{x}) + \bm{\epsilon}$. We make an assumption that this can be approximated as a matrix multiplication $X\bm{\beta}$ plus an error vector $\bm{\epsilon}$. This means that each element of the vector
$\bm{y}$ can be expressed as follows:

\begin{equation}
y_i = \sum_j x_{ij}\beta_j + \epsilon_i
\end{equation}

If we take the expectation value of this expression we get the following:

\begin{gather}
E[y_i] = E\left[\sum_j x_{ij}\beta_j + \epsilon_i\right] \\
E[y_i] = E\left[\sum_j x_{ij}\beta_j\right] + E[\epsilon_i]
\end{gather}

However, the elements of $X$ are not stochastic, and neither are the elements of $\beta$ and therefore the first expectation value is simply the sum itself. Furthermore, $\epsilon$ is explicitly defined as a normal distribution $N(0,\sigma^2)$, and will by definition have the expectation value $0$. Therefore, we end up with the final expression:

\begin{equation}
E[y_i] = \sum_j x_{ij}\beta_j = \bm{X}_{i,*}\bm{\beta}
\end{equation}

Next, we show that the variance of $y_i$ is equal to the variance of the noise:

\begin{gather}
var[y_i] = E\left[(y_i - E[y_i])^2\right]  \\
= E\left[y_i^2 - 2E[y_i]y_i + E[y_i]^2\right]
\end{gather}

Distributing the outer expectation value function:

\begin{gather}
var[y_i] = E[y_i^2] - 2E\left[E[y_i]\right]E[y_i] + E\left[E[y]^2\right]  \\
= E[y_i^2] - 2E\left[E[y_i]\right]E[y_i] + E\left[E[y]\right]E\left[E[y]\right]
\end{gather}

The result of calculating the expectation value is non-stochastic. This means that $E[E[X]] = E[X]$. From this it follows that

\begin{gather}
var[y_i] = E[y_i^2] - 2E[y_i]E[y_i] + E[y_i]E[y_i]  \\
= E[y_i^2] - E[y_i]E[y_i]  \\
= E[y_i^2] - E[y_i]^2
\end{gather}

The expectation value in the second summand has been proven to be equal to $\bm{X}_{i,*}\bm{\beta}$ above. We therefore have

\begin{gather}
var[y_i] = E[y_i^2] - (\bm{X}_{i,*}\bm{\beta})^2  \\
= E\left[(X_{i,*}\bm{\beta})^2 + X_{i,*}\bm{\beta}\epsilon_i + \epsilon_i^2\right] - (\bm{X}_{i,*}\bm{\beta})^2  \\
= E\left[(X_{i,*}\bm{\beta})^2\right] + E[X_{i,*}\bm{\beta}\epsilon_i] + E[\epsilon_i^2] - (\bm{X}_{i,*}\bm{\beta})^2
\end{gather}

$X_{i,*}\bm{\beta}$ and $\epsilon_i$ are both scalars. Therefore the expectation value can be written as $E[X_{i,*}\bm{\beta}]E[\epsilon_i]$. However, $E[\epsilon_i]$ is by definition $0$, because $\epsilon_i$ is defined as a normal distribution of mean $0$ and variance $\sigma^2$. Furthermore, the expectation value of the non-stochastic $(X_{i,*}\bm{\beta})^2$ is simply the expression itself. We therefore have

\begin{gather}
var[y_i] = (X_{i,*}\bm{\beta})^2 + E[X_{i,*}\bm{\beta}]\cdot 0 + E[\epsilon_i^2] - (\bm{X}_{i,*}\bm{\beta})^2  \\
= E[\epsilon_i^2]
\end{gather}

We can prove that this is equal to the variance of $\epsilon_i$:

\begin{gather}
E[\epsilon_i^2] = \frac{1}{n} \sum_i \epsilon_i^2 \\
var[\epsilon_i] = \frac{1}{n} \sum_i (\epsilon_i - \bar{\epsilon_i})^2 \\
var[\epsilon_i] = \frac{1}{n} \sum_i (\epsilon_i - 0)^2 \\
var[\epsilon_i] = \frac{1}{n} \sum_i \epsilon_i^2 = E[\epsilon_i^2]
\end{gather}

And of course, we know that the variance of $\epsilon_i$ by definition is $\sigma^2$.

\begin{gather}
var[y_i] = E[\epsilon_i^2] = var[\epsilon_i] = \sigma^2 
\end{gather}

We should also prove some qualities about or parameters. We start by proving that $E[\hat{\beta}] = \beta$

Because we are assuming $(X^{T}X)^{-1}X^{T}$ to be deterministic, $E[(X^{T}X)^{-1}X^{T}] = (X^{T}X)^{-1}X^{T}$. Moreover, we know $E[y] = y$, and we can thus rewrite the equation as follows:

\begin{equation}
E[(X^{T}X)^{-1}X^{T}y] = (X^{T}X)^{-1}X^{T}y = \beta
\end{equation}

Next, we prove that $var(\beta) = \sigma^{2}(X^{T}X)^{-1}$.

$$var(\beta) = var((X^{T}X)^{-1}X^{T}y)$$

Because we are assuming $(X^{T}X)^{-1}X^{T}$ to be deterministic and $y$ to be stochastic, we can rewrite the $var(\beta)$ as following:

\begin{gather}
var(\beta) = (X^{T}X)^{-1}X^{T}var(y)((X^{T}X)^{-1}X^{T})^{T} \\
= (X^{T}X)^{-1}X^{T}var(y)(X^{T})^{T}((X^{T}X)^{-1})^{T}
\end{gather}

Since $(X^{T})^{T} = X$ and $(X^{T}X)^{-1})^{T}$ = $(X^{T}X)^{T})^{-1} = (X^{T}X)^{-1}$

\begin{gather}
= (X^{T}X)^{-1}X^{T}var(y)X(X^{T}X)^{-1}
\end{gather}

We know var(y) = $\sigma^{2}$, and since it is a scalar it is commutative. Thus we may move it freely

\begin{gather}
= \sigma^{2}(X^{T}X)^{-1}X^{T}X(X^{T}X)^{-1} \\
= \sigma^{2}(X^{T}X)^{-1}
\end{gather}

\subsubsection{Ridge Regression}

A common problem with OLS is what is known as overfitting (we will discuss this in greater detail in later sections). A symptom of overfitting is that the values in the parameter vector $\bm{\beta}$ grow very large, and vary wildly for small changes in our dataset. To counteract this, we can add to the cost function a term which increases with the values in $\bm{\beta}$. Optimizing the cost function would now not only mean finding the smallest difference between target and predicted values, but also minimizing the absolute value of the parameters. We now define a new cost function

\begin{equation}
	C_{lasso}(\bm{\beta}) = \sum_{i=0}^{n-1}(y_i-X_{i*}\bm{\beta})^2 + \lambda \sum_{j=0}^{p-1} \beta_j^2
\end{equation}

$||\beta||_2^2$ is equal to $\beta^T\beta$ and is simply the norm of the parameter vector, squared. With the new hyperparameter $\lambda \in (0, \inf)$ , we can decide how harshly we should punish large values in our parameter vector. Finding an optimal $\lambda$ is an important step in the process of finding the best parameters for our model.

A similar derivation as done in the previous section gives us the normal equation for Ridge regression:

\begin{equation}
	\bm{\hat{\beta}} = (X^TX + \lambda I)^{-1}X^T\bm{y}
\end{equation}

We see that this equation, with a $\lambda > 0$ avoids the problem of an inversion of a non-singular matrix, and gives us an analytical solution for $\bm{\hat{\beta}}$ yet again.\footnote{We should cherish this analytical solution, as it may well be the last time we see a cost function that can be optimized analytically.} By using ridge regresison, we can control the complexity of our model and effectively hinder overfitting.

\subsubsection{Lasso Regression}

Another cost function is one where we add the L1-norm (the sum of the absolute values of the parameters), instead of the L2-norm. The new cost function is as follows.

\begin{equation}
	C_{lasso}(\bm{\beta}) = \sum_{i=0}^{n-1}(y_i-X_{i*}\bm{\beta})^2 + \lambda \sum_{j=0}^{p-1} |\beta_j|
\end{equation}

However, the derivative of the absolute value is not continous, so there is no analytical solution for the optimal parameters with this form of linear regression. We must then solve the problem numerically, which can be done by using Gradient Descent. Like Ridge regression, this form of regression controls the complexity of the model, with the added benefit that Lasso can drive parameters to exactly zero.

\subsubsection{Resampling methods: The bootstrap}

Bootstrap is a resampling method divided into simple steps. First, we reshuffle our dataset with replacement, which will provide us with a new dataset. Note that we will already have split our data into train and test data before we begin the bootstrap, so when we reshuffle, we will only reshuffle the training data. The reshuffled datasets act as our samples.

Secondly, we will perform our linear regression on this new dataset and store the predictions of the test data. Thirdly, we repeat for a given large number of iterations. Since we use all the test predictions to calculate the MSE, bias and variance, we will then according to the central limit theorem \cite{CLT} obtain a normal distribution of these performance metrics. The law of large numbers \cite{LLN} then tells us that the mean sample values of the MSE, bias and variance approach the true value of MSE, bias and variance. This is dependent on our assumption that our data is independently and identically distributed.

These approximations of the true value of MSE, bias and variance will become useful when performing our bias-variance trade-off analysis.

\subsubsection{Cross validation}

Cross validation is another resampling technique, in which we reshuffle our data and then split them into k-folds, where k specifies the number of folds. One fold is used as our test data, and the rest is used as training data. We run our linear regression, then switch up which fold is used for testing until all folds have been used. As in bootstrapping, we will approach the true values of for example the mean squared error for our model. 
Theoretically, because both two resampling techniques use the central limit theorem and the law of large numbers, they should return the same values for the MSE, bias and variance given the same dataset. Though resampling techniques can be computationally expensive, they make up for it by aiding in model evaluation by supplying us with performance metrics obtained from a larger number of samples.

\subsubsection{Bias-variance trade-off}

A crucial part of model evaluation is ensuring that the model is neither underfitted nor overfitted, which we can determine by observing the relationship between the bias and the variance. Bias is defined as $(Bias[\tilde{y}])^2 = (y-E[\tilde{y}])^2$, the difference between the expected value of our prediction and the observed value. Models with high bias are usually underfitted, whilst models with low bias are the opposite. Variance on the other hand is defined as $var[\tilde{f}] = \frac{1}{n}\sum(\tilde{y_i}-E[\tilde{y}])^2$, how far away predictions are from the mean prediction. Models with high variance are usually overfitted to the training data, whilst low variance models are the opposite.

A good rule of thumb for finding a good fit is looking at the intercept between bias and variance. This is because the cost function can be decomposed into these two elements. The proof for this is as follows:

Given data $\bm{y} = f(\bm{x}) + \bm{\epsilon}$ and a OLS model $\bm{\tilde{y}} = X\bm{\beta}$, we define
the Mean Squared Error as follows

\begin{gather}
C(X,\beta) = \frac{1}{n}\sum_{i=0}^{n-1}(y_i - \tilde{y_i})^2
\end{gather}

This is simply the expectation value of $(\bm{y} - \bm{\tilde{y}})^2$

\begin{gather}
C(X,\beta) = E\left[(\bm{y} - \bm{\tilde{y}})^2\right]
\end{gather}

To derive the bias variance decomposition, we now add $E[\bm{\tilde{y}}] - E[\bm{\tilde{y}}]$ 
as a sneaky trick. We will also replace $\bm{y}$ with our definition $\bm{y} = \bm{f} + \bm{\epsilon}$

\begin{gather}
C(X,\beta) = E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}] + E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2\right]
\end{gather}
%%%%%%%%%%%%%%%%%5
\begin{multline}
= E[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2 + 2[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])(E[\bm{\tilde{y}}] \\
	- \bm{\tilde{y}})] +(E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2]
\end{multline}

\begin{multline}
= E[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2] + \\
	2E[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])(E[\bm{\tilde{y}}] - \bm{\tilde{y}})] + E[(E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2]
\end{multline}

Let us focus our attention on the middle term first.

\begin{gather}
2E\left[(\bm{f} + \bm{\epsilon} + E[\bm{\tilde{y}}])(E[\bm{\tilde{y}}] - \bm{\tilde{y}})\right] 
\end{gather}

Expanding the multiplication:

\begin{gather}
2E\left[\bm{f}E[\bm{\tilde{y}}] + \bm{\epsilon}E[\bm{\tilde{y}}] - E[\bm{\tilde{y}}]E[\bm{\tilde{y}}] - \bm{f}\bm{\tilde{y}} - \bm{\epsilon}\bm{\tilde{y}} + E[\bm{\tilde{y}}]\bm{\tilde{y}}\right] 
\end{gather}

Distributing the expectation value:

\begin{gather}
2\left[E[\bm{f}E[\bm{\tilde{y}}]] + E[\bm{\epsilon}E[\bm{\tilde{y}}]] - E[E[\bm{\tilde{y}}]E[\bm{\tilde{y}}]] - E[\bm{f}\bm{\tilde{y}}] - E[\bm{\epsilon}\bm{\tilde{y}}] + E[E[\bm{\tilde{y}}]\bm{\tilde{y}}]\right] 
\end{gather}

We know that $\bm{\tilde{y}}$ and $\bm{f}$ are non-stochastic, and we know that the expectation value of $\bm{\epsilon}$ is $0$. We also know that 
$\bm{\epsilon}$ and $\bm{\tilde{y}}$ are independent, which means we can write $E[\bm{\epsilon}\bm{\tilde{y}}]$ as $E[\bm{\epsilon}]E[\bm{\tilde{y}}]$ We therefore have

\begin{gather}
2\left[\bm{f}\bm{\tilde{y}} + 0 - \bm{\tilde{y}}\bm{\tilde{y}} - \bm{f}\bm{\tilde{y}} - 0 + \bm{\tilde{y}}\bm{\tilde{y}}\right] = 0 
\end{gather}

We then have

\begin{gather}
C(X,\beta) = E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2\right] + E\left[(E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2\right] 
\end{gather}

The second term is by definition the variance of model.

\begin{gather}
C(X,\beta) = E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2\right] + var[\bm{\tilde{f}}] 
\end{gather}

And now the final term

\begin{gather}
E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2\right]
\end{gather}

\begin{multline}
= E[\bm{f}^2 + \bm{f}\bm{\epsilon} - \bm{f}E[\bm{\tilde{y}}] + \bm{f\epsilon} + \bm{\epsilon}^2 - \\ 
	\bm{\epsilon}E[\bm{\tilde{y}}] - \bm{f}E[\bm{\tilde{y}}] - \bm{\epsilon}E[\bm{\tilde{y}}] + E[\bm{\tilde{y}}]^2]
\end{multline}

After distributing the expectation value we get the following, keeping in mind that $E[\bm{f}] = E[\bm{y} - \bm{\epsilon}] = E[\bm{y}] - E[\bm{\epsilon}] = \bm{y}$

\begin{gather}
\bm{y}^2 - 2\bm{y}\bm{\tilde{y}} + E[\bm{\tilde{y}}]^2 + \sigma^2 \\
(\bm{y} + E[\bm{\tilde{y}}]^2 + \sigma^2 
\end{gather}

We see that the quadratic form is equal to $bias[\bm{\tilde{y}}]^2$. We can then conclude that the cost function can be rewritten as follows:

\begin{gather}
C(X,\beta) = bias[\bm{\tilde{y}}]^2 + var[\bm{\tilde{f}}] + \sigma^2 
\end{gather}

Because the cost function is a sum of the bias and the variance, the lowest error will generally occur when both bias and variance are low. This also fits our understanding of the definitions of bias and variance, as we wish to obtain a model that is neither high in bias (underfitted) or high in variance (overfitted).

With the theory explained, we can go on to describe how we have applied it to our problem.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Implementation}
The source code for our implementation can be found here: \href{url}{https://github.com/jonatan-hanssen/3155-project1.git}.
\subsubsection{Preprocessing: Creating the design matrix and scaling our data}
We have two datasets, one synthetic and one real. Regardless of the source of the data, we create the design matrix in a similar manner, as the relation between our input data and our target values are the same for both datasets, that is, for each combination of $x$ and $y$, we have a height $z$. For our real data, this value represents the elevation in meters above ocean level for a terrain in Norway. As we wish to perform a polynomial fit, we create our design matrix such that each row $i$ has the following values:

\begin{equation}
X_{i*} = 1 \;\; x_i \;\; y_i \; x_i^2 \;\; x_i y_i \;\; y_i^2 \;\; ... \;\; x_i^p y_i^p
\end{equation}

Below is the python code used to create the design matrix, which can be found in \texttt{src/utils.py}.

\lstinputlisting[language=Python, firstline=30, lastline=44]{../src/utils.py}

While the creating of the design matrix is equivalent for the two datasets, the preprocessing of the data afterwards is not. For the Franke function no scaling is needed, as we generate the data ourselves to be between $0$ and $1$. Given that we are doing a polynomial fit, every element in our design matrix will be a multiplication of $x^p$ and $y^k$ for different $p$'s and $k$'s, which, given our choice of range for $x$ and $y$, guarantees that every element is between $0$ and $1$. However, we have included a centering of the data, which has no effect on the Ordinary Least Squares regression, but which is used when applying Ridge and Lasso regression, to avoid punishing the intercept. The centering is done as follows: First, the first coloumn of the design matrix (the intercept coloumn) is discarded. Then we calculate the mean for the target values and the design matrix (the mean for the design matrix is a vector with mean values for each coloumn). We subtract this from the design matrix and the target values when performing the fit giving us a vector $\bm{\beta}$. Afterwards we calculate the intercept with the following equation

\begin{equation}
	intercept = \bar{y} - \bm{\bar{x}_c}^T \bm{\beta}
\end{equation}

	where $\bar{y}$ is the mean value of $\bm{y}$ and $\bm{\bar{x}_c}$ is a vector of size $p$ with $\bar{X}_i = mean(X_i)$ (the mean of each coloumn in $X$). We finally create our prediction by using the $\bm{\beta}$, $X$ and intercept we calculated above:

\begin{equation}
	\bm{\hat{y}} = X \bm{\beta} + intercept
\end{equation}

The code for this, taken from the \texttt{evaluate\_model} function in \texttt{src/utils.py} is as follows:\footnote{Some intermediate parts of our implementation not related to this have been removed for readability}

\begin{lstlisting}[language=Python]
X_train = X_train[:, 1:]
z_train_mean = np.mean(z_train, axis=0)
X_train_mean = np.mean(X_train, axis=0)
beta = model((X_train - X_train_mean), (z_train - z_train_mean))
intercept = np.mean(z_train_mean - X_train_mean @ beta)
z_pred_train = X_train @ beta + intercept
\end{lstlisting}

Where \texttt{model} calculates \texttt{beta} using the normal equations for OLS.

For the real elevation data, scaling is definitely required, as the values for $x$, $y$ reach several hundred and $z$ reaches over a thousand. For polynomial degree $10$ and an $x$ value which reaches $400$, we will have a design matrix with values from $0$ to $400^{10}$. This large range of values makes our linear regression perform poorly. We have therefore chosen to scale our design matrix using Scikit's \texttt{MinMaxScaler}, which scales each column to be in the range of $(0,1)$. The scaling coefficients are calculated on the training data, and applied to both the training and test data. We have done the same with the values for $z$. This gives a much more stable model. The code for doing this on the design matrix is as follows, taken from \texttt{src/utils.py}:

\lstinputlisting[language=Python, firstline=242, lastline=249]{../src/utils.py}

Another form of "preprocessing" which is strictly practical in nature, is that we have reduced the dimensions of the data to 320 by 640. This was done by downscaling the image.\footnote{The interpolation was done using "nearest neighbor interpolation", which does not introduce new values into the data \cite[77]{dip}. An equivalent programmatic solution would be to for instance only look at every 4 pixels} This was simply done because the design matrix becomes several gigabytes in size if one uses the whole dataset, and model fitting then takes an unreasonable amount of time.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Implementation of linear regression}

Linear regression for OLS and Ridge are implemented by using NumPy's \texttt{numpy.linalg.pinv} and the normal equations. We have implemented them as functions which take in the target values and the design matrix and return $\bm{\hat{\beta}}$. They are called from a wrapper function called \texttt{evaluate\_model} which takes care of the scaling and works with both SciKit models and our own. Below are the functions, taken from \texttt{src/utils.py}:

\lstinputlisting[language=Python, firstline=56, lastline=64]{../src/utils.py}

For repeated calls of these functions with different polynomial degrees, we use the function \texttt{linreg\_to\_N}, which passes the correctly sliced design matrix to the regression functions and saves the predictions and various performance metrics for each polynomial degree for easy plotting.

For Lasso regression, we use SciKit's implementation. We also use SciKit's implementation for OLS and Ridge for comparisons with our own implementation.

The correctness of our Ordinary Least Squares regression can be confirmed by running our debug function (the so called SkrankeFunction)\footnote{this can be done by running \texttt{\$ python3 task\_b.py --debug --noscale -n 2}}, which is a second degree polynomial where $b_0 = 0, \; b_1 = 1, \; b_2 = 2$ and so on. The SkrankeFunction (taken from \texttt{src/utils.py}) looks like this

\lstinputlisting[language=Python, firstline=26, lastline=27]{../src/utils.py}

and the result of our regression is correctly given as $\bm{\beta} = (0 \; 1 \; 2 \; 3 \; 4 \; 5\;)^T$.

\subsubsection{Implementation of resampling methods}

We use resampling methods to better estimate our performance metrics. We use both cross validation and the bootstrap method. Cross validation is implemented in the function \texttt{crossval} in \texttt{src/utils.py}. We split the data into $K$ folds by slicing with NumPy taking consecutive rows of the design matrix and target vector. But before we do this, we resample the data with Scikit's \texttt{resample}. This is important because our $x$'s and $y$'s are not normal distributions but consecutive values between $0$ and $1$, and therefore consecutive rows correspond specific areas of our data, which is not wanted. We then fit on $K-1$ folds, and test on the remaining fold. We return the average error.

We implement bootstrap by repeatedly resampling our data with replacement and performing the prediction. We use the predictions returned from hundreds of resamplings to calculate the bias and variance of our model.

\subsubsection{Franke Function}

We use the Franke Function with an added stochastic noise of to test our models. We explore how higher and lower levels of noise affect the model. Below is the Franke function:

\begin{multline}
f(x,y) = \frac{3}{4}\exp{\left(-\frac{(9x-2)^2}{4}-\frac{(9y-2)^2}{4}\right)} + \\ \frac{3}{4}\exp{\left(-\frac{(9x+1)^2}{49}-\frac{(9y+1)}{10}\right)} + \\
\frac{1}{2}\exp{\left(-\frac{(9x-7)^2}{4}-\frac{(9y-3)^2}{4}\right)} - \\
\frac{1}{5}\exp{\left(-(9x-4)^2-(9y-7)^2\right)}
\end{multline}

As we can see, it is simply a weighted sum of exponentials. As such it has an infinite Taylor expansion into polynomials.\footnote{The proof for this is left as an exercise for the reader}

\section{Results}

\subsection{Resampling}

As stated in the method section, we suspect that bootstrap and cross validation both yield approximations to the true performance metrics due to the central limit theorem, and thus should give identical results. We can observe from Fig.~\ref{frankemsebyresamplingmethod} that our implementation of bootstrap and cross validation yield almost exactly the same results. (Fig.~\ref{realmsebyresamplingmethod} in the appendix shows a similar result for the real data)

\begin{figure}[h]
\centerline{\includegraphics[width=3.25in]{figure/frankemsebyresamplingmethod.png}}
\caption{Comparison between approximated MSE for different resampling methods for the Franke function}
\label{frankemsebyresamplingmethod}
\end{figure}

\subsection{Franke Function}

To gain insight into how the three different regression methods work, we experiment with synthetic data generated by the Franke function.

\subsubsection{Ordinary Least Squares}

We find that OLS creates a good fit for the Franke Function. With 400 datapoints, we can increase our polynomial degree to 20 before we see signs of overfitting.\footnote{Here we also see something a bit surprising: We see that our own implementation of OLS follows Scikit's implementation up to around polynomial degree 10, at which point scikit's MSE increases rapidly. We chalk this up to numerical errors and differing implementation of the normal equations, as we simply use the normal equation exactly, while scikit uses an implementation which is probably computationally more efficient but gives slight differences when comparing to the normal equation. These differences seem to create a large error after a certain size of the design matrix is reached, which we can not adequately explain.} This is most likely because the Franke function can be expanded into a sum of polynomials. We would expect to see higher polynomial degrees give us better and better prediction, up to a certain point.\footnote{And indeed, this seems to be the case, as if we increase the availiable datapoints considerably, as seen in Fig.~\ref{frankiehigherdatapoints} in the appendix, we can reach much higher polynomial degrees without overfitting} When we reach degrees higher than around 20, we see that our test error increases slightly, a sign of overfitting. This is also to be expected, as even if the continous Franke function could be expanded into a series of polynomials, we are working with a limited dataset of only 400 samples (and only 80\% of those in our training set). As we reach higher and higher degrees, we are more and more likely to see unfortunate train-test-splits where the best fit for the training data leads to large errors for the test data. Regardless, our fit for lower model complexity is remarkably good. See Fig.~\ref{frankenonoise} for a comparison between the Franke function and our best prediction, which shows that they are essentially identical. Fig.~\ref{frankenonoisemse} shows the train and test MSE by polynomial degree. The MSE for test in these plots is the average MSE for 100 bootstraps. For a comprehensive list of the shell command required to reproduce this and every other plot, see Table~\ref{allparamstable} in the appendix.

\begin{figure}[h]
\centerline{\includegraphics[width=3.25in]{figure/frankenonoise.png}}
\caption{A comparison of our best OLS prediction and the Franke function with no noise, at polynomial degree 10}
\label{frankenonoise}
\end{figure}

\begin{figure}[h]
\centerline{\includegraphics[width=3.25in]{figure/frankenonoisemse.png}}
\caption{Test and train MSE for the Franke function with no noise, using OLS}
\label{frankenonoisemse}
\end{figure}

However, when we introduce a stochastic noise, we see that overfitting happens rapidly. Without a regularization hyperparameter, parameters fluctuate wildly and our prediction suffers. Fig.~\ref{frankenoisemse} shows our MSE up to polynomial degree 24 with an added stochastic noise $\epsilon \sim N(0,0.05)$, where we see that we are overfitting severely starting at around degree 7. See Fig.~\ref{frankienoisecomparison} in the appendix for test MSE for more and more noisy target values.

\begin{figure}[h]
\centerline{\includegraphics[width=3.25in]{figure/frankenoisemse.png}}
\caption{Test and train MSE for the Franke function with noise, using OLS}
\label{frankenoisemse}
\end{figure}

By using bootstrap resampling, we can study the bias-variance-tradeoff, which shows how the variance of our model between different bootstraps iterations increases as our model overfits our noisy data. See Fig.~\ref{frankiebiasvariance}. Clearly, we need to introduce regularization if we want a model that does not overfit. Fig. ~\ref{frankenoisybeta} in the appendix shows the beta progression for selected betas, clearly showing that we are overfitting.

\begin{figure}[h]
\centerline{\includegraphics[width=3.25in]{figure/frankiebiasvariance.png}}
\caption{Bias, variance and test MSE plotted over polynomial degree for our OLS prediction on the Franke function with an added stochastic noise $\epsilon \sim N(0,0.05)$}
\label{frankiebiasvariance}
\end{figure}

Table ~\ref{ols_mse_table} shows the best polynomial degree and MSE for regular and noisy data. Here we see that the best degree without noise is 9, which shows that even without noise, the low amount of data leads to a slight overfit even at very small polynomial degrees. This is not surprising, as at $N = 9$ we have $55$ features, which is a substantial amount compared to our measly 320 training datapoints. We could of course arbitrarily increase our dataset to ludicrous levels, but doing linear regression with this dataset gives valuable insight into how our performance is dependent on our availiable data.

\begin{table}[h]
\caption{Best MSE and polynomial degree for OLS}
\begin{center}
\label{ols_mse_table}
\begin{tabular}{c l l}
% & & \\ % put some space after the caption
% \hline
Std of noise & Best MSE & Best degree \\
\hline
0 & 0.0005109 & 9\\
0.05 & 0.0038302 & 7 \\
0.10 & 0.0139833 & 7 \\
\hline
\end{tabular}
\end{center}
\end{table}

\subsubsection{Ridge}

Using Ridge regression, we see from Fig.~\ref{frankiedifflambdas} in the appendix that larger and larger values of $\lambda$ efficiently prevents overfitting and reduces the variance between predictions. Looking at the beta progression for the same values of $\lambda$ (Fig.~\ref{frankeridgebetas} in the appendix), we see that the regularization parameter is successfully shrinking the beta values. Using Scikit's gridsearch, we find the optimal $\lambda$ to be $6.10 \cdot 10^{-8}$ and the optimal polynomial degree to be 13, with an average MSE of $0.003027$. This is a good improvement of our OLS fit, and shows that we are able to push our model complexity significantly higher and gain a moderate MSE improvement with a correctly chosen regularization parameter. Looking at the beta progression for this value of lambda (see Fig.~\ref{frankebestlambdabetas}), we see that the beta values settle and do not vary greatly between polynomial degrees. Fig.~\ref{frankebestlambdamse} shows the MSE score, which shows how we are successfully avoiding overfitting. Table~\ref{ols_vs_ridge_vs_lasso_table} shows the relevant values.

\begin{figure}[h]
\centerline{\includegraphics[width=3.25in]{figure/frankebestlambdabetas.png}}
\caption{Beta progression for our Ridge prediction with the optimal lambda}
\label{frankebestlambdabetas}
\end{figure}

\begin{figure}[h]
\centerline{\includegraphics[width=3.25in]{figure/frankebestlambdamse.png}}
\caption{Test and train MSE for ridge}
\label{frankebestlambdamse}
\end{figure}

% We see from this that Ridge regression has not really had much of an effect in decreasing the MSE. We must now remember what Ridge aims to do. With the regularization parameter, we can prevent overfitting if our model is too complex for our data. Looking at Fig.~\ref{frankeridgenoisecomparison} in comparison to Fig.~\ref{frankienoisecomparison} (both in the appendix), we can see that Ridge really has achieved this, reducing overfitting as our data becomes more and more noisy. But we should recall that we decide the complexity of our model, by deciding which polynomial degree we use when creating the design matrix. In fact, by choosing polynomial degree 7 for our OLS method to get the best MSE for noisy data, we are in fact doing a primitive form of best subset selection, as described in Hastie et Al.\cite[57]{hastie}. Almost by definition, the polynomial degree with the best MSE is not overfitted, because if it was, it would not be the model with the lowest MSE among all polynomial degrees. Therefore it is not so unexpected that Ridge regression has a relatively low impact on the MSE. We could expect that Ridge would allow us to reach higher polynomial degrees without overfitting, by shrinking features of the higher polynomial degrees that are not supported but allowing us to keep other high degree terms (for example, allowing us to use $x^5y^4$, but discard $x^9$, if only $x^5y^4$ was supported by the data). But we can see from the Table ~\ref{ols_vs_ridge_vs_lasso_table} that we are not able to push into higher degrees. A possible explanation for this is that every feature is useful, due to the polynomial expansion of the Franke function.

\subsubsection{Lasso}

Looking at Lasso regression, we find that it has the same capability to reduce overfitting as Ridge, as can be seen in Fig.~\ref{frankedifflambdaslasso} in the appendix. However, looking at Table~\ref{ols_vs_ridge_vs_lasso_table}, we see that Lasso falls short when finding a good MSE. Intuitively, it does not make sense for Lasso to be worse than OLS, because Lasso with $\lambda = 0$ is equivalent to OLS. But we cannot avoid considering the fact that Lasso does not have an analytical solution, and as such we must use numerical methods to find the optimal parameters. Using scikit's implementation, we have not been able to get the numerical method to converge, which means that we have not found a global minimum.\footnote{Scikit explicitly mentions that using $\lambda = 0$ does not converge well with their implementation \cite{lasso}} For this reason, Lasso regression has not given satisfying results.

\begin{table*}[t]
\caption{Best MSE for OLS, Ridge and Lasso with 400 datapoints}
\begin{center}
\label{ols_vs_ridge_vs_lasso_table}
\begin{tabular}{c | l l l l}
% & & \\ % put some space after the caption
% \hline
Regression & Std of noise & MSE & Degree & $\lambda$ \\
\hline
OLS & 0.05 & 0.0038302 & 7 & - \\
Ridge & 0.05 & 0.003027 & 13 & $6.10 \cdot 10^{-8}$ \\
Lasso & 0.05 & 0.008298 & 12 & $1.38 \cdot 10^{-5}$ \\
\hline
\end{tabular}
\end{center}
\end{table*}

\subsection{Real elevation data}

With the confidence that our methods work as expected on synthetic data, we turn our attention to the real elevation data. First, let us confirm that it is necessary to scale our values using MinMax-scaling, as we have described in our implementation. Fig.~\ref{realunscaled} and Fig.~\ref{realunscaledmse} in the appendix shows our predicted terrain and associated MSE, showing that we reach a very poor prediction without scaling.

\subsubsection{Ordinary Least Squares}

Here we find some limitations of linear regression. We see from Fig.~\ref{real1msetraintest} that the we are able to fit the terrain up to a certain point, gaining an MSE of $0.0142683$ with OLS with polynomial degree 19, our highest polynomial. This is about two orders of magnitude higher than our lowest MSE for the Franke function without noise, which had values on a similar scale.\footnote{The Franke function without noise has a range of about $(0,1.2)$ and our MinMaxed terrain data has been scaled to a range of $(0,1)$. While not exactly the same, these are on the same order of magnitude, and we believe that comparing the MSE scores is reasonable} In fact, our MSE is comparable to the Franke function with a stochastic noise $\epsilon \sim N(0,0.1)$ (Table~\ref{ols_mse_table_dif_data}), which implies that our terrain is very irregular and difficult to approximate. Looking at the plot for polynomial degree 19 (Fig.~\ref{real1surfacecomp} in the appendix), we see that our poor MSE is reflected visually in our fit as well, as we can only make out general similarities in shape, not finer details. Comparing the best polynomial degree of the best fit with the synthetic data, we see that our best fit is reached much later. This implies that we are not overfitting yet, and indeed the bias variance decomposition confirms this, as we see that variance is essentially inconsequential (Fig.~\ref{realbiasvarianceOLS}). The fact that we are not overfitting can be seen when we increase the polynomial degree to very large amounts, seeing only a very modest decrease in MSE to $0.01381$ at degree 40 (Fig.~\ref{realOLSmse40} in the appendix). Comparing this to the tenfold decrease in MSE when increasing the polynomial degree for the Franke function to 50, we see a clear difference. Looking at the beta progression for selected betas (Fig.\ref{realOLSbetaprog40} in the appendix), we see that the parameters stabilize, which again implies that we are not overfitting, but simply reaching our best possible prediction for polynomial regression. It is reasonable to expect that our model will not overfit, due to the large increase in data we now use for the fit. Our synthetic data used $20 x 20 = 400$ datapoints, while we now have $320 x 640 = 204 800$ datapoints. This means that for a 19th order polynomial, which will have $210$ features, we will have around a thousand datapoints per feature. If we decrease the number of datapoints availiable considerably, we can see overfitting start to occur (Fig.~\ref{real1mselessdata} in the appendix). So we are not overfitting, but we are not decreasing our MSE either. This is likely because the elevation data has a jagged and complex shape, which cannot be accurately recreated, even with a polynomial of degree $19$. We should remember that the Franke function can be expanded into a series of polynomials, which is why we can expect to get a lower and lower MSE as the polynomial degree increases (so long as we have enough data and it is not noisy). The same is not true for our elevation data, and as such we have little to gain from increasing our model complexity after a certain point. Clearly, linear regression does not offer a very good tradeoff between computational cost and performance for this data.

\begin{table}
\caption{Best MSE and polynomial degree for OLS on synthetic and real data}
\begin{center}
\label{ols_mse_table_dif_data}
\begin{tabular}{c l l l}
Dataset & Std of noise & Best MSE & Best degree \\
\hline
Franke & 0 & 0.0005109 & 9\\
Franke & 0.05 & 0.0038302 & 7 \\
Franke & 0.10 & 0.0139833 & 7 \\
Terrain 1 & - & 0.0142683 & 19 \\
\hline
\end{tabular}
\end{center}
\end{table}

\begin{figure}[h]
\centerline{\includegraphics[width=3.25in]{figure/realbiasvarianceOLS.png}}
\caption{Bias, variance and test MSE plotted over polynomial degree for our OLS prediction of Terrain 1}
\label{realbiasvarianceOLS}
\end{figure}

% \begin{figure} 
% \centerline{\includegraphics[width=3.25in]{figure/real1msetraintest.png}}
% \caption{Test and train MSE for OLS on Terrain 1}
% \label{real1msetraintest}
% \end{figure}

\begin{figure}[h]
\centerline{\includegraphics[width=3.25in]{figure/real1msetraintest.png}}
\caption{Test and train MSE for OLS on Terrain 1}
\label{real1msetraintest}
\end{figure}

\subsubsection{Ridge Regression}

As our model is not overfitting, we do not expect Ridge to offer much improvement. This proves to be the case, as we see from Fig.~\ref{real1ridgedifflambdas} in the appendix. Here we see that different values of lambda have very little impact, as there are no parameters that need regularization, and no overfitting to supress. For very large values, we see that we are shrinking the parameters too much, and we see a slight increase in MSE, but overall we do not see much of an effect. We could expect to see that Ridge shows improvement at higher polynomial degrees, when OLS starts to overfit, but due to computational constraints we have not been able to push our grid search much further than 9 degrees. Using scikit's GridSearch, we find that the optimal lambda is 0, which is exactly what we would expect on a model that does not overfit, as Ridge regression with a lambda equal to 0 is equivalent with OLS. See Table~\ref{ols_vs_ridge_vs_lasso_table_real}.

\subsubsection{Lasso Regression}

With lasso regression, we see similar results as with Ridge. Fig.~\ref{real1lassobiasvariance} in the appendix shows little change, except for when our $\lambda$ is too high and we force all parameters to $0$. Comparing lasso to the two other methods for the real dataset has proven to be difficult, as the numerical methods for finding the best parameters are very time consuming. We have not been able to gridsearch up to polynomial degree 19, but up to degree 9 we found the best lambda to be $2.154^{-7}$ for polynomial degree 9, giving an MSE of $0.01685$.

Comparing all three methods up to degree 9 for different lambdas (using the same lambda for Lasso and Ridge, just to get an overview) in Fig.~\ref{real1msecomparison} in the appendix, we see the same trend as we have been describing. Regularization offers little reward, which is to be expected when we are not overfitting.

\begin{table*}[t]
\caption{Best MSE for OLS, Ridge and Lasso regression on real elevation data. We only fit up to degree 9 due to computational constraints}
\begin{center}
\label{ols_vs_ridge_vs_lasso_table_real}
\begin{tabular}{c | l l l}
Regression & MSE & Degree & $\lambda$ \\
\hline
OLS & 0.01538 & 9 & - \\
Ridge & 0.01538 & 9 & 0 \\
Lasso & 0.01685 & 9 & $2.154 \cdot 10^{-7}$ \\
\hline
\end{tabular}
\end{center}
\end{table*}

\section{Conclusion}

In this paper we have compared Ordinary Least Squares, Ridge and Lasso regression for both synthetic and real data. We have seen the value of using resampling methods such as bootstrap and cross validation to gain valuable insights about the quality of our predictions. Using resampling methods, we have found that linear regression was well suited for approximating the synthetic Franke function, but less suited for the real elevation data. We have seen that a correctly chosen regularization parameter can allow us to push our model complexity higher than we can with OLS, and thus give us a better MSE than OLS. Looking at the real data, we have seen that regularization does not offer much value when we are not overfitting our dataset. Thus all three linear regression methods perform similarly on this dataset.

\bibliographystyle{apalike}
\bibliography{bibliography}

\section*{Appendix A: Plots and tables}

The next pages contain plots and tables that are of interest.

\begin{table*}
\caption{This table shows the command to execute to reproduce every figure in the report. (More info about the scripts and their parameters can be found in the README)}
\begin{center}
\label{allparamstable}
\begin{tabular}{c | l l l}
Figure & Shell command (leading \texttt{python3} omitted) \\
\hline
\ref{frankenonoise} & \texttt{task\_b.py --noise 0 -n 25}\\
\ref{frankebestlambdamse} & \texttt{task\_b.py --noise 0 -n 25} \footnotesize{(\it Model has been changed to ridge with lambda=$1.43 \cdot 10^{-7}$)}\\
\ref{frankebestlambdabetas} & \texttt{task\_b.py --noise 0 -n 25} \footnotesize{(\it Model has been changed to ridge with lambda=$1.43 \cdot 10^{-7}$)}\\
\ref{realunscaled} & \texttt{task\_b.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 19 --noscale}\\
\ref{realunscaledmse} & \texttt{task\_b.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 19 --noscale}\\
\ref{frankenonoisemse} & \texttt{task\_c.py --noise 0 -n 24}\\
\ref{frankenoisemse} & \texttt{task\_c.py -n 24}\\
\ref{frankiebiasvariance} & \texttt{task\_c.py -n 24 --step 0.1}\\
\ref{frankiedifflambdas} & \texttt{task\_e\_biasvariance.py -n 24 --step 0.1}\\
\ref{frankedifflambdaslasso} & \texttt{task\_f\_biasvariance.py -n 9 --step 0.15}\\
\ref{real1ridgedifflambdas} & \texttt{task\_e\_biasvariance.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 24}\\
\ref{frankeridgebetas} & \texttt{task\_e\_betas.py -n 24}\\
\ref{frankenoisybeta} & \texttt{task\_c.py -n 24}\\
\ref{frankienoisecomparison} & \texttt{noise\_plot.py}\\
\ref{frankiehigherdatapoints} & \texttt{task\_c.py --noise 0 -n 30 --step 0.01}\\
\ref{real1surfacecomp} & \texttt{task\_b.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 19}\\
\ref{real1mselessdata} & \texttt{task\_c.py --file ../data/tiny\_SRTM\_data\_Norway\_1.tif -n 19}\\
\ref{real1msetraintest} & \texttt{task\_c.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 19}\\
\ref{realbiasvarianceOLS} & \texttt{task\_c.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 19}\\
\ref{realOLSmse40} & \texttt{task\_b.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 40}\\
\ref{realOLSbetaprog40} & \texttt{task\_b.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 40}\\
\ref{frankemsebyresamplingmethod} & \texttt{task\_d.py -n 15 --step 0.02}\\
\ref{real1msecomparison} & \texttt{task\_f\_msecomparison.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 9}\\
\hline
\end{tabular}
\end{center}
\end{table*}

%%%%%%%%%%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%5

\begin{figure*}
\centerline{\includegraphics[width=3.25in]{figure/realmsebyresamplingmethod.png}}
\caption{Comparison between approximated MSE for different resampling methods for the real terrain data}
\label{realmsebyresamplingmethod}
\end{figure*}

\begin{figure*}
\centerline{\includegraphics[width=3.25in]{figure/frankiehigherdatapoints.png}}
\caption{Franke function fitted up to polynomial degree 30, with 10000 datapoints instead of 400. As we can see, we do not overfit even at this polynomial degree when the availiable data is plentiful. The highest MSE was reached at $N=29$, and was $0.00003511$, one tenth of our best prediction with only 400 datapoints}
\label{frankiehigherdatapoints}
\end{figure*}

\begin{figure*}
% \begin{figure}[t]
\centerline{\includegraphics[width=3.25in]{figure/frankienoisecomparison.png}}
\caption{Test and train MSE for the OLS prediction of Franke function with an increasing amount of noise}
\label{frankienoisecomparison}
\end{figure*}

\begin{figure*}
\centerline{\includegraphics[width=3.25in]{figure/frankenoisybeta.png}}
\caption{Beta progression for OLS}
\label{frankenoisybeta}
\end{figure*}

\begin{figure*} 
\centerline{\includegraphics[width=3.25in]{figure/frankiedifflambdas.png}}
\caption{Bias, variance and test MSE plotted over polynomial degree for our Ridge prediction on the Franke function with an added stochastic noise $\epsilon \sim N(0,0.05)$, using lambdas of increasing size. Here we have reduced the number of datapoints to only 100 to force overfitting earlier.}
\label{frankiedifflambdas}
\end{figure*}

\begin{figure*} 
\centerline{\includegraphics[width=3.25in]{figure/frankeridgebetas.png}}
\caption{Beta progression for different values of lambda in Ridge regression. Note the different scales on each of the plots, showing how increasing the regularization parameter shrinks the betas}
\label{frankeridgebetas}
\end{figure*}

% \begin{figure*}
% \centerline{\includegraphics[width=3.25in]{figure/frankeridgenoisecomparison.png}}
% \caption{Test and train MSE for the Ridge prediction with $\lambda = 10^{-8}$ of the Franke function with an increasing amount of noise}
% \label{frankeridgenoisecomparison}
% \end{figure*}

\begin{figure*}
\centerline{\includegraphics[width=3.25in]{figure/frankedifflambdaslasso.png}}
\caption{Bias, variance and test MSE plotted over polynomial degree for our Lasso prediction on the Franke function with an added stochastic noise $\epsilon \sim N(0,0.05)$, using lambdas of increasing size. Here we have reduced the number of datapoints to less than 100 to force overfitting earlier.}
\label{frankedifflambdaslasso}
\end{figure*}

\begin{figure*}
\centerline{\includegraphics[width=3.25in]{figure/realunscaled.png}}
\caption{Our prediction for the real terrain using unscaled data, showing that we reach a very poor result}
\label{realunscaled}
\end{figure*}

\begin{figure*}
\centerline{\includegraphics[width=3.25in]{figure/realunscaledmse.png}}
\caption{Test and train MSE for our prediction for the real terrain using unscaled data, showing that we reach a very poor result}
\label{realunscaledmse}
\end{figure*}

\begin{figure*} 
\centerline{\includegraphics[width=6.85in]{figure/real1surfacecomp.png}}
\caption{A comparison of our best OLS prediction and actual terrain data (terrain 1), at polynomial degree 19}
\label{real1surfacecomp}
\end{figure*}

\begin{figure*}
\centerline{\includegraphics[width=3.25in]{figure/realOLSmse40.png}}
\caption{Test and train MSE on our terrain, this time up to degree 40. We see only a slight decrease in MSE, down to $0.01381$, even though we double the amount of degrees. This has not been bootstrapped due to computational limitations}
\label{realOLSmse40}
\end{figure*}

\begin{figure*}
\centerline{\includegraphics[width=3.25in]{figure/realOLSbetaprog40.png}}
\caption{Beta progression up to polynomial degree 40 for real terrain data, showing that the are relatively constant, implying that we are not overfitting}
\label{realOLSbetaprog40}
\end{figure*}

\begin{figure*}
\centerline{\includegraphics[width=3.25in]{figure/real1mselessdata.png}}
\caption{Test and train MSE for the OLS prediction of terrain 1, this time downscaled to only $50x100$. Here the optimal degree was 11, showing that with lower data we start to overfit more rapidly.}
\label{real1mselessdata}
\end{figure*}

\begin{figure} 
\centerline{\includegraphics[width=3.25in]{figure/real1ridgedifflambdas.png}}
\caption{Bias variance tradeoff for ridge regression with different lambdas, for the real dataset}
\label{real1ridgedifflambdas}
\end{figure}

\begin{figure} 
\centerline{\includegraphics[width=3.25in]{figure/real1lassobiasvariance.png}}
\caption{Bias variance tradeoff for lasso regression with different lambdas, for the real dataset}
\label{real1lassobiasvariance}
\end{figure}

\begin{figure*}
\centerline{\includegraphics[width=3.25in]{figure/real1msecomparison.png}}
\caption{Comparison between OLS, Ridge and Lasso for different values of lambda, for real terrain data}
\label{real1msecomparison}
\end{figure*}

\end{document}

### 2022-10-18 src/utils.py

@@ -1,0 +1,415 @@
+from mpl_toolkits.mplot3d import Axes3D
+import matplotlib.pyplot as plt
+from matplotlib import cm
+from matplotlib.ticker import LinearLocator, FormatStrFormatter
+import numpy as np
+from random import random, seed
+from sklearn.model_selection import train_test_split, GridSearchCV, KFold
+from sklearn.linear_model import LinearRegression
+from sklearn.preprocessing import MinMaxScaler
+from sklearn.utils import resample
+from typing import Tuple, Callable
+from imageio import imread
+import sys
+import argparse
+
+
+def FrankeFunction(x, y):
+    term1 = 0.75 * np.exp(-(0.25 * (9 * x - 2) ** 2) - 0.25 * ((9 * y - 2) ** 2))
+    term2 = 0.75 * np.exp(-((9 * x + 1) ** 2) / 49.0 - 0.1 * (9 * y + 1))
+    term3 = 0.5 * np.exp(-((9 * x - 7) ** 2) / 4.0 - 0.25 * ((9 * y - 3) ** 2))
+    term4 = -0.2 * np.exp(-((9 * x - 4) ** 2) - (9 * y - 7) ** 2)
+    return term1 + term2 + term3 + term4
+
+
+# debug function
+def SkrankeFunction(x, y):
+    return 0 + 1*x + 2*y + 3*x**2 + 4*x*y + 5*y**2
+
+
+def create_X(x, y, n):
+    if len(x.shape) > 1:
+        x = np.ravel(x)
+        y = np.ravel(y)
+
+    N = len(x)
+    l = int((n + 1) * (n + 2) / 2)  # Number of elements in beta
+    X = np.ones((N, l))
+
+    for i in range(1, n + 1):
+        q = int((i) * (i + 1) / 2)
+        for k in range(i + 1):
+            X[:, q + k] = (x ** (i - k)) * (y**k)
+
+    return X
+
+
+def R2(y_data, y_model):
+    return 1 - np.sum((y_data - y_model) ** 2) / np.sum((y_data - np.mean(y_data)) ** 2)
+
+
+def MSE(y_data, y_model):
+    n = np.size(y_model)
+    return np.sum((y_data - y_model) ** 2) / n
+
+
+def OLS(X_train: np.ndarray, z_train: np.ndarray):
+    beta = np.linalg.pinv(X_train.T @ X_train) @ X_train.T @ z_train
+    return beta
+
+
[…]

### 2022-11-03 src/testing/diabetes.py
Add testing files

@@ -1,0 +1,57 @@
+from utils import *
+from sklearn.datasets import load_diabetes
+
+np.random.seed(42069)
+
+
+inputs = 10
+diabetes = load_diabetes()
+
+X = diabetes.data
+t = diabetes.target
+
+X_train, X_test, t_train, t_test = train_test_split(X, t)
+
+t_train = t_train.reshape(t_train.shape[0], 1)
+t_test = t_test.reshape(t_test.shape[0], 1)
+
+t_train /= 10
+t_test /= 10
+
+scaler = MinMaxScaler()
+scaler.fit(X_train)
+
+X_train_sc = scaler.transform(X_train)
+X_test_sc = scaler.transform(X_test)
+
+eta = 0.008
+momentum = 0.5
+rho = 0.9
+rho2 = 0.99
+
+sched = Adam
+params = [eta, rho, rho2]
+
+dims = (inputs, 10, 1)
+neural = FFNN(
+    dims,
+    # output_func=sigmoid,
+    # cost_func=CostLogReg,
+)
+train_errors, test_errors = neural.fit(
+    X_train,
+    t_train,
+    sched,
+    *params,
+    batches=10,
+    epochs=10000,
+    X_test=X_test,
+    t_test=t_test,
+)
+plt.plot(train_errors, label="Train")
+plt.plot(test_errors, label="Train")
+plt.legend()
+plt.xlabel("Epochs")
+plt.ylabel("MSE")
+plt.title("MSE over Epochs")
+plt.show()

### 2022-11-03 src/testing/ionosphere.py

@@ -1,0 +1,55 @@
+from utils import *
+
+np.random.seed(42069)
+
+
+inputs = 34
+data = np.genfromtxt("../data/ionosphere.csv", delimiter=",")
+
+data = resample(data)
+X = data[:, :inputs]
+t = data[:, inputs]
+
+
+X_train, X_test, t_train, t_test = train_test_split(X, t)
+
+t_train = t_train.reshape(t_train.shape[0], 1)
+t_test = t_test.reshape(t_test.shape[0], 1)
+
+scaler = MinMaxScaler()
+scaler.fit(X_train)
+
+X_train_sc = scaler.transform(X_train)
+X_test_sc = scaler.transform(X_test)
+
+eta = 0.008
+momentum = 0.5
+rho = 0.9
+rho2 = 0.99
+
+sched = Adam
+params = [eta, rho, rho2]
+
+dims = (inputs, 1)
+neural = FFNN(
+    dims,
+    output_func=sigmoid,
+    cost_func=CostLogReg,
+)
+train_errors, test_errors = neural.fit(
+    X_train,
+    t_train,
+    sched,
+    *params,
+    batches=10,
+    epochs=10000,
+    X_test=X_test,
+    t_test=t_test,
+)
+plt.plot(train_errors, label="Train")
+plt.plot(test_errors, label="Train")
+plt.legend()
+plt.xlabel("Epochs")
+plt.ylabel("MSE")
+plt.title("MSE over Epochs")
+plt.show()

### 2022-11-03 src/testing/mlptest.py

@@ -1,0 +1,92 @@
+"""
+task b (and task g): plot terrain, approximate terrain with OLS (own implementation and scikit) and calculate MSE, R2 &
+                     beta over model complexity for real data. Performs task_b, so no resampling.
+"""
+# Our own library of functions
+from utils import *
+from sklearn.neural_network import MLPRegressor
+
+np.random.seed(42069)
+
+# tests schedulers for a given model
+def test_scheduler():
+    # define model (should probably be sent in)
+    (
+        betas_to_plot,
+        N,
+        X,
+        X_train,
+        X_test,
+        z,
+        z_train,
+        z_test,
+        centering,
+        x,
+        y,
+        z,
+    ) = read_from_cmdline()
+
+    dummy_x = np.array(
+        [
+            [0.134, 0.91827, 0.1982, 0.34654],
+            [0.7246, 0.8887, 0.1513, 0.97716],
+            [0.441, 0.123, 0.321, 0.71],
+        ]
+    )
+    dims = (2, 20, 20, 20, 1)
+
+    mlp = MLPRegressor(
+        hidden_layer_sizes=dims,
+        activation="tanh",
+        solver="adam",
+        max_iter=10000,
+        learning_rate="adaptive",
+        tol=0,
+        n_iter_no_change=10000000,
+    )
+
+    z_train = FrankeFunction(X_train[:, 1], X_train[:, 2])
+    z_train = z_train.reshape(z_train.shape[0], 1)
+
+    mlp.fit(X_train, np.ravel(z_train))
+    z_pred = mlp.predict(X)
+    print(mlp.n_iter_)
+
+    pred_map = z_pred.reshape(z.shape)
+
+    return pred_map, x, y, z
+
+
+pred_map, x, y, z = test_scheduler()
+print(pred_map)
+
+# ------------ PLOTTING 3D -----------------------
+fig = plt.figure(figsize=plt.figaspect(0.3))
+
+# Subplot for terrain
+ax = fig.add_subplot(121, projection="3d")
+# Plot the surface.
+surf = ax.plot_surface(x, y, z, cmap=cm.coolwarm, linewidth=0, antialiased=False)
+ax.zaxis.set_major_locator(LinearLocator(10))
+ax.zaxis.set_major_formatter(FormatStrFormatter("%.02f"))
+ax.set_title("Scaled terrain", size=24)
[…]

### 2022-11-03 src/testing/sonar.py

@@ -1,0 +1,55 @@
+from utils import *
+
+np.random.seed(42069)
+
+
+inputs = 60
+data = np.genfromtxt("../data/sonar.csv", delimiter=",")
+
+data = resample(data)
+X = data[:, :inputs]
+t = data[:, inputs]
+
+
+X_train, X_test, t_train, t_test = train_test_split(X, t)
+
+t_train = t_train.reshape(t_train.shape[0], 1)
+t_test = t_test.reshape(t_test.shape[0], 1)
+
+scaler = MinMaxScaler()
+scaler.fit(X_train)
+
+X_train_sc = scaler.transform(X_train)
+X_test_sc = scaler.transform(X_test)
+
+eta = 0.008
+momentum = 0.5
+rho = 0.9
+rho2 = 0.99
+
+sched = Adam
+params = [eta, rho, rho2]
+
+dims = (inputs, 30, 30, 1)
+neural = FFNN(
+    dims,
+    output_func=sigmoid,
+    cost_func=CostLogReg,
+)
+train_errors, test_errors = neural.fit(
+    X_train,
+    t_train,
+    sched,
+    *params,
+    batches=10,
+    epochs=10000,
+    X_test=X_test,
+    t_test=t_test,
+)
+plt.plot(train_errors, label="Train")
+plt.plot(test_errors, label="Train")
+plt.legend()
+plt.xlabel("Epochs")
+plt.ylabel("MSE")
+plt.title("MSE over Epochs")
+plt.show()

### 2022-11-03 src/testing/ionosphere.py
Fix

@@ -4,7 +4,7 @@ np.random.seed(42069)
 
 
 inputs = 34
-data = np.genfromtxt("../data/ionosphere.csv", delimiter=",")
+data = np.genfromtxt("../../data/ionosphere.csv", delimiter=",")
 
 data = resample(data)
 X = data[:, :inputs]

### 2022-11-03 src/testing/sonar.py

@@ -4,7 +4,7 @@ np.random.seed(42069)
 
 
 inputs = 60
-data = np.genfromtxt("../data/sonar.csv", delimiter=",")
+data = np.genfromtxt("../../data/sonar.csv", delimiter=",")
 
 data = resample(data)
 X = data[:, :inputs]

### 2022-11-03 doc/paper.tex
Add more on the paper

@@
-\documentclass[twocolumn,10pt,cleanfoot]{asme2ej}
+\documentclass[onecolumn,10pt,cleanfoot]{asme2ej}
 
@@
 \usepackage{xcolor}
-\usepackage{hyperref}
 
@@
 
-\title{Linear regression for approximating real and synthetic data}
+\title{Neural networks for solving regression and classification problems}
 
@@
     \affiliation{
-	Bachelor Student, Robotics and Intelligent Systems\\
-	Department of Informatics\\
-	The faculty of Mathmatics and Natural Sciences\\
+	Bachelor Student, Robotics and \\
+	Intelligent Systems\\ \\[-10pt]
+	Department of Informatics\\ \\[-10pt]
+	The faculty of Mathematics and \\
+	Natural Sciences\\ \\[-10pt]
     Email: jonatahh@ifi.uio.no
@@
     \affiliation{
-	Bachelor Student, Robotics and Intelligent Systems\\
-	Department of Informatics\\
-	The faculty of Mathmatics and Natural Sciences\\
+	Bachelor Student, Robotics and \\
+	Intelligent Systems\\ \\[-10pt]
+	Department of Informatics\\ \\[-10pt]
+	The faculty of Mathematics and \\
+	Natural Sciences\\ \\[-10pt]
     Email: ericer@ifi.uio.no
+    }
+}
+
+\author{Gregor Kajda
+    \affiliation{
+	Bachelor Student, Robotics and \\
+	Intelligent Systems\\ \\[-10pt]
+	Department of Informatics\\ \\[-10pt]
+	The faculty of Mathematics and \\
+	Natural Sciences\\ \\[-10pt]
+    Email: grzegork@ifi.uio.no
     }
@@
 \section{Abstract}
-We used linear regression to create a fit of a synthetic dataset generated by the Franke function, and of a real dataset of elevation data from a terrain in Norway.
-More specifically, we used Ordinary Least Squares (OLS), Ridge and Lasso regression to explore how the different methods perform on different datasets.
-On the synthetic dataset, we found that Ridge regression allowed us to increase our model complexity and achieve lower MSE scores than OLS, but that Lasso struggled due to computational constraints.
-On the real dataset, we found that regularization had little impact, as we did not reach model complexities in which overfitting occured.
-% Comparing the Mean Squared Error (MSE) for the different methods, we found that Ridge and Lasso outperform OLS for small datasets and high polynomial degrees, but that OLS, given the optimal polynomial degree for the dataset, yields the lowest MSE.
-Overall we found that the different forms of linear regression fit the synthetic data well.
-However, when looking at the real datasets, we found that all forms of linear regression performed relatively poorly.
-We concluded that linear regression perhaps is not the best way of predicting terrain data with rapid and irregular changes in elevation.
+
+We developed a Feed Forward Neural Network (FFNN) and applied it to many common datasets used for benchmarking machine learning algorithms.
+Furthermore, we compared our results to those attained by linear and logistic regression.
+To optimize our network, we explored the use of different schedulers in our gradient descent method, and compared their performance against eachother.
+Overall, we found that our FFNN performed well in all classification tasks we tested, attaining test accuracies between $95$ and $100$ percent for all classification problem.
+However, we also found that simple logistic regression performed similarly, with a much lower computational cost.
+For regression on terrain data, we found our neural net lacking, not being able to beat our polynomial fit from our previous project, and only attaining an MSE of 1 BILLION, which is 2000 BILLION lower than our best polynomial fit of the Franke Function.
+In our gradient descent, we found that SCHEDULER performed best, being 1 BILLION percent better than the average of all other methods.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % \begin{nomenclature}
@@
 
-One of the core challenges in machine learning is model selection, because we cannot make a priori assessments about how well the model will fit our dataset.
-A multitude of factors determine the efficacy of our model, the machine learning method we use, the hyperparameters of said method, and the model complexity we select.
-An uninformed decision in model selection could result in an under- or overfitting to the data, yielding poor results.
-Thus, it is important that we rigorously test our tuning such that these factors coincide at an intersection, providing the minimal error in unity.
-
-In this paper, we will provide a detailed inspection of the Ordinary Least Squares (OLS), Ridge and Lasso linear regression methods, their hyperparameters and varying model complexities with the aim of making an informed decision about model selection when fitting to different datasets.
-The datasets consist of synthetic data created by the Franke function, and a real dataset of terrain in Norway.
-We inform our decisions by plotting the mean squared error (MSE), our cost function, over different polynomial degrees describing model complexity, over different resampling techniques and over different values of lambda, which is the hyperparameter responsible for a suitable regularization.
-We will perform a bias-variance decomposition for the different linear regression methods to check for under- and overfitting and plot our implementations of the methods up against popular implementations such as scikit learn to ensure that the behaviour of our models is expected.
-
-First, we will go more into depth about the methods and the theory behind them, along with a brief description of our implementation.
-We will further continue to the results, alongside a critical discussion of their implications.
-Finally, we will give a conclusion on the optimal model selection for the different datasets.
+Neural Networks are all the rage these days, everyone wants to have a neural network in their home. Ive even seen people install neural networks in their dogs.
 
@@
 
-This sections covers the theoretic background which underlies the methods used in this paper. We use the following shorthands for our mathematical notation.
+This sections covers the theoretic background which underlies the methods used in this paper.
 
-\subsubsection{Linear regression}
+\subsubsection{Logistic Regression}
 
-To describe how linear regression works, let us first describe the problem it aims to solve.
-Assume we have a dataset, which contains a series of inputs $\{x_i\}$, and corresponding target values $\{y_i\}$.
-Our assumption is that there is some relationship $y_i = f(x_i) + \epsilon_i$ where $\epsilon_i \sim N(0,\sigma^2)$, and we wish to create a model which approximates this relationship.
-More specifically, we wish to find a vector $\beta$ which gives us the "best" approximation $\bm{\tilde{y}}$ of $\bm{y}$, by using the parameters in the following equation:
+As part of the research conducted in this project, we introduce a new type of regression model, known as Logistic regression.
+This is a powerful, yet simple algorithm used in machine learning for the purpose of approximating a probability associated with the occurrence of a specific event given measured data.
+The fundamental idea behind logistic regression is based on the exact approach used in Linear Regression, meaning that we wish to describe a relationship between dependent and explanatory variables in terms of a set of estimated parameters.
+These methods differ however in how they resolve the said matter, with Linear regression modeling a set of parameters by solving a set of linear equations, whilst Logistic regression uses the gradient descent algorithm and the sigmoid function for this purpose.
+
+Logistic regression is thus just plain linear regression where the weighted sum of the LHS of the linear equation - with randomly initialized parameters - is fed into an activation function.
+This in turn allows for the computation of probability related to an event taking place.
+Rounding of said probability leads to labeling of each input as belonging to one of two classes, meaning that each object fed into the model will be classified as being of class 0 or 1.
+To begin with however, the probability output by the activation function will be incorrect for majority of inputs, due to the random nature of the initialized parameters.
+While this may seem to be an issue, the weights can be adjusted to give better classification results through the application of the gradient descent coupled with an adjustable variable called learning rate.
+As we also shall see, the implementation of an activation function at the output gives a significant advantage over the classic Linear regression in terms of classification power.
+
+\subsubsection{Gradient Descent}
+
+In our previous project, we have primarily dealt with problems that have an analytical solution.
+This means that finding the optimal solution in these cases is a deterministic process which does not require fine tuning of parameters involved.
+However, for the methods which are applied to problems in this research - such as Logistic Regression - no analytical solutions exist, thus we will have to turn to numerical techniques to find the ‘ideal’ solution.
+Our method of choice in this case is gradient descent, an optimization technique which reduces the residuals by utilizing the idea of moving downhill a convex function until a globally minimal scalar value is reached.
+The gradient itself is nothing more than a first order derivative of a given function, which points in the direction of the greatest change.
+
+The application of gradient descent to our problems as means of finding the ideal solution implies that the parameters be initialized randomly.
+The most common approach is to set each parameter equal to an arbitrary number ranging between 0 and 1, and iteratively update the parameters by calculating the gradient at each given point in the residual space \cite[preface p.
+x]{dds}.
+For any one-layer model, the gradient function used to tune the weights will take the form of the product of the partial derivative of the cost function with respect to the activation function, and the partial derivative of the activation function with respect to the weights:
 
 \begin{equation}
-X \bm{\beta} = \bm{\tilde{y}}
-\label{first}
+\frac{\partial{\cal C}}{\partial w_{jk}}  = \frac{\partial C}{\partial a_{j}}\frac{\partial a_j}{\partial w_{jk}}
 \end{equation}
 
-$X$ is called the design matrix, and is an $n$ by $p$ matrix created by our input data and our choice of basis functions, where $n$ is the amount of different input states we have and $p$ is the number of features we use for our fit.
-For our problem, $n$ is the number of distinct $x$ values times the number of distinct $y$ values, and $p$ is based on the polynomial degree we use.
-The basis functions are the summands of a two variable polynomial of degree $N$.
-The element $x_{ij}$ is calculated by evaluating the basis function corresponding the coloumn $j$ for the input data corresponding to the row $i$.
-As we can see from Eq.~\ref{first}, each row of this matrix is scalar multiplied by the vector $\beta$, giving us a linear combination of the values of $X_{i*}$ for each $y_i$ in the vector $\bm{\tilde{y}}$.
-The vector $\bm{\beta}$ is called the parameter vector, and finding the optimal values $\hat{\bm{\beta}}$ is the problem we wish to solve.
-But for this, we need to define what an "optimal" prediction $\bm{\tilde{y}}$ is.
+Although gradient decent is an excellent algorithm for finding optimal solutions, it can also become quite computationally expensive for problems where large numbers of datapoints are available.
+However, this issue may be effectively countered by applying a variant of standard gradient approach known as Stochastic Gradient Decent or simply SGD.
+The SGD is based on randomly choosing one or a small number of instances from the dataset to update the parameters, hence the name stochastic.
+While it may seem like a poor idea to approximate the gradient instead of using the true gradient, the SGD has proven to give a large computational advantage and will often allow the model to converge into a global minima faster than GD.
+Furthermore, the stochastic nature of this method can allow SGD to avoid getting stuck in local minima \cite[46]{sr}.
 
-\subsubsection{The cost function}
+\subsubsection{Schedulers}
 
-To define the optimal prediction, we use a cost function, and the choice of cost function greatly influences what optimal parameters we get.\footnote{It is here important to keep in mind that the cost function used when fitting the model (training) rarely is the same as the one used for testing our prediction} An intuitive cost function is the mean squared error (MSE).
-This function simply computes the average distance between our prediction and the target value, squared:
+Whilst SGD itself gives a significant boost to the learning pace of machine learning models, it is possible to improve the algorithm further by applying a special type of optimization method called scheduling.
+These methods work by adaptively adjusting the learning rate, thus allowing us to avoid local minima and converge even faster.
+One of the simpler methods involves adding a fraction of the previous change to the current value used to update the weights, thus keeping a sort of memory of the previous iterations.
+This is called momentum, due to the physical analogue of a moving particle having momentum.
+Other more complex methods include keeping running averages of the first and second moments of the gradient, and using this to inform the direction of change in the current update.
+
+\subsubsection{Feedforward Neural Networks}
+
+For some problems, like approximating a simple function like the Franke Function in project 1, it is sufficient to choose a set of basis functions and find their optimal linear combination.
+Given that our problem can be reasonably approximated by this basis, we can achieve good results.
+However, by limiting ourselves to linear functions, we are possibly limiting our model, as our dataset may well be better approximated by a different basis, our perhaps not even by a linear combination of functions at all.
+Because of this, linear regression is not able to understand the interaction between to arbitrary input variables \cite[165]{gbc}.
+A better approach would be to simply feed our model our features directly, and hope that it may somehow learn the relationship organically.
+
+To learn any arbitrary relationship between our features may at first seem like an intractable problem, but we are actually able to achieve this with relatively small changes to our linear regression methods.
+Instead of having our prediction vector be a linear combination of our input vector, we use the linear combinations of the inputs to create an intermediate vector, and apply an activation function to this vector.
+The resulting values from this function are either passed to the output as a linear combination again, or sent to another intermediate vector.
+By having the activation function be a nonlinear function, such as the sigmoid function, we are able to approximate any continous function to arbitrary accuracy \cite[230]{cmb}.
+By introducing these intermediate vectors, we have created a layered structure, where information is fed forward from layer to layer.
+Because of this, we call this model a Multilayer Perceptron or a Feedforward Neural Network.
+
+Mathematically, each node on a layer is a linear combination of all the nodes of the previous layer, fed into the activation function for the current layer.
 
 \begin{equation}
-	C_{mse}(\tilde{\bm{y}}) = \sum_{i=0}^{n-1}(y_i-\tilde{y_i})^2
+a_i^l = f^l(z_i^l) = f^l\left(\sum_{j=1}^{N_{l-1}} w_{ij}^l a_j^{l-1} + b_i^l\right).
 \end{equation}
 
-We may now plug in the definition of $\bm{\tilde{y}}$.
+% For an FFNN with an arbitrary amount of hidden layers, the output layer is thus defined as
+%
+% \begin{equation}
+% a^{l+1}_i = f^{l+1}\left[\!\sum_{j=1}^{N_l} w_{ij}^3 f^l\left(\sum_{k=1}^{N_{l-1}}w_{jk}^{l-1}\left(\dots f^1\left(\sum_{n=1}^{N_0} w_{mn}^1 x_n+ b_m^1\right)\dots\right)+b_k^2\right)+b_1^3\right].
+% \end{equation}
+%
+% \begin{equation}
+% 	a^{L}_i = f^{L}\left[\!\sum_{j=1}^{N_{L-1}} w_{ij}^3 f^{L-1}\left(\sum_{k=1}^{N_{L-2}}w_{jk}^{L-2}\left(\dots f^1\left(\sum_{n=1}^{N_0} w_{mn}^1 x_n+ b_m^1\right)\dots\right)+b_k^2\right)+b_1^3\right].
+% \end{equation}
+
+\subsubsection{Activation functions}
+
+As we have seen, the behaviour of our neural network is defined by the amount of hidden layers and the amount of nodes in each one (our architecture), but also by our choice of activation functions.
+For the output layer, the activation is usually easily decided based on the problem (identity for regression problems, sigmoid for binary classification, SoftMax for multiclass).
+For the hidden layers, the choice of activation function is often decided through a process of trial and error, as this is an area without definitive guiding theoretical principles \cite[188]{gbc}.
+Below are the definitions of some of the most common activation functions.
+
+The sigmoid is defined as follows:
 
 \begin{equation}
-	C_{mse}(\bm{\beta}) = \sum_{i=0}^{n-1}(y_i-X_{i*}\bm{\beta})^2
+y = \frac{1}{1 + exp(x)}
 \end{equation}
 
-We see that we have a function of our parameters $\bm{\beta}$, which describe how much our prediction deviates from our target values.
-Our problem now is to optimize this function, that is, find the values of $\beta$ which gives the MSE the lowest value.
-Luckily for us, it can be proven that the second derivative of this function is always positive \cite[106]{linreg}, which means that we can find the global minimum by solving the equation
+As we can see from the equation, this function has a range of $y \in (0,1)$, and for high values of x, the rate of change is very small.
+This saturation is a problem for gradient descent methods, and especially for backpropagation, as the repeated multiplication of very small gradients can lead to weights barely being updated and training slowing down (or even stopping completely).
+Because of this saturation, the sigmoid is generally not used as an activation function for the hidden layers \cite[191]{gbc}.
+
+Another activation function is the Rectified Linear Unit, and its cousin the Leaky Rectified Linear Unit:
+
+\begin{equation}
+RELU(x) = \left\{\begin{array}{cc} 0 & \; \; if \; \; x < 0, \\  x & \; \; if \; \; x \ge 0.\end{array}\right.
+\end{equation}
+
+\begin{equation}
+LRELU(x) = \left\{\begin{array}{cc} -\alpha \cdot x & \; \; if \; \; x < 0, \\  x & \; \; if \; \; x \ge 0.\end{array}\right.
+\end{equation}
+
+These do not suffer the same saturation problems displayed by the sigmoid function, and as such are often the default choice of activation for the hidden layers \cite[188]{gbc}.
+
+\subsubsection{Backpropagation}
+
+Like logistic regression, the optimal weights for each layer in an FFNN can not be found analytically, and we have to turn to gradient descent.
+By repeated use of the chain rule, we are able to calculate the gradient for each layer iteratively, using the gradient of the layer after it and starting at the output.
+This is simply the reverse mode of automatic differentiation, and is computationally preferred compared to forward mode, where we start at the input layer \cite[416]{sr}.
+The equations for calculating the gradients in an FFNN, which we will not derive here, are as follows \cite{morten}:
+
+First we calculate the so called delta terms for all layers, starting with the output layer:
+
+\begin{equation}
+y_i^l = f^l(u_i^l) = f^l\left(\sum_{j=1}^{N_{l-1}} w_{ij}^l y_j^{l-1} + b_i^l\right).
+\end{equation}
+
+We use this term in an iterative process, working backwards through the network with the following equation:
+
+\begin{equation}
+\delta_j^l = \sum_k \delta_k^{l+1}w_{kj}^{l+1}f'(z_j^l).
+\end{equation}
+
+Using these delta terms, we can calculate the gradients for the weights in each layer:
 
 \begin{gather}
-\frac{\delta}{\delta \beta} \sum_{i=0}^{n-1}(y_i-X_{i*}\bm{\beta})^2 = 0 \\
- - \frac{2}{n} X^T(\bm{y} - X\bm{\beta}) = 0
+w_{jk}^l\leftarrow  = w_{jk}^l- \eta \delta_j^la_k^{l-1} \\
+b_j^l \leftarrow b_j^l-\eta \frac{\partial {\cal C}}{\partial b_j^l}=b_j^l-\eta \delta_j^l
 \end{gather}
-
-Solving this for $\beta$ gives us the so called normal equation, an analytical solution\footnote{This solution requires that $X^TX$ is non-singular, which may not be the case.
-This is solved by using the singular value decomposition of the design matrix, which we will not go into here} for the optimal $\hat{\beta}$.
-
-% \begin{equation}
-\begin{gather}
-- \frac{2}{n} X^T(\bm{y} - X\bm{\beta}) = 0 \\
-X^T(\bm{y} - X\bm{\beta}) = 0 \\
-X^T\bm{y} - X^TX\bm{\beta} = 0 \\
-X^T\bm{y} = X^TX\bm{\beta} \\
-\bm{\hat{\beta}} = (X^TX)^{-1}X^T\bm{y}
-\end{gather}
-
-The process we have just described, with MSE as the cost function, is what is known as Ordinary Least Squares (OLS) regression.
-Using different cost functions when fitting the model leads to different equations for $\bm{\hat{\beta}}$.
-This leads us to the two next forms of linear regression we use in this paper.
-But before this, a...
-
-\subsubsection{Brief statistical intermission}
-
-We must remember that our target values are defined as $y_i = f + \epsilon_i$, that is, they are distorted by a stochastic noise $\epsilon_i \sim N(0,\sigma^2)$.
-We should therefore take a brief intermission to prove some of the statistical qualities of the method we have just described, so as to reassure ourselves that we are justified in applying linear regression to noisy data.
-We will start by showing that
-
-\begin{equation}
-E[y_i] = \sum_j X_{ij}\beta_j
-\end{equation}
-
-The proof is as follows:
-
-$\bm{y}$ is defined as the $f(\bm{x}) + \bm{\epsilon}$.
-We make an assumption that this can be approximated as a matrix multiplication $X\bm{\beta}$ plus an error vector $\bm{\epsilon}$.
-This means that each element of the vector
-$\bm{y}$ can be expressed as follows:
-
-\begin{equation}
-y_i = \sum_j x_{ij}\beta_j + \epsilon_i
-\end{equation}
-
-If we take the expectation value of this expression we get the following:
-
-\begin{gather}
-E[y_i] = E\left[\sum_j x_{ij}\beta_j + \epsilon_i\right] \\
-E[y_i] = E\left[\sum_j x_{ij}\beta_j\right] + E[\epsilon_i]
-\end{gather}
-
-However, the elements of $X$ are not stochastic, and neither are the elements of $\beta$ and therefore the first expectation value is simply the sum itself.
-Furthermore, $\epsilon$ is explicitly defined as a normal distribution $N(0,\sigma^2)$, and will by definition have the expectation value $0$.
-Therefore, we end up with the final expression:
-
-\begin{equation}
-E[y_i] = \sum_j x_{ij}\beta_j = \bm{X}_{i,*}\bm{\beta}
-\end{equation}
-
-Next, we show that the variance of $y_i$ is equal to the variance of the noise:
-
-\begin{gather}
-var[y_i] = E\left[(y_i - E[y_i])^2\right]  \\
-= E\left[y_i^2 - 2E[y_i]y_i + E[y_i]^2\right]
-\end{gather}
-
-Distributing the outer expectation value function:
-
-\begin{gather}
-var[y_i] = E[y_i^2] - 2E\left[E[y_i]\right]E[y_i] + E\left[E[y]^2\right]  \\
-= E[y_i^2] - 2E\left[E[y_i]\right]E[y_i] + E\left[E[y]\right]E\left[E[y]\right]
-\end{gather}
-
-The result of calculating the expectation value is non-stochastic. This means that $E[E[X]] = E[X]$. From this it follows that
-
-\begin{gather}
-var[y_i] = E[y_i^2] - 2E[y_i]E[y_i] + E[y_i]E[y_i]  \\
-= E[y_i^2] - E[y_i]E[y_i]  \\
-= E[y_i^2] - E[y_i]^2
-\end{gather}
-
-The expectation value in the second summand has been proven to be equal to $\bm{X}_{i,*}\bm{\beta}$ above. We therefore have
-
-\begin{gather}
-var[y_i] = E[y_i^2] - (\bm{X}_{i,*}\bm{\beta})^2  \\
-= E\left[(X_{i,*}\bm{\beta})^2 + X_{i,*}\bm{\beta}\epsilon_i + \epsilon_i^2\right] - (\bm{X}_{i,*}\bm{\beta})^2  \\
-= E\left[(X_{i,*}\bm{\beta})^2\right] + E[X_{i,*}\bm{\beta}\epsilon_i] + E[\epsilon_i^2] - (\bm{X}_{i,*}\bm{\beta})^2
-\end{gather}
-
-$X_{i,*}\bm{\beta}$ and $\epsilon_i$ are both scalars.
-Therefore the expectation value can be written as $E[X_{i,*}\bm{\beta}]E[\epsilon_i]$.
-However, $E[\epsilon_i]$ is by definition $0$, because $\epsilon_i$ is defined as a normal distribution of mean $0$ and variance $\sigma^2$.
-Furthermore, the expectation value of the non-stochastic $(X_{i,*}\bm{\beta})^2$ is simply the expression itself.
-We therefore have
-
-\begin{gather}
-var[y_i] = (X_{i,*}\bm{\beta})^2 + E[X_{i,*}\bm{\beta}]\cdot 0 + E[\epsilon_i^2] - (\bm{X}_{i,*}\bm{\beta})^2  \\
-= E[\epsilon_i^2]
-\end{gather}
-
-We can prove that this is equal to the variance of $\epsilon_i$:
-
-\begin{gather}
-E[\epsilon_i^2] = \frac{1}{n} \sum_i \epsilon_i^2 \\
-var[\epsilon_i] = \frac{1}{n} \sum_i (\epsilon_i - \bar{\epsilon_i})^2 \\
-var[\epsilon_i] = \frac{1}{n} \sum_i (\epsilon_i - 0)^2 \\
-var[\epsilon_i] = \frac{1}{n} \sum_i \epsilon_i^2 = E[\epsilon_i^2]
-\end{gather}
-
-And of course, we know that the variance of $\epsilon_i$ by definition is $\sigma^2$.
-
-\begin{gather}
-var[y_i] = E[\epsilon_i^2] = var[\epsilon_i] = \sigma^2 
-\end{gather}
-
-We should also prove some qualities about or parameters. We start by proving that $E[\hat{\beta}] = \beta$
-
-Because we are assuming $(X^{T}X)^{-1}X^{T}$ to be deterministic, $E[(X^{T}X)^{-1}X^{T}] = (X^{T}X)^{-1}X^{T}$. Moreover, we know $E[y] = y$, and we can thus rewrite the equation as follows:
-
-\begin{equation}
-E[(X^{T}X)^{-1}X^{T}y] = (X^{T}X)^{-1}X^{T}y = \beta
-\end{equation}
-
-Next, we prove that $var(\beta) = \sigma^{2}(X^{T}X)^{-1}$.
-
-$$var(\beta) = var((X^{T}X)^{-1}X^{T}y)$$
-
-Because we are assuming $(X^{T}X)^{-1}X^{T}$ to be deterministic and $y$ to be stochastic, we can rewrite the $var(\beta)$ as following:
-
-\begin{gather}
-var(\beta) = (X^{T}X)^{-1}X^{T}var(y)((X^{T}X)^{-1}X^{T})^{T} \\
-= (X^{T}X)^{-1}X^{T}var(y)(X^{T})^{T}((X^{T}X)^{-1})^{T}
-\end{gather}
-
-Since $(X^{T})^{T} = X$ and $(X^{T}X)^{-1})^{T}$ = $(X^{T}X)^{T})^{-1} = (X^{T}X)^{-1}$
-
-\begin{gather}
-= (X^{T}X)^{-1}X^{T}var(y)X(X^{T}X)^{-1}
-\end{gather}
-
-We know var(y) = $\sigma^{2}$, and since it is a scalar it is commutative. Thus we may move it freely
-
-\begin{gather}
-= \sigma^{2}(X^{T}X)^{-1}X^{T}X(X^{T}X)^{-1} \\
-= \sigma^{2}(X^{T}X)^{-1}
-\end{gather}
-
-\subsubsection{Ridge Regression}
-
-A common problem with OLS is what is known as overfitting (we will discuss this in greater detail in later sections).
-A symptom of overfitting is that the values in the parameter vector $\bm{\beta}$ grow very large, and vary wildly for small changes in our dataset.
-To counteract this, we can add to the cost function a term which increases with the values in $\bm{\beta}$.
-Optimizing the cost function would now not only mean finding the smallest difference between target and predicted values, but also minimizing the absolute value of the parameters.
-We now define a new cost function
-
-\begin{equation}
-	C_{lasso}(\bm{\beta}) = \sum_{i=0}^{n-1}(y_i-X_{i*}\bm{\beta})^2 + \lambda \sum_{j=0}^{p-1} \beta_j^2
-\end{equation}
-
-$||\beta||_2^2$ is equal to $\beta^T\beta$ and is simply the norm of the parameter vector, squared.
-With the new hyperparameter $\lambda \in (0, \inf)$ , we can decide how harshly we should punish large values in our parameter vector.
-Finding an optimal $\lambda$ is an important step in the process of finding the best parameters for our model.
-
-A similar derivation as done in the previous section gives us the normal equation for Ridge regression:
-
-\begin{equation}
-	\bm{\hat{\beta}} = (X^TX + \lambda I)^{-1}X^T\bm{y}
-\end{equation}
-
-We see that this equation, with a $\lambda > 0$ avoids the problem of an inversion of a non-singular matrix, and gives us an analytical solution for $\bm{\hat{\beta}}$ yet again.\footnote{We should cherish this analytical solution, as it may well be the last time we see a cost function that can be optimized analytically.} By using ridge regresison, we can control the complexity of our model and effectively hinder overfitting.
-
-\subsubsection{Lasso Regression}
-
-Another cost function is one where we add the L1-norm (the sum of the absolute values of the parameters), instead of the L2-norm. The new cost function is as follows.
-
-\begin{equation}
-	C_{lasso}(\bm{\beta}) = \sum_{i=0}^{n-1}(y_i-X_{i*}\bm{\beta})^2 + \lambda \sum_{j=0}^{p-1} |\beta_j|
-\end{equation}
-
-However, the derivative of the absolute value is not continous, so there is no analytical solution for the optimal parameters with this form of linear regression.
-We must then solve the problem numerically, which can be done by using Gradient Descent.
-Like Ridge regression, this form of regression controls the complexity of the model, with the added benefit that Lasso can drive parameters to exactly zero.
-
-\subsubsection{Resampling methods: The bootstrap}
-
-Bootstrap is a resampling method divided into simple steps.
-First, we reshuffle our dataset with replacement, which will provide us with a new dataset.
-Note that we will already have split our data into train and test data before we begin the bootstrap, so when we reshuffle, we will only reshuffle the training data.
-The reshuffled datasets act as our samples.
-
-Secondly, we will perform our linear regression on this new dataset and store the predictions of the test data.
-Thirdly, we repeat for a given large number of iterations.
-Since we use all the test predictions to calculate the MSE, bias and variance, we will then according to the central limit theorem \cite{CLT} obtain a normal distribution of these performance metrics.
-The law of large numbers \cite{LLN} then tells us that the mean sample values of the MSE, bias and variance approach the true value of MSE, bias and variance.
-This is dependent on our assumption that our data is independently and identically distributed.
-
-These approximations of the true value of MSE, bias and variance will become useful when performing our bias-variance trade-off analysis.
-
-\subsubsection{Cross validation}
-
-Cross validation is another resampling technique, in which we reshuffle our data and then split them into k-folds, where k specifies the number of folds.
-One fold is used as our test data, and the rest is used as training data.
-We run our linear regression, then switch up which fold is used for testing until all folds have been used.
-As in bootstrapping, we will approach the true values of for example the mean squared error for our model.
-Theoretically, because both two resampling techniques use the central limit theorem and the law of large numbers, they should return the same values for the MSE, bias and variance given the same dataset.
-Though resampling techniques can be computationally expensive, they make up for it by aiding in model evaluation by supplying us with performance metrics obtained from a larger number of samples.
-
-\subsubsection{Bias-variance trade-off}
-
-A crucial part of model evaluation is ensuring that the model is neither underfitted nor overfitted, which we can determine by observing the relationship between the bias and the variance.
-Bias is defined as $(Bias[\tilde{y}])^2 = (y-E[\tilde{y}])^2$, the difference between the expected value of our prediction and the observed value.
-Models with high bias are usually underfitted, whilst models with low bias are the opposite.
-Variance on the other hand is defined as $var[\tilde{f}] = \frac{1}{n}\sum(\tilde{y_i}-E[\tilde{y}])^2$, how far away predictions are from the mean prediction.
-Models with high variance are usually overfitted to the training data, whilst low variance models are the opposite.
-
-A good rule of thumb for finding a good fit is looking at the intercept between bias and variance.
-This is because the cost function can be decomposed into these two elements.
-The proof for this is as follows:
-
-Given data $\bm{y} = f(\bm{x}) + \bm{\epsilon}$ and a OLS model $\bm{\tilde{y}} = X\bm{\beta}$, we define
-the Mean Squared Error as follows
-
-\begin{gather}
-C(X,\beta) = \frac{1}{n}\sum_{i=0}^{n-1}(y_i - \tilde{y_i})^2
-\end{gather}
-
-This is simply the expectation value of $(\bm{y} - \bm{\tilde{y}})^2$
-
-\begin{gather}
-C(X,\beta) = E\left[(\bm{y} - \bm{\tilde{y}})^2\right]
-\end{gather}
-
-To derive the bias variance decomposition, we now add $E[\bm{\tilde{y}}] - E[\bm{\tilde{y}}]$ 
-as a sneaky trick. We will also replace $\bm{y}$ with our definition $\bm{y} = \bm{f} + \bm{\epsilon}$
-
-\begin{gather}
-C(X,\beta) = E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}] + E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2\right]
-\end{gather}
-%%%%%%%%%%%%%%%%%5
-\begin{multline}
-= E[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2 + 2[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])(E[\bm{\tilde{y}}] \\
-	- \bm{\tilde{y}})] +(E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2]
-\end{multline}
-
-\begin{multline}
-= E[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2] + \\
-	2E[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])(E[\bm{\tilde{y}}] - \bm{\tilde{y}})] + E[(E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2]
-\end{multline}
-
-Let us focus our attention on the middle term first.
-
-\begin{gather}
-2E\left[(\bm{f} + \bm{\epsilon} + E[\bm{\tilde{y}}])(E[\bm{\tilde{y}}] - \bm{\tilde{y}})\right] 
-\end{gather}
-
-Expanding the multiplication:
-
-\begin{gather}
-2E\left[\bm{f}E[\bm{\tilde{y}}] + \bm{\epsilon}E[\bm{\tilde{y}}] - E[\bm{\tilde{y}}]E[\bm{\tilde{y}}] - \bm{f}\bm{\tilde{y}} - \bm{\epsilon}\bm{\tilde{y}} + E[\bm{\tilde{y}}]\bm{\tilde{y}}\right] 
-\end{gather}
-
-Distributing the expectation value:
-
-\begin{gather}
-2\left[E[\bm{f}E[\bm{\tilde{y}}]] + E[\bm{\epsilon}E[\bm{\tilde{y}}]] - E[E[\bm{\tilde{y}}]E[\bm{\tilde{y}}]] - E[\bm{f}\bm{\tilde{y}}] - E[\bm{\epsilon}\bm{\tilde{y}}] + E[E[\bm{\tilde{y}}]\bm{\tilde{y}}]\right]
-\end{gather}
-
-We know that $\bm{\tilde{y}}$ and $\bm{f}$ are non-stochastic, and we know that the expectation value of $\bm{\epsilon}$ is $0$. We also know that 
-$\bm{\epsilon}$ and $\bm{\tilde{y}}$ are independent, which means we can write $E[\bm{\epsilon}\bm{\tilde{y}}]$ as $E[\bm{\epsilon}]E[\bm{\tilde{y}}]$ We therefore have
-
-\begin{gather}
-2\left[\bm{f}\bm{\tilde{y}} + 0 - \bm{\tilde{y}}\bm{\tilde{y}} - \bm{f}\bm{\tilde{y}} - 0 + \bm{\tilde{y}}\bm{\tilde{y}}\right] = 0 
-\end{gather}
-
-We then have
-
-\begin{gather}
-C(X,\beta) = E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2\right] + E\left[(E[\bm{\tilde{y}}] - \bm{\tilde{y}})^2\right] 
-\end{gather}
-
-The second term is by definition the variance of model.
-
-\begin{gather}
-C(X,\beta) = E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2\right] + var[\bm{\tilde{f}}] 
-\end{gather}
-
-And now the final term
-
-\begin{gather}
-E\left[(\bm{f} + \bm{\epsilon} - E[\bm{\tilde{y}}])^2\right]
-\end{gather}
-
-\begin{multline}
-= E[\bm{f}^2 + \bm{f}\bm{\epsilon} - \bm{f}E[\bm{\tilde{y}}] + \bm{f\epsilon} + \bm{\epsilon}^2 - \\ 
-	\bm{\epsilon}E[\bm{\tilde{y}}] - \bm{f}E[\bm{\tilde{y}}] - \bm{\epsilon}E[\bm{\tilde{y}}] + E[\bm{\tilde{y}}]^2]
-\end{multline}
-
-After distributing the expectation value we get the following, keeping in mind that $E[\bm{f}] = E[\bm{y} - \bm{\epsilon}] = E[\bm{y}] - E[\bm{\epsilon}] = \bm{y}$
-
-\begin{gather}
-\bm{y}^2 - 2\bm{y}\bm{\tilde{y}} + E[\bm{\tilde{y}}]^2 + \sigma^2 \\
-(\bm{y} + E[\bm{\tilde{y}}]^2 + \sigma^2 
-\end{gather}
-
-We see that the quadratic form is equal to $bias[\bm{\tilde{y}}]^2$. We can then conclude that the cost function can be rewritten as follows:
-
-\begin{gather}
-C(X,\beta) = bias[\bm{\tilde{y}}]^2 + var[\bm{\tilde{f}}] + \sigma^2 
-\end{gather}
-
-Because the cost function is a sum of the bias and the variance, the lowest error will generally occur when both bias and variance are low.
-This also fits our understanding of the definitions of bias and variance, as we wish to obtain a model that is neither high in bias (underfitted) or high in variance (overfitted).
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Implementation}
 
-\subsection{Implementation}
-The source code for our implementation can be found here: \href{url}{https://github.com/jonatan-hanssen/3155-project1.git}.
-\subsubsection{Preprocessing: Creating the design matrix and scaling our data}
-We have two datasets, one synthetic and one real.
-Regardless of the source of the data, we create the design matrix in a similar manner, as the relation between our input data and our target values are the same for both datasets, that is, for each combination of $x$ and $y$, we have a height $z$.
-For our real data, this value represents the elevation in meters above ocean level for a terrain in Norway.
-As we wish to perform a polynomial fit, we create our design matrix such that each row $i$ has the following values:
+We implement the FFNN as a class, using NumPy's many y
 
-\begin{equation}
-X_{i*} = 1 \;\; x_i \;\; y_i \; x_i^2 \;\; x_i y_i \;\; y_i^2 \;\; ... \;\; x_i^p y_i^p
-\end{equation}
-
-Below is the python code used to create the design matrix, which can be found in \texttt{src/utils.py}.
-
-\lstinputlisting[language=Python, firstline=30, lastline=44]{../src/utils.py}
-
-While the creating of the design matrix is equivalent for the two datasets, the preprocessing of the data afterwards is not.
-For the Franke function no scaling is needed, as we generate the data ourselves to be between $0$ and $1$.
-Given that we are doing a polynomial fit, every element in our design matrix will be a multiplication of $x^p$ and $y^k$ for different $p$'s and $k$'s, which, given our choice of range for $x$ and $y$, guarantees that every element is between $0$ and $1$.
-However, we have included a centering of the data, which has no effect on the Ordinary Least Squares regression, but which is used when applying Ridge and Lasso regression, to avoid punishing the intercept.
-The centering is done as follows: First, the first coloumn of the design matrix (the intercept coloumn) is discarded.
-Then we calculate the mean for the target values and the design matrix (the mean for the design matrix is a vector with mean values for each coloumn).
-We subtract this from the design matrix and the target values when performing the fit giving us a vector $\bm{\beta}$.
-Afterwards we calculate the intercept with the following equation
-
-\begin{equation}
-	intercept = \bar{y} - \bm{\bar{x}_c}^T \bm{\beta}
-\end{equation}
-
-	where $\bar{y}$ is the mean value of $\bm{y}$ and $\bm{\bar{x}_c}$ is a vector of size $p$ with $\bar{X}_i = mean(X_i)$ (the mean of each coloumn in $X$).
-We finally create our prediction by using the $\bm{\beta}$, $X$ and intercept we calculated above:
-
-\begin{equation}
-	\bm{\hat{y}} = X \bm{\beta} + intercept
-\end{equation}
-
-The code for this, taken from the \texttt{evaluate\_model} function in \texttt{src/utils.py} is as follows:\footnote{Some intermediate parts of our implementation not related to this have been removed for readability}
-
-\begin{lstlisting}[language=Python]
-X_train = X_train[:, 1:]
-z_train_mean = np.mean(z_train, axis=0)
-X_train_mean = np.mean(X_train, axis=0)
-beta = model((X_train - X_train_mean), (z_train - z_train_mean))
-intercept = np.mean(z_train_mean - X_train_mean @ beta)
-z_pred_train = X_train @ beta + intercept
-\end{lstlisting}
-
-Where \texttt{model} calculates \texttt{beta} using the normal equations for OLS.
-
-For the real elevation data, scaling is definitely required, as the values for $x$, $y$ reach several hundred and $z$ reaches over a thousand.
-For polynomial degree $10$ and an $x$ value which reaches $400$, we will have a design matrix with values from $0$ to $400^{10}$.
-This large range of values makes our linear regression perform poorly.
-We have therefore chosen to scale our design matrix using Scikit's \texttt{MinMaxScaler}, which scales each column to be in the range of $(0,1)$.
-The scaling coefficients are calculated on the training data, and applied to both the training and test data.
-We have done the same with the values for $z$.
-This gives a much more stable model.
-The code for doing this on the design matrix is as follows, taken from \texttt{src/utils.py}:
-
-\lstinputlisting[language=Python, firstline=242, lastline=249]{../src/utils.py}
-
-Another form of "preprocessing" which is strictly practical in nature, is that we have reduced the dimensions of the data to 320 by 640.
-This was done by downscaling the image.\footnote{The interpolation was done using "nearest neighbor interpolation", which does not introduce new values into the data \cite[77]{dip}.
-An equivalent programmatic solution would be to for instance only look at every 4 pixels} This was simply done because the design matrix becomes several gigabytes in size if one uses the whole dataset, and model fitting then takes an unreasonable amount of time.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Implementation of linear regression}
-
-Linear regression for OLS and Ridge are implemented by using NumPy's \texttt{numpy.linalg.pinv} and the normal equations.
-We have implemented them as functions which take in the target values and the design matrix and return $\bm{\hat{\beta}}$.
-They are called from a wrapper function called \texttt{evaluate\_model} which takes care of the scaling and works with both SciKit models and our own.
-Below are the functions, taken from \texttt{src/utils.py}:
-
-\lstinputlisting[language=Python, firstline=56, lastline=64]{../src/utils.py}
-
-For repeated calls of these functions with different polynomial degrees, we use the function \texttt{linreg\_to\_N}, which passes the correctly sliced design matrix to the regression functions and saves the predictions and various performance metrics for each polynomial degree for easy plotting.
-
-For Lasso regression, we use SciKit's implementation. We also use SciKit's implementation for OLS and Ridge for comparisons with our own implementation.
-
-The correctness of our Ordinary Least Squares regression can be confirmed by running our debug function (the so called SkrankeFunction)\footnote{this can be done by running \texttt{\$ python3 task\_b.py --debug --noscale -n 2}}, which is a second degree polynomial where $b_0 = 0, \; b_1 = 1, \; b_2 = 2$ and so on.
-The SkrankeFunction (taken from \texttt{src/utils.py}) looks like this
-
-\lstinputlisting[language=Python, firstline=26, lastline=27]{../src/utils.py}
-
-and the result of our regression is correctly given as $\bm{\beta} = (0 \; 1 \; 2 \; 3 \; 4 \; 5\;)^T$.
-
-\subsubsection{Implementation of resampling methods}
-
-We use resampling methods to better estimate our performance metrics.
-We use both cross validation and the bootstrap method.
-Cross validation is implemented in the function \texttt{crossval} in \texttt{src/utils.py}.
-We split the data into $K$ folds by slicing with NumPy taking consecutive rows of the design matrix and target vector.
-But before we do this, we resample the data with Scikit's \texttt{resample}.
-This is important because our $x$'s and $y$'s are not normal distributions but consecutive values between $0$ and $1$, and therefore consecutive rows correspond specific areas of our data, which is not wanted.
-We then fit on $K-1$ folds, and test on the remaining fold.
-We return the average error.
-
-We implement bootstrap by repeatedly resampling our data with replacement and performing the prediction.
-We use the predictions returned from hundreds of resamplings to calculate the bias and variance of our model.
-
-\subsubsection{Franke Function}
-
-We use the Franke Function with an added stochastic noise of to test our models. We explore how higher and lower levels of noise affect the model. Below is the Franke function:
-
-\begin{multline}
-f(x,y) = \frac{3}{4}\exp{\left(-\frac{(9x-2)^2}{4}-\frac{(9y-2)^2}{4}\right)} + \\ \frac{3}{4}\exp{\left(-\frac{(9x+1)^2}{49}-\frac{(9y+1)}{10}\right)} + \\
-\frac{1}{2}\exp{\left(-\frac{(9x-7)^2}{4}-\frac{(9y-3)^2}{4}\right)} - \\
-\frac{1}{5}\exp{\left(-(9x-4)^2-(9y-7)^2\right)}
-\end{multline}
-
-As we can see, it is simply a weighted sum of exponentials. As such it has an infinite Taylor expansion into polynomials.\footnote{The proof for this is left as an exercise for the reader}
+We set the bias to 0.1 as recommended in Goodfellow et Al. \cite[189]{gbc}.
 
@@
 
-\subsection{Resampling}
-
-As stated in the method section, we suspect that bootstrap and cross validation both yield approximations to the true performance metrics due to the central limit theorem, and thus should give identical results.
-We can observe from Fig.~\ref{frankemsebyresamplingmethod} that our implementation of bootstrap and cross validation yield almost exactly the same results.
-(Fig.~\ref{realmsebyresamplingmethod} in the appendix shows a similar result for the real data)
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=3.25in]{figure/frankemsebyresamplingmethod.png}}
-\caption{Comparison between approximated MSE for different resampling methods for the Franke function}
-\label{frankemsebyresamplingmethod}
-\end{figure}
-
-\subsection{Franke Function}
+\subsection{Logistic Regression}
 
@@
 
-We find that OLS creates a good fit for the Franke Function.
-With 400 datapoints, we can increase our polynomial degree to 20 before we see signs of overfitting.\footnote{Here we also see something a bit surprising: We see that our own implementation of OLS follows Scikit's implementation up to around polynomial degree 10, at which point scikit's MSE increases rapidly.
-We chalk this up to numerical errors and differing implementation of the normal equations, as we simply use the normal equation exactly, while scikit uses an implementation which is probably computationally more efficient but gives slight differences when comparing to the normal equation.
-These differences seem to create a large error after a certain size of the design matrix is reached, which we can not adequately explain.} This is most likely because the Franke function can be expanded into a sum of polynomials.
-We would expect to see higher polynomial degrees give us better and better prediction, up to a certain point.\footnote{And indeed, this seems to be the case, as if we increase the availiable datapoints considerably, as seen in Fig.~\ref{frankiehigherdatapoints} in the appendix, we can reach much higher polynomial degrees without overfitting} When we reach degrees higher than around 20, we see that our test error increases slightly, a sign of overfitting.
-This is also to be expected, as even if the continous Franke function could be expanded into a series of polynomials, we are working with a limited dataset of only 400 samples (and only 80\% of those in our training set).
-As we reach higher and higher degrees, we are more and more likely to see unfortunate train-test-splits where the best fit for the training data leads to large errors for the test data.
-Regardless, our fit for lower model complexity is remarkably good.
-See Fig.~\ref{frankenonoise} for a comparison between the Franke function and our best prediction, which shows that they are essentially identical.
-Fig.~\ref{frankenonoisemse} shows the train and test MSE by polynomial degree.
-The MSE for test in these plots is the average MSE for 100 bootstraps.
-For a comprehensive list of the shell command required to reproduce this and every other plot, see Table~\ref{allparamstable} in the appendix.
+\subsubsection{Ridge}
 
 \begin{figure}[h]
-\centerline{\includegraphics[width=3.25in]{figure/frankenonoise.png}}
-\caption{A comparison of our best OLS prediction and the Franke function with no noise, at polynomial degree 10}
-\label{frankenonoise}
-\end{figure}
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=3.25in]{figure/frankenonoisemse.png}}
-\caption{Test and train MSE for the Franke function with no noise, using OLS}
-\label{frankenonoisemse}
-\end{figure}
-
-However, when we introduce a stochastic noise, we see that overfitting happens rapidly.
-Without a regularization hyperparameter, parameters fluctuate wildly and our prediction suffers.
-Fig.~\ref{frankenoisemse} shows our MSE up to polynomial degree 24 with an added stochastic noise $\epsilon \sim N(0,0.05)$, where we see that we are overfitting severely starting at around degree 7.
-See Fig.~\ref{frankienoisecomparison} in the appendix for test MSE for more and more noisy target values.
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=3.25in]{figure/frankenoisemse.png}}
-\caption{Test and train MSE for the Franke function with noise, using OLS}
-\label{frankenoisemse}
-\end{figure}
-
-By using bootstrap resampling, we can study the bias-variance-tradeoff, which shows how the variance of our model between different bootstraps iterations increases as our model overfits our noisy data.
-See Fig.~\ref{frankiebiasvariance}.
-Clearly, we need to introduce regularization if we want a model that does not overfit.
-Fig.
-~\ref{frankenoisybeta} in the appendix shows the beta progression for selected betas, clearly showing that we are overfitting.
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=3.25in]{figure/frankiebiasvariance.png}}
-\caption{Bias, variance and test MSE plotted over polynomial degree for our OLS prediction on the Franke function with an added stochastic noise $\epsilon \sim N(0,0.05)$}
-\label{frankiebiasvariance}
-\end{figure}
-
-Table ~\ref{ols_mse_table} shows the best polynomial degree and MSE for regular and noisy data.
-Here we see that the best degree without noise is 9, which shows that even without noise, the low amount of data leads to a slight overfit even at very small polynomial degrees.
-This is not surprising, as at $N = 9$ we have $55$ features, which is a substantial amount compared to our measly 320 training datapoints.
-We could of course arbitrarily increase our dataset to ludicrous levels, but doing linear regression with this dataset gives valuable insight into how our performance is dependent on our availiable data.
-
-\begin{table}[h]
-\caption{Best MSE and polynomial degree for OLS}
-\begin{center}
-\label{ols_mse_table}
-\begin{tabular}{c l l}
-% & & \\ % put some space after the caption
-% \hline
-Std of noise & Best MSE & Best degree \\
-\hline
-0 & 0.0005109 & 9\\
-0.05 & 0.0038302 & 7 \\
-0.10 & 0.0139833 & 7 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsubsection{Ridge}
-
-Using Ridge regression, we see from Fig.~\ref{frankiedifflambdas} in the appendix that larger and larger values of $\lambda$ efficiently prevents overfitting and reduces the variance between predictions.
-Looking at the beta progression for the same values of $\lambda$ (Fig.~\ref{frankeridgebetas} in the appendix), we see that the regularization parameter is successfully shrinking the beta values.
-Using Scikit's gridsearch, we find the optimal $\lambda$ to be $6.10 \cdot 10^{-8}$ and the optimal polynomial degree to be 13, with an average MSE of $0.003027$.
-This is a good improvement of our OLS fit, and shows that we are able to push our model complexity significantly higher and gain a moderate MSE improvement with a correctly chosen regularization parameter.
-Looking at the beta progression for this value of lambda (see Fig.~\ref{frankebestlambdabetas}), we see that the beta values settle and do not vary greatly between polynomial degrees.
-Fig.~\ref{frankebestlambdamse} shows the MSE score, which shows how we are successfully avoiding overfitting.
-Table~\ref{ols_vs_ridge_vs_lasso_table} shows the relevant values.
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=3.25in]{figure/frankebestlambdabetas.png}}
-\caption{Beta progression for our Ridge prediction with the optimal lambda}
-\label{frankebestlambdabetas}
-\end{figure}
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=3.25in]{figure/frankebestlambdamse.png}}
+\centerline{\includegraphics[width=5in]{figure/frankebestlambdamse.png}}
 \caption{Test and train MSE for ridge}
@@
 
-% We see from this that Ridge regression has not really had much of an effect in decreasing the MSE.
-We must now remember what Ridge aims to do.
-With the regularization parameter, we can prevent overfitting if our model is too complex for our data.
-Looking at Fig.~\ref{frankeridgenoisecomparison} in comparison to Fig.~\ref{frankienoisecomparison} (both in the appendix), we can see that Ridge really has achieved this, reducing overfitting as our data becomes more and more noisy.
-But we should recall that we decide the complexity of our model, by deciding which polynomial degree we use when creating the design matrix.
-In fact, by choosing polynomial degree 7 for our OLS method to get the best MSE for noisy data, we are in fact doing a primitive form of best subset selection, as described in Hastie et Al.\cite[57]{hastie}.
-Almost by definition, the polynomial degree with the best MSE is not overfitted, because if it was, it would not be the model with the lowest MSE among all polynomial degrees.
-Therefore it is not so unexpected that Ridge regression has a relatively low impact on the MSE.
-We could expect that Ridge would allow us to reach higher polynomial degrees without overfitting, by shrinking features of the higher polynomial degrees that are not supported but allowing us to keep other high degree terms (for example, allowing us to use $x^5y^4$, but discard $x^9$, if only $x^5y^4$ was supported by the data).
-But we can see from the Table ~\ref{ols_vs_ridge_vs_lasso_table} that we are not able to push into higher degrees.
-A possible explanation for this is that every feature is useful, due to the polynomial expansion of the Franke function.
-
-\subsubsection{Lasso}
-
-Looking at Lasso regression, we find that it has the same capability to reduce overfitting as Ridge, as can be seen in Fig.~\ref{frankedifflambdaslasso} in the appendix.
-However, looking at Table~\ref{ols_vs_ridge_vs_lasso_table}, we see that Lasso falls short when finding a good MSE.
-Intuitively, it does not make sense for Lasso to be worse than OLS, because Lasso with $\lambda = 0$ is equivalent to OLS.
-But we cannot avoid considering the fact that Lasso does not have an analytical solution, and as such we must use numerical methods to find the optimal parameters.
-Using scikit's implementation, we have not been able to get the numerical method to converge, which means that we have not found a global minimum.\footnote{Scikit explicitly mentions that using $\lambda = 0$ does not converge well with their implementation \cite{lasso}} For this reason, Lasso regression has not given satisfying results.
-
-\begin{table*}[t]
-\caption{Best MSE for OLS, Ridge and Lasso with 400 datapoints}
-\begin{center}
-\label{ols_vs_ridge_vs_lasso_table}
-\begin{tabular}{c | l l l l}
-% & & \\ % put some space after the caption
-% \hline
-Regression & Std of noise & MSE & Degree & $\lambda$ \\
-\hline
-OLS & 0.05 & 0.0038302 & 7 & - \\
-Ridge & 0.05 & 0.003027 & 13 & $6.10 \cdot 10^{-8}$ \\
-Lasso & 0.05 & 0.008298 & 12 & $1.38 \cdot 10^{-5}$ \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
-\subsection{Real elevation data}
-
-With the confidence that our methods work as expected on synthetic data, we turn our attention to the real elevation data.
-First, let us confirm that it is necessary to scale our values using MinMax-scaling, as we have described in our implementation.
-Fig.~\ref{realunscaled} and Fig.~\ref{realunscaledmse} in the appendix shows our predicted terrain and associated MSE, showing that we reach a very poor prediction without scaling.
-
-\subsubsection{Ordinary Least Squares}
-
-Here we find some limitations of linear regression.
-We see from Fig.~\ref{real1msetraintest} that the we are able to fit the terrain up to a certain point, gaining an MSE of $0.0142683$ with OLS with polynomial degree 19, our highest polynomial.
-This is about two orders of magnitude higher than our lowest MSE for the Franke function without noise, which had values on a similar scale.\footnote{The Franke function without noise has a range of about $(0,1.2)$ and our MinMaxed terrain data has been scaled to a range of $(0,1)$.
-While not exactly the same, these are on the same order of magnitude, and we believe that comparing the MSE scores is reasonable} In fact, our MSE is comparable to the Franke function with a stochastic noise $\epsilon \sim N(0,0.1)$ (Table~\ref{ols_mse_table_dif_data}), which implies that our terrain is very irregular and difficult to approximate.
-Looking at the plot for polynomial degree 19 (Fig.~\ref{real1surfacecomp} in the appendix), we see that our poor MSE is reflected visually in our fit as well, as we can only make out general similarities in shape, not finer details.
-Comparing the best polynomial degree of the best fit with the synthetic data, we see that our best fit is reached much later.
-This implies that we are not overfitting yet, and indeed the bias variance decomposition confirms this, as we see that variance is essentially inconsequential (Fig.~\ref{realbiasvarianceOLS}).
-The fact that we are not overfitting can be seen when we increase the polynomial degree to very large amounts, seeing only a very modest decrease in MSE to $0.01381$ at degree 40 (Fig.~\ref{realOLSmse40} in the appendix).
-Comparing this to the tenfold decrease in MSE when increasing the polynomial degree for the Franke function to 50, we see a clear difference.
-Looking at the beta progression for selected betas (Fig.\ref{realOLSbetaprog40} in the appendix), we see that the parameters stabilize, which again implies that we are not overfitting, but simply reaching our best possible prediction for polynomial regression.
-It is reasonable to expect that our model will not overfit, due to the large increase in data we now use for the fit.
-Our synthetic data used $20 x 20 = 400$ datapoints, while we now have $320 x 640 = 204 800$ datapoints.
-This means that for a 19th order polynomial, which will have $210$ features, we will have around a thousand datapoints per feature.
-If we decrease the number of datapoints availiable considerably, we can see overfitting start to occur (Fig.~\ref{real1mselessdata} in the appendix).
-So we are not overfitting, but we are not decreasing our MSE either.
-This is likely because the elevation data has a jagged and complex shape, which cannot be accurately recreated, even with a polynomial of degree $19$.
-We should remember that the Franke function can be expanded into a series of polynomials, which is why we can expect to get a lower and lower MSE as the polynomial degree increases (so long as we have enough data and it is not noisy).
-The same is not true for our elevation data, and as such we have little to gain from increasing our model complexity after a certain point.
-Clearly, linear regression does not offer a very good tradeoff between computational cost and performance for this data.
-
 \begin{table}
-\caption{Best MSE and polynomial degree for OLS on synthetic and real data}
 \begin{center}
@@
 \end{table}
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=3.25in]{figure/realbiasvarianceOLS.png}}
-\caption{Bias, variance and test MSE plotted over polynomial degree for our OLS prediction of Terrain 1}
-\label{realbiasvarianceOLS}
-\end{figure}
-
-% \begin{figure} 
-% \centerline{\includegraphics[width=3.25in]{figure/real1msetraintest.png}}
-% \caption{Test and train MSE for OLS on Terrain 1}
-% \label{real1msetraintest}
-% \end{figure}
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=3.25in]{figure/real1msetraintest.png}}
-\caption{Test and train MSE for OLS on Terrain 1}
-\label{real1msetraintest}
-\end{figure}
 

### 2022-11-07 src/plot_a1.py
for merge

@@ -1,0 +1,97 @@
+from Schedulers import *
+from FFNN import FFNN
+from utils import *
+
+np.random.seed(42069)
+
+(
+    betas_to_plot,
+    N,
+    X,
+    X_train,
+    X_test,
+    z,
+    z_train,
+    z_test,
+    centering,
+    x,
+    y,
+    z,
+) = read_from_cmdline()
+z_train = z_train.reshape(z_train.shape[0], 1)
+z_test = z_test.reshape(z_test.shape[0], 1)
+
+dims = (X.shape[1], 1)
+eta = np.logspace(-5, -1, 5)
+lam = np.logspace(-5, -1, 5)
+lam[0] = 0
+
+epochs = 1000
+neural = FFNN(dims)
+
+loss_heatmap = np.zeros((eta.shape[0], lam.shape[0]))
+for y in range(eta.shape[0]):
+    for x in range(lam.shape[0]):
+        _, test_error, _, _ = neural.fit(
+            X_train,
+            z_train,
+            Constant,
+            eta[y],
+            epochs=epochs,
+            batches=1,
+            lam=lam[x],
+            X_test=X_test,
+            t_test=z_test,
+        )
+        loss_heatmap[y, x] = test_error[-1]
+        neural._reset_weights()
+
+# select optimal eta, lambda
+y, x = (
+    loss_heatmap.argmin() // loss_heatmap.shape[1],
+    loss_heatmap.argmin() % loss_heatmap.shape[1],
+)
+
+optimal_eta = eta[y]
+optimal_lambda = lam[x]
+
+
+train_error, test_error, _, _ = neural.fit(
+    X_train,
+    z_train,
+    Constant,
+    optimal_eta,
+    epochs=epochs,
+    batches=1,
+    lam=optimal_lambda,
+    X_test=X_test,
+    t_test=z_test,
+)
+
+ax = sns.heatmap(loss_heatmap, xticklabels=lam, yticklabels=eta, annot=True)
+ax.add_patch(
+    Rectangle(
+        (x, y),
+        width=1,
+        height=1,
+        fill=False,
+        edgecolor="crimson",
+        lw=4,
+        clip_on=False,
+    )
+)
+plt.xlabel("lambda", fontsize=18)
+plt.ylabel("eta", fontsize=18)
+plt.title(f"Gridsearch lambda, eta for Constant scheduler", fontsize=22)
+plt.show()
+
[…]

### 2022-11-07 doc/paper.tex
Add implementation part

@@
 
-We implement the FFNN as a class, using NumPy's many y
+We implement the FFNN as a class, using NumPy's many matrix functions.
+In the instantiation of the class, one is able to specify an arbitrary architecture, and choose activation functions for the output and the hidden layer.
+The cost function can also be specified, and this, along with the choice of output function, allows the network to be used for regression, binary classification and multiclass classification.
+By specifying no hidden layers in the architecture, the model can be used for linear and logistic regression as well.
+Because of this, no explicit implementation of linear and logistic regression has been made, which means that almost all relevant code is contained in \texttt{src/FFNN.py} and \texttt{src/Schedulers.py}.
+When instatiating the model, we set the weights to normal distributions with mean zero and variance one.
+We set the bias to $0.1$ as recommended in Goodfellow et Al.
+\cite[189]{gbc}.
 
-We set the bias to 0.1 as recommended in Goodfellow et Al. \cite[189]{gbc}.
+The feedforward part is implemented as a simple matrix vector multiplication, followed by an elementwise application of the chosen activation function.
+For processing batches of inputs, the same multiplication code is used, but this time the input is the design matrix instead of a single row.
+Each hidden layer will now contain matrices themselves, consisting of a number of rows equivalent to the batch size and a number of columns equal to the number of nodes in this layer.
+
+The backpropagation is implemented by differentiating all the functions declared when the class is instatiated, and following the algorithm laid out in Hjort-Jensens lectures \cite{morten}.
+For batch backpropagation, we accumulate the gradient.
+In practice this means that the gradient for the weights, normally a matrix the same size as the relevant weight matrix, is now one matrix per element in our batch (in practice implemented as a NumPy array of three dimensions).
+At the end of the batch, we sum all the gradients for each weight matrix together and subtract this from the corresponding weight matrix.
+
+Training a model is implemented as repeated calls to feedforward and backpropagation, once per batch per epoch.
+During training, one is able to specify which scheduler to use.
+These are implemented as a series of classes all inheriting from a base class, and are all implemented following Hjort-Jensens lectures \cite{mortensched}.
+For finding the optimal parameters, we implement our own gridsearch which finds the lowest error among combinations of $\lambda$, $\eta$, and parameters used in for the scheduling.
 
@@
 
+\subsection{Different schedulers used with linear regression}
+
+\subsection{Neural Network for regression}
+
+For the organic terrain data used in our previous project, we found that our neural net performed slightly better, gaining an MSE of NUMBER, which was about PERCENTAGE better than our best polynomial fit.
+However, looking at our prediction visually, and also comparing the MSE to the one gained on the Franke Function, we see that we are still lacking.
+For terrain data like this, which contain a lot of spatial information, we believe that Convolutional Neural Networks could be a better fit.
+
+\subsection{Neural Network for binary and multiclass classification}
+
+To test our networks ability to classify data, we trained it on many common datasets used in the development of neural networks.
+The datasets were the Wisconsin Breast Cancer dataset and the 8x8 MNIST dataset, as well as other datasets\footnote{Sonar dataset, House dataset}.
+We found good results on all datasets achieving a test accuracy of over 97\% on all
+
 \subsection{Logistic Regression}
 
-To gain insight into how the three different regression methods work, we experiment with synthetic data generated by the Franke function.
+Although our neural network performed well, we also found very similar results with logistic regression, at a much lower computational cost.
+This implies that our classification datasets were perhaps too simple, and that using Neural Networks was perhaps not needed to achieve good results.
 

### 2022-11-07 src/FFNN.py
fix weight initialization

@@ -42,6 +42,7 @@ class FFNN:
         self.checkpoint_file = checkpoint_file
 
         for i in range(len(self.dimensions) - 1):
+            np.random.seed(42069)
             weight_array = np.random.randn(self.dimensions[i] + 1, self.dimensions[i + 1])
             weight_array[0, :] = np.random.randn(self.dimensions[i + 1]) * 0.01
 
@@ -85,6 +86,7 @@ class FFNN:
         test_accs.fill(np.nan)
 
         batch_size = X.shape[0] // batches
+        np.random.seed(42069)
         X, t = resample(X, t)
 
         checkpoint_length = epochs // 10
@@ -219,7 +221,9 @@ class FFNN:
             lam: list[float],
             batch_sizes: list[int],
             *args,
+            batches = 1,
             epochs: int = 1000,
+            # classify: bool = False
     ):
         """
         Optimizes neural network by gridsearching optimal parameters
@@ -239,19 +243,56 @@ class FFNN:
 
         :return: optimal parameters for fit
         """
-        if scheduler is not Momentum:
+        # if classify:
+        #     if scheduler is not Momentum:
+        #         loss_heatmap, optimal_params, optimal_lambda = self._gridsearch_scheduler_classify(
+        #             X, t, X_test, t_test, scheduler, eta, lam, *args epochs=epochs
+        #         )
+        #     else:
+        #         loss_heatmap, optimal_params, optimal_lambda = self._gridsearch_momentum_classify(
+        #             X, t, X_test, t_test, eta, lam, *args, epochs=epochs
+        #         )
+        #     optimal_batch = 0
+        #     batch_size_search = np.zeros(len(batch_sizes))
+        #     for i in range(len(batch_sizes)):
+        #         _, _, _, test_accs = self.fit(
+        #             X,
+        #             t,
+        #             scheduler,
+        #             *optimal_params,
+        #             batches=batch_sizes[i],
+        #             epochs=epochs,
+        #             lam=optimal_lambda,
+        #             X_test=X_test,
+        #             t_test=t_test,
+        #         )
+        #         self._reset_weights()
+        #         batch_size_search[i] = test_error[-1]
+        #     minimal_error = np.min(batch_size_search)
+        #     optimal_batch = batch_sizes[np.argmax(batch_size_search)]
+        #
+        #     plotting_data = [loss_heatmap, batch_size_search]
+        #
+        #     return (
+        #         optimal_params,
+        #         optimal_lambda,
+        #         optimal_batch,
+        #         minimal_error,
[…]

### 2022-11-07 src/task_a.py

@@ -27,11 +27,11 @@ np.random.seed(42069)
 z_train = z_train.reshape(z_train.shape[0], 1)
 z_test = z_test.reshape(z_test.shape[0], 1)
 
-epochs = 200
+epochs = 20
 
 # no hidden layers, no activation function
 dims = (X.shape[1], 1)
-# dims = (2, 20, 1)
+# dims = (2, 20, 20, 1)
 
 eta = np.logspace(-5, -1, 5)
 lam = np.logspace(-5, -1, 5)
@@ -41,9 +41,11 @@ rho2 = 0.999
 
 batch_sizes = np.linspace(1, X.shape[0] // 2, 5, dtype=int)
 schedulers = [Constant, Momentum, Adagrad, RMS_prop, Adam]
+# schedulers = [Constant, Momentum, AdagradMomentum]
 
 constant_params = []
 momentum_params = np.linspace(0, 1, 5)
+adagrad_momentum_params = np.linspace(0, 1, 5)
 adagrad_params = []
 rms_params = [rho]
 adam_params = [rho, rho2]
@@ -55,6 +57,7 @@ params_list = [
     rms_params,
     adam_params,
 ]
+# params_list = [constant_params, momentum_params, adagrad_momentum_params]
 optimal_params_list = []
 
 optimal_eta = np.zeros(len(schedulers))
@@ -73,8 +76,10 @@ for i in range(len(schedulers)):
         minimal_error,
         plotting_data,
     ) = neural.optimize_scheduler(
+        # X_train[:, 1:3],
         X_train,
         z_train,
+        # X_test[:, 1:3],
         X_test,
         z_test,
         schedulers[i],
@@ -82,6 +87,7 @@ for i in range(len(schedulers)):
         lam,
         batch_sizes,
         params_list[i],
+        batches=X.shape[0]//8,
         epochs=epochs // 2,
     )
 
@@ -115,14 +121,17 @@ for i in range(len(schedulers)):
     )
     # neural.read(f"comparison{i}")
     neural = FFNN(dims)
-    _, test_error = neural.fit(
+    _, test_error, _, _ = neural.fit(
+        # X_train[:, 1:3],
         X_train,
         z_train,
         schedulers[i],
         *optimal_params_list[i],
-        batches=optimal_batches[i],
+        # batches=optimal_batches[i],
+        batches=X.shape[0]//8,
         epochs=epochs,
         lam=optimal_lambdas[i],
[…]

### 2022-11-08 doc/paper.tex
More on schedulers

@@
 
+% Neural networks are powerful tools, able to solve many different problems with
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The fundamental idea behind logistic regression is based on the exact approach used in Linear Regression, meaning that we wish to describe a relationship between dependent and explanatory variables in terms of a set of estimated parameters.
-These methods differ however in how they resolve the said matter, with Linear regression modeling a set of parameters by solving a set of linear equations, whilst Logistic regression uses the gradient descent algorithm and the sigmoid function for this purpose.
+However, in this case we are not using the weighted sum of our basis functions directly, but feeding them into an activation function.
+This function, often the sigmoid function, maps the reals onto a range $y \in (0,1)$.
+This in turn allows for the computation of probability related to an event taking place.
+Rounding of said probability leads to labeling of each input as belonging to one of two classes, meaning that each object fed into the model will be classified as being of class 0 or 1 (in the case of binary classification).
 
-Logistic regression is thus just plain linear regression where the weighted sum of the LHS of the linear equation - with randomly initialized parameters - is fed into an activation function.
-This in turn allows for the computation of probability related to an event taking place.
-Rounding of said probability leads to labeling of each input as belonging to one of two classes, meaning that each object fed into the model will be classified as being of class 0 or 1.
-To begin with however, the probability output by the activation function will be incorrect for majority of inputs, due to the random nature of the initialized parameters.
+% To begin with however, the probability output by the activation function will be incorrect for majority of inputs, due to the random nature of the initialized parameters.
 While this may seem to be an issue, the weights can be adjusted to give better classification results through the application of the gradient descent coupled with an adjustable variable called learning rate.
@@
 In our previous project, we have primarily dealt with problems that have an analytical solution.
-This means that finding the optimal solution in these cases is a deterministic process which does not require fine tuning of parameters involved.
-However, for the methods which are applied to problems in this research - such as Logistic Regression - no analytical solutions exist, thus we will have to turn to numerical techniques to find the ‘ideal’ solution.
+This means that once a model has been decided on (polynomial degree and value for lambda for example), finding the optimal solution is a deterministic process which does not require further tuning.
+However, for the methods we use in this project - such as Logistic Regression - no analytical solutions exist, thus we will have to turn to numerical techniques to find the ‘ideal’ solution.
 Our method of choice in this case is gradient descent, an optimization technique which reduces the residuals by utilizing the idea of moving downhill a convex function until a globally minimal scalar value is reached.
@@
 The application of gradient descent to our problems as means of finding the ideal solution implies that the parameters be initialized randomly.
-The most common approach is to set each parameter equal to an arbitrary number ranging between 0 and 1, and iteratively update the parameters by calculating the gradient at each given point in the residual space \cite[preface p.
+The most common approach is to set each parameter equal to a normal distribution with mean zero and variance equal to one, and iteratively update the parameters by calculating the gradient at each given point in the residual space \cite[preface p.
 x]{dds}.
@@
 
-Whilst SGD itself gives a significant boost to the learning pace of machine learning models, it is possible to improve the algorithm further by applying a special type of optimization method called scheduling.
+Whilst SGD itself gives a significant boost to the learning pace of machine learning models, it is possible to improve the algorithm further by performing what is known as scheduling.
 These methods work by adaptively adjusting the learning rate, thus allowing us to avoid local minima and converge even faster.
@@
 This is called momentum, due to the physical analogue of a moving particle having momentum.
-Other more complex methods include keeping running averages of the first and second moments of the gradient, and using this to inform the direction of change in the current update.
+With this addition, we will keep moving in parameter space even if our current gradient is small, because we have gained momentum in a certain direction based on our previous iterations.
+Furthermore, we will keep moving in the direction where most gradients point, even if outliers point in other directions.
+Other, more complex methods, also keep track of approximations of the second moment of the cost function.
+These methods aim to adaptively change the learning rate for different directions in parameter space, so that small steps are taking in steep, narrow directions and large steps taken in shallow, flat directions \cite{mortensched}.
+Examples of schedulers using the second moment are RMS-prop, Adam and Adagrad.
+
+% Adaptive optimization methods don't always have good generalization.
+Recent studies have shown that adaptive methods such as ADAM, RMSPorp, and AdaGrad tend to have poor generalization compared to SGD or SGD with momentum, particularly in the high-dimensional limit (i.e.
+the number of parameters exceeds the number of data points).
+Although it is not clear at this stage why these methods perform so well in training deep neural networks, simpler procedures like properly-tuned SGD may work as well or better in these applications.
 
@@
 However, by limiting ourselves to linear functions, we are possibly limiting our model, as our dataset may well be better approximated by a different basis, our perhaps not even by a linear combination of functions at all.
-Because of this, linear regression is not able to understand the interaction between to arbitrary input variables \cite[165]{gbc}.
+Because of this, linear regression is not able to understand the interaction between two arbitrary input variables \cite[165]{gbc}.
 A better approach would be to simply feed our model our features directly, and hope that it may somehow learn the relationship organically.
@@
 The resulting values from this function are either passed to the output as a linear combination again, or sent to another intermediate vector.
-By having the activation function be a nonlinear function, such as the sigmoid function, we are able to approximate any continous function to arbitrary accuracy \cite[230]{cmb}.
+By having the activation function be a nonlinear function, such as the sigmoid or RELU function, we are able to approximate any continous function to arbitrary accuracy \cite[230]{cmb}.
 By introducing these intermediate vectors, we have created a layered structure, where information is fed forward from layer to layer.
@@
 
-Another activation function is the Rectified Linear Unit, and its cousin the Leaky Rectified Linear Unit:
+Another activation function is the Rectified Linear Unit, and its cousin, the Leaky Rectified Linear Unit:
 
@@
 
-Using these delta terms, we can calculate the gradients for the weights in each layer:
+Using these delta terms, we can calculate the gradients for the weights and biases in each layer:
 
@@
 These are implemented as a series of classes all inheriting from a base class, and are all implemented following Hjort-Jensens lectures \cite{mortensched}.
-For finding the optimal parameters, we implement our own gridsearch which finds the lowest error among combinations of $\lambda$, $\eta$, and parameters used in for the scheduling.
+For finding the optimal parameters, we implement our own gridsearch which finds the lowest error among combinations of lambda, eta, and parameters used for the scheduling.
 

### 2022-11-08 src/FFNN.py
Add seed parameter

@@ -29,6 +29,7 @@ class FFNN:
             output_func: Callable = lambda x: x,
             cost_func: Callable = CostOLS,
             checkpoint_file: str = None,
+            seed: int = None,
     ):
         self.weights = list()
         self.schedulers_weight = list()
@@ -40,9 +41,11 @@ class FFNN:
         self.output_func = output_func
         self.cost_func = cost_func
         self.checkpoint_file = checkpoint_file
+        self.seed = seed
 
         for i in range(len(self.dimensions) - 1):
-            np.random.seed(42069)
+            if self.seed:
+                np.random.seed(self.seed)
             weight_array = np.random.randn(self.dimensions[i] + 1, self.dimensions[i + 1])
             weight_array[0, :] = np.random.randn(self.dimensions[i + 1]) * 0.01
 
@@ -86,7 +89,8 @@ class FFNN:
         test_accs.fill(np.nan)
 
         batch_size = X.shape[0] // batches
-        np.random.seed(42069)
+        if self.seed:
+            np.random.seed(self.seed)
         X, t = resample(X, t)
 
         checkpoint_length = epochs // 10
@@ -372,7 +376,8 @@ class FFNN:
         """
         Resets weights in order to reuse FFNN object for training
         """
-        np.random.seed(42069)
+        if self.seed:
+            np.random.seed(self.seed)
         self.weights = list()
         for i in range(len(self.dimensions) - 1):
             weight_array = np.random.randn(self.dimensions[i] + 1, self.dimensions[i + 1])

### 2022-11-08 src/FFNN.py
Return scores dict

@@ -2,6 +2,7 @@ import numpy as np
 from Schedulers import *
 from utils import *
 
+
 class FFNN:
     """
     Feed Forward Neural Network
@@ -23,13 +24,13 @@ class FFNN:
     """
 
     def __init__(
-            self,
-            dimensions: tuple[int],
-            hidden_func: Callable = sigmoid,
-            output_func: Callable = lambda x: x,
-            cost_func: Callable = CostOLS,
-            checkpoint_file: str = None,
-            seed: int = None,
+        self,
+        dimensions: tuple[int],
+        hidden_func: Callable = sigmoid,
+        output_func: Callable = lambda x: x,
+        cost_func: Callable = CostOLS,
+        checkpoint_file: str = None,
+        seed: int = None,
     ):
         self.weights = list()
         self.schedulers_weight = list()
@@ -44,24 +45,26 @@ class FFNN:
         self.seed = seed
 
         for i in range(len(self.dimensions) - 1):
-            if self.seed:
+            if self.seed is not None:
                 np.random.seed(self.seed)
-            weight_array = np.random.randn(self.dimensions[i] + 1, self.dimensions[i + 1])
+            weight_array = np.random.randn(
+                self.dimensions[i] + 1, self.dimensions[i + 1]
+            )
             weight_array[0, :] = np.random.randn(self.dimensions[i + 1]) * 0.01
 
             self.weights.append(weight_array)
 
     def fit(
-            self,
-            X: np.ndarray,
-            t: np.ndarray,
-            scheduler_class,
-            *args,  # arguments for the scheduler (sans batchsize)
-            batches: int = 1,
-            epochs: int = 1000,
-            lam: float = 0,
-            X_test: np.ndarray = None,
-            t_test: np.ndarray = None,
+        self,
+        X: np.ndarray,
+        t: np.ndarray,
+        scheduler_class,
+        *args,  # arguments for the scheduler (sans batchsize)
+        batches: int = 1,
+        epochs: int = 1000,
+        lam: float = 0,
+        X_test: np.ndarray = None,
+        t_test: np.ndarray = None,
     ):
         """
         Trains the neural network via feedforward and backpropagation
@@ -89,7 +92,7 @@ class FFNN:
         test_accs.fill(np.nan)
 
         batch_size = X.shape[0] // batches
-        if self.seed:
+        if self.seed is not None:
             np.random.seed(self.seed)
         X, t = resample(X, t)
 
@@ -139,8 +142,8 @@ class FFNN:
[…]

### 2022-11-08 src/task_a.py

@@ -65,7 +65,7 @@ optimal_lambdas = np.zeros(len(schedulers))
 optimal_batches = np.zeros(len(schedulers), dtype=int)
 minimal_errors = np.zeros(len(schedulers))
 
-neural = FFNN(dims)
+neural = FFNN(dims, seed=1337)
 for i in range(len(schedulers)):
     plt.subplot(321 + i)
     plt.suptitle("Test loss for eta, lambda grid", fontsize=22)
@@ -87,7 +87,7 @@ for i in range(len(schedulers)):
         lam,
         batch_sizes,
         params_list[i],
-        batches=X.shape[0]//8,
+        batches=X.shape[0] // 8,
         epochs=epochs // 2,
     )
 
@@ -120,21 +120,22 @@ for i in range(len(schedulers)):
         f"{schedulers[i].__name__} \n {optimal_eta[i]=}\n {optimal_lambdas[i]=} \n {optimal_batches[i]=}\n{minimal_errors[i]=}"
     )
     # neural.read(f"comparison{i}")
-    neural = FFNN(dims)
-    _, test_error, _, _ = neural.fit(
+    neural = FFNN(dims, seed=1337)
+    scores = neural.fit(
         # X_train[:, 1:3],
         X_train,
         z_train,
         schedulers[i],
         *optimal_params_list[i],
         # batches=optimal_batches[i],
-        batches=X.shape[0]//8,
+        batches=X.shape[0] // 8,
         epochs=epochs,
         lam=optimal_lambdas[i],
         # X_test=X_test[:, 1:3],
         X_test=X_test,
         t_test=z_test,
     )
+    test_error = scores["test_error"]
     plt.plot(test_error, label=f"{schedulers[i].__name__}")
     plt.legend(loc=(1.04, 0))
 best_MSE_analytically = np.zeros(epochs)

### 2022-11-08 src/task_a.py
for pulL

@@ -27,7 +27,7 @@ np.random.seed(42069)
 z_train = z_train.reshape(z_train.shape[0], 1)
 z_test = z_test.reshape(z_test.shape[0], 1)
 
-epochs = 20
+epochs = 200
 
 # no hidden layers, no activation function
 dims = (X.shape[1], 1)
@@ -87,7 +87,7 @@ for i in range(len(schedulers)):
         lam,
         batch_sizes,
         params_list[i],
-        batches=X.shape[0]//8,
+        batches=X.shape[0] // 8,
         epochs=epochs // 2,
     )
 
@@ -128,7 +128,7 @@ for i in range(len(schedulers)):
         schedulers[i],
         *optimal_params_list[i],
         # batches=optimal_batches[i],
-        batches=X.shape[0]//8,
+        batches=X.shape[0] // 8,
         epochs=epochs,
         lam=optimal_lambdas[i],
         # X_test=X_test[:, 1:3],

### 2022-11-09 doc/paper.tex
Remove old text

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Abstract}
@@
 To optimize our network, we explored the use of different schedulers in our gradient descent method, and compared their performance against eachother.
-Overall, we found that our FFNN performed well in all classification tasks we tested, attaining test accuracies between $95$ and $100$ percent for all classification problem.
+Overall, we found that our FFNN performed well in all classification tasks we tested, attaining test accuracies between $95$ and $100$ percent for all classification problems.
 However, we also found that simple logistic regression performed similarly, with a much lower computational cost.
-For regression on terrain data, we found our neural net lacking, not being able to beat our polynomial fit from our previous project, and only attaining an MSE of 1 BILLION, which is 2000 BILLION lower than our best polynomial fit of the Franke Function.
+For regression on real terrain data, we found that our neural network slightly outperformed our previous polynomial fit, attaining an MSE of MONKEY, PERCENT percent lower than simple linear regression.
 In our gradient descent, we found that SCHEDULER performed best, being 1 BILLION percent better than the average of all other methods.
+
+%we found our neural net lacking, not being able to beat our polynomial fit from our previous project, and only attaining an MSE of 1 BILLION, which is 2000 BILLION lower than our best polynomial fit of the Franke Function.
 
@@
 This means that once a model has been decided on (polynomial degree and value for lambda for example), finding the optimal solution is a deterministic process which does not require further tuning.
-However, for the methods we use in this project - such as Logistic Regression - no analytical solutions exist, thus we will have to turn to numerical techniques to find the ‘ideal’ solution.
+However, for the methods we use in this project - such as Logistic Regression - no analytical solutions exist, thus we will have to turn to numerical techniques to find the 'ideal' solution.
 Our method of choice in this case is gradient descent, an optimization technique which reduces the residuals by utilizing the idea of moving downhill a convex function until a globally minimal scalar value is reached.
@@
 In practice this means that the gradient for the weights, normally a matrix the same size as the relevant weight matrix, is now one matrix per element in our batch (in practice implemented as a NumPy array of three dimensions).
-At the end of the batch, we sum all the gradients for each weight matrix together and subtract this from the corresponding weight matrix.
+At the end of the batch, we average all the gradients for each weight matrix together and feed this into our scheduler calculation to determine how much our weights should change.
 
@@
 
-\subsubsection{Ordinary Least Squares}
+% \begin{figure}[h]
+% \centerline{\includegraphics[width=5in]{figure/frankebestlambdamse.png}}
+% \caption{Test and train MSE for ridge}
+% \label{frankebestlambdamse}
+% \end{figure}
 
-\subsubsection{Ridge}
+% \begin{table}
+% \begin{center}
+% \label{ols_mse_table_dif_data}
+% \begin{tabular}{c l l l}
+% Dataset & Std of noise & Best MSE & Best degree \\
+% \hline
+% Franke & 0 & 0.0005109 & 9\\
+% Franke & 0.05 & 0.0038302 & 7 \\
+% Franke & 0.10 & 0.0139833 & 7 \\
+% Terrain 1 & - & 0.0142683 & 19 \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \end{table}
 
-\begin{figure}[h]
-\centerline{\includegraphics[width=5in]{figure/frankebestlambdamse.png}}
-\caption{Test and train MSE for ridge}
-\label{frankebestlambdamse}
-\end{figure}
-
-\begin{table}
-\begin{center}
-\label{ols_mse_table_dif_data}
-\begin{tabular}{c l l l}
-Dataset & Std of noise & Best MSE & Best degree \\
-\hline
-Franke & 0 & 0.0005109 & 9\\
-Franke & 0.05 & 0.0038302 & 7 \\
-Franke & 0.10 & 0.0139833 & 7 \\
-Terrain 1 & - & 0.0142683 & 19 \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
-
-\subsubsection{Ridge Regression}
-
-As our model is not overfitting, we do not expect Ridge to offer much improvement.
-This proves to be the case, as we see from Fig.~\ref{real1ridgedifflambdas} in the appendix.
-Here we see that different values of lambda have very little impact, as there are no parameters that need regularization, and no overfitting to supress.
-For very large values, we see that we are shrinking the parameters too much, and we see a slight increase in MSE, but overall we do not see much of an effect.
-We could expect to see that Ridge shows improvement at higher polynomial degrees, when OLS starts to overfit, but due to computational constraints we have not been able to push our grid search much further than 9 degrees.
-Using scikit's GridSearch, we find that the optimal lambda is 0, which is exactly what we would expect on a model that does not overfit, as Ridge regression with a lambda equal to 0 is equivalent with OLS.
-See Table~\ref{ols_vs_ridge_vs_lasso_table_real}.
-
-\subsubsection{Lasso Regression}
-
-With lasso regression, we see similar results as with Ridge.
-Fig.~\ref{real1lassobiasvariance} in the appendix shows little change, except for when our $\lambda$ is too high and we force all parameters to $0$.
-Comparing lasso to the two other methods for the real dataset has proven to be difficult, as the numerical methods for finding the best parameters are very time consuming.
-We have not been able to gridsearch up to polynomial degree 19, but up to degree 9 we found the best lambda to be $2.154^{-7}$ for polynomial degree 9, giving an MSE of $0.01685$.
-
-Comparing all three methods up to degree 9 for different lambdas (using the same lambda for Lasso and Ridge, just to get an overview) in Fig.~\ref{real1msecomparison} in the appendix, we see the same trend as we have been describing.
-Regularization offers little reward, which is to be expected when we are not overfitting.
-
-\begin{table*}[t]
-\caption{Best MSE for OLS, Ridge and Lasso regression on real elevation data. We only fit up to degree 9 due to computational constraints}
-\begin{center}
-\label{ols_vs_ridge_vs_lasso_table_real}
-\begin{tabular}{c | l l l}
-Regression & MSE & Degree & $\lambda$ \\
-\hline
-OLS & 0.01538 & 9 & - \\
-Ridge & 0.01538 & 9 & 0 \\
-Lasso & 0.01685 & 9 & $2.154 \cdot 10^{-7}$ \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
+% \begin{table*}[t]
+% \caption{Best MSE for OLS, Ridge and Lasso regression on real elevation data. We only fit up to degree 9 due to computational constraints}
+% \begin{center}
+% \label{ols_vs_ridge_vs_lasso_table_real}
+% \begin{tabular}{c | l l l}
+% Regression & MSE & Degree & $\lambda$ \\
+% \hline
+% OLS & 0.01538 & 9 & - \\
+% Ridge & 0.01538 & 9 & 0 \\
+% Lasso & 0.01685 & 9 & $2.154 \cdot 10^{-7}$ \\
+% \hline
+% \end{tabular}
+% \end{center}
+% \end{table*}
 
@@
 
-In this paper we have compared Ordinary Least Squares, Ridge and Lasso regression for both synthetic and real data.
-We have seen the value of using resampling methods such as bootstrap and cross validation to gain valuable insights about the quality of our predictions.
-Using resampling methods, we have found that linear regression was well suited for approximating the synthetic Franke function, but less suited for the real elevation data.
-We have seen that a correctly chosen regularization parameter can allow us to push our model complexity higher than we can with OLS, and thus give us a better MSE than OLS.
-Looking at the real data, we have seen that regularization does not offer much value when we are not overfitting our dataset.
-Thus all three linear regression methods perform similarly on this dataset.
+In this paper we have explored the use of a Feed Forward Neural Network for both regression and classification. We have studied how different schedulers,
 
@@
 \ref{frankenonoise} & \texttt{task\_b.py --noise 0 -n 25}\\
-\ref{frankebestlambdamse} & \texttt{task\_b.py --noise 0 -n 25} \footnotesize{(\it Model has been changed to ridge with lambda=$1.43 \cdot 10^{-7}$)}\\
-\ref{frankebestlambdabetas} & \texttt{task\_b.py --noise 0 -n 25} \footnotesize{(\it Model has been changed to ridge with lambda=$1.43 \cdot 10^{-7}$)}\\
-\ref{realunscaled} & \texttt{task\_b.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 19 --noscale}\\
-\ref{realunscaledmse} & \texttt{task\_b.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 19 --noscale}\\
-\ref{frankenonoisemse} & \texttt{task\_c.py --noise 0 -n 24}\\
-\ref{frankenoisemse} & \texttt{task\_c.py -n 24}\\
-\ref{frankiebiasvariance} & \texttt{task\_c.py -n 24 --step 0.1}\\
-\ref{frankiedifflambdas} & \texttt{task\_e\_biasvariance.py -n 24 --step 0.1}\\
-\ref{frankedifflambdaslasso} & \texttt{task\_f\_biasvariance.py -n 9 --step 0.15}\\
-\ref{real1ridgedifflambdas} & \texttt{task\_e\_biasvariance.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 24}\\
-\ref{frankeridgebetas} & \texttt{task\_e\_betas.py -n 24}\\
-\ref{frankenoisybeta} & \texttt{task\_c.py -n 24}\\
-\ref{frankienoisecomparison} & \texttt{noise\_plot.py}\\
-\ref{frankiehigherdatapoints} & \texttt{task\_c.py --noise 0 -n 30 --step 0.01}\\
-\ref{real1surfacecomp} & \texttt{task\_b.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 19}\\
-\ref{real1mselessdata} & \texttt{task\_c.py --file ../data/tiny\_SRTM\_data\_Norway\_1.tif -n 19}\\
-\ref{real1msetraintest} & \texttt{task\_c.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 19}\\
-\ref{realbiasvarianceOLS} & \texttt{task\_c.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 19}\\
-\ref{realOLSmse40} & \texttt{task\_b.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 40}\\
-\ref{realOLSbetaprog40} & \texttt{task\_b.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 40}\\
-\ref{frankemsebyresamplingmethod} & \texttt{task\_d.py -n 15 --step 0.02}\\
-\ref{real1msecomparison} & \texttt{task\_f\_msecomparison.py --file ../data/small\_SRTM\_data\_Norway\_1.tif -n 9}\\
 \hline
@@
 
-\begin{figure*}
-\centerline{\includegraphics[width=3.25in]{figure/realmsebyresamplingmethod.png}}
-\caption{Comparison between approximated MSE for different resampling methods for the real terrain data}
-\label{realmsebyresamplingmethod}
-\end{figure*}
-
-\begin{figure*}
-\centerline{\includegraphics[width=3.25in]{figure/frankiehigherdatapoints.png}}
-\caption{Franke function fitted up to polynomial degree 30, with 10000 datapoints instead of 400.
-As we can see, we do not overfit even at this polynomial degree when the availiable data is plentiful.
-The highest MSE was reached at $N=29$, and was $0.00003511$, one tenth of our best prediction with only 400 datapoints}
-\label{frankiehigherdatapoints}
-\end{figure*}
-
-\begin{figure*}
-% \begin{figure}[t]
-\centerline{\includegraphics[width=3.25in]{figure/frankienoisecomparison.png}}
-\caption{Test and train MSE for the OLS prediction of Franke function with an increasing amount of noise}
-\label{frankienoisecomparison}
-\end{figure*}
-
-\begin{figure*}
-\centerline{\includegraphics[width=3.25in]{figure/frankenoisybeta.png}}
-\caption{Beta progression for OLS}
-\label{frankenoisybeta}
-\end{figure*}
-
-\begin{figure*} 
-\centerline{\includegraphics[width=3.25in]{figure/frankiedifflambdas.png}}
-\caption{Bias, variance and test MSE plotted over polynomial degree for our Ridge prediction on the Franke function with an added stochastic noise $\epsilon \sim N(0,0.05)$, using lambdas of increasing size.
-Here we have reduced the number of datapoints to only 100 to force overfitting earlier.}
-\label{frankiedifflambdas}
-\end{figure*}
-
-\begin{figure*} 
-\centerline{\includegraphics[width=3.25in]{figure/frankeridgebetas.png}}
-\caption{Beta progression for different values of lambda in Ridge regression. Note the different scales on each of the plots, showing how increasing the regularization parameter shrinks the betas}
-\label{frankeridgebetas}
-\end{figure*}
-
 % \begin{figure*} 
-% \centerline{\includegraphics[width=3.25in]{figure/frankeridgenoisecomparison.png}}
-% \caption{Test and train MSE for the Ridge prediction with $\lambda = 10^{-8}$ of the Franke function with an increasing amount of noise}
-% \label{frankeridgenoisecomparison}
+% \centerline{\includegraphics[width=3.25in]{figure/frankeridgebetas.png}}
+% \caption{Beta progression for different values of lambda in Ridge regression. Note the different scales on each of the plots, showing how increasing the regularization parameter shrinks the betas}
+% \label{frankeridgebetas}
 % \end{figure*}
 
-\begin{figure*}
-\centerline{\includegraphics[width=3.25in]{figure/frankedifflambdaslasso.png}}
-\caption{Bias, variance and test MSE plotted over polynomial degree for our Lasso prediction on the Franke function with an added stochastic noise $\epsilon \sim N(0,0.05)$, using lambdas of increasing size.
-Here we have reduced the number of datapoints to less than 100 to force overfitting earlier.}
-\label{frankedifflambdaslasso}
-\end{figure*}
-
-\begin{figure*}
-\centerline{\includegraphics[width=3.25in]{figure/realunscaled.png}}
-\caption{Our prediction for the real terrain using unscaled data, showing that we reach a very poor result}
-\label{realunscaled}
-\end{figure*}
-
-\begin{figure*}
-\centerline{\includegraphics[width=3.25in]{figure/realunscaledmse.png}}
-\caption{Test and train MSE for our prediction for the real terrain using unscaled data, showing that we reach a very poor result}
-\label{realunscaledmse}
-\end{figure*}
-
-\begin{figure*} 
-\centerline{\includegraphics[width=6.85in]{figure/real1surfacecomp.png}}
-\caption{A comparison of our best OLS prediction and actual terrain data (terrain 1), at polynomial degree 19}
-\label{real1surfacecomp}
-\end{figure*}
-
-\begin{figure*}
-\centerline{\includegraphics[width=3.25in]{figure/realOLSmse40.png}}
-\caption{Test and train MSE on our terrain, this time up to degree 40.
-We see only a slight decrease in MSE, down to $0.01381$, even though we double the amount of degrees.
-This has not been bootstrapped due to computational limitations}
-\label{realOLSmse40}
-\end{figure*}
-
-\begin{figure*}
-\centerline{\includegraphics[width=3.25in]{figure/realOLSbetaprog40.png}}
-\caption{Beta progression up to polynomial degree 40 for real terrain data, showing that the are relatively constant, implying that we are not overfitting}
-\label{realOLSbetaprog40}
-\end{figure*}
-
-\begin{figure*}
-\centerline{\includegraphics[width=3.25in]{figure/real1mselessdata.png}}
-\caption{Test and train MSE for the OLS prediction of terrain 1, this time downscaled to only $50x100$. Here the optimal degree was 11, showing that with lower data we start to overfit more rapidly.}
-\label{real1mselessdata}
-\end{figure*}
-
-\begin{figure} 
-\centerline{\includegraphics[width=3.25in]{figure/real1ridgedifflambdas.png}}
-\caption{Bias variance tradeoff for ridge regression with different lambdas, for the real dataset}
-\label{real1ridgedifflambdas}
-\end{figure}
-
-\begin{figure} 
-\centerline{\includegraphics[width=3.25in]{figure/real1lassobiasvariance.png}}
-\caption{Bias variance tradeoff for lasso regression with different lambdas, for the real dataset}
-\label{real1lassobiasvariance}
-\end{figure}
-
-\begin{figure*}
-\centerline{\includegraphics[width=3.25in]{figure/real1msecomparison.png}}
-\caption{Comparison between OLS, Ridge and Lasso for different values of lambda, for real terrain data}
-\label{real1msecomparison}
-\end{figure*}
-
 \end{document}

### 2022-11-10 src/FFNN.py
task_b and task_d updated to fit teh changes done in optimize scheduler

@@ -231,7 +231,7 @@ class FFNN:
         *args,
         batches=1,
         epochs: int = 1000,
-        classify: bool = False
+        classify: bool = False,
     ):
         """
         Optimizes neural network by gridsearching optimal parameters
@@ -264,7 +264,7 @@ class FFNN:
                 *args,
                 batches=batches,
                 epochs=epochs,
-                classify=classify
+                classify=classify,
             )
         else:
             loss_heatmap, optimal_params, optimal_lambda = self._gridsearch_momentum(
@@ -278,7 +278,7 @@ class FFNN:
                 *args,
                 batches=batches,
                 epochs=epochs,
-                classify=classify
+                classify=classify,
             )
 
         return optimal_params, optimal_lambda, loss_heatmap
@@ -314,7 +314,7 @@ class FFNN:
             self.reset_weights()
             # todo would be interesting to see how much time / how fast it happens
             batches_list_search[i, :] = test_error
-        optimal_batch = batches_list[np.argmin(batches_list_search[:,-1])]
+        optimal_batch = batches_list[np.argmin(batches_list_search[:, -1])]
 
         return optimal_batch, batches_list_search
 
@@ -501,7 +501,18 @@ class FFNN:
         return len(line)
 
     def _gridsearch_scheduler(
-        self, X, t, X_test, t_test, scheduler, eta, lam, *args, batches=1, epochs=1000, classify=False,
+        self,
+        X,
+        t,
+        X_test,
+        t_test,
+        scheduler,
+        eta,
+        lam,
+        *args,
+        batches=1,
+        epochs=1000,
+        classify=False,
     ):
         """
         Help function for optimize_scheduler
@@ -524,7 +535,7 @@ class FFNN:
                     t_test=t_test,
                 )
                 if classify:
-                    test_accs = scores["test_accs"]
+                    test_accs = scores["test_acc"]
                     loss_heatmap[y, x] = test_accs[-1]
                 else:
                     test_error = scores["test_error"]
@@ -556,7 +567,7 @@ class FFNN:
         momentums,
         batches=1,
         epochs=1000,
-        classify=False
+        classify=False,
     ):
         """
         Help function for optimize_scheduler
@@ -581,7 +592,7 @@ class FFNN:
                         t_test=t_test,
                     )
                     if classify:
-                        test_accs = scores["test_accs"]
+                        test_accs = scores["test_acc"]
                         loss_heatmap[y, x, z] = test_accs[-1]
                     else:
                         test_error = scores["test_error"]

### 2022-11-10 src/task_b.py

@@ -34,7 +34,7 @@ opt_params = [rho, rho2]
 # params = [eta, momentum]
 
 dims = (2, 224, 112, 1)
-train_epochs = 2000
+train_epochs = 1000
 
 eta = np.logspace(-5, -1, 5)
 lams = np.logspace(-5, -1, 5)
@@ -52,8 +52,8 @@ optimal_params, optimal_lambda, _ = neural.optimize_scheduler(
     eta,
     lams,
     opt_params,
-    batches=10,
-    epochs=20,
+    batches=25,
+    epochs=1000,
 )
 
 params = [optimal_params[0], rho, rho2]
@@ -67,7 +67,7 @@ optimal_batch = neural.optimize_batch(
     optimal_lambda,
     *params,
     batches_list=batch_sizes,
-    epochs=20,
+    epochs=30,
 )
 
 scores = neural.fit(

### 2022-11-10 src/task_d.py

@@ -32,9 +32,10 @@ X_test_sc = scaler.transform(X_test)
 
 
 # parameters
-neural_dims = (30, 100, 1)
+neural_dims = (30, 60, 1)
 logreg_dims = (30, 1)
-eta = 0.001
+eta = np.logspace(-5, -1, 5)
+lam = np.logspace(-5, -1, 5)
 rho = 0.90
 rho2 = 0.999
 z_train = z_train.reshape(z_train.shape[0], 1)
@@ -50,9 +51,26 @@ sched = RMS_prop
 sched = Adam
 # params = [eta, momentum]
 # params = [eta]
+opt_params = [rho, rho2]
 params = [eta, rho, rho2]
 # params = [eta]
 # params = [eta, rho]
+
+optimal_params, optimal_lambda, _ = neural.optimize_scheduler(
+    X_train_sc,
+    z_train,
+    X_test_sc,
+    z_test,
+    sched,
+    eta,
+    lam,
+    opt_params,
+    batches=20,
+    epochs=100,
+    classify=True,
+)
+
+params = [optimal_params[0], rho, rho2]
 
 scores = neural.fit(
     X_train_sc,
@@ -61,7 +79,7 @@ scores = neural.fit(
     *params,
     batches=batches,
     epochs=1000,
-    # lam=0.01,
+    lam=optimal_lambda,
     X_test=X_test_sc,
     t_test=z_test,
 )
@@ -76,5 +94,4 @@ plt.title("LogLoss over Epochs")
 plt.show()
 
 prediction = neural.predict(X_test_sc)
-
 plot_confusion(prediction, z_test)

### 2022-11-10 src/FFNN.py
copy-pasted FFNN from branch marriage to this one due to rebase confilct

@@ -90,6 +90,8 @@ class FFNN:
         train_accs.fill(np.nan)  # makes for better plots if we cancel early
         test_accs = np.empty(epochs)
         test_accs.fill(np.nan)
+        if X_test is not None:
+            test_preds = np.zeros((X_test.shape[0], epochs))
 
         if self.seed is not None:
             np.random.seed(self.seed)
@@ -134,11 +136,17 @@ class FFNN:
                 for scheduler in self.schedulers_bias:
                     scheduler.reset()
 
-                train_error = cost_function_train(self.predict(X, raw=True))
+                prediction = self.predict(X, raw=True)
+                train_error = cost_function_train(prediction)
+                if train_error > 10e20:
+                    # if this happens, we have a problem
+                    break
                 if X_test is not None and t_test is not None:
-                    test_error = cost_function_test(self.predict(X_test, raw=True))
+                    prediction_test = self.predict(X_test, raw=True)
+                    test_error = cost_function_test(prediction_test)
+                    test_preds[:, e] = prediction_test.ravel()
                 else:
-                    test_error = 0
+                    test_error = np.nan  # a
 
                 train_acc = None
                 test_acc = None
@@ -192,7 +200,61 @@ class FFNN:
             "test_error": test_errors,
             "train_acc": train_accs,
             "test_acc": test_accs,
+            "test_pred": test_preds,
         }
+
+    def bootstrap(
+        self,
+        num_bootstraps: int,
+        X: np.ndarray,
+        t: np.ndarray,
+        scheduler_class,
+        *args,  # arguments for the scheduler (sans batchsize)
+        batches: int = 1,
+        epochs: int = 1000,
+        lam: float = 0,
+        X_test: np.ndarray = None,
+        t_test: np.ndarray = None,
+    ):
+        """Bootstrap
+        Parameters:
+            num_bootstraps (int): number of bootstraps
+            args: positional arguments for FFNN.fit()
+            kwargs: keyword arguments for FFNN.fit()
+        Returns:
+            bootstrapped test error, bias, variance by epochs
+        """
+        matrices = bootstrap(X, t, num_bootstraps)
+
+        test_errors = np.zeros(epochs)
+        predictions = np.zeros((X_test.shape[0], num_bootstraps, epochs))
+        biases = np.zeros(epochs)
+        variances = np.zeros(epochs)
+
+        for i in range(len(matrices)):
+            scores = self.fit(
+                matrices[i][0],
+                matrices[i][1],
+                scheduler_class,
+                *args,
[…]

### 2022-11-10 doc/paper.tex
Add verification

@@
 \usepackage{xcolor}
+\usepackage{xurl}
 
@@
 \definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
+  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
+  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
+  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
+  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
+  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
+  \do\Y\do\Z}
 
@@
 
-Neural Networks are all the rage these days, everyone wants to have a neural network in their home. Ive even seen people install neural networks in their dogs.
+Neural Networks are all the rage these days, everyone wants to have a neural network in their home. Ive even seen people install neural networks in their dogs. \cite{ericjona}
 
@@
 
-The backpropagation is implemented by differentiating all the functions declared when the class is instatiated, and following the algorithm laid out in Hjort-Jensens lectures \cite{morten}.
+The backpropagation is implemented by differentiating all the functions declared when the class is instatiated, and following the algorithm laid out in Hjorth-Jensens lectures \cite{morten}.
 For batch backpropagation, we accumulate the gradient.
 In practice this means that the gradient for the weights, normally a matrix the same size as the relevant weight matrix, is now one matrix per element in our batch (in practice implemented as a NumPy array of three dimensions).
-At the end of the batch, we average all the gradients for each weight matrix together and feed this into our scheduler calculation to determine how much our weights should change.
+At the end of the batch, we average all the gradients for each weight matrix together and feed this into our scheduling calculation to determine how much our weights should change.
 
@@
 During training, one is able to specify which scheduler to use.
-These are implemented as a series of classes all inheriting from a base class, and are all implemented following Hjort-Jensens lectures \cite{mortensched}.
+These are implemented as a series of classes all inheriting from a base class, and are all implemented following Hjorth-Jensens lectures \cite{mortensched}.
 For finding the optimal parameters, we implement our own gridsearch which finds the lowest error among combinations of lambda, eta, and parameters used for the scheduling.
+
+For verification that our implementation is working correctly, we train on a simple second degree polynomial, and see that our predicted y value is very close to the actual y value\footnote{see \texttt{src/verification.py}}.
+See Fig.~\ref{verification}.
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/verification.png}}
+\caption{Simple verification case}
+\label{verification}
+\end{figure}
 
@@
 Although our neural network performed well, we also found very similar results with logistic regression, at a much lower computational cost.
-This implies that our classification datasets were perhaps too simple, and that using Neural Networks was perhaps not needed to achieve good results.
-
-% \begin{figure}[h]
-% \centerline{\includegraphics[width=5in]{figure/frankebestlambdamse.png}}
-% \caption{Test and train MSE for ridge}
-% \label{frankebestlambdamse}
-% \end{figure}
+This implies that our classification datasets were perhaps too simple, and that using Neural Networks was not needed to achieve good results.
 
@@
 
-%%%%%%%%%%% FIGURES %%%%%%%%%%%%%%%%%%%%%%%%%5
-
-% \begin{figure*} 
-% \centerline{\includegraphics[width=3.25in]{figure/frankeridgebetas.png}}
-% \caption{Beta progression for different values of lambda in Ridge regression. Note the different scales on each of the plots, showing how increasing the regularization parameter shrinks the betas}
-% \label{frankeridgebetas}
-% \end{figure*}
-
 \end{document}

### 2022-11-10 src/FFNN.py
Clean up code, more comments

@@ -81,6 +81,9 @@ class FFNN:
 
         :return: error over epochs for training and test runs for plotting
         """
+
+
+        # --------- setup ---------
         train_errors = np.empty(epochs)
         train_errors.fill(np.nan)  # makes for better plots if we cancel early
         test_errors = np.empty(epochs)
@@ -90,20 +93,24 @@ class FFNN:
         train_accs.fill(np.nan)  # makes for better plots if we cancel early
         test_accs = np.empty(epochs)
         test_accs.fill(np.nan)
-        if X_test is not None:
-            test_preds = np.zeros((X_test.shape[0], epochs))
+
+        self.schedulers_weight = list()
+        self.schedulers_bias = list()
+
+        batch_size = X.shape[0] // batches
 
         if self.seed is not None:
             np.random.seed(self.seed)
+
+        if X_test is not None:
+            test_preds = np.zeros((X_test.shape[0], epochs))
+
+
+
         X, t = resample(X, t)
 
         checkpoint_length = epochs // 10
         checkpoint_num = 0
-
-        batch_size = X.shape[0] // batches
-
-        self.schedulers_weight = list()
-        self.schedulers_bias = list()
 
         # this function returns a function valued only at X
         cost_function_train = self.cost_func(t)
@@ -115,10 +122,12 @@ class FFNN:
             self.schedulers_bias.append(scheduler_class(*args))
 
         print(scheduler_class.__name__)
+        # this try is only so that we may cancel early by hitting ctrl+c
         try:
             for e in range(epochs):
                 for i in range(batches):
-                    # print(f"Batch: {i}")
+
+                    # -------- minibatch gradient descent ---------
                     if i == batches - 1:
                         # if we are on the last, take all thats left
                         X_batch = X[i * batch_size :, :]
@@ -130,11 +139,15 @@ class FFNN:
                     self._feedforward(X_batch)
                     self._backpropagate(X_batch, t_batch, lam)
 
+                # reset schedulers every epoch
                 for scheduler in self.schedulers_weight:
                     scheduler.reset()
 
                 for scheduler in self.schedulers_bias:
                     scheduler.reset()
+
+
+                # --------- performance metrics -------
 
                 prediction = self.predict(X, raw=True)
                 train_error = cost_function_train(prediction)
@@ -164,6 +177,7 @@ class FFNN:
                 test_errors[e] = test_error
                 progression = e / epochs
 
+                 # ----- printing progress bar ------------
                 length = self._progress_bar(
                     progression,
[…]

### 2022-11-10 src/FFNN.py
bootstrap

@@ -118,7 +118,6 @@ class FFNN:
         try:
             for e in range(epochs):
                 for i in range(batches):
-                    # print(f"Batch: {i}")
                     if i == batches - 1:
                         # if we are on the last, take all thats left
                         X_batch = X[i * batch_size :, :]
@@ -247,8 +246,8 @@ class FFNN:
             predictions[:, i, :] = scores["test_pred"]
 
         for i in range(epochs):
-            if not i:
-                print(predictions[:, :, i])
+            # if not i:
+            #     print(predictions[:, :, i])
             error, bias, variance = bias_variance(t_test, predictions[:, :, i])
             biases[i] = bias
             variances[i] = variance
@@ -403,10 +402,6 @@ class FFNN:
                             self.z_matrices[i + 1]
                         ) * cost_func_derivative(self.a_matrices[i + 1])
                     except ZeroDivisionError:
-                        print(f"{self.a_matrices=}")
-                        print(f"{self.weights=}")
-                        print(f"{self.a_matrices[i+1]=}")
-                        print(f"{self.a_matrices[i+1].shape=}")
                         exit()
 
             else:
@@ -441,7 +436,6 @@ class FFNN:
                     self.schedulers_weight[i].update_change(gradient_weights),
                 ]
             )
-            # print(f"{update_matrix=}")
             update_list.insert(0, update_matrix)
 
         self._update_w_and_b(update_list)
@@ -483,6 +477,7 @@ class FFNN:
         batches=1,
         epochs: int = 1000,
         classify: bool = False,
+        bootstraps=1,
     ):
         """
         Optimizes neural network by gridsearching optimal parameters
@@ -521,6 +516,7 @@ class FFNN:
                 batches=batches,
                 epochs=epochs,
                 classify=classify,
+                bootstraps=bootstraps,
             )
         else:
             (
@@ -540,6 +536,7 @@ class FFNN:
                 batches=batches,
                 epochs=epochs,
                 classify=classify,
+                bootstraps=bootstraps,
             )
 
         string = (
@@ -574,7 +571,6 @@ class FFNN:
         optimal_batch = 0
         batches_list_search = np.zeros((len(batches_list), epochs))
         for i in range(len(batches_list)):
-            print(batches_list[i])
             scores = self.fit(
                 X,
                 t,
@@ -607,6 +603,7 @@ class FFNN:
         batches=1,
         epochs=1000,
         classify=False,
+        bootstraps=1,
     ):
         """
[…]

### 2022-11-10 src/FFNN.py
NO TEST_ACCS IN FFNNga task_d.py ga task_d.py

@@ -630,7 +630,7 @@ class FFNN:
                     t_test=t_test,
                 )
                 if classify:
-                    test_accs = scores["test_accs"]
+                    test_accs = scores["test_acc"]
                     loss_heatmap[y, x] = test_accs[-1]
                     min_heatmap[y, x] = np.min(test_accs)
                 else:
@@ -689,7 +689,7 @@ class FFNN:
                         t_test=t_test,
                     )
                     if classify:
-                        test_accs = scores["test_accs"]
+                        test_accs = scores["test_acc"]
                         loss_heatmap[y, x, z] = test_accs[-1]
                         min_heatmap[y, x, z] = np.min(test_accs)
                     else:

### 2022-11-10 src/task_d.py

@@ -55,6 +55,8 @@ opt_params = [rho, rho2]
 params = [eta, rho, rho2]
 # params = [eta]
 # params = [eta, rho]
+batch_sizes = np.linspace(1, X.shape[0] // 2, 5, dtype=int)
+
 
 optimal_params, optimal_lambda, _ = neural.optimize_scheduler(
     X_train_sc,
@@ -77,7 +79,7 @@ scores = neural.fit(
     z_train,
     sched,
     *params,
-    batches=batches,
+    batches=optimal_batch,
     epochs=1000,
     lam=optimal_lambda,
     X_test=X_test_sc,

### 2022-11-10 src/FFNN.py
Add ability to use the best weights

@@ -1,6 +1,7 @@
 import numpy as np
 from Schedulers import *
 from utils import *
+from copy import deepcopy
 
 
 class FFNN:
@@ -65,6 +66,7 @@ class FFNN:
         lam: float = 0,
         X_test: np.ndarray = None,
         t_test: np.ndarray = None,
+        use_best_weights: bool = False,
     ):
         """
         Trains the neural network via feedforward and backpropagation
@@ -76,8 +78,8 @@ class FFNN:
         :param batches: batches to split data into
         :param epochs: how many epochs for training
         :param lam: regularization parameter
-        :param X_test: pass -> return test plotting data
-        :param t_test: pass -> return test plotting data
+        :param use_best_weights: save the best weights and only use them. This runs slower because
+        it has to check every epoch if it is better than the best. Only use with X_test set
 
         :return: error over epochs for training and test runs for plotting
         """
@@ -85,12 +87,12 @@ class FFNN:
 
         # --------- setup ---------
         train_errors = np.empty(epochs)
-        train_errors.fill(np.nan)  # makes for better plots if we cancel early
+        train_errors.fill(np.nan)  # nan makes for better plots if we cancel early (they are not plotted)
         test_errors = np.empty(epochs)
         test_errors.fill(np.nan)
 
         train_accs = np.empty(epochs)
-        train_accs.fill(np.nan)  # makes for better plots if we cancel early
+        train_accs.fill(np.nan)
         test_accs = np.empty(epochs)
         test_accs.fill(np.nan)
 
@@ -105,15 +107,19 @@ class FFNN:
         if X_test is not None:
             test_preds = np.zeros((X_test.shape[0], epochs))
 
-
+        if use_best_weights and X_test is not None:
+            best_weights = deepcopy(self.weights)
+            best_error = 10e20
+            best_acc = 0
 
         X, t = resample(X, t)
 
         checkpoint_length = epochs // 10
         checkpoint_num = 0
 
+
         # this function returns a function valued only at X
-        cost_function_train = self.cost_func(t)
+        cost_function_train = self.cost_func(t) # used for performance metrics
         if X_test is not None and t_test is not None:
             cost_function_test = self.cost_func(t_test)
 
@@ -154,12 +160,19 @@ class FFNN:
                 if train_error > 10e20:
                     # if this happens, we have a problem
                     break
-                if X_test is not None and t_test is not None:
+                if X_test is not None:
                     prediction_test = self.predict(X_test, raw=True)
[…]

### 2022-11-10 src/utils.py

@@ -196,6 +196,7 @@ def LRELU(x: np.ndarray):
 
 
 def accuracy(prediction: np.ndarray, target: np.ndarray):
+    assert prediction.size == target.size
     return np.average((target == prediction))
 
 

### 2022-11-11 src/FFNN.py
i gotta go biology man

@@ -267,7 +267,7 @@ class FFNN:
 
         return scores
 
-    def crossval(
+    def cross_val(
         self,
         folds: int,
         X: np.ndarray,
@@ -290,20 +290,21 @@ class FFNN:
         """
         if self.seed:
             np.random.seed(self.seed)
-        matrices = crossval(X, t, folds)
+        cv_data = crossval(X, t, folds, batches)
 
         # avg_weights = [x * 0 for x in self.weights]
-
         avg_scores = None
-        for i in range(len(matrices)):
+        for i in range(len(cv_data)):
+            print(f"{cv_data[i][0]=}")
+            print(f"{cv_data[i][0]=}")
             self.reset_weights()
             scores = self.fit(
-                matrices[i][0],
-                matrices[i][2],
+                cv_data[i][0],
+                cv_data[i][2],
                 scheduler_class,
                 *args,
-                X_test=matrices[i][1],
-                t_test=matrices[i][3],
+                X_test=cv_data[i][1],
+                t_test=cv_data[i][3],
                 lam=lam,
                 batches=batches,
                 epochs=epochs
@@ -427,6 +428,7 @@ class FFNN:
         # put a coloumn of ones as the first coloumn of the design matrix, so that
         # we have a bias term
         X = np.hstack([np.ones((X.shape[0], 1)), X])
+        print(f"{X=}")
 
         # a^0, the nodes in the input layer (one a^0 for each row in X)
         a = X
@@ -443,10 +445,13 @@ class FFNN:
                 self.a_matrices.append(a)
             else:
                 # a^L, the nodes in our output layer
+                print(f"{a=}")
                 z = a @ self.weights[i]
                 a = self.output_func(z)
                 self.a_matrices.append(a)
                 self.z_matrices.append(z)
+                print("OUTPUT")
+                print(f"{z=}")
 
         # this will be a^L
         return a
@@ -464,11 +469,16 @@ class FFNN:
                     delta_matrix = self.a_matrices[i + 1] - t
                 else:
                     try:
+                        
                         cost_func_derivative = grad(self.cost_func(t))
+                        print(f"{self.z_matrices[i+1]=}")
+                        print(f"{self.z_matrices[i+1]=}")
                         delta_matrix = out_derivative(
                             self.z_matrices[i + 1]
                         ) * cost_func_derivative(self.a_matrices[i + 1])
+
                     except ZeroDivisionError:
+                        print("ZeroDivisionError occured in backpropagation")
                         exit()
 
             else:

### 2022-11-11 src/task_a.py

@@ -30,6 +30,7 @@ z_test = z_test.reshape(z_test.shape[0], 1)
 
 # epochs to run for
 epochs = 20
+folds = 1
 
 # no hidden layers, no activation function
 dims = (X.shape[1], 1)
@@ -90,7 +91,7 @@ for i in range(len(schedulers)):
         params_list[i],
         batches=X_train.shape[0],
         epochs=epochs // 2,
-        bootstraps=1,
+        folds=folds
     )
 
     optimal_eta[i] = optimal_params[0]
@@ -142,15 +143,11 @@ for i in range(len(schedulers)):
     plt.title(schedulers[i].__name__, fontsize=22)
 plt.show()
 
-test_errors = np.zeros((len(schedulers), epochs))
-all_biases = np.zeros((len(schedulers), epochs))
-all_variances = np.zeros((len(schedulers), epochs))
-
 # plot best run for each scheduler
 for i in range(len(schedulers)):
     neural = FFNN(dims, seed=1337)
-    test_error, biases, variances = neural.bootstrap(
-        3,
+    scores = neural.cross_val(
+        folds,
         X_train,
         z_train,
         schedulers[i],
@@ -162,9 +159,7 @@ for i in range(len(schedulers)):
         t_test=z_test,
     )
 
-    test_errors[i] = test_error
-    all_biases[i] = biases
-    all_variances[i] = variances
+    test_errors = scores["test_errors"]
 
     plt.plot(test_error, label=f"{schedulers[i].__name__}")
     plt.legend(loc=(1.04, 0))

### 2022-11-11 src/utils.py

@@ -204,33 +204,35 @@ def crossval(
     X: np.ndarray,
     z: np.ndarray,
     K: int,
+    batches: int = 1
 ):
-    chunksize = X.shape[0] // K
-    np.random.seed(3)
+    batch_size = X.shape[0] // batches // K
+    np.random.seed(1337)
     X, z = resample(X, z)
     tuples = list()
 
     for k in range(K):
         if k == K - 1:
             # if we are on the last, take all thats left
-            X_test = X[k * chunksize :, :]
-            z_test = z[k * chunksize :]
+            X_left_out = X[k * batch_size :, :]
+            z_left_out = z[k * batch_size :]
         else:
-            X_test = X[k * chunksize : (k + 1) * chunksize, :]
-            z_test = z[k * chunksize : (k + 1) * chunksize :]
+            X_left_out = X[k * batch_size : (k + 1) * batch_size :]
+            z_left_out = z[k * batch_size : (k + 1) * batch_size :]
 
+        print(f"{z=}")
         X_train = np.delete(
             X,
-            [i for i in range(k * chunksize, k * chunksize + X_test.shape[0])],
+            [i for i in range(k * batch_size, k * batch_size + X_left_out.shape[0])],
             axis=0,
         )
         z_train = np.delete(
             z,
-            [i for i in range(k * chunksize, k * chunksize + z_test.shape[0])],
+            [i for i in range(k * batch_size, k * batch_size + z_left_out.shape[0])],
             axis=0,
         )
 
-        tuples.append((X_train, X_test, z_train, z_test))
+        tuples.append((X_train, X_left_out, z_train, z_left_out))
 
     return tuples
 

### 2022-11-11 src/FFNN.py
i gotta go biology man

@@ -295,8 +295,8 @@ class FFNN:
         # avg_weights = [x * 0 for x in self.weights]
         avg_scores = None
         for i in range(len(cv_data)):
-            print(f"{cv_data[i][0]=}")
-            print(f"{cv_data[i][0]=}")
+            print(f"{cv_data[i][1].shape=}")
+            print(f"{cv_data[i][3].shape=}")
             self.reset_weights()
             scores = self.fit(
                 cv_data[i][0],

### 2022-11-11 src/task_a.py

@@ -30,7 +30,7 @@ z_test = z_test.reshape(z_test.shape[0], 1)
 
 # epochs to run for
 epochs = 20
-folds = 1
+folds = 5
 
 # no hidden layers, no activation function
 dims = (X.shape[1], 1)

### 2022-11-11 src/utils.py

@@ -220,7 +220,7 @@ def crossval(
             X_left_out = X[k * batch_size : (k + 1) * batch_size :]
             z_left_out = z[k * batch_size : (k + 1) * batch_size :]
 
-        print(f"{z=}")
+        print(f"{X.shape=}")
         X_train = np.delete(
             X,
             [i for i in range(k * batch_size, k * batch_size + X_left_out.shape[0])],

### 2022-11-11 src/task_d.py
plotting error or accuracy over model complexity implemented through method optimie_arch in utils.py

@@ -103,8 +103,8 @@ scores = neural.fit(
     t_test=z_test,
 ) """
 
-optimize_arch(FFNN, 
-              60, 
+o1, o2, t1, t2 = optimize_arch(FFNN, 
+              15, 
               [RELU, sigmoid, CostLogReg],
               X_train_sc,
               z_train,
@@ -119,15 +119,16 @@ optimize_arch(FFNN,
 
 
 
-train_errors = scores["train_error"]
-test_errors = scores["test_error"]
-plt.plot(train_errors, label="train")
-plt.plot(test_errors, label="test")
+plt.plot(o1, label="o1")
+plt.plot(o2, label="o2")
 plt.legend()
-plt.xlabel("Epochs")
-plt.ylabel("LogLoss")
-plt.title("LogLoss over Epochs")
 plt.show()
+
+plt.plot(t1, label="t1")
+plt.plot(t2, label="t2")
+plt.legend()
+plt.show()
+
 
 prediction = neural.predict(X_test_sc)
 plot_confusion(prediction, z_test)

### 2022-11-11 src/utils.py

@@ -372,9 +372,9 @@ def optimize_arch(
     two_hid_train = np.zeros(max_nodes)
     two_hid_test = np.zeros(max_nodes)
 
-    for i in range(1, max_nodes + 1, 1):
+    for i in range(1, max_nodes + 1, 3):
         neural = model(
-            (X.shape[1], i+29, t.shape[1]), hidden_func=funcs[0], output_func=funcs[1], cost_func=funcs[2]
+            (X.shape[1], i, t.shape[1]), hidden_func=funcs[0], output_func=funcs[1], cost_func=funcs[2]
         )
         print(neural.dimensions)
         scores = neural.fit(
@@ -387,16 +387,16 @@ def optimize_arch(
             lam=lams,
             X_test=X_test,
             t_test=t_test,
+            use_best_weights=True,
         )
         if classify:
-            one_hid_test[i] = scores["test_acc"]
-            one_hid_train[i] = score["train_acc"]
+            one_hid_test[i] = scores["final_test_acc"]
+            one_hid_train[i] = scores["final_train_acc"]
         else: 
-            one_hid_test[i] = scores["test_acc"]
-            one_hid_train[i] = scores["train_acc"]
-        self.reset_weights()
+            one_hid_test[i] = scores["final_test_error"]
+            one_hid_train[i] = scores["final_train_error"]
 
-    for j in range(1, (max_nodes//2)+1, 1): 
+    for j in range(1, (max_nodes//2)+1, 3): 
         neural = model(
             (X.shape[1], j, j, t.shape[1]), hidden_func=funcs[0], output_func=funcs[1], cost_func=funcs[2]
         )
@@ -412,12 +412,11 @@ def optimize_arch(
             t_test=t_test,
         )
         if classify:
-            two_hid_test[j] = scores["test_acc"]
-            two_hid_train[j] = score["train_acc"]
+            two_hid_test[j] = scores["final_test_acc"]
+            two_hid_train[j] = scores["final_train_acc"]
         else: 
-            two_hid_test[j] = scores["test_acc"]
-            two_hid_train[j] = scores["train_acc"]
-        self.reset_weights()
+            two_hid_test[j] = scores["final_test_error"]
+            two_hid_train[j] = scores["final_train_error"]
 
     return one_hid_train, one_hid_test, two_hid_train, two_hid_test
 

### 2022-11-11 src/FFNN.py
Fix bugs related to crossvalidation

@@ -151,8 +151,9 @@ class FFNN:
                         X_batch = X[i * batch_size :, :]
                         t_batch = t[i * batch_size :, :]
                     else:
+                        # print(f"{t=}")
                         X_batch = X[i * batch_size : (i + 1) * batch_size, :]
-                        t_batch = t[i * batch_size : (i + 1) * batch_size :, :]
+                        t_batch = t[i * batch_size : (i + 1) * batch_size, :]
 
                     self._feedforward(X_batch)
                     self._backpropagate(X_batch, t_batch, lam)
@@ -173,16 +174,16 @@ class FFNN:
                     break
                 if test_set:
                     prediction_test = self.predict(X_test, raw=True)
-                    test_errors = cost_function_test(prediction_test)
+                    test_error = cost_function_test(prediction_test)
                     # test_preds[:, e] = prediction_test.ravel()
 
                     if use_best_weights:
-                        if test_errors < best_test_error:
-                            best_test_error = test_errors
+                        if test_error < best_test_error:
+                            best_test_error = test_error
                             best_train_error = train_error
                             best_weights = deepcopy(self.weights)
                     else:
-                        best_test_error = test_errors
+                        best_test_error = test_error
                         best_train_error = train_error
 
                 else:
@@ -206,14 +207,14 @@ class FFNN:
                             best_train_acc = train_acc
 
                 train_errors[e] = train_error
-                test_errors[e] = test_errors
+                test_errors[e] = test_error
                 progression = e / epochs
 
                 # ----- printing progress bar ------------
                 length = self._progress_bar(
                     progression,
                     train_error=train_error,
-                    test_errors=test_errors,
+                    test_error=test_error,
                     train_acc=train_acc,
                     test_acc=test_acc,
                 )
@@ -237,7 +238,7 @@ class FFNN:
         self._progress_bar(
             1,
             train_error=train_error,
-            test_errors=test_errors,
+            test_error=test_error,
             train_acc=train_acc,
             test_acc=test_acc,
         )
@@ -279,6 +280,7 @@ class FFNN:
         lam: float = 0,
         X_test: np.ndarray = None,
         t_test: np.ndarray = None,
+        use_best_weights=False
     ):
         """Bootstrap
         Parameters:
[…]

### 2022-11-11 src/task_a.py

@@ -51,6 +51,7 @@ batches_list = np.logspace(
 
 # schedulers to test for
 schedulers = [Constant, Momentum, Adagrad, AdagradMomentum, RMS_prop, Adam]
+schedulers = [Constant]
 
 # parameters for schedulers
 constant_params = []
@@ -63,11 +64,11 @@ adam_params = [rho, rho2]
 # list of scheduler parameters
 params_list = [
     constant_params,
-    momentum_params,
-    adagrad_params,
-    adagrad_momentum_params,
-    rms_params,
-    adam_params,
+    # momentum_params,
+    # adagrad_params,
+    # adagrad_momentum_params,
+    # rms_params,
+    # adam_params,
 ]
 
 # results
@@ -91,7 +92,7 @@ for i in range(len(schedulers)):
         params_list[i],
         batches=X_train.shape[0],
         epochs=epochs // 2,
-        folds=folds
+        folds=folds,
     )
 
     optimal_eta[i] = optimal_params[0]
@@ -161,7 +162,7 @@ for i in range(len(schedulers)):
 
     test_errors = scores["test_errors"]
 
-    plt.plot(test_error, label=f"{schedulers[i].__name__}")
+    plt.plot(test_errors, label=f"{schedulers[i].__name__}")
     plt.legend(loc=(1.04, 0))
 
 best_MSE_analytically = np.zeros(epochs)
@@ -174,14 +175,14 @@ plt.ylabel("MSE", fontsize=18)
 plt.title("MSE over Epochs for different schedulers", fontsize=22)
 plt.show()
 
-# plot bias-variance trade-off
-for i in range(len(schedulers)):
-    plt.subplot(321 + i)
-    plt.suptitle("Bias-variance trade-off", fontsize=22)
-
-    plt.title(schedulers[i].__name__, fontsize=22)
-    plt.plot(test_errors[i], label="MSE score")
-    plt.plot(all_biases[i], label="bias")
-    plt.plot(all_variances[i], label="variance")
-    plt.legend(loc=(1.04, 0))
-plt.show()
+# # plot bias-variance trade-off
+# for i in range(len(schedulers)):
+#     plt.subplot(321 + i)
+#     plt.suptitle("Bias-variance trade-off", fontsize=22)
+#
+#     plt.title(schedulers[i].__name__, fontsize=22)
+#     plt.plot(test_errors[i], label="MSE score")
+#     plt.plot(all_biases[i], label="bias")
[…]

### 2022-11-11 src/utils.py

@@ -206,7 +206,7 @@ def crossval(
     K: int,
     batches: int = 1
 ):
-    batch_size = X.shape[0] // batches // K
+    batch_size = X.shape[0] // K
     np.random.seed(1337)
     X, z = resample(X, z)
     tuples = list()
@@ -215,12 +215,11 @@ def crossval(
         if k == K - 1:
             # if we are on the last, take all thats left
             X_left_out = X[k * batch_size :, :]
-            z_left_out = z[k * batch_size :]
+            z_left_out = z[k * batch_size :, :]
         else:
-            X_left_out = X[k * batch_size : (k + 1) * batch_size :]
-            z_left_out = z[k * batch_size : (k + 1) * batch_size :]
+            X_left_out = X[k * batch_size : (k + 1) * batch_size, :]
+            z_left_out = z[k * batch_size : (k + 1) * batch_size, :]
 
-        print(f"{X.shape=}")
         X_train = np.delete(
             X,
             [i for i in range(k * batch_size, k * batch_size + X_left_out.shape[0])],
@@ -232,7 +231,7 @@ def crossval(
             axis=0,
         )
 
-        tuples.append((X_train, X_left_out, z_train, z_left_out))
+        tuples.append((X_train, z_train, X_left_out, z_left_out))
 
     return tuples
 

### 2022-11-11 src/task_a.py
Update crossval

@@ -156,8 +156,6 @@ for i in range(len(schedulers)):
         batches=optimal_batch,
         epochs=epochs,
         lam=optimal_lambdas[i],
-        X_test=X_test,
-        t_test=z_test,
     )
 
     test_errors = scores["test_errors"]

### 2022-11-11 src/task_a.py
for merge

@@ -51,7 +51,7 @@ batches_list = np.logspace(
 
 # schedulers to test for
 schedulers = [Constant, Momentum, Adagrad, AdagradMomentum, RMS_prop, Adam]
-schedulers = [Constant]
+# schedulers = [Constant]
 
 # parameters for schedulers
 constant_params = []
@@ -64,11 +64,11 @@ adam_params = [rho, rho2]
 # list of scheduler parameters
 params_list = [
     constant_params,
-    # momentum_params,
-    # adagrad_params,
-    # adagrad_momentum_params,
-    # rms_params,
-    # adam_params,
+    momentum_params,
+    adagrad_params,
+    adagrad_momentum_params,
+    rms_params,
+    adam_params,
 ]
 
 # results

### 2022-11-11 src/task_d.py
Small print improvement

@@ -83,4 +83,6 @@ plt.ylabel("LogLoss")
 plt.title("LogLoss over Epochs")
 plt.show()
 
-plot_confusion(scores["confusion"])
+plot_confusion(
+    scores["confusion"], title_text="Confusion matrix for the cancer dataset"
+)

### 2022-11-11 src/utils.py

@@ -283,7 +283,7 @@ def confusion(prediction: np.ndarray, target: np.ndarray):
     return np.array([[true_neg_perc, false_neg_perc], [false_pos_perc, true_pos_perc]])
 
 
-def plot_confusion(confusion_matrix: np.ndarray):
+def plot_confusion(confusion_matrix: np.ndarray, title_text=None):
     fontsize = 40
 
     sns.set(font_scale=4)
@@ -293,6 +293,11 @@ def plot_confusion(confusion_matrix: np.ndarray):
         fmt=".2%",
         cmap="Blues",
     )
+    if title_text:
+        plt.title(title_text)
+    else:
+        plt.title("Confusion matrix")
+
     plt.xlabel("Predicted class")
     plt.ylabel("True class")
     plt.show()

### 2022-11-11 src/task_d.py
Something Something

@@ -1,134 +1,68 @@
-"""
-task b (and task g): plot terrain, approximate terrain with OLS (own implementation and scikit) and calculate MSE, R2 &
-                     beta over model complexity for real data. Performs task_b, so no resampling.
-"""
 # Our own library of functions
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn as sns
+
+from matplotlib.patches import Rectangle
 from utils import *
 from Schedulers import *
 from FFNN import FFNN
-from sklearn import datasets
-from sklearn.model_selection import train_test_split
-from sklearn.preprocessing import StandardScaler
-import matplotlib.pyplot as plt
-from matplotlib.colors import ListedColormap
-from sklearn import metrics
-
+from sklearn.datasets import load_breast_cancer
+from sklearn.model_selection import train_test_split, GridSearchCV, KFold
 
 np.random.seed(1337)
 
-cancer = datasets.load_breast_cancer()
+# read in data
+cancer = load_breast_cancer()
 
 X = cancer.data
 z = cancer.target
 
-
 X_train, X_test, z_train, z_test = train_test_split(X, z)
 
-scaler = StandardScaler()
-scaler.fit(X_train)
-
-scaler = StandardScaler()
-scaler.fit(X)
-X = scaler.transform(X)
-
-X_train_sc = scaler.transform(X_train)
-X_test_sc = scaler.transform(X_test)
-
-
-# parameters
-neural_dims = (30, 60, 1)
-logreg_dims = (30, 1)
-eta = 0.001
-#lam = np.logspace(-5, -1, 5)
-rho = 0.90
-rho2 = 0.999
 z_train = z_train.reshape(z_train.shape[0], 1)
 z_test = z_test.reshape(z_test.shape[0], 1)
-z = z.reshape(z.shape[0], 1)
-batches = 10
 
-#neural = FFNN(neural_dims, hidden_func=RELU, output_func=sigmoid, cost_func=CostLogReg)
-# neural = FFNN(logreg_dims, output_func=sigmoid, cost_func=CostLogReg)
+# epochs to run for
+epochs = 200
+folds = 5
 
-momentum = 0.5
+# no hidden layers, no activation function
+dims = (X.shape[1], 1)
+neural = FFNN(dims, seed=1337, hidden_func=)
 
-sched = RMS_prop
+# parameters to test for
[…]

### 2022-11-11 src/FFNN.py
Fix scaling of the data

@@ -304,8 +304,8 @@ class FFNN:
             t_test = cv_data[i][3]
 
             scaler.fit(X_train)
-            scaler.transform(X_train)
-            scaler.transform(X_test)
+            X_train = scaler.transform(X_train)
+            X_test = scaler.transform(X_test)
 
 
             ratio = X.shape[0] / X_train.shape[0]
@@ -557,8 +557,6 @@ class FFNN:
         self,
         X: np.ndarray,
         t: np.ndarray,
-        X_test: np.ndarray,
-        t_test: np.ndarray,
         scheduler: Scheduler,
         eta: list[float],
         lam: list[float],
@@ -596,8 +594,6 @@ class FFNN:
             ) = self._gridsearch_scheduler(
                 X,
                 t,
-                X_test,
-                t_test,
                 scheduler,
                 eta,
                 lam,
@@ -616,8 +612,6 @@ class FFNN:
             ) = self._gridsearch_momentum(
                 X,
                 t,
-                X_test,
-                t_test,
                 scheduler,
                 eta,
                 lam,
@@ -688,8 +682,6 @@ class FFNN:
         self,
         X,
         t,
-        X_test,
-        t_test,
         scheduler,
         eta,
         lam,
@@ -747,8 +739,6 @@ class FFNN:
         self,
         X,
         t,
-        X_test,
-        t_test,
         scheduler,
         eta,
         lam,

### 2022-11-11 src/task_d.py
More on task d

@@ -17,19 +17,16 @@ cancer = load_breast_cancer()
 
 X = cancer.data
 z = cancer.target
-
-X_train, X_test, z_train, z_test = train_test_split(X, z)
-
-z_train = z_train.reshape(z_train.shape[0], 1)
-z_test = z_test.reshape(z_test.shape[0], 1)
+z = z.reshape(z.shape[0], 1)
 
 # epochs to run for
 epochs = 200
 folds = 5
 
+dims = (30, 60, 1)
+
 # no hidden layers, no activation function
-dims = (X.shape[1], 1)
-neural = FFNN(dims, seed=1337, hidden_func=)
+neural = FFNN(dims, hidden_func=RELU, output_func=sigmoid, cost_func=CostLogReg)
 
 # parameters to test for
 eta = np.logspace(-5, -1, 5)
@@ -40,29 +37,34 @@ rho = 0.9
 rho2 = 0.999
 
 # batches to test for
-batches_list = np.logspace(
-    0, np.log(X_train.shape[0] + 1), 7, base=np.exp(1), dtype=int
-)
+batches_list = np.logspace(0, np.log(X.shape[0] + 1), 7, base=np.exp(1), dtype=int)
 
 # schedulers to test for
 sched = Adam
 adam_params = [rho, rho2]
+adam_params = [0.005, rho, rho2]
 
-optimal_params, optimal_lambda, loss_heatmap = neural.optimize_scheduler(
-    X_train,
-    z_train,
-    X_test,
-    z_test,
-    sched,
-    eta,
-    lam,
-    adam_params,
+# optimal_params, optimal_lambda, loss_heatmap = neural.optimize_scheduler(
+#     X,
+#     z,
+#     sched,
+#     eta,
+#     lam,
+#     adam_params,
+#     batches=7,
+#     epochs=epochs // 2,
+#     folds=folds,
+# )
+
+neural.cross_val(
+    5,
+    X,
+    z,
+    Adam,
+    *adam_params,
     batches=7,
-    epochs=epochs // 2,
-    folds=folds,
+    epochs=1000,
 )
 
 print(optimal_params)
 print(optimal_lambda)
-
-

### 2022-11-11 src/FFNN.py
Fix bug in crossval

@@ -325,7 +325,7 @@ class FFNN:
                 )
 
             if self.cost_func.__name__ == "CostLogReg":
-                avg_confusion += confusion(self.predict(cv_data[i][2]), cv_data[i][3]) / folds
+                avg_confusion += confusion(self.predict(X_test), t_test) / folds
 
             # avg_weights = [avg_weights[i] + self.weights[i] for i in range(len(avg_weights))]
             # print(avg_weights)

### 2022-11-11 src/task_d.py

@@ -42,29 +42,29 @@ batches_list = np.logspace(0, np.log(X.shape[0] + 1), 7, base=np.exp(1), dtype=i
 # schedulers to test for
 sched = Adam
 adam_params = [rho, rho2]
-adam_params = [0.005, rho, rho2]
+# adam_params = [0.005, rho, rho2]
 
-# optimal_params, optimal_lambda, loss_heatmap = neural.optimize_scheduler(
-#     X,
-#     z,
-#     sched,
-#     eta,
-#     lam,
-#     adam_params,
-#     batches=7,
-#     epochs=epochs // 2,
-#     folds=folds,
-# )
-
-neural.cross_val(
-    5,
+optimal_params, optimal_lambda, loss_heatmap = neural.optimize_scheduler(
     X,
     z,
-    Adam,
-    *adam_params,
+    sched,
+    eta,
+    lam,
+    adam_params,
     batches=7,
-    epochs=1000,
+    epochs=epochs // 2,
+    folds=folds,
 )
+
+# neural.cross_val(
+#     5,
+#     X,
+#     z,
+#     Adam,
+#     *adam_params,
+#     batches=7,
+#     epochs=40,
+# )
 
 print(optimal_params)
 print(optimal_lambda)

### 2022-11-11 src/utils.py

@@ -355,7 +355,7 @@ def fmt(value, N=4):
     return f"{value:.{N-n-1}f}"
 
 
-def optimize_arch(
+def plot_arch(
     model, 
     max_nodes,
     funcs,

### 2022-11-11 src/bycomplexity.py
Work on mse by model complexity

@@ -1,0 +1,38 @@
+# Our own library of functions
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn as sns
+
+from matplotlib.patches import Rectangle
+from utils import *
+from Schedulers import *
+from FFNN import FFNN
+from sklearn.datasets import load_breast_cancer
+from sklearn.model_selection import train_test_split, GridSearchCV, KFold
+
+np.random.seed(1337)
+
+# read in data
+cancer = load_breast_cancer()
+
+X = cancer.data
+z = cancer.target
+z = z.reshape(z.shape[0], 1)
+
+# epochs to run for
+epochs = 200
+folds = 5
+
+plot_arch(
+    FFNN, 
+    max_nodes,
+    funcs,
+    X,
+    t,
+    scheduler,
+    *args,
+    lam: float = 0,
+    batches: int = 1,
+    epochs: int = 1000,
+    classify: bool = False,
+):

### 2022-11-11 src/utils.py

@@ -361,13 +361,11 @@ def plot_arch(
     funcs,
     X,
     t,
-    X_test,
-    t_test,
     scheduler,
-    args,
-    lams: float,
-    batch: int,
-    epochs: int,
+    *args,
+    lam: float = 0,
+    batches: int = 1,
+    epochs: int = 1000,
     classify: bool = False,
 ):
 
@@ -381,16 +379,14 @@ def plot_arch(
             (X.shape[1], i, t.shape[1]), hidden_func=funcs[0], output_func=funcs[1], cost_func=funcs[2]
         )
         print(neural.dimensions)
-        scores = neural.fit(
+        scores = neural.cross_val(
             X,
             t,
             scheduler,
             *args,
-            batches=batch,
+            batches=batches,
             epochs=epochs,
-            lam=lams,
-            X_test=X_test,
-            t_test=t_test,
+            lam=lam,
             use_best_weights=True,
         )
         if classify:
@@ -404,16 +400,15 @@ def plot_arch(
         neural = model(
             (X.shape[1], j, j, t.shape[1]), hidden_func=funcs[0], output_func=funcs[1], cost_func=funcs[2]
         )
-        scores = neural.fit(    
+        scores = neural.cross_val(
             X,
             t,
             scheduler,
             *args,
-            batches=batch,
+            batches=batches,
             epochs=epochs,
-            lam=lams,
-            X_test=X_test,
-            t_test=t_test,
+            lam=lam,
+            use_best_weights=True,
         )
         if classify:
             two_hid_test[j] = scores["final_test_acc"]

### 2022-11-11 doc/paper.tex
More on the paper
Add discusison on cancer data

@@
 
-\expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
-  \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
-  \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
-  \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
-  \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
-  \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
-  \do\Y\do\Z}
+% \expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
+%   \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
+%   \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
+%   \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
+%   \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
+%   \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
+%   \do\Y\do\Z}
 
@@
 
+To test our neural networks ability to solve classification problems, we first applied it to the Wisconsin Breast Cancer dataset.
+Using our gridsearch method, we found that the best scheduler was Adam, with an eta of NUMBER and lambda of NUMBER.
+With this, we were able to achieve a crossvalidated test accuracy of $98.6 \%$, which is very good.
+Looking at the confusion matrix (Fig.~\ref{cancerconf}), we see that we are achieving good predictions for both true positives and true negatives.
+Confident
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/cancerconf.png}}
+\caption{Confusion matrix for the cancer dataset using a neural net with a single hidden layer of 100 nodes}
+\label{verification}
+\end{figure}
+
 To test our networks ability to classify data, we trained it on many common datasets used in the development of neural networks.

### 2022-11-11 doc/paper.tex
add introduction

@@
 \section{Introduction}
+Many problems in machine learning cannot be solved analytically.
+Thus, we turn to numerical methods, but this change implicitly comes with computationally heavy caveats as they require large matrix calculations over many iterations.
+Finding a solution becomes a waiting game, finding the many solutions of varying hyperparameters in order to make an informed model selection becomes virtually impossible.
+The problem of model selection is at the heart machine learning, as optimal model selection allows us to make predictions with the greatest possible accuracy.
+Finding ways around the liminations of numerical methods becomes a must for proper usage of these powerful machine learning tools.
 
-Neural Networks are all the rage these days, everyone wants to have a neural network in their home. Ive even seen people install neural networks in their dogs. \cite{ericjona}
+In this paper we will investigate implantations of the numerical optimization technique of gradient descent by researching stochastic gradient descent, along with constant, adaptive and momentum-enhanced schedulers with the aim of decreasing the computational cost of our calculations.
+We will study different supervised learning models which utilize gradient descent such as linear regression, logistic regression and a feed forward neural network (FFNN) in order to gain an understanding of model selection across regression, binary- and multi-class classification problems.
+
+First, we will give an explanation of the methods and theory used in this paper, followed up by a presentation and critical discussion of our results.
+Finally, we will provide a conclusion summarizing the core results and lessons.
+
+% Neural Networks are all the rage these days, everyone wants to have a neural network in their home. Ive even seen people install neural networks in their dogs. \cite{ericjona}
 
@@
 
+-hessian vs constant?
+-constant vs momentum
+-adaptive methods
+
+In order to study the attributes of the different scheduler methods we performed a simple linear regression using gradient descent on the Franke Function.
+IF HESSIAN: Using Newtons method, the computational cost is much greater than simply replacing the Hessian matrix inversion with a constant learning rate as can be seen by TIME COMPARISON.
+Furthermore, we can observe from FIGURE COMPARING HESSIAN TO CONSTANT that the constant scheduler does a good job at reducing the MSE, given that the final MSE score is CONSTANT_SCORE compared to the Newton method's HESSIAN_SCORE.
+
+From FIGURE MOMENTUM v CONSTANT we can observe that adding momentum reduces the amount of epochs necessary to reach a lower MSE score.
+
 \subsection{Neural Network for regression}
@@
 \subsection{Neural Network for binary and multiclass classification}
-
-To test our neural networks ability to solve classification problems, we first applied it to the Wisconsin Breast Cancer dataset.
-Using our gridsearch method, we found that the best scheduler was Adam, with an eta of NUMBER and lambda of NUMBER.
-With this, we were able to achieve a crossvalidated test accuracy of $98.6 \%$, which is very good.
-Looking at the confusion matrix (Fig.~\ref{cancerconf}), we see that we are achieving good predictions for both true positives and true negatives.
-Confident
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=5in]{figure/cancerconf.png}}
-\caption{Confusion matrix for the cancer dataset using a neural net with a single hidden layer of 100 nodes}
-\label{verification}
-\end{figure}
 

### 2022-11-14 src/FFNN.py
Functionality for Hessian added in line 159 in FFNN.py

@@ -155,6 +155,9 @@ class FFNN:
                         # print(f"{t=}")
                         X_batch = X[i * batch_size : (i + 1) * batch_size, :]
                         t_batch = t[i * batch_size : (i + 1) * batch_size, :]
+                    
+                    if scheduler_class.__name__ = "Hessian": 
+                        scheduler_class.set_invH(X_batch, t_batch)
 
                     self._feedforward(X_batch)
                     self._backpropagate(X_batch, t_batch, lam)
@@ -516,6 +519,7 @@ class FFNN:
 
             # gradient_weights = self.a_matrices[i][:, 1:].T @ delta_matrix
             gradient_weights += self.weights[i][1:, :] * lam
+
 
             update_matrix = np.vstack(
                 [

### 2022-11-14 src/task_a.py
minor fix

@@ -89,7 +89,7 @@ for i in range(len(schedulers)):
         eta,
         lam,
         params_list[i],
-        batches=X_train.shape[0],
+        batches=batches_list[3],
         epochs=epochs // 2,
         folds=folds,
     )

### 2022-11-14 src/FFNN.py
small fix

@@ -156,7 +156,7 @@ class FFNN:
                         X_batch = X[i * batch_size : (i + 1) * batch_size, :]
                         t_batch = t[i * batch_size : (i + 1) * batch_size, :]
                     
-                    if scheduler_class.__name__ = "Hessian": 
+                    if scheduler_class.__name__ == "Hessian": 
                         scheduler_class.set_invH(X_batch, t_batch)
 
                     self._feedforward(X_batch)

### 2022-11-14 src/FFNN.py
Changes to Hessian

@@ -156,8 +156,13 @@ class FFNN:
                         X_batch = X[i * batch_size : (i + 1) * batch_size, :]
                         t_batch = t[i * batch_size : (i + 1) * batch_size, :]
                     
-                    if scheduler_class.__name__ = "Hessian": 
-                        scheduler_class.set_invH(X_batch, t_batch)
+                    print(X_batch[:5])
+
+                    if scheduler_class.__name__ == "Hessian": 
+                        for scheduler in self.schedulers_weight: 
+                            scheduler.set_invH(X_batch)
+                        for scheduler in self.schedulers_bias: 
+                            scheduler.set_invH(X_batch)
 
                     self._feedforward(X_batch)
                     self._backpropagate(X_batch, t_batch, lam)

### 2022-11-14 src/Schedulers.py

@@ -17,19 +17,23 @@ class Scheduler:
     def reset(self):
         pass
 
-def Hessian(Scheduler): 
+class Hessian(Scheduler): 
+    def __init__(self, eta):
+        super().__init__(eta)
+
+    def update_change(self, gradient):
+        print(self.invH)
+        return self.invH @ gradient  
+
     
-    def update_change(self, gradient, inv_H):
-        return inv_H @ gradients
-    
-    def newtons_method(X, z): 
+    def set_invH(self, X): 
 
         # concat bias
         # X = np.c_[np.ones((X.shape[0], 1))*0.1, X]
         XT_X = X.T @ X 
     
         H = (2.0/X.shape[0])*XT_X
-        return np.linalg.pinv(H)
+        self.invH = np.linalg.pinv(H)
 
 class Constant(Scheduler):
     def __init__(self, eta):

### 2022-11-14 src/task_d.py
reverted task_d

@@ -38,10 +38,10 @@ rho2 = 0.999
 
 # batches to test for
 batches_list = np.logspace(0, np.log(X.shape[0] + 1), 7, base=np.exp(1), dtype=int)
-
 # schedulers to test for
 sched = Adam
 adam_params = [rho, rho2]
+
 # adam_params = [0.005, rho, rho2]
 
 optimal_params, optimal_lambda, loss_heatmap = neural.optimize_scheduler(

### 2022-11-14 src/FFNN.py
Problems with Hessian continue

@@ -156,11 +156,12 @@ class FFNN:
                         X_batch = X[i * batch_size : (i + 1) * batch_size, :]
                         t_batch = t[i * batch_size : (i + 1) * batch_size, :]
                     
-                    if scheduler_class.__name__ == "Hessian": 
-                        scheduler_class.set_invH(X_batch, t_batch)
+                    if scheduler_class.__name__ == "Hessian":
+                        for scheduler in self.schedulers_weight: 
+                            scheduler.set_invH(X_batch)
 
                     self._feedforward(X_batch)
-                    self._backpropagate(X_batch, t_batch, lam)
+                    self._backpropagate(X_batch, t_batch, lam, scheduler=scheduler_class.__name__)
 
                 # reset schedulers every epoch
                 for scheduler in self.schedulers_weight:
@@ -477,7 +478,7 @@ class FFNN:
         # this will be a^L
         return a
 
-    def _backpropagate(self, X, t, lam):
+    def _backpropagate(self, X, t, lam, scheduler=''):
         out_derivative = derivate(self.output_func)
         hidden_derivative = derivate(self.hidden_func)
         update_list = list()
@@ -519,16 +520,19 @@ class FFNN:
 
             # gradient_weights = self.a_matrices[i][:, 1:].T @ delta_matrix
             gradient_weights += self.weights[i][1:, :] * lam
+            
+            if scheduler == 'Hessian': 
+                update_matrix = self.schedulers_weight[i].update_change(gradient_weights)
+            else: 
+                update_matrix = np.vstack(
+                    [
+                        self.schedulers_bias[i].update_change(
+                            delta_accumulated.reshape(1, delta_accumulated.shape[0])
+                        ),
+                        self.schedulers_weight[i].update_change(gradient_weights),
+                    ]
+                )
 
-
-            update_matrix = np.vstack(
-                [
-                    self.schedulers_bias[i].update_change(
-                        delta_accumulated.reshape(1, delta_accumulated.shape[0])
-                    ),
-                    self.schedulers_weight[i].update_change(gradient_weights),
-                ]
-            )
             update_list.insert(0, update_matrix)
 
         self._update_w_and_b(update_list)

### 2022-11-14 src/Schedulers.py

@@ -20,12 +20,11 @@ class Scheduler:
 class Hessian(Scheduler): 
     def __init__(self, eta):
         super().__init__(eta)
-
+        self.invH = 0
+    
     def update_change(self, gradient):
-        print(self.invH)
         return self.invH @ gradient  
 
-    
     def set_invH(self, X): 
 
         # concat bias

### 2022-11-14 doc/paper.tex
Add more papertime

@@
 
+\subsection{Datasets}
+
+Below is a short description of each dataset used in this report, along with a table (Table.~\ref{datasettable}) listing their characteristics.
+
+\subsubsection{Franke Function}
+
+The Franke Function is a simple weighted sum of polynomials, taking two inputs and returning a real value. It is commonly used to test machine learning algorithms for regression problems.
+
+\subsubsection{Elevation data over Norway}
+
+This data is similar to the Franke Function in that it takes two inputs (x and y coordinates) and returns a real value, which is the elevation of a specific point in Norway.
+
+\subsubsection{Wisconsin Breast Cancer Dataset}
+
+This dataset contains 30 features describing important characteristics of tumors.
+The features are real numbers.
+Each instance has an associated class, indicating whether the tumor was benign or not.
+This leads to a binary classification problem.
+
+\subsubsection{MNIST 8x8}
+
+This dataset contains a preprocessed set of black and white images of hand drawn digits.
+The images are 8 by 8 pixels, leading to 64 features, where each feature is an integer in the range $[0-16]$ describing the intensity of the pixel at that position\footnote{in reality, this describes the number of "on" pixels in a 4x4 region in the original image}.
+Each image has an associated class, which is simply the digit depicted.
+This leads to a multiclass classification problem with 10 classes.
+
+\begin{table}[h]
+\caption{Characteristics of each dataset used in the report}
+\begin{center}
+\label{datasettable}
+\begin{tabular}{| c | l | l | l |}
+\hline
+Dataset & Number of instances & Features & Type \\
+\hline
+Franke Function & synthetic & 2 & regression \\
+Terrain (downscaled) & 204 800 & 2 & regression \\
+Wisconsin Breast Cancer & 569 & 30 & binary classification \\
+MNIST 8x8 & 1797 & 64 & multiclass classification (10 classes) \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \section{Results}
@@
 IF HESSIAN: Using Newtons method, the computational cost is much greater than simply replacing the Hessian matrix inversion with a constant learning rate as can be seen by TIME COMPARISON.
-Furthermore, we can observe from FIGURE COMPARING HESSIAN TO CONSTANT that the constant scheduler does a good job at reducing the MSE, given that the final MSE score is CONSTANT_SCORE compared to the Newton method's HESSIAN_SCORE.
+Furthermore, we can observe from FIGURE COMPARING HESSIAN TO CONSTANT that the constant scheduler does a good job at reducing the MSE, given that the final MSE score is CONSTANTSCORE compared to the Newton method's HESSIANSCORE.
 
@@
 To test our neural networks ability to solve classification problems, we first applied it to the Wisconsin Breast Cancer dataset.
-Using our gridsearch method, we found that the best scheduler was Adam, with an eta of NUMBER and lambda of NUMBER.
+Using our gridsearch method, we found that the best scheduler was Adam, with an eta of $0.01$ and no regularization.
+For our architecture, we chose to use only one hidden layer with 60 nodes, based on our testing.
 With this, we were able to achieve a crossvalidated test accuracy of $98.6 \%$, which is very good.
 Looking at the confusion matrix (Fig.~\ref{cancerconf}), we see that we are achieving good predictions for both true positives and true negatives.
-Confident
 
@@
 \centerline{\includegraphics[width=5in]{figure/cancerconf.png}}
-\caption{Confusion matrix for the cancer dataset using a neural net with a single hidden layer of 100 nodes}
+\caption{Confusion matrix for the cancer dataset using a neural net with a single hidden layer of 60 nodes}
 \label{verification}
@@
 
-To test our networks ability to classify data, we trained it on many common datasets used in the development of neural networks.
-The datasets were the Wisconsin Breast Cancer dataset and the 8x8 MNIST dataset, as well as other datasets\footnote{Sonar dataset, House dataset}.
-We found good results on all datasets achieving a test accuracy of over 97\% on all
+Confident in our networks's ability to to binary classification, we turned to the MNIST dataset.
+Using our gridsearch, we again found that the best scheduler was Adam.
+Our architecture in this case was a single hidden layer with 100 nodes.
+Being a multiclass problem with ten different classes, one might expect our neural network to struggle more with this dataset.
+However, this was not the case, as we achieved a crossvalidated test accuracy of $99.07$.
+From these two datasets, we can see that a correctly optimized neural network is adept at solving classification problems.
+On the other hand, these good results, and the easew with which our network attained them, could imply that the problems we have dealt with are not as complicated as they may seem.
 
@@
 
-Although our neural network performed well, we also found very similar results with logistic regression, at a much lower computational cost.
+Based on our good results with the neural net, we attempt to solve the same problems using simple logistic regression.
+Using logistic regression on the cancer dataset, we found through our gridsearch that Adam with an eta $0.1$ gave the best results.
+Here we found that logistic regression performed essentially equivalent or even better, at a much lower computational cost.
+Looking at Fig.~\ref{cancerlogneur}, we see that logistic regression keeps up and even surpasses our neural network in accuracy per epoch.
+The value of logistic regression becomes even greater when we consider the computational cost of each method.
+In this case, the logistic regression was about 65 \% faster\footnote{The time taken to complete a crossvalidation with 5 folds up to 200 epochs was about 52 seconds for our neural network, but only 18 seconds for logistic regression}, with results surpassing our neural network (see Fig.~\ref{cancerconflog}).
 This implies that our classification datasets were perhaps too simple, and that using Neural Networks was not needed to achieve good results.
 
-% \begin{table}
-% \begin{center}
-% \label{ols_mse_table_dif_data}
-% \begin{tabular}{c l l l}
-% Dataset & Std of noise & Best MSE & Best degree \\
-% \hline
-% Franke & 0 & 0.0005109 & 9\\
-% Franke & 0.05 & 0.0038302 & 7 \\
-% Franke & 0.10 & 0.0139833 & 7 \\
-% Terrain 1 & - & 0.0142683 & 19 \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \end{table}
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/cancerlogneur.png}}
+\caption{Comparison between the test accuracy of logistic regression and a neural network on classifying the cancer dataset}
+\label{cancerlogneur}
+\end{figure}
 
-% \begin{table*}[t]
-% \caption{Best MSE for OLS, Ridge and Lasso regression on real elevation data. We only fit up to degree 9 due to computational constraints}
-% \begin{center}
-% \label{ols_vs_ridge_vs_lasso_table_real}
-% \begin{tabular}{c | l l l}
-% Regression & MSE & Degree & $\lambda$ \\
-% \hline
-% OLS & 0.01538 & 9 & - \\
-% Ridge & 0.01538 & 9 & 0 \\
-% Lasso & 0.01685 & 9 & $2.154 \cdot 10^{-7}$ \\
-% \hline
-% \end{tabular}
-% \end{center}
-% \end{table*}
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/cancerconflog.png}}
+\caption{Confusion matrix for the cancer dataset using logistic regression}
+\label{cancerconflog}
+\end{figure}
 

### 2022-11-14 doc/paper.tex
scheduler results

--- 2022-11-11 doc/paper.tex
@@
 -adaptive methods
+-batch sizes
 
-In order to study the attributes of the different scheduler methods we performed a simple linear regression using gradient descent on the Franke Function.
+In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function.
 IF HESSIAN: Using Newtons method, the computational cost is much greater than simply replacing the Hessian matrix inversion with a constant learning rate as can be seen by TIME COMPARISON.
@@
 From FIGURE MOMENTUM v CONSTANT we can observe that adding momentum reduces the amount of epochs necessary to reach a lower MSE score.
+Already after EPOCHS epochs, the momentum scheduler has achieved MOMENTUM MSE AT EPOCHS, a notable improvement over CONSTANT MSE AT EPOCHS achieved by a constant learning rate.
+Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs.
+We can also observe that the different schedulers that they all reach REASONABLE MSE within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data.
+Looking at TIME TABLE FOR ALL SCHEDULERS, BEST took the least time, beating out our baseline constant scheduler by TIME DIFF.
+Surprisingly, RMS prop performed more poorly than expected.
+
+Before continuing, it is interesting to note that all the schedulers work best with a combination of similar eta and lambda values when gridsearched for achieving the minimal MSE.
+It is also interesting that they in the long term reach more or less the same error as can be observed in HEATMAP FIG.
+This could be because given more epochs, the other schedulers are given more time to catch up as the best performing schedulers, such as Adam, only decrease their MSE slightly.
+However, as can be seen in HEATMAP FIG, Adam offers a $\frac{0.018}{0.016} \dot 100 = 11.11\%$ decrease in MSE compared to the constant schedulers.
+
+However, when trained for 500 epochs, Adam with optimal parameters does not yield any improvement in MSE.
+Scikits Adam scheduler does perform better, further decreasing the MSE slightly.
+Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model.
+Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of ADAM FINAL MSE and Scikits final MSE of SCIKIT FINAL MSE both falling short of the analytical MSE of 0,003.
+In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions.
+This result might be due to 500 epochs not being enough training time.
 

### 2022-11-14 doc/paper.tex
windows user reading the terminal challenge

--- 2022-11-14 doc/paper.tex
@@
 -adaptive methods
+-batch sizes
 
-In order to study the attributes of the different scheduler methods we performed a simple linear regression using gradient descent on the Franke Function.
+In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function.
 IF HESSIAN: Using Newtons method, the computational cost is much greater than simply replacing the Hessian matrix inversion with a constant learning rate as can be seen by TIME COMPARISON.
-Furthermore, we can observe from FIGURE COMPARING HESSIAN TO CONSTANT that the constant scheduler does a good job at reducing the MSE, given that the final MSE score is CONSTANTSCORE compared to the Newton method's HESSIANSCORE.
+Furthermore, we can observe from FIGURE COMPARING HESSIAN TO CONSTANT that the constant scheduler does a good job at reducing the MSE, given that the final MSE score is CONSTANT SCORE compared to the Newton method's HESSIAN SCORE.
 
 From FIGURE MOMENTUM v CONSTANT we can observe that adding momentum reduces the amount of epochs necessary to reach a lower MSE score.
+Already after EPOCHS epochs, the momentum scheduler has achieved MOMENTUM MSE AT EPOCHS, a notable improvement over CONSTANT MSE AT EPOCHS achieved by a constant learning rate.
+Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs.
+We can also observe that the different schedulers that they all reach REASONABLE MSE within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data.
+Looking at TIME TABLE FOR ALL SCHEDULERS, BEST took the least time, beating out our baseline constant scheduler by TIME DIFF.
+Surprisingly, RMS prop performed more poorly than expected.
+
+Before continuing, it is interesting to note that all the schedulers work best with a combination of similar eta and lambda values when gridsearched for achieving the minimal MSE.
+It is also interesting that they in the long term reach more or less the same error as can be observed in HEATMAP FIG.
+This could be because given more epochs, the other schedulers are given more time to catch up as the best performing schedulers, such as Adam, only decrease their MSE slightly.
+However, as can be seen in HEATMAP FIG, Adam offers a $\frac{0.018}{0.016} \dot 100 = 11.11\%$ decrease in MSE compared to the constant schedulers.
+
+However, when trained for 500 epochs, Adam with optimal parameters does not yield any improvement in MSE.
+Scikits Adam scheduler does perform better, further decreasing the MSE slightly.
+Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model.
+Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of ADAM FINAL MSE and Scikits final MSE of SCIKIT FINAL MSE both falling short of the analytical MSE of 0,003.
+In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions.
+This result might be due to 500 epochs not being enough training time.
 

### 2022-11-14 src/FFNN.py
Small fixes

@@ -141,7 +141,7 @@ class FFNN:
             self.schedulers_weight.append(scheduler_class(*args))
             self.schedulers_bias.append(scheduler_class(*args))
 
-        print(scheduler_class.__name__)
+        print(f"{scheduler_class.__name__}: Eta={args[0]}, Lambda={lam}")
         # this try is only so that we may cancel early by hitting ctrl+c
         try:
             for e in range(epochs):
@@ -288,14 +288,13 @@ class FFNN:
         cv_data = crossval(X, t, folds)
 
         if self.cost_func.__name__ == "CostLogReg":
-            avg_confusion = np.array([[0.,0.],[0.,0.]])
+            avg_confusion = np.array([[0.0, 0.0], [0.0, 0.0]])
 
         avg_weights = [x * 0 for x in self.weights]
         # print(avg_weights)
         avg_scores = None
         for i in range(len(cv_data)):
             scaler = MinMaxScaler()
-
 
             X_train = cv_data[i][0]
             t_train = cv_data[i][1]
@@ -306,7 +305,6 @@ class FFNN:
             scaler.fit(X_train)
             X_train = scaler.transform(X_train)
             X_test = scaler.transform(X_test)
-
 
             ratio = X.shape[0] / X_train.shape[0]
             scaled_batches = int(batches / ratio)
@@ -322,7 +320,7 @@ class FFNN:
                 batches=scaled_batches,
                 epochs=epochs,
                 use_best_weights=use_best_weights,
-                )
+            )
 
             if self.cost_func.__name__ == "CostLogReg":
                 avg_confusion += confusion(self.predict(X_test), t_test) / folds
@@ -330,12 +328,12 @@ class FFNN:
             # avg_weights = [avg_weights[i] + self.weights[i] for i in range(len(avg_weights))]
             # print(avg_weights)
             if not avg_scores:
-                avg_scores = scores 
+                avg_scores = scores
                 for key in avg_scores:
-                    avg_scores[key] /=  folds
+                    avg_scores[key] /= folds
             else:
                 for key in avg_scores:
-                    avg_scores[key] +=  scores[key] / folds
+                    avg_scores[key] += scores[key] / folds
 
         # avg_weights = [avg_weights[i] / folds for i in range(len(avg_weights))]
         # self.weights = avg_weights
@@ -479,7 +477,6 @@ class FFNN:
         hidden_derivative = derivate(self.hidden_func)
         update_list = list()
 
-
         for i in range(len(self.weights) - 1, -1, -1):
 
[…]

### 2022-11-14 src/task_d.py

@@ -9,6 +9,7 @@ from Schedulers import *
 from FFNN import FFNN
 from sklearn.datasets import load_breast_cancer
 from sklearn.model_selection import train_test_split, GridSearchCV, KFold
+import time
 
 np.random.seed(1337)
 
@@ -23,16 +24,12 @@ z = z.reshape(z.shape[0], 1)
 epochs = 200
 folds = 5
 
-dims = (30, 60, 1)
-
-# no hidden layers, no activation function
-neural = FFNN(dims, hidden_func=RELU, output_func=sigmoid, cost_func=CostLogReg)
 
 # parameters to test for
 eta = np.logspace(-5, -1, 5)
 lam = np.logspace(-5, -1, 5)
-momentums = np.linspace(0, 0.1, 5)
 lam[0] = 0
+momentums = np.linspace(0, 0.1, 5)
 rho = 0.9
 rho2 = 0.999
 
@@ -44,27 +41,93 @@ sched = Adam
 adam_params = [rho, rho2]
 # adam_params = [0.005, rho, rho2]
 
-optimal_params, optimal_lambda, loss_heatmap = neural.optimize_scheduler(
-    X,
-    z,
-    sched,
-    eta,
-    lam,
-    adam_params,
-    batches=7,
-    epochs=epochs // 2,
-    folds=folds,
+dims = (30, 60, 1)
+neural = FFNN(
+    dims, hidden_func=RELU, output_func=sigmoid, cost_func=CostLogReg, seed=1337
 )
-
-# neural.cross_val(
-#     5,
+# optimal_params, optimal_lambda, loss_heatmap = neural.optimize_scheduler(
 #     X,
 #     z,
-#     Adam,
-#     *adam_params,
+#     sched,
+#     eta,
+#     lam,
+#     adam_params,
 #     batches=7,
-#     epochs=40,
+#     epochs=epochs // 2,
+#     folds=folds,
 # )
+
+# optimal params found to be:
+optimal_params = [0.01, 0.9, 0.999]
+optimal_lambda = 0
+
+
+start = time.time()
 
 print(optimal_params)
 print(optimal_lambda)
+
+neuralscores = neural.cross_val(
+    folds,
+    X,
+    z,
+    Adam,
+    *optimal_params,
+    batches=7,
+    epochs=epochs,
+    lam=optimal_lambda,
+    use_best_weights=True,
+)
+print(f"Time taken: {time.time() - start:.4f}")
+plot_confusion(
+    neuralscores["confusion"],
[…]

### 2022-11-14 src/task_d_digits.py

@@ -1,0 +1,77 @@
+# Our own library of functions
+import matplotlib.pyplot as plt
+import numpy as np
+import seaborn as sns
+
+from matplotlib.patches import Rectangle
+from utils import *
+from Schedulers import *
+from FFNN import FFNN
+from sklearn.datasets import load_digits
+from sklearn.model_selection import train_test_split, GridSearchCV, KFold
+
+np.random.seed(1337)
+
+# read in data
+digits = load_digits()
+
+X = digits.data
+z = digits.target
+
+print(X[0])
+z = onehot(z)
+# z = z.reshape(z.shape[0], 1)
+
+# epochs to run for
+epochs = 200
+folds = 5
+
+dims = (64, 100, 10)
+
+# no hidden layers, no activation function
+neural = FFNN(dims, hidden_func=RELU, output_func=softmax, cost_func=CostCrossEntropy)
+
+# parameters to test for
+eta = np.logspace(-5, -1, 5)
+lam = np.logspace(-5, -1, 5)
+momentums = np.linspace(0, 0.1, 5)
+lam[0] = 0
+rho = 0.9
+rho2 = 0.999
+
+# batches to test for
+batches_list = np.logspace(0, np.log(X.shape[0] + 1), 7, base=np.exp(1), dtype=int)
+
+# schedulers to test for
+sched = Adam
+adam_params = [rho, rho2]
+# adam_params = [0.005, rho, rho2]
+
+# optimal_params, optimal_lambda, loss_heatmap = neural.optimize_scheduler(
+#     X,
+#     z,
+#     sched,
+#     eta,
+#     lam,
+#     adam_params,
+#     batches=7,
+#     epochs=epochs // 2,
+#     folds=folds,
+# )
+
+eta = 0.001
+adam_params = [eta, rho, rho2]
+
+scores = neural.cross_val(
+    5,
+    X,
+    z,
+    Adam,
+    *adam_params,
+    batches=7,
+    epochs=200,
+    use_best_weights=True,
+)
+
+print(f"{scores['final_train_acc']=}")
+print(f"{scores['final_test_acc']=}")

### 2022-11-14 src/FFNN.py
Remove Hessian from FFNN
Too complicated

@@ -155,13 +155,9 @@ class FFNN:
                         # print(f"{t=}")
                         X_batch = X[i * batch_size : (i + 1) * batch_size, :]
                         t_batch = t[i * batch_size : (i + 1) * batch_size, :]
-                    
-                    if scheduler_class.__name__ == "Hessian":
-                        for scheduler in self.schedulers_weight: 
-                            scheduler.set_invH(X_batch)
 
                     self._feedforward(X_batch)
-                    self._backpropagate(X_batch, t_batch, lam, scheduler=scheduler_class.__name__)
+                    self._backpropagate(X_batch, t_batch, lam)
 
                 # reset schedulers every epoch
                 for scheduler in self.schedulers_weight:
@@ -476,7 +472,7 @@ class FFNN:
         # this will be a^L
         return a
 
-    def _backpropagate(self, X, t, lam, scheduler=''):
+    def _backpropagate(self, X, t, lam):
         out_derivative = derivate(self.output_func)
         hidden_derivative = derivate(self.hidden_func)
         update_list = list()
@@ -516,18 +512,15 @@ class FFNN:
 
             # gradient_weights = self.a_matrices[i][:, 1:].T @ delta_matrix
             gradient_weights += self.weights[i][1:, :] * lam
-            
-            if scheduler == 'Hessian': 
-                update_matrix = self.schedulers_weight[i].update_change(gradient_weights)
-            else: 
-                update_matrix = np.vstack(
-                    [
-                        self.schedulers_bias[i].update_change(
-                            delta_accumulated.reshape(1, delta_accumulated.shape[0])
-                        ),
-                        self.schedulers_weight[i].update_change(gradient_weights),
-                    ]
-                )
+
+            update_matrix = np.vstack(
+                [
+                    self.schedulers_bias[i].update_change(
+                        delta_accumulated.reshape(1, delta_accumulated.shape[0])
+                    ),
+                    self.schedulers_weight[i].update_change(gradient_weights),
+                ]
+            )
 
             update_list.insert(0, update_matrix)
 

### 2022-11-14 doc/paper.tex
more scheduler

@@
 This could be because given more epochs, the other schedulers are given more time to catch up as the best performing schedulers, such as Adam, only decrease their MSE slightly.
-However, as can be seen in HEATMAP FIG, Adam offers a $\frac{0.018}{0.016} \dot 100 = 11.11\%$ decrease in MSE compared to the constant schedulers.
+However, as can be seen in HEATMAP FIG, Adam offers a $\frac{0.018}{0.016} \dot 100 = 11.11\%$ decrease in MSE compared to the constant scheduler.
 
@@
 In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions.
-This result might be due to 500 epochs not being enough training time.
+This result might be due to 500 epochs not being enough training time, as the Franke function is quite complex and has added noise.
 

### 2022-11-14 src/testing/verification_hessian.py
Add verification file

@@ -1,0 +1,27 @@
+from FFNN import FFNN
+from utils import *
+from Schedulers import *
+
+
+# data
+x = np.arange(0, 1, 0.01)
+y = 3 + x + x**2
+
+x = x.reshape(x.shape[0], 1)
+y = y.reshape(y.shape[0], 1)
+
+X = np.c_[np.ones((x.shape[0], 1)), x, x**2]
+
+x_train, x_test, y_train, y_test = train_test_split(X, y, test_size=0.2)
+
+scores, beta = hessian(
+    x_train,
+    y_train,
+    epochs=1000,
+    X_test=x_test,
+    t_test=y_test,
+)
+
+plt.plot(scores["train_errors"], label="train")
+plt.plot(scores["test_errors"], label="test")
+plt.show()

### 2022-11-14 doc/paper.tex
heatmap text

--- 2022-11-14 doc/paper.tex
@@
 Already after EPOCHS epochs, the momentum scheduler has achieved MOMENTUM MSE AT EPOCHS, a notable improvement over CONSTANT MSE AT EPOCHS achieved by a constant learning rate.
+
+- GD vs SGD
+
 Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs.
@@
 
-Before continuing, it is interesting to note that all the schedulers work best with a combination of similar eta and lambda values when gridsearched for achieving the minimal MSE.
-It is also interesting that they in the long term reach more or less the same error as can be observed in HEATMAP FIG.
-This could be because given more epochs, the other schedulers are given more time to catch up as the best performing schedulers, such as Adam, only decrease their MSE slightly.
-However, as can be seen in HEATMAP FIG, Adam offers a $\frac{0.018}{0.016} \dot 100 = 11.11\%$ decrease in MSE compared to the constant schedulers.
+An interesting insight from FIGURE NUM is that all the schedulers prefer high learning rates and regularization.
+As this plot was generated for 100 epochs due to the computational complexity of gridsearching the six schedulers, the bias towards higher learning rates is certainly due to the high learning rates being able to reach a low MSE in the short training time.
+With the lambda = 0 column representing the ordinary least squares solutions, these are clearly consistantly outperformed by Ridge regression as the best MSE is achieved with regularization.
+For this specific plot, momentum does not seem to be an important parameter.
+Both the momentum and adagrad with momentum heatmaps are identical to their momentumless counterparts, constant and adagrad, implying that the best gridsearched momentum is 0.
+That momentum is irrelevant in this instance is not a generalizable result, as momentum has played a key role in previous plots and results.
 
-However, when trained for 500 epochs, Adam with optimal parameters does not yield any improvement in MSE.
-Scikits Adam scheduler does perform better, further decreasing the MSE slightly.
-Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model.
-Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of ADAM FINAL MSE and Scikits final MSE of SCIKIT FINAL MSE both falling short of the analytical MSE of 0,003.
+Another observation from FIG NUM  is that the schedulers final error for100 epochs are more or less the same.
+As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adam, starts slowing down the rate at which the MSE is decreased.
+However, as can be seen in HEATMAP FIG, Adam offers a 11.11 \% \footnote{$\frac{constant MSE}{adam MSE} * 100 = \frac{0.018}{0.016} * 100 = 11.11\%$} decrease in MSE compared to the constant scheduler.
+It is clear from the results thus far that Adam is the best scheduler for approximating the Franke function when using linear regression with gradient descent.
+
+Furthermore, when trained for 5000 epochs, Adam with optimal parameters keeps improving the MSE.
+Scikit learn's Adam scheduler does perform better, offering a greater decrease the MSE.
+Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model, in addition to other optimizations present in the scikit learn implementation.
+Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of an MSE of 0.012 on the test data and Scikits final MSE of 0.004 (granted this is on the train data, as scikit learn does not provide the loss curve for the test data).
+both falling short of the analytical MSE of 0,003.
 In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions.
-This result might be due to 500 epochs not being enough training time.
+This result might be due to 5000 epochs not being enough training time, as the Franke function is a quite complex function with added noise.
 

### 2022-11-14 src/time_plot.py
add time plot

@@ -1,0 +1,67 @@
+from utils import *
+from Schedulers import *
+from FFNN import FFNN
+import timeit
+from tabulate import tabulate
+ 
+np.random.seed(1337)
+
+# read in data
+(
+    betas_to_plot,
+    N,
+    X,
+    X_train,
+    X_test,
+    z,
+    z_train,
+    z_test,
+    centering,
+    x,
+    y,
+    z,
+) = read_from_cmdline()
+z_train = z_train.reshape(z_train.shape[0], 1)
+z_test = z_test.reshape(z_test.shape[0], 1)
+
+# epochs to run for
+epochs = 100
+folds = 5
+
+# batches to test for
+batches_list = np.logspace(
+    0, np.log(X_train.shape[0] + 1), 7, base=np.exp(1), dtype=int
+)
+
+schedulers = [Constant, Momentum, Adagrad, AdagradMomentum, RMS_prop, Adam]
+# parameters dont matter as we just want to test runtime
+optimal_params = [[0.01], [0.01, 0.1], [0.1], [0.1, 0.075], [0.001, 0.9], [0.01, 0.9, 0.999]]
+optimal_lambda = 0.1
+
+# no hidden layers, no activation function
+dims = (X.shape[1], 1)
+neural = FFNN(dims, seed=1337)
+
+time = np.zeros(len(schedulers))
+
+for i in range(len(schedulers)):
+    start = timeit.default_timer()
+    scores = neural.cross_val(
+        folds,
+        X_train,
+        z_train,
+        schedulers[i],
+        *optimal_params[i],
+        batches=batches_list[3],
+        epochs=epochs,
+        lam=optimal_lambda,
+    )
+    stop = timeit.default_timer()
+    time[i] = stop - start
+
+time /= folds
+table = [schedulers, time]
+
+string = tabulate(table)
+with open(f"time_comparison", "w") as file:
+    file.write(string)

### 2022-11-14 src/linreg_franke.py
small change

@@ -53,6 +53,7 @@ batches_list = np.logspace(
 best_MSE_analytically = np.zeros(epochs)
 best_MSE_analytically[:] = 0.00328322949832417
 
+# optimize Adam
 optimal_params, optimal_lambda, _ = neural.optimize_scheduler(
     X_train,
     z_train,
@@ -65,6 +66,7 @@ optimal_params, optimal_lambda, _ = neural.optimize_scheduler(
     folds=folds,
 )
 
+# fit using optimal params
 scores = neural.cross_val(
     folds,
     X_train,
@@ -78,6 +80,7 @@ scores = neural.cross_val(
 test_errors = scores["test_errors"]
 train_errors = scores["train_errors"]
 
+# test against sklearn
 scikit_MLP = MLPRegressor(
     hidden_layer_sizes=[],
     activation="identity",
@@ -97,6 +100,7 @@ for i in range(folds):
     scikit_train_errors[:] += scikit_MLP.loss_curve_
 scikit_train_errors /= folds
 
+# plot
 plt.plot(scikit_train_errors, label="scikit Adam train")
 plt.plot(train_errors, label="Adam train")
 plt.plot(test_errors, label="Adam test")

### 2022-11-14 src/FFNN.py
change write file

@@ -212,7 +212,7 @@ class FFNN:
                 progression = e / epochs
 
                 # ----- printing progress bar ------------
-                length = progress_bar(
+                length = self._progress_bar(
                     progression,
                     train_error=train_error,
                     test_error=test_error,
@@ -236,7 +236,7 @@ class FFNN:
 
         # overwrite last print so that we dont get 99.9 %
         print(" " * length, end="\r")
-        progress_bar(
+        self._progress_bar(
             1,
             train_error=train_error,
             test_error=test_error,
@@ -376,7 +376,8 @@ class FFNN:
         """
         # print(f'Writing weights to file "{path}"')
         np.set_printoptions(threshold=np.inf)
-        with open(path, "w") as file:
+        with i
+         open(path, "w") as file:
             text = str(self.dimensions) + "\n"
             for weight in self.weights:
                 text += str(weight.shape) + "\n"
@@ -534,6 +535,21 @@ class FFNN:
         for i in range(len(self.weights)):
             self.weights[i] -= update_list[i]
 
+    def _progress_bar(self, progression, **kwargs):
+        length = 40
+        num_equals = int(progression * length)
+        num_not = length - num_equals
+        arrow = ">" if num_equals > 0 else ""
+        bar = "[" + "=" * (num_equals - 1) + arrow + "-" * num_not + "]"
+        perc_print = fmt(progression * 100, N=5)
+        line = f"  {bar} {perc_print}% "
+
+        for key in kwargs:
+            if kwargs[key]:
+                value = fmt(kwargs[key], N=4)
+                line += f"| {key}: {value} "
+        print(line, end="\r")
+        return len(line)
 
     def optimize_scheduler(
         self,
@@ -615,9 +631,11 @@ class FFNN:
             + "\n"
             + f"minimal MSE={min_heatmap[np.unravel_index(loss_heatmap.argmin(), loss_heatmap.shape)[0], np.unravel_index(loss_heatmap.argmin(), loss_heatmap.shape)[1]]}"
             + "\n"
+            + f"for epochs={epochs}"
+            + "\n"
         )
         print(string)
-        with open(f"{scheduler.__name__}_optimal_params.txt", "w") as file:
+        with open(f"{scheduler.__name__}_optimal_params.txt", "a") as file:
             file.write(string)
         return optimal_params, optimal_lambda, loss_heatmap
 

### 2022-11-14 src/time_plot.py

@@ -25,7 +25,7 @@ z_train = z_train.reshape(z_train.shape[0], 1)
 z_test = z_test.reshape(z_test.shape[0], 1)
 
 # epochs to run for
-epochs = 100
+epochs = 50
 folds = 5
 
 # batches to test for
@@ -34,17 +34,29 @@ batches_list = np.logspace(
 )
 
 schedulers = [Constant, Momentum, Adagrad, AdagradMomentum, RMS_prop, Adam]
-# parameters dont matter as we just want to test runtime
+# optimal parameters from task_a.py
 optimal_params = [[0.01], [0.01, 0.1], [0.1], [0.1, 0.075], [0.001, 0.9], [0.01, 0.9, 0.999]]
-optimal_lambda = 0.1
+optimal_lambdas = [0.1, 0.1, 0.001, 0.001, 0.001, 0.01]
 
 # no hidden layers, no activation function
 dims = (X.shape[1], 1)
 neural = FFNN(dims, seed=1337)
 
 time = np.zeros(len(schedulers))
+MSE_to_reach = 0.2
 
 for i in range(len(schedulers)):
+    optimal_batch, _ = neural.optimize_batch(
+        X_train,
+        z_train,
+        X_test,
+        z_test,
+        schedulers[i],
+        optimal_lambdas[i],
+        *optimal_params[i],
+        batches_list=batches_list,
+        epochs=epochs,
+    )
     start = timeit.default_timer()
     scores = neural.cross_val(
         folds,
@@ -52,12 +64,20 @@ for i in range(len(schedulers)):
         z_train,
         schedulers[i],
         *optimal_params[i],
-        batches=batches_list[3],
+        batches=optimal_batch,
         epochs=epochs,
-        lam=optimal_lambda,
+        lam=optimal_lambdas[i],
     )
+    index=0
+    test_errors = scores["test_errors"]
+    for j in range(epochs):
+        if test_errors[j] <= MSE_to_reach:
+            index = j
+            break
     stop = timeit.default_timer()
-    time[i] = stop - start
+    runtime = stop - start
+    runtime /= epochs
+    time[i] = runtime * index
 
 time /= folds
 table = [schedulers, time]

### 2022-11-14 doc/paper.tex
update Franke dataset

@@
 
-The Franke Function is a simple weighted sum of polynomials, taking two inputs and returning a real value. It is commonly used to test machine learning algorithms for regression problems.
+The Franke Function is a simple weighted sum of polynomials, taking two inputs and returning a real value.
+It is commonly used to test machine learning algorithms for regression problems.
+For our linear regression problems, our training data will be based on a 9th degree polynomial approximation to the Franke Function, with 55 features.
+For the feed forward neural network, we will use 2 of the features (x, y) of the Franke Function.
 
@@
 \hline
+Franke Function & synthetic & 55 & regression \\
 Franke Function & synthetic & 2 & regression \\

### 2022-11-14 src/FFNN.py
adagrad supremecy

@@ -376,8 +376,7 @@ class FFNN:
         """
         # print(f'Writing weights to file "{path}"')
         np.set_printoptions(threshold=np.inf)
-        with i
-         open(path, "w") as file:
+        with open(path, "w") as file:
             text = str(self.dimensions) + "\n"
             for weight in self.weights:
                 text += str(weight.shape) + "\n"
@@ -635,7 +634,7 @@ class FFNN:
             + "\n"
         )
         print(string)
-        with open(f"{scheduler.__name__}_optimal_params.txt", "a") as file:
+        with open(f"{scheduler.__name__}_optimal_params.txt", "w") as file:
             file.write(string)
         return optimal_params, optimal_lambda, loss_heatmap
 

### 2022-11-14 src/time_plot.py

@@ -25,7 +25,7 @@ z_train = z_train.reshape(z_train.shape[0], 1)
 z_test = z_test.reshape(z_test.shape[0], 1)
 
 # epochs to run for
-epochs = 50
+epochs = 100
 folds = 5
 
 # batches to test for
@@ -46,17 +46,6 @@ time = np.zeros(len(schedulers))
 MSE_to_reach = 0.2
 
 for i in range(len(schedulers)):
-    optimal_batch, _ = neural.optimize_batch(
-        X_train,
-        z_train,
-        X_test,
-        z_test,
-        schedulers[i],
-        optimal_lambdas[i],
-        *optimal_params[i],
-        batches_list=batches_list,
-        epochs=epochs,
-    )
     start = timeit.default_timer()
     scores = neural.cross_val(
         folds,
@@ -64,17 +53,17 @@ for i in range(len(schedulers)):
         z_train,
         schedulers[i],
         *optimal_params[i],
-        batches=optimal_batch,
+        batches=batches_list[3],
         epochs=epochs,
         lam=optimal_lambdas[i],
     )
+    stop = timeit.default_timer()
     index=0
     test_errors = scores["test_errors"]
     for j in range(epochs):
         if test_errors[j] <= MSE_to_reach:
             index = j
             break
-    stop = timeit.default_timer()
     runtime = stop - start
     runtime /= epochs
     time[i] = runtime * index

### 2022-11-15 src/bycomplexity.py
Hessian and plot over model complexity

@@ -27,9 +27,9 @@ args = [0.01, 0.9, 0.999]
 
 funcs = [RELU, sigmoid, CostLogReg]
 
-one_hid_train, one_hid_test, two_hid_train, two_hid_test = plot_arch(
+results = plot_arch(
     FFNN,
-    50,
+    400,
     funcs,
     X,
     z,
@@ -37,19 +37,38 @@ one_hid_train, one_hid_test, two_hid_train, two_hid_test = plot_arch(
     *args,
     lam=0,
     batches=7,
-    epochs=10,
+    epochs=50,
     classify=True,
+    step_size=20
 )
 
-one_hid_train2 = list()
 
-for item in one_hid_train:
-    if not np.isnan(item):
-        one_hid_train2.append(item)
-
-
-print(f"{one_hid_train2=}")
-print(f"{one_hid_test=}")
-plt.plot(one_hid_train2)
-plt.plot(one_hid_test)
+sns.set(font_scale=2)
+plt.title("Accuracy by model complexity for cancer data")
+plt.xlabel("Total amount of hidden nodes")
+plt.ylabel("Accuracy")
+plt.plot(
+    results["node_sizes"],
+    results["one_hid_train"],
+    label="One hidden layer: train",
+    lw=4,
+)
+plt.plot(
+    results["node_sizes"], results["one_hid_test"], label="One hidden layer: test", lw=4
+)
+plt.plot(
+    results["node_sizes"],
+    results["two_hid_train"],
+    "--",
+    label="Two hidden layers: train",
+    lw=4,
+)
+plt.plot(
+    results["node_sizes"],
+    results["two_hid_test"],
+    "--",
+    label="Two hidden layers: test",
+    lw=4,
+)
+plt.legend()
 plt.show()

### 2022-11-15 src/task_d.py

@@ -21,7 +21,7 @@ z = cancer.target
 z = z.reshape(z.shape[0], 1)
 
 # epochs to run for
-epochs = 200
+epochs = 100
 folds = 5
 
 
@@ -41,6 +41,17 @@ adam_params = [rho, rho2]
 
 # adam_params = [0.005, rho, rho2]
 
+X_train, X_test, t_train, t_test = train_test_split(X, z)
+
+hessianscores, _ = hessian(
+    X_train,
+    t_train,
+    epochs=epochs,
+    X_test=X_test,
+    t_test=t_test,
+)
+
+# dims = (30, 60, 1)
 dims = (30, 60, 1)
 neural = FFNN(
     dims, hidden_func=RELU, output_func=sigmoid, cost_func=CostLogReg, seed=1337
@@ -118,6 +129,7 @@ logregscores = logreg.cross_val(
 )
 print(f"Time taken: {time.time() - start}")
 
+
 plot_confusion(
     logregscores["confusion"],
     title="Confusion matrix for cancer data using logistic regression",
@@ -127,6 +139,7 @@ plot_confusion(
 plt.title("Crossvalidated test accuracy per epoch")
 plt.plot(neuralscores["test_accs"], label="Neural", lw=4)
 plt.plot(logregscores["test_accs"], label="Logreg", lw=4)
+plt.plot(hessianscores["test_accs"], label="Hessian", lw=4)
 plt.xlabel("Epochs")
 plt.ylabel("Test accuracy")
 plt.legend()

### 2022-11-15 src/testing/terrain_testing.py

@@ -22,32 +22,39 @@ np.random.seed(42069)
 z_train = z_train.reshape(z_train.shape[0], 1)
 z_test = z_test.reshape(z_test.shape[0], 1)
 
-eta = 0.00005
+eta = 0.001
 momentum = 0.5
 rho = 0.9
 rho2 = 0.99
+epochs = 1000
 
 sched = Adam
-# sched = Momentum
 params = [eta, rho, rho2]
-# params = [eta, momentum]
+
+hessianscores, _ = hessian(
+    X_train,
+    z_train,
+    epochs=epochs,
+    X_test=X_test,
+    t_test=z_test,
+)
 
 dims = (2, 80, 80, 1)
-neural = FFNN(dims, checkpoint_file="weights", hidden_func=RELU)
-# neural.read("weights")
-train_errors, test_errors, _, _ = neural.fit(
+neural = FFNN(dims, hidden_func=RELU)
+scores = neural.fit(
     X_train[:, 1:3],
     z_train,
     sched,
     *params,
     batches=30,
-    epochs=10000,
+    epochs=epochs,
     lam=0.08,
     X_test=X_test[:, 1:3],
     t_test=z_test,
 )
-plt.plot(train_errors, label="Train")
-plt.plot(test_errors, label="Train")
+
+plt.plot(scores["train_errors"], label="Train")
+plt.plot(scores["test_errors"], label="Train")
 plt.legend()
 plt.xlabel("Epochs")
 plt.ylabel("MSE")

### 2022-11-15 src/utils.py

@@ -329,7 +329,7 @@ def hessian(
     train_errors.fill(np.nan)
 
     cost_func_derivative = grad(CostOLS)
-    
+
     for e in range(epochs):
         train_error = CostOLS(beta)
         train_errors[e] = train_error
@@ -341,7 +341,7 @@ def hessian(
 
         gradient = cost_func_derivative(beta)
         beta -= inv_hessian @ gradient
-        
+
         progression = e / epochs
         length = progress_bar(
             progression,
@@ -454,28 +454,26 @@ def plot_arch(
     lam: float = 0,
     batches: int = 1,
     epochs: int = 1000,
-    size_step: int = 10,
+    step_size: int = 10,
     classify=False,
 ):
 
-    one_hid_train = np.empty(max_nodes)
-    one_hid_train.fill(np.nan)
-    one_hid_test = np.empty(max_nodes)
-    one_hid_test.fill(np.nan)
-    two_hid_train = np.empty(max_nodes)
-    two_hid_train.fill(np.nan)
-    two_hid_test = np.empty(max_nodes)
-    two_hid_test.fill(np.nan)
-
-    node_sizes = np.arange(0, max_nodes, size_step)
+    node_sizes = np.arange(0, max_nodes, step_size)
     node_sizes[0] = 1
+    node_sizes[-1] = max_nodes
 
-    for i in range(1, max_nodes, 10):
+    one_hid_train = np.zeros(node_sizes.shape[0])
+    one_hid_test = np.zeros(node_sizes.shape[0])
+    two_hid_train = np.zeros(node_sizes.shape[0])
+    two_hid_test = np.zeros(node_sizes.shape[0])
+
+    for i in range(len(node_sizes)):
         neural = model(
-            (X.shape[1], i, t.shape[1]),
+            (X.shape[1], node_sizes[i], t.shape[1]),
             hidden_func=funcs[0],
             output_func=funcs[1],
             cost_func=funcs[2],
+            seed=1337
         )
         print(neural.dimensions)
         scores = neural.cross_val(
@@ -497,12 +495,14 @@ def plot_arch(
             one_hid_test[i] = scores["final_test_error"]
             one_hid_train[i] = scores["final_train_error"]
 
-    for j in range(1, (max_nodes // 2) + 1, 5):
+    for i in range(len(node_sizes)):
+        node_size = node_sizes[i] // 2 or 1
         neural = model(
-            (X.shape[1], j, j, t.shape[1]),
[…]

### 2022-11-15 doc/paper.tex
adagrad

@@
 Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs.
-We can also observe that the different schedulers that they all reach REASONABLE MSE within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data.
-Looking at TIME TABLE FOR ALL SCHEDULERS, BEST took the least time, beating out our baseline constant scheduler by TIME DIFF.
-Surprisingly, RMS prop performed more poorly than expected.
+We can also observe that the different schedulers all reach an MSE of 0.02 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.2.
+Looking at TIME TABLE FOR ALL SCHEDULERS, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds.
+Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.
 

### 2022-11-15 src/bycomplexity.py
Make hessian crossvalidated

@@ -70,5 +70,5 @@ plt.plot(
     label="Two hidden layers: test",
     lw=4,
 )
-plt.legend()
+plt.legend(loc=(1.04, 0))
 plt.show()

### 2022-11-15 src/testing/verification_hessian.py

@@ -5,22 +5,23 @@ from Schedulers import *
 
 # data
 x = np.arange(0, 1, 0.01)
-y = 3 + x + x**2
+y = 3 + x + x**2 + x**4 + x**5 + x
 
 x = x.reshape(x.shape[0], 1)
 y = y.reshape(y.shape[0], 1)
 
 X = np.c_[np.ones((x.shape[0], 1)), x, x**2]
 
-x_train, x_test, y_train, y_test = train_test_split(X, y, test_size=0.2)
-
-scores, beta = hessian(
-    x_train,
-    y_train,
-    epochs=1000,
-    X_test=x_test,
-    t_test=y_test,
+print(X.shape)
+print(y.shape)
+scores, beta = hessian_cv(
+    5,
+    X,
+    y,
+    epochs=5,
 )
+
+print(scores)
 
 plt.plot(scores["train_errors"], label="train")
 plt.plot(scores["test_errors"], label="test")

### 2022-11-15 src/utils.py
Added wrong utils

@@ -301,11 +301,37 @@ def progress_bar(progression, **kwargs):
     return len(line)
 
 
+def hessian_cv(
+    K,
+    X,
+    t,
+    epochs: int = 1000,
+):
+    matrices = crossval(X, t, K)
+
+    avgbeta = np.zeros((X.shape[1], 1))
+
+    test_errors = np.zeros(epochs)
+    train_errors = np.zeros(epochs)
+
+    for cv in matrices:
+        X_train = cv[0]
+        t_train = cv[1]
+        X_test = cv[2]
+        t_test = cv[3]
+
+        scores, beta = hessian(X_train, t_train, epochs=epochs, X_test=X_test, t_test=t_test)
+        test_errors += scores["test_errors"] / K
+        train_errors += scores["train_errors"] / K
+        avgbeta += beta / K
+
+    return {"test_errors": test_errors, "train_errors": train_errors}, avgbeta
+
+
 def hessian(
     X,
     t,
     epochs: int = 1000,
-    lam: float = 0,
     X_test: np.ndarray = None,
     t_test: np.ndarray = None,
 ):
@@ -350,6 +376,7 @@ def hessian(
         )
 
     scores = dict()
+    print()
 
     scores["train_errors"] = train_errors
 
@@ -473,7 +500,7 @@ def plot_arch(
             hidden_func=funcs[0],
             output_func=funcs[1],
             cost_func=funcs[2],
-            seed=1337
+            seed=1337,
         )
         print(neural.dimensions)
         scores = neural.cross_val(
@@ -502,7 +529,7 @@ def plot_arch(
             hidden_func=funcs[0],
             output_func=funcs[1],
             cost_func=funcs[2],
-            seed=1337
+            seed=1337,
         )
         print(neural.dimensions)
         scores = neural.cross_val(

### 2022-11-15 src/FFNN.py
Work more on task d and complexity

@@ -253,7 +253,7 @@ class FFNN:
         scores = dict()
 
         scores["train_errors"] = train_errors
-        scores["final_train_error"] = best_test_error
+        scores["final_train_error"] = best_train_error
 
         if test_set:
             scores["test_errors"] = test_errors

### 2022-11-15 src/bycomplexityfranke.py

@@ -13,38 +13,51 @@ from sklearn.model_selection import train_test_split, GridSearchCV, KFold
 np.random.seed(1337)
 
 # read in data
-cancer = load_breast_cancer()
-
-X = cancer.data
-z = cancer.target
+(
+    betas_to_plot,
+    N,
+    X,
+    X_train,
+    X_test,
+    z,
+    z_train,
+    z_test,
+    centering,
+    x,
+    y,
+    z,
+) = read_from_cmdline()
+z_train = z_train.reshape(z_train.shape[0], 1)
+z_test = z_test.reshape(z_test.shape[0], 1)
+z = z.ravel()
 z = z.reshape(z.shape[0], 1)
 
 # epochs to run for
-epochs = 200
+epochs = 10
 folds = 5
 scheduler = Adam
-args = [0.01, 0.9, 0.999]
+args = [0.001, 0.9, 0.999]
 
-funcs = [sigmoid, sigmoid, CostLogReg]
+funcs = [sigmoid, lambda x: x, CostOLS]
 
 results = plot_arch(
     FFNN,
     400,
     funcs,
-    X,
+    X[:, 1:3],
     z,
     scheduler,
     *args,
     lam=0,
     batches=7,
-    epochs=50,
-    classify=True,
-    step_size=60
+    epochs=800,
+    step_size=40,
+    folds=3,
 )
 
 
 sns.set(font_scale=2)
-plt.title("Accuracy by model complexity for cancer data")
+plt.title("Accuracy by model complexity for franke")
 plt.xlabel("Total amount of hidden nodes")
 plt.ylabel("Accuracy")
 plt.plot(
@@ -54,12 +67,15 @@ plt.plot(
     lw=4,
 )
 plt.plot(
-    results["node_sizes"], results["one_hid_test"], label="One hidden layer: test", lw=4
+    results["node_sizes"],
+    results["one_hid_test"],
+    "--",
+    label="One hidden layer: test",
+    lw=4,
 )
 plt.plot(
     results["node_sizes"],
     results["two_hid_train"],
-    "--",
     label="Two hidden layers: train",
     lw=4,
 )
@@ -70,5 +86,19 @@ plt.plot(
     label="Two hidden layers: test",
     lw=4,
 )
-plt.legend(loc=(1.04, 0))
+plt.plot(
+    results["node_sizes"],
+    results["three_hid_test"],
+    "--",
+    label="Three hidden layers: test",
+    lw=4,
[…]

### 2022-11-15 src/utils.py

@@ -200,7 +200,7 @@ def derivate(func):
     elif func.__name__ == "RELU":
 
         def func(x):
-            return np.where(x > np.zeros(x.shape), np.ones(x.shape), np.zeros(x.shape))
+            return np.where(x, 1, 0)
 
         return func
 
@@ -208,9 +208,7 @@ def derivate(func):
 
         def func(x):
             delta = 10e-4
-            return np.where(
-                x > np.zeros(x.shape), np.ones(x.shape), np.full((x.shape), delta)
-            )
+            return np.where(x, 1, delta)
 
         return func
 
@@ -483,6 +481,7 @@ def plot_arch(
     epochs: int = 1000,
     step_size: int = 10,
     classify=False,
+    folds: int = 5,
 ):
 
     node_sizes = np.arange(0, max_nodes, step_size)
@@ -493,6 +492,64 @@ def plot_arch(
     one_hid_test = np.zeros(node_sizes.shape[0])
     two_hid_train = np.zeros(node_sizes.shape[0])
     two_hid_test = np.zeros(node_sizes.shape[0])
+    three_hid_train = np.zeros(node_sizes.shape[0])
+    three_hid_test = np.zeros(node_sizes.shape[0])
+
+    for i in range(len(node_sizes)):
+        node_size = node_sizes[i] // 3 or 1
+        neural = model(
+            (X.shape[1], node_size, node_size, node_size, t.shape[1]),
+            hidden_func=funcs[0],
+            output_func=funcs[1],
+            cost_func=funcs[2],
+            seed=1337,
+        )
+        print(neural.dimensions)
+        scores = neural.cross_val(
+            folds,
+            X,
+            t,
+            scheduler,
+            *args,
+            batches=batches,
+            epochs=epochs,
+            lam=lam,
+        )
+        if classify:
+            three_hid_test[i] = scores["final_test_acc"]
+            three_hid_train[i] = scores["final_train_acc"]
+        else:
+            three_hid_test[i] = scores["final_test_error"]
+            print(f"{scores['final_test_error']=}")
+            print(f"{scores['final_train_error']=}")
+            three_hid_train[i] = scores["final_train_error"]
+
+    for i in range(len(node_sizes)):
+        node_size = node_sizes[i] // 2 or 1
+        neural = model(
+            (X.shape[1], node_size, node_size, t.shape[1]),
+            hidden_func=funcs[0],
+            output_func=funcs[1],
+            cost_func=funcs[2],
[…]

### 2022-11-15 doc/paper.tex
Paper

--- 2022-11-14 doc/paper.tex
@@
 Already after EPOCHS epochs, the momentum scheduler has achieved MOMENTUM MSE AT EPOCHS, a notable improvement over CONSTANT MSE AT EPOCHS achieved by a constant learning rate.
-
-- GD vs SGD
-
 Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs.
@@
 
-An interesting insight from FIGURE NUM is that all the schedulers prefer high learning rates and regularization.
-As this plot was generated for 100 epochs due to the computational complexity of gridsearching the six schedulers, the bias towards higher learning rates is certainly due to the high learning rates being able to reach a low MSE in the short training time.
-With the lambda = 0 column representing the ordinary least squares solutions, these are clearly consistantly outperformed by Ridge regression as the best MSE is achieved with regularization.
-For this specific plot, momentum does not seem to be an important parameter.
-Both the momentum and adagrad with momentum heatmaps are identical to their momentumless counterparts, constant and adagrad, implying that the best gridsearched momentum is 0.
-That momentum is irrelevant in this instance is not a generalizable result, as momentum has played a key role in previous plots and results.
+Before continuing, it is interesting to note that all the schedulers work best with a combination of similar eta and lambda values when gridsearched for achieving the minimal MSE.
+It is also interesting that they in the long term reach more or less the same error as can be observed in HEATMAP FIG.
+This could be because given more epochs, the other schedulers are given more time to catch up as the best performing schedulers, such as Adam, only decrease their MSE slightly.
+However, as can be seen in HEATMAP FIG, Adam offers a $\frac{0.018}{0.016} \dot 100 = 11.11\%$ decrease in MSE compared to the constant schedulers.
 
-Another observation from FIG NUM  is that the schedulers final error for100 epochs are more or less the same.
-As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adam, starts slowing down the rate at which the MSE is decreased.
-However, as can be seen in HEATMAP FIG, Adam offers a 11.11 \% \footnote{$\frac{constant MSE}{adam MSE} * 100 = \frac{0.018}{0.016} * 100 = 11.11\%$} decrease in MSE compared to the constant scheduler.
-It is clear from the results thus far that Adam is the best scheduler for approximating the Franke function when using linear regression with gradient descent.
-
-Furthermore, when trained for 5000 epochs, Adam with optimal parameters keeps improving the MSE.
-Scikit learn's Adam scheduler does perform better, offering a greater decrease the MSE.
-Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model, in addition to other optimizations present in the scikit learn implementation.
-Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of an MSE of 0.012 on the test data and Scikits final MSE of 0.004 (granted this is on the train data, as scikit learn does not provide the loss curve for the test data).
-both falling short of the analytical MSE of 0,003.
+However, when trained for 500 epochs, Adam with optimal parameters does not yield any improvement in MSE.
+Scikits Adam scheduler does perform better, further decreasing the MSE slightly.
+Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model.
+Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of ADAM FINAL MSE and Scikits final MSE of SCIKIT FINAL MSE both falling short of the analytical MSE of 0.003.
 In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions.
-This result might be due to 5000 epochs not being enough training time, as the Franke function is a quite complex function with added noise.
+This result might be due to 500 epochs not being enough training time.
 
 \subsection{Neural Network for regression}
+
+Having gained a great understanding of how the different schedulers function, and having created the necessary infrastructure to study our models performance based on a number of different factors, we use these methods to optimize our neural network for linear regression.
+More specifically, we attempt to fit the Franke function.
+First, we perform a preliminary test on MSE over model complexity.
+For comparisons sake, we use the same scheduler for each architecture, and fix the regularization parameter at 0.
+By fixing the regularization parameter at 0, we are of course favouring less complex models, but this a good way to get a general sense of when the model starts to overfit.
+Using the sigmoid activation function, we can study the MSE as a function of the total hidden node count (Fig.~\ref{msebymodelfranke}).
+Here we can see that we are not overfitting
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=6in]{figure/msebymodelfranke.png}}
+\caption{Doge}
+\label{msebymodelfranke}
+\end{figure}
 
@@
 To test our neural networks ability to solve classification problems, we first applied it to the Wisconsin Breast Cancer dataset.
-Using our gridsearch method, we found that the best scheduler was Adam, with an eta of $0.01$ and no regularization.
-For our architecture, we chose to use only one hidden layer with 60 nodes, based on our testing.
-With this, we were able to achieve a crossvalidated test accuracy of $98.6 \%$, which is very good.
+But to get the best results, there are a large amount of parameters that need to be decided on.
+We must find the best architecture, the best, activation functions, best scheduler and the best parameters for the scheduler.
+Finding the absolute best combination of all these choices is impossible, so compromises must be made.
+
+We start by trying to find a good model complexity, i.e the best number of hidden layers, and the amount of nodes in them.
+To do this, we plotted the accuracy over different model architectures.
+Like we did during regression, we fix all the other parameters and study the architecture by itself.
+Using the sigmoid activation function (Fig.~\ref{cancerarchcompsigmoid}), our results are somewhat surprising.
+We see that even at 800 total hidden nodes, we are not overfitting.
+Regardless of this peculiarity, we can see that the gain from using more layers and more nodes is miniscule.
+Using the RELU activation function, the results are quite different, as can be seen in Fig.~\ref{cancerarchcomp}.
+Here we can observe that the neural net with two hidden layers starts performing very inconsistently when the number of nodes increases.
+Looking more closely at the training process, we can see that training stops completely in some of our crossvalidations.
+Clearly, we are running into the "dying RELU problem", where too many of our RELU nodes output zero, leading to our gradient descent slowing down and eventually stopping completely.
+Here we can clearly see some of the benefits of the sigmoid, as even if the gradients become very small, they will never be zero, as with RELU.
+Based on these findings, we decided to use a simple model with a single hidden layer of 100 nodes.
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=6in]{figure/cancerarchcompsigmoid.png}}
+\caption{Comparison between a neural net with one hidden layer, and one with two hidden layers, for different totals of hidden nodes.
+For the network with two hidden layers, the total is evenly divided between the layers, such that at $x = 60$ we have a network with two hidden layers of 30 nodes each.
+In this case, the hidden layers used the sigmoid activation function.}
+\label{cancerarchcompsigmoid}
+\end{figure}
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=6in]{figure/cancerarchcomp.png}}
+\caption{Comparison between a neural net with one hidden layer, and one with two hidden layers, for different totals of hidden nodes.
+For the network with two hidden layers, the total is evenly divided between the layers, such that at $x = 60$ we have a network with two hidden layers of 30 nodes each.
+In this case, the hidden layers used the RELU activation function.}
+\label{cancerarchcomp}
+\end{figure}
+
+Using our gridsearch method on this model, we found that the best scheduler was Adam, with an eta of $0.001$ and a lambda of $0.0001$.
+With this combination of architecture and scheduler, we were able to achieve a crossvalidated test accuracy of $98.76\%$, which is a great result.
+This outperforms SciKit's MLP classifier, which achieved a crossvalidated test accuracy of around 95\%.
 Looking at the confusion matrix (Fig.~\ref{cancerconf}), we see that we are achieving good predictions for both true positives and true negatives.
@@
 \caption{Confusion matrix for the cancer dataset using a neural net with a single hidden layer of 60 nodes}
-\label{verification}
+\label{cancerconf}
 \end{figure}
 
-Confident in our networks's ability to to binary classification, we turned to the MNIST dataset.
-Using our gridsearch, we again found that the best scheduler was Adam.
-Our architecture in this case was a single hidden layer with 100 nodes.
+Confident in our networks's ability to do binary classification, we turned to the MNIST dataset.
+Here, we perform an educated guess on the different parameters, based on our previous testing.
+We decided upon an architecture of a single layer with 100 nodes using the RELU activation function.
+For our scheduler, we used Adam with an eta of $0.001$ and no regularization.
 Being a multiclass problem with ten different classes, one might expect our neural network to struggle more with this dataset.
-However, this was not the case, as we achieved a crossvalidated test accuracy of $99.07$.
-From these two datasets, we can see that a correctly optimized neural network is adept at solving classification problems.
-On the other hand, these good results, and the easew with which our network attained them, could imply that the problems we have dealt with are not as complicated as they may seem.
+However, this was not the case, as we achieved a crossvalidated test accuracy of $99.07\%$.
+Based on these results, we conclude that the initial guess for our model was a good one, and that further optimization has little room for improvement.
+From these two datasets, we can see that a neural network is adept at solving classification problems.
+On the other hand, these good results, and the ease with which our network attained them, could imply that the problems we have dealt with are not as complicated as they may seem.
 
@@
 Here we found that logistic regression performed essentially equivalent or even better, at a much lower computational cost.
-Looking at Fig.~\ref{cancerlogneur}, we see that logistic regression keeps up and even surpasses our neural network in accuracy per epoch.
+Looking at Fig.~\ref{cancerlogneur}, we see that logistic regression keeps up and even surpasses our neural network in accuracy over epochs.
 The value of logistic regression becomes even greater when we consider the computational cost of each method.
-In this case, the logistic regression was about 65 \% faster\footnote{The time taken to complete a crossvalidation with 5 folds up to 200 epochs was about 52 seconds for our neural network, but only 18 seconds for logistic regression}, with results surpassing our neural network (see Fig.~\ref{cancerconflog}).
+In this case, the logistic regression was about 65\% faster\footnote{The time taken to complete a crossvalidation with 5 folds up to 200 epochs was about 52 seconds for our neural network, but only 18 seconds for logistic regression}, with results surpassing our neural network (see Fig.~\ref{cancerconflog}).
 This implies that our classification datasets were perhaps too simple, and that using Neural Networks was not needed to achieve good results.

### 2022-11-15 doc/paper.tex
constand vs newton

--- 2022-11-15 doc/paper.tex
@@
 
-In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function.
-IF HESSIAN: Using Newtons method, the computational cost is much greater than simply replacing the Hessian matrix inversion with a constant learning rate as can be seen by TIME COMPARISON.
-Furthermore, we can observe from FIGURE COMPARING HESSIAN TO CONSTANT that the constant scheduler does a good job at reducing the MSE, given that the final MSE score is CONSTANT SCORE compared to the Newton method's HESSIAN SCORE.
+In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function with 55 features.
+We started off by comparing Newtons method for gradient descent to a constant learning rate using full batch gradient descent.
+
+From observing FIG CONSTANT V HESSIAN we see that Newton's method using Hessian inversion greatly outperforms the constant scheduler by reaching an MSE of 0.02 within the first epoch.
+This is due to the Hessian inversion in the first epoch in Newton's method being equivalent to the analytical solution.
+By comparison, the constant scheduler given 200 epochs reaches an MSE of 0.21.
+Given this result, it would seem indefensible to claim that Newtons method requires improvements, especially considering the time it took the Newton method 0.1512 seconds to compute 200 epochs, again beating the result from the constant scheduler which took 0.4991 seconds \footnote{this result should be taken with a grain of salt, given that the constant scheduler is run as a part of a larger architecture, whilst the Hessian calculation is done in a seperate script.
+Still, this provides an intuition that Newton's method does not take much time to compute this specific problem}.
+This result is due to the Hessian matrix only being a mere 55x55 matrix for this problem, such that the matrix inversion will not be computationally complex.
+However, if we were to increase the complexity of the problem, Newtons method would rapidly increase in computational complexity as it is a large matrix inversion with many floating point operations.
+For problems of such a small scale, when our features are based on a polynomial approximation Franke Function, it can be important to remember that we don't necessarily have to turn to these fancier solutions as the simple Newton method will do the job perfectly.
+Keeping this lesson in mind, we wish to further investigate the various scheduler methods in order to gain a better understanding of how well they perform, which will become useful for larger problems.
 
@@
 
-- GD vs SGD
+By introducing mini-batch gradient descent of different batch sizes, we can observe from BATCH SIZE FIG that the scheduler methods perform better when optimized for different batch sizes.
+Both the constant and the momentum scheduler seem to vary equally over batch sizes, with the optimal batch size being equal to one.
+This means that our schedulers work the best for a fully stochastic gradient descent.
+The MSE with full gradient descent, as can be seen in BATCH SIZE FIG when batch size equals 320, provides an MSE of
+A full comparison of batch sizes for all the different schedulers can be viewed in APPENDIX in the appendix.
 
@@
 Looking at TIME TABLE FOR ALL SCHEDULERS, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds.
+Based on FIG ALL, Adam seems to reach an MSE of 0.2 the fastest when strictly looking at the number of epochs required, however, whilst Adagrad requires more epochs the computations are computed faster in our implementation.
 Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.
@@
 With the lambda = 0 column representing the ordinary least squares solutions, these are clearly consistantly outperformed by Ridge regression as the best MSE is achieved with regularization.
-For this specific plot, momentum does not seem to be an important parameter.
-Both the momentum and adagrad with momentum heatmaps are identical to their momentumless counterparts, constant and adagrad, implying that the best gridsearched momentum is 0.
-That momentum is irrelevant in this instance is not a generalizable result, as momentum has played a key role in previous plots and results.
+For this specific plot, momentum does not seem to make out a noticeable difference for the best solutions, as the optimal solution equals their momentumless counterparts.
+That being said, in FIG ALL it is clear that the added momentum aids in a quicker convergence when compared to the constant scheduler.
 
 Another observation from FIG NUM  is that the schedulers final error for100 epochs are more or less the same.
-As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adam, starts slowing down the rate at which the MSE is decreased.
-However, as can be seen in HEATMAP FIG, Adam offers a 11.11 \% \footnote{$\frac{constant MSE}{adam MSE} * 100 = \frac{0.018}{0.016} * 100 = 11.11\%$} decrease in MSE compared to the constant scheduler.
-It is clear from the results thus far that Adam is the best scheduler for approximating the Franke function when using linear regression with gradient descent.
+As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased.
+However, as can be seen in HEATMAP FIG, Adagrad offers a 12\% \footnote{$\frac{constant MSE}{adagrad MSE} * 100 = \frac{0.018}{0.015} * 100 = 12\%$} decrease in MSE compared to the constant scheduler.
+It is clear from the results thus far that Adagrad is the best scheduler for approximating the Franke function when using linear regression with gradient descent.
+We believe that the reason Adagrad wins out is that these result are based on training for 100 epochs.
+The downfall of Adagrad in the literature seems to be that the aggressive decaying of the learning rate hinders a quick convergence deeper into training SOURCE https://towardsdatascience.com/learning-parameters-part-5-65a2f3583f7d.
+We are thus cautionary to claim that Adagrad will do the best job over a much larger amount of epochs when compared to adaptive schedulers that claim to improve this aspect.
+
+To test this, we compared Adagrad with Adam over 5000 epochs.
+Both schedulers performed equally, both resulting in an MSE of 0.017 with Adagrad taking the edge with an MSE 0.0002 lower.
+Scikit MSE 0.005.
 
@@
 Scikit learn's Adam scheduler does perform better, offering a greater decrease the MSE.
-Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model, in addition to other optimizations present in the scikit learn implementation.
+We compared Adagrad to scikit learn's Adam MLPRegressor as this seemed to be the closest option for comparing our results.
+Even though we have cross validated both our own Adagrad implementation and Scikit's with 5 folds, these differences in results might be due to the difference in solver, a more favorable weight initiation in Scikits model or additional optimizations present in the scikit learn implementation.
 Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of an MSE of 0.012 on the test data and Scikits final MSE of 0.004 (granted this is on the train data, as scikit learn does not provide the loss curve for the test data).

### 2022-11-15 doc/paper.tex
poetry, shakespeare even

@@
 
-From observing FIG CONSTANT V HESSIAN we see that Newton's method using Hessian inversion greatly outperforms the constant scheduler by reaching an MSE of 0.02 within the first epoch.
+FIG CONSTANT V HESSIAN
+
+From observing FIG CONSTANT V HESSIAN we see that Newton's method using Hessian inversion greatly outperforms the constant scheduler by reaching an MSE of 0.02 for the testing data within the first epoch.
 This is due to the Hessian inversion in the first epoch in Newton's method being equivalent to the analytical solution.
-By comparison, the constant scheduler given 200 epochs reaches an MSE of 0.21.
+By comparison, the constant scheduler given 200 epochs reaches an MSE of 0.198 for the testing data.
 Given this result, it would seem indefensible to claim that Newtons method requires improvements, especially considering the time it took the Newton method 0.1512 seconds to compute 200 epochs, again beating the result from the constant scheduler which took 0.4991 seconds \footnote{this result should be taken with a grain of salt, given that the constant scheduler is run as a part of a larger architecture, whilst the Hessian calculation is done in a seperate script.
@@
 
-By introducing mini-batch gradient descent of different batch sizes, we can observe from BATCH SIZE FIG that the scheduler methods perform better when optimized for different batch sizes.
+BATCH SIZE FIG (should mention in text that 320 is the size of the training data. Should mention base batch size for grid search was 18)
+
+By introducing mini-batch gradient descent of different batch sizes, we can observe from BATCH SIZE FIG that the scheduler methods perform better when optimized for different batch sizes over 100 epochs.
 Both the constant and the momentum scheduler seem to vary equally over batch sizes, with the optimal batch size being equal to one.
-This means that our schedulers work the best for a fully stochastic gradient descent.
-The MSE with full gradient descent, as can be seen in BATCH SIZE FIG when batch size equals 320, provides an MSE of
-A full comparison of batch sizes for all the different schedulers can be viewed in APPENDIX in the appendix.
+This means that our schedulers work the best for a fully stochastic gradient descent (SGD).
+The MSE with full gradient descent, as can be seen in BATCH SIZE FIG when batch size equals 320, provides an MSE of 4.211 of the test data, whilst SGD provides an MSE of 0.28 for the test data in an equal amount of time.
+Because we first gridsearched the optimal values of eta and lambda first with a predefined base batch size, we must acknowledge that BATCH SIZE FIG will have some bias towards the batch size used when found the optimal hyperparameters.
+However, given that we have gridsearched with different base batch sizes accross our testing, we have found the results to be similar accross different base batch sizes, and that this bias which we induced to save computational time was insignificant.
+A full comparison of MSE over batch sizes for all the different schedulers can be viewed in APPENDIX in the appendix.
+
+FIG ALL
+TIME TABLE FOR ALL SCHEDULERS
 
 Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs.
-We can also observe that the different schedulers all reach an MSE of 0.02 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.2.
+We can also observe that the different schedulers all reach an MSE of 0.02 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.2 of the test data.
 Looking at TIME TABLE FOR ALL SCHEDULERS, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds.
@@
 Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.
+
+FIGURE NUM (heatmap)
 
@@
 
-Another observation from FIG NUM  is that the schedulers final error for100 epochs are more or less the same.
+Another observation from FIG NUM is that the schedulers final error for 100 epochs are more or less the same.
 As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased.
@@
 Both schedulers performed equally, both resulting in an MSE of 0.017 with Adagrad taking the edge with an MSE 0.0002 lower.
-Scikit MSE 0.005.
+Scikit obtained a test data MSE of 0.005, but followed the same trend in performance of dipping early and stabilizing after just a 100 epochs.
 

### 2022-11-16 src/FFNN.py
Fix momentum shenanigans

@@ -172,6 +172,11 @@ class FFNN:
                 train_error = cost_function_train(prediction)
                 if train_error > 10e20:
                     # if this happens, we have a problem
+                    length = 10
+                    train_error=None
+                    test_error=None
+                    train_acc=None
+                    test_acc=None
                     break
                 if test_set:
                     prediction_test = self.predict(X_test, raw=True)
@@ -189,6 +194,7 @@ class FFNN:
 
                 else:
                     test_errors = np.nan  # a
+
 
                 train_acc = None
                 test_acc = None

### 2022-11-16 doc/paper.tex
figs, many figs

@@
 It is commonly used to test machine learning algorithms for regression problems.
-For our linear regression problems, our training data will be based on a 9th degree polynomial approximation to the Franke Function, with 55 features.
+For our linear regression problems, our training data will be based on a 9th degree polynomial approximation to the Franke Function, with 56 features.
 For the feed forward neural network, we will use 2 of the features (x, y) of the Franke Function.
@@
 \hline
-Franke Function & synthetic & 55 & regression \\
+Franke Function & synthetic & 56 & regression \\
 Franke Function & synthetic & 2 & regression \\
@@
 Still, this provides an intuition that Newton's method does not take much time to compute this specific problem}.
-This result is due to the Hessian matrix only being a mere 55x55 matrix for this problem, such that the matrix inversion will not be computationally complex.
+This result is due to the Hessian matrix only being a mere 56x56 matrix for this problem, such that the matrix inversion will not be computationally complex.
 However, if we were to increase the complexity of the problem, Newtons method would rapidly increase in computational complexity as it is a large matrix inversion with many floating point operations.
@@
 
+From FIGURE MOMENTUM  v CONSTANT, we can observe that adding momentum does not improve upon the solutions found by the constant scheduler when using full gradient descent, reaching once again an MSE of 0.21 for the test data.
+This implies that when both schedulers are gridsearched for optimal hyperparameters, the optimal value for the momentum hyperparameter is 0.
+This result makes sense, as the momentum hyperparameter is used to reduce noise from derivatives calculated from smaller batch sizes, so-called mini batches.
+When utilizing a full gradient descent, where the batch size equals the size of the training data, momentum will not improve our results.
+
 BATCH SIZE FIG (should mention in text that 320 is the size of the training data. Should mention base batch size for grid search was 18)
+
+By introducing mini-batch gradient descent, we can observe that
 
@@
 The downfall of Adagrad in the literature seems to be that the aggressive decaying of the learning rate hinders a quick convergence deeper into training SOURCE https://towardsdatascience.com/learning-parameters-part-5-65a2f3583f7d.
-We are thus cautionary to claim that Adagrad will do the best job over a much larger amount of epochs when compared to adaptive schedulers that claim to improve this aspect.
+We are thus cautionary to claim that Adagrad will do the best job over a much larger amount of epochs when compared to adaptive schedulers that claim to improve this aspect such as Adam.
 
-To test this, we compared Adagrad with Adam over 5000 epochs.
-Both schedulers performed equally, both resulting in an MSE of 0.017 with Adagrad taking the edge with an MSE 0.0002 lower.
-Scikit obtained a test data MSE of 0.005, but followed the same trend in performance of dipping early and stabilizing after just a 100 epochs.
+5000e_adam_v_adagrad
 
-Furthermore, when trained for 5000 epochs, Adam with optimal parameters keeps improving the MSE.
-Scikit learn's Adam scheduler does perform better, offering a greater decrease the MSE.
-We compared Adagrad to scikit learn's Adam MLPRegressor as this seemed to be the closest option for comparing our results.
-Even though we have cross validated both our own Adagrad implementation and Scikit's with 5 folds, these differences in results might be due to the difference in solver, a more favorable weight initiation in Scikits model or additional optimizations present in the scikit learn implementation.
-Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of an MSE of 0.012 on the test data and Scikits final MSE of 0.004 (granted this is on the train data, as scikit learn does not provide the loss curve for the test data).
-both falling short of the analytical MSE of 0,003.
-In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions.
-This result might be due to 5000 epochs not being enough training time, as the Franke function is a quite complex function with added noise.
+To test this, we compared Adagrad with Adam over 5000 epochs, cross validated for 5 folds.
+Both schedulers perform almost equally, with Adagrad reaching a slightly lower MSE of 0.011 for the test data when compared to Adam's 0.012.
+It could be argued that 5000 epochs is not enough for the improvements of Adam to be noticeable, as the MSE could still be pushed much lower.
+However, for these computationally expensive runs it is clear that Adagrad is the scheduler of choice, especially considering our previous conclusion that Adagrad used less time to train.
+It should be noted that scikit learn's implementation outperforms both our implementation of Adagrad and Adam, with an MSE of 0.004.
+This could be due to a more favorable wight initialization in scikits model, along with additional optimizations implemented in their library.
+That we don't beat the analytical solution is no surprise, as we have limited ourself to a polynomial approximation of the Franke Function with 56 features.
+Our MSE might approach the analytical MSE, but this might take an unreasonable amount of epochs considering the computational runtime.
+Next, we investigate the use of linear regression in a neural network with hidden nodes, such that we can free ourselves of the limitations of the polynomial approximation.
 

### 2022-11-16 src/task_b.py
Update this to work

@@ -10,7 +10,7 @@ np.random.seed(1337)
     X,
     X_train,
     X_test,
-    t,
+    z,
     z_train,
     z_test,
     centering,
@@ -32,10 +32,7 @@ params = [eta, rho, rho2]
 #dims = (2, 266, 133, 1) # second best
 #dims = (2, 133, 133, 1)
 #dims = (2, 320, 160, 1) # second best 
-#dims = (2, 133, 133, 133, 1) # best so far 
-#dims = (2, 133, 133, 133, 1)
-dims = (2, 66,66,66, 1) 
-
+dims = (2, 133, 133, 133, 1) # best so far 
 train_epochs = 100
 
 
@@ -69,20 +66,17 @@ optimal_batch = neural.optimize_batch(
     epochs=30,
 )
 """
-scores = neural.cross_val(
-    5,
-    X[:, 1:3],
-    z.reshape(400, 1) ,
+scores = neural.fit(
+    X_train[:, 1:3],
+    z_train,
     sched,
     *params,
     batches=X_train.shape[0],
     epochs=train_epochs,
-    lam=10e-6,
-    #X_test=X_test[:, 1:3],
-    #t_test=z_test,
+    lam=0.0001,
+    X_test=X_test[:, 1:3],
+    t_test=z_test,
 )
-
-print(f"Final test_error: {scores['final_test_error']}")
 
 # ------------------------- MLPRegressor -------------------------
 mlp = MLPRegressor(

### 2022-11-16 src/FFNN.py
fix NaN

@@ -721,17 +721,19 @@ class FFNN:
                     test_accs = scores["test_accs"]
                     loss_heatmap[y, x] = test_accs[-1]
                     min_heatmap[y, x] = scores["final_test_acc"]
+                    min_heatmap[y, x] = np.nanmax(test_accs)
                 else:
                     test_scores = scores["test_errors"]
                     loss_heatmap[y, x] = test_scores[-1]
                     min_heatmap[y, x] = scores["final_test_error"]
+                    min_heatmap[y, x] = np.nanmin(test_scores)
                 self.reset_weights()
 
         # select optimal eta, lambda
         if classify:
-            y, x = np.unravel_index(loss_heatmap.argmax(), loss_heatmap.shape)
+            y, x = np.unravel_index(loss_heatmap.nanargmax(), loss_heatmap.shape)
         else:
-            y, x = np.unravel_index(loss_heatmap.argmin(), loss_heatmap.shape)
+            y, x = np.unravel_index(loss_heatmap.nanargmin(), loss_heatmap.shape)
 
         optimal_eta = eta[y]
         optimal_lambda = lam[x]
@@ -778,24 +780,23 @@ class FFNN:
                         # todo wont work with bootstrap
                         test_accs = scores["test_accs"]
                         loss_heatmap[y, x, z] = test_accs[-1]
-                        min_heatmap[y, x, z] = np.min(test_accs)
+                        min_heatmap[y, x, z] = np.nanmax(test_accs)
                     else:
                         test_errors = scores["test_errors"]
                         loss_heatmap[y, x, z] = test_errors[-1]
-                        min_heatmap[y, x, z] = np.min(test_errors)
+                        min_heatmap[y, x, z] = np.nanmin(test_errors)
                     self.reset_weights()
 
         if classify:
-            y, x, z = np.unravel_index(loss_heatmap.argmax(), loss_heatmap.shape)
+            y, x, z = np.unravel_index(np.nanargmax(loss_heatmap), loss_heatmap.shape)
         else:
-            y, x, z = np.unravel_index(loss_heatmap.argmin(), loss_heatmap.shape)
+            y, x, z = np.unravel_index(np.nanargmin(loss_heatmap), loss_heatmap.shape)
 
         optimal_eta = eta[y]
         optimal_lambda = lam[x]
         optimal_momentum = momentums[z]
 
         optimal_params = [optimal_eta, optimal_momentum]
[…]

### 2022-11-16 src/FFNN.py
fix NaN

@@ -731,9 +731,9 @@ class FFNN:
 
         # select optimal eta, lambda
         if classify:
-            y, x = np.unravel_index(loss_heatmap.nanargmax(), loss_heatmap.shape)
+            y, x = np.unravel_index(np.nanargmax(loss_heatmap), loss_heatmap.shape)
         else:
-            y, x = np.unravel_index(loss_heatmap.nanargmin(), loss_heatmap.shape)
+            y, x = np.unravel_index(np.nanargmin(loss_heatmap), loss_heatmap.shape)
 
         optimal_eta = eta[y]
         optimal_lambda = lam[x]

### 2022-11-16 src/utils.py
This saved github from being destroyed by greg

@@ -200,7 +200,7 @@ def derivate(func):
     elif func.__name__ == "RELU":
 
         def func(x):
-            return np.where(x, 1, 0)
+            return np.where(x > 0, 1, 0)
 
         return func
 

### 2022-11-16 src/utils.py
REALY LEAKY

@@ -208,7 +208,7 @@ def derivate(func):
 
         def func(x):
             delta = 10e-4
-            return np.where(x, 1, delta)
+            return np.where(x > 0, 1, delta)
 
         return func
 

### 2022-11-16 src/bycomplexitycancer.py
Titel change

@@ -25,7 +25,7 @@ folds = 5
 scheduler = Adam
 args = [0.01, 0.9, 0.999]
 
-funcs = [sigmoid, sigmoid, CostLogReg]
+funcs = [RELU, sigmoid, CostLogReg]
 
 results = plot_arch(
     FFNN,
@@ -44,7 +44,9 @@ results = plot_arch(
 
 
 sns.set(font_scale=2)
-plt.title("Accuracy by model complexity for cancer data")
+plt.title(
+    "Accuracy by model complexity for cancer data, \n using RELU as the activation function"
+)
 plt.xlabel("Total amount of hidden nodes")
 plt.ylabel("Accuracy")
 plt.plot(

### 2022-11-16 src/constant_v_hessian.py
fontsize changes

@@ -79,10 +79,11 @@ print(f"Hessian time to train {epochs} epochs: {time[1]}")
 hessian_test_errors = hessian_scores["test_errors"]
 
 # plot
-plt.plot(constant_test_errors, label="Constant test")
-plt.plot(hessian_test_errors, label="Hessian test")
+sns.set(font_scale=3)
+plt.plot(constant_test_errors, label="Constant test", linewidth=5)
+plt.plot(hessian_test_errors, label="Hessian test", linewidth=5)
 plt.legend(loc="upper right")
-plt.xlabel("epochs", fontsize=18)
-plt.ylabel("MSE", fontsize=18)
-plt.title("Constant vs Hessian MSE over epochs", fontsize=22)
+plt.xlabel("epochs", fontsize=32)
+plt.ylabel("MSE", fontsize=32)
+plt.title("Constant vs Hessian MSE over epochs", fontsize=42)
 plt.show()

### 2022-11-16 src/task_a.py

@@ -29,8 +29,8 @@ z_train = z_train.reshape(z_train.shape[0], 1)
 z_test = z_test.reshape(z_test.shape[0], 1)
 
 # epochs to run for
-epochs = 100
-folds = 5
+epochs = 10
+folds = 2
 
 # no hidden layers, no activation function
 dims = (X.shape[1], 1)
@@ -39,7 +39,7 @@ neural = FFNN(dims, seed=1337)
 # parameters to test for
 eta = np.logspace(-5, -1, 5)
 lam = np.logspace(-5, -1, 5)
-momentums = np.linspace(0, 1, 5)
+momentums = np.linspace(0, 0.1, 5)
 lam[0] = 0
 rho = 0.9
 rho2 = 0.999
@@ -51,7 +51,6 @@ batches_list = np.logspace(
 
 # schedulers to test for
 schedulers = [Constant, Momentum, Adagrad, AdagradMomentum, RMS_prop, Adam]
-schedulers = [Momentum]
 #
 # parameters for schedulers
 constant_params = []
@@ -63,12 +62,12 @@ adam_params = [rho, rho2]
 
 # list of scheduler parameters
 params_list = [
-    # constant_params,
+    constant_params,
     momentum_params,
-#     adagrad_params,
-#     adagrad_momentum_params,
-#     rms_params,
-#     adam_params,
+    adagrad_params,
+    adagrad_momentum_params,
+    rms_params,
+    adam_params,
 ]
 
 # results
@@ -77,10 +76,11 @@ optimal_eta = np.zeros(len(schedulers))
 optimal_lambdas = np.zeros(len(schedulers))
 optimal_batches = np.zeros(len(schedulers), dtype=int)
 
+sns.set(font_scale=2)
 # gridsearch eta, lambda
 for i in range(len(schedulers)):
     plt.subplot(321 + i)
-    plt.suptitle("Test loss for eta, lambda grid", fontsize=22)
+    plt.suptitle("Test loss for eta, lambda grid", fontsize=32)
     optimal_params, optimal_lambda, loss_heatmap = neural.optimize_scheduler(
         X_train,
         z_train,
@@ -98,6 +98,8 @@ for i in range(len(schedulers)):
     optimal_params_list.append(optimal_params)
 
     # plot heatmap
+    color = plt.get_cmap("magma_r")
+    color.set_bad("lightblue")
     ax = sns.heatmap(loss_heatmap, xticklabels=lam, yticklabels=eta, annot=True)
     ax.add_patch(
[…]

### 2022-11-16 doc/paper.tex
Add stuff

--- 2022-11-15 doc/paper.tex
@@
 
-We developed a Feed Forward Neural Network (FFNN) and applied it to many common datasets used for benchmarking machine learning algorithms.
+We developed a Feed Forward Neural Network (FFNN) and applied it to both regression and classification problems.
 Furthermore, we compared our results to those attained by linear and logistic regression.
 To optimize our network, we explored the use of different schedulers in our gradient descent method, and compared their performance against eachother.
-Overall, we found that our FFNN performed well in all classification tasks we tested, attaining test accuracies between $95$ and $100$ percent for all classification problems.
+Amongst our schedulers, we found that Adam and Adagrad performed the best.
+Overall, we found that our FFNN performed well in both classification tasks we tested, attaining a test accuracy of $98.76$ on the Wisconsin Breast Cancer Dataset and a test accuracy of $99.07$ on the MNIST 8x8 dataset.
 However, we also found that simple logistic regression performed similarly, with a much lower computational cost.
-For regression on real terrain data, we found that our neural network slightly outperformed our previous polynomial fit, attaining an MSE of MONKEY, PERCENT percent lower than simple linear regression.
-In our gradient descent, we found that SCHEDULER performed best, being 1 BILLION percent better than the average of all other methods.
+For regression applied to the Franke Function, we found that our neural net performed worse than a simple polynomial fit using linear regression, attaining a MSE of $0.007$, more than twice as high as the one achieved using linear regression.
 
@@
 Finally, we will provide a conclusion summarizing the core results and lessons.
-
-% Neural Networks are all the rage these days, everyone wants to have a neural network in their home. Ive even seen people install neural networks in their dogs. \cite{ericjona}
-
-% Neural networks are powerful tools, able to solve many different problems with
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-The Franke Function is a simple weighted sum of polynomials, taking two inputs and returning a real value.
-It is commonly used to test machine learning algorithms for regression problems.
-For our linear regression problems, our training data will be based on a 9th degree polynomial approximation to the Franke Function, with 55 features.
-For the feed forward neural network, we will use 2 of the features (x, y) of the Franke Function.
-
-\subsubsection{Elevation data over Norway}
-
-This data is similar to the Franke Function in that it takes two inputs (x and y coordinates) and returns a real value, which is the elevation of a specific point in Norway.
+The Franke Function is a simple weighted sum of polynomials, taking two inputs and returning a real value. It is commonly used to test machine learning algorithms for regression problems.
 
@@
 \hline
-Franke Function & synthetic & 55 & regression \\
 Franke Function & synthetic & 2 & regression \\
-Terrain (downscaled) & 204 800 & 2 & regression \\
 Wisconsin Breast Cancer & 569 & 30 & binary classification \\
@@
 
-In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function with 55 features.
-We started off by comparing Newtons method for gradient descent to a constant learning rate using full batch gradient descent.
-
-From observing FIG CONSTANT V HESSIAN we see that Newton's method using Hessian inversion greatly outperforms the constant scheduler by reaching an MSE of 0.02 within the first epoch.
-This is due to the Hessian inversion in the first epoch in Newton's method being equivalent to the analytical solution.
-By comparison, the constant scheduler given 200 epochs reaches an MSE of 0.21.
-Given this result, it would seem indefensible to claim that Newtons method requires improvements, especially considering the time it took the Newton method 0.1512 seconds to compute 200 epochs, again beating the result from the constant scheduler which took 0.4991 seconds \footnote{this result should be taken with a grain of salt, given that the constant scheduler is run as a part of a larger architecture, whilst the Hessian calculation is done in a seperate script.
-Still, this provides an intuition that Newton's method does not take much time to compute this specific problem}.
-This result is due to the Hessian matrix only being a mere 55x55 matrix for this problem, such that the matrix inversion will not be computationally complex.
-However, if we were to increase the complexity of the problem, Newtons method would rapidly increase in computational complexity as it is a large matrix inversion with many floating point operations.
-For problems of such a small scale, when our features are based on a polynomial approximation Franke Function, it can be important to remember that we don't necessarily have to turn to these fancier solutions as the simple Newton method will do the job perfectly.
-Keeping this lesson in mind, we wish to further investigate the various scheduler methods in order to gain a better understanding of how well they perform, which will become useful for larger problems.
+In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function.
+IF HESSIAN: Using Newtons method, the computational cost is much greater than simply replacing the Hessian matrix inversion with a constant learning rate as can be seen by TIME COMPARISON.
+Furthermore, we can observe from FIGURE COMPARING HESSIAN TO CONSTANT that the constant scheduler does a good job at reducing the MSE, given that the final MSE score is CONSTANT SCORE compared to the Newton method's HESSIAN SCORE.
 
@@
 Already after EPOCHS epochs, the momentum scheduler has achieved MOMENTUM MSE AT EPOCHS, a notable improvement over CONSTANT MSE AT EPOCHS achieved by a constant learning rate.
+Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs.
+We can also observe that the different schedulers that they all reach REASONABLE MSE within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data.
+Looking at TIME TABLE FOR ALL SCHEDULERS, BEST took the least time, beating out our baseline constant scheduler by TIME DIFF.
+Surprisingly, RMS prop performed more poorly than expected.
 
-By introducing mini-batch gradient descent of different batch sizes, we can observe from BATCH SIZE FIG that the scheduler methods perform better when optimized for different batch sizes.
-Both the constant and the momentum scheduler seem to vary equally over batch sizes, with the optimal batch size being equal to one.
-This means that our schedulers work the best for a fully stochastic gradient descent.
-The MSE with full gradient descent, as can be seen in BATCH SIZE FIG when batch size equals 320, provides an MSE of
-A full comparison of batch sizes for all the different schedulers can be viewed in APPENDIX in the appendix.
+Before continuing, it is interesting to note that all the schedulers work best with a combination of similar eta and lambda values when gridsearched for achieving the minimal MSE.
+It is also interesting that they in the long term reach more or less the same error as can be observed in HEATMAP FIG.
+This could be because given more epochs, the other schedulers are given more time to catch up as the best performing schedulers, such as Adam, only decrease their MSE slightly.
+However, as can be seen in HEATMAP FIG, Adam offers a $\frac{0.018}{0.016} \dot 100 = 11.11\%$ decrease in MSE compared to the constant schedulers.
 
-Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs.
-We can also observe that the different schedulers all reach an MSE of 0.02 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.2.
-Looking at TIME TABLE FOR ALL SCHEDULERS, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds.
-Based on FIG ALL, Adam seems to reach an MSE of 0.2 the fastest when strictly looking at the number of epochs required, however, whilst Adagrad requires more epochs the computations are computed faster in our implementation.
-Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.
-
-An interesting insight from FIGURE NUM is that all the schedulers prefer high learning rates and regularization.
-As this plot was generated for 100 epochs due to the computational complexity of gridsearching the six schedulers, the bias towards higher learning rates is certainly due to the high learning rates being able to reach a low MSE in the short training time.
-With the lambda = 0 column representing the ordinary least squares solutions, these are clearly consistantly outperformed by Ridge regression as the best MSE is achieved with regularization.
-For this specific plot, momentum does not seem to make out a noticeable difference for the best solutions, as the optimal solution equals their momentumless counterparts.
-That being said, in FIG ALL it is clear that the added momentum aids in a quicker convergence when compared to the constant scheduler.
-
-Another observation from FIG NUM  is that the schedulers final error for100 epochs are more or less the same.
-As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased.
-However, as can be seen in HEATMAP FIG, Adagrad offers a 12\% \footnote{$\frac{constant MSE}{adagrad MSE} * 100 = \frac{0.018}{0.015} * 100 = 12\%$} decrease in MSE compared to the constant scheduler.
-It is clear from the results thus far that Adagrad is the best scheduler for approximating the Franke function when using linear regression with gradient descent.
-We believe that the reason Adagrad wins out is that these result are based on training for 100 epochs.
-The downfall of Adagrad in the literature seems to be that the aggressive decaying of the learning rate hinders a quick convergence deeper into training SOURCE https://towardsdatascience.com/learning-parameters-part-5-65a2f3583f7d.
-We are thus cautionary to claim that Adagrad will do the best job over a much larger amount of epochs when compared to adaptive schedulers that claim to improve this aspect.
-
-To test this, we compared Adagrad with Adam over 5000 epochs.
-Both schedulers performed equally, both resulting in an MSE of 0.017 with Adagrad taking the edge with an MSE 0.0002 lower.
-Scikit MSE 0.005.
-
-Furthermore, when trained for 5000 epochs, Adam with optimal parameters keeps improving the MSE.
-Scikit learn's Adam scheduler does perform better, offering a greater decrease the MSE.
-We compared Adagrad to scikit learn's Adam MLPRegressor as this seemed to be the closest option for comparing our results.
-Even though we have cross validated both our own Adagrad implementation and Scikit's with 5 folds, these differences in results might be due to the difference in solver, a more favorable weight initiation in Scikits model or additional optimizations present in the scikit learn implementation.
-Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of an MSE of 0.012 on the test data and Scikits final MSE of 0.004 (granted this is on the train data, as scikit learn does not provide the loss curve for the test data).
-both falling short of the analytical MSE of 0,003.
+However, when trained for 500 epochs, Adam with optimal parameters does not yield any improvement in MSE.
+Scikits Adam scheduler does perform better, further decreasing the MSE slightly.
+Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model.
+Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of ADAM FINAL MSE and Scikits final MSE of SCIKIT FINAL MSE both falling short of the analytical MSE of 0.003.
 In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions.
-This result might be due to 5000 epochs not being enough training time, as the Franke function is a quite complex function with added noise.
+This result might be due to 500 epochs not being enough training time.
 
@@
 
-For the organic terrain data used in our previous project, we found that our neural net performed slightly better, gaining an MSE of NUMBER, which was about PERCENTAGE better than our best polynomial fit.
-However, looking at our prediction visually, and also comparing the MSE to the one gained on the Franke Function, we see that we are still lacking.
-For terrain data like this, which contain a lot of spatial information, we believe that Convolutional Neural Networks could be a better fit.
+Having gained an understanding of how the different schedulers function, and having created the necessary infrastructure to study our models performance based on a number of different factors, we use these methods to optimize our neural network for linear regression.
+More specifically, we attempt to fit the Franke function.
+First, we perform a preliminary test on MSE over model complexity.
+For comparisons sake, we use the same scheduler for each architecture, and fix the regularization parameter at 0.
+By fixing the regularization parameter at 0, we are of course favouring less complex models, but this a good way to get a general sense of when the model starts to overfit.
+Using the sigmoid activation function, we can study the MSE as a function of the total hidden node count (Fig.~\ref{msebymodelfranke}).
+Here we see, somewhat surprisingly, that our train and test MSE are not deviating particularly much from eachother, even with very complex networks.
+There are many reasons that this may be the case.
+As we have stated before, once we start using numerical methods, computational effort becomes a large factor in any test one does, and many decisions have to made regarding how much time can be spent on each test.
+Due to these computational constraints, we have not been able to grid search for the best parameters for each model complexity, nor have we been able to train for more than 200 epochs for each model.
+These shortcomings will greatly affect how our models perform, as it is unlikely that each model uses the same learning rate and scheduler, or that more complex models use the same amount of epochs to train their weights.
+Based on this, we can assume that the higher complexity models have not been trained enough, and that this explains why both their train and test MSE are higher than the less complex models.
+Regardless, it seems that for up to about 200 total nodes, three hidden layers does the best, and as such we chose an architecture of three hidden layers with 66 nodes each (which is where we achieved our lowest MSE).
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=6in]{figure/msebymodelfranke.png}}
+\caption{Comparison between a neural net with one hidden layer, and one with two hidden layers, for different totals of hidden nodes.
+For the network with two hidden layers, the total is evenly divided between the layers, such that at $x = 60$ we have a network with two hidden layers of 30 nodes each.
+In this case, the hidden layers used the sigmoid activation function.}
+\label{msebymodelfranke}
+\end{figure}
+
+Using this architecture, we performed a gridsearch.
+Due to time constraints, we have felt it necessary to limit our gridsearch to only two schedulers.
+Based on our testing with linear regression with gradient descent, we have seen that Adam and Adagrad seems to be the best performers.
+Based on Goodfellows recommendations, we started with the RELU activation function \cite[188]{gbc}.
+With this, we found Adam to be the best scheduler, with an eta of 0.01 and no regularization.
+However, when using these parameters, we encountered the so called "dying RELU problem", where too many of our RELU nodes output zero, leading to our gradient descent slowing down and eventually stopping completely.
+A better weight initialization, for example using Xavier Initilization, could solve this problem, but another way to circumvent this issue is to use Leaky RELU.
+This activation function has a small but non-zero gradient for negative values, instead of zero as with RELU.
+Gridsearching with Leaky RELU instead gave the exact same parameters.
+
+Using these parameters along with Adam and LRELU, we acheived and MSE of VERYNICE, which is better than the results gained using gradient descent on our polynomial fit.
+Furthermore, we beat our gradient descended polynomial fit after a very short amount of epochs, achieving MSEs lower than 0.011 (the best gradient descended polynomial fit) after only EPOCH epochs.
+We can see that by not forcing our model to try and find a polynomial fit, we are achieving better results more quickly.
+However, we are still not able to beat the analytical solution to the best polynomial fit at polynomial degree 9, which is somewhat disappointing.
+A possible explanation for this is that the Franke function can be represented as a weighted sum of polynomials, and as such a polynomial fit is especially suited to approximate this function.
+Our neural net must learn the relation between the inputs and outputs on its own, a process which is computationally intensive and difficult.
+Looking at our predicted terrain (Fig.~\ref{predterrain}), we can see that our prediction is somewhat reasonable, but not entirely convincing.
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=6in]{figure/Franke_fit_66x3_lam0.png}}
+\caption{Comparison of our prediction of }
+\label{predterrain}
+\end{figure}
 
@@
 
+With mediocre results on regression, we turn to classification.
 To test our neural networks ability to solve classification problems, we first applied it to the Wisconsin Breast Cancer dataset.
-Using our gridsearch method, we found that the best scheduler was Adam, with an eta of $0.01$ and no regularization.
-For our architecture, we chose to use only one hidden layer with 60 nodes, based on our testing.
-With this, we were able to achieve a crossvalidated test accuracy of $98.6 \%$, which is very good.
+But to get the best results, there are a large amount of parameters that need to be decided on.
+We must find the best architecture, the best activation functions, best scheduler and the best parameters for the scheduler.
+Finding the absolute best combination of all these choices is impossible, so compromises must be made.
+
+We start by trying to find a good model complexity, i.e the best number of hidden layers, and the amount of nodes in them.
+To do this, we plotted the accuracy over different model architectures.
+Like we did during regression, we fix all the other parameters and study the architecture by itself.
+Using the sigmoid activation function (Fig.~\ref{cancerarchcompsigmoid}), our results are somewhat surprising.
+We see that even at 400 total hidden nodes, we are not overfitting.
+This implies that the dataset is quite homogenous, and that the test set is very similar to our training set.
+With such a dataset, it is possible to gain a very good fit using more and more complex models, although the computational effort obviously increases as well.
+Moreover, it is not given that instances outside this dataset are as homogenous as the relatively small (569) sample we have here, so one should be careful to use an overcomplex model even if it does not hurt the test accuracy for this particular dataset.
+Using the RELU activation function, the results are quite different, as can be seen in Fig.~\ref{cancerarchcomp}.
+Here we can observe that a neural net with two or three hidden layers starts performing very inconsistently when the number of nodes increases.
+Looking more closely at the training process, we can see that training stops completely in some of our crossvalidations.
+Clearly, we are again running into the "dying RELU problem".
+Here we can clearly see some of the benefits of the sigmoid, as even if the gradients become very small, they will never be zero, as with RELU.
+Regardless, we see similar performance for RELU using only one hidden layer.
+Based on these findings, we decided to use a model with a single hidden layer of 100 nodes, using the RELU activation function.
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=6in]{figure/cancerarchcompsigmoid.png}}
+\caption{Comparison between a neural net with one hidden layer, and one with two hidden layers, for different totals of hidden nodes.
+For the network with two hidden layers, the total is evenly divided between the layers, such that at $x = 60$ we have a network with two hidden layers of 30 nodes each.
+In this case, the hidden layers used the sigmoid activation function.}
+\label{cancerarchcompsigmoid}
+\end{figure}
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=6in]{figure/cancerarchcomp.png}}
+\caption{Comparison between a neural net with one hidden layer, and one with two hidden layers, for different totals of hidden nodes.
+For the network with two hidden layers, the total is evenly divided between the layers, such that at $x = 60$ we have a network with two hidden layers of 30 nodes each.
+In this case, the hidden layers used the RELU activation function.}
+\label{cancerarchcomp}
+\end{figure}
+
+Using our gridsearch method on this model, we found that the best scheduler was Adam, with an eta of $0.001$ and a lambda of $0.0001$.
+With this combination of architecture and scheduler, we were able to achieve a crossvalidated test accuracy of $98.76\%$, which is a great result.
+This outperforms SciKit's MLP classifier, which achieved a crossvalidated test accuracy of around 95\%.
 Looking at the confusion matrix (Fig.~\ref{cancerconf}), we see that we are achieving good predictions for both true positives and true negatives.
@@
 
-In this paper we have explored the use of a Feed Forward Neural Network for both regression and classification. We have studied how different schedulers,
+In this paper we have explored the use of a Feed Forward Neural Network for both regression and classification. We have studied how different schedulers, affect the gradient descent
 
@@
 \begin{table*}
-\caption{This table shows the command to execute to reproduce every figure in the report. (More info about the scripts and their parameters can be found in the README)}
+\caption{This table shows the file which created every plot in the report}
 \begin{center}
@@
 \begin{tabular}{c | l l l}
-Figure & Shell command (leading \texttt{python3} omitted) \\
+Figure & Shell command \\
 \hline
-\ref{frankenonoise} & \texttt{task\_b.py --noise 0 -n 25}\\
+\ref{verification} & \texttt{python3 testing/verification.py}\\
+\ref{cancerconflog} & \texttt{python3 task\_d.py}\\
+\ref{cancerconf} & \texttt{python3 task\_d.py}\\
+\ref{cancerlogneur} & \texttt{python3 task\_d.py}\\
+\ref{cancerarchcomp} & \texttt{python3 bycomplexitycancer.py}\\
+\ref{cancerarchcompsigmoid} & \texttt{python3 bycomplexitycancer.py}\\
+\ref{msebymodelfranke} & \texttt{python3 bycomplexityfranke.py}\\
 \hline

### 2022-11-16 doc/paper.tex
more changes

--- 2022-11-16 doc/paper.tex
@@
 
--hessian vs constant?
--constant vs momentum
--adaptive methods
--batch sizes
-
 In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function with 55 features.
@@
 
-FIG CONSTANT V HESSIAN
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/constant_v_hessian.png}}
+\caption{Constant scheduler with gridsearched hyperparameters vs hessian (Newton's method) MSE for test data plotted for 200 epochs}
+\label{constant_v_hessian}
+\end{figure}
 
-From observing FIG CONSTANT V HESSIAN we see that Newton's method using Hessian inversion greatly outperforms the constant scheduler by reaching an MSE of 0.02 for the testing data within the first epoch.
-This is due to the Hessian inversion in the first epoch in Newton's method being equivalent to the analytical solution.
+From observing Fig.~\ref{constant_v_hessian} we see that Newton's method using inversion of the  Hessian matrix greatly outperforms the constant scheduler by reaching an MSE of 0.020 for the testing data within the first epoch.
+Newton's method quickly converges, but does so with some error SULI P121, meaning it will not improve past this result.
+The analytical solution provides an MSE of approximately 0.003 REBER ET HANSSEN, a whole order of magnitude lower than the MSE Newton's method achieves.
 By comparison, the constant scheduler given 200 epochs reaches an MSE of 0.198 for the testing data.
@@
 Still, this provides an intuition that Newton's method does not take much time to compute this specific problem}.
-This result is due to the Hessian matrix only being a mere 56x56 matrix for this problem, such that the matrix inversion will not be computationally complex.
+This result is due to the Hessian matrix only being a mere 56x56 matrix for this problem, such that the Hessian matrix inversion will not be computationally complex.
 However, if we were to increase the complexity of the problem, Newtons method would rapidly increase in computational complexity as it is a large matrix inversion with many floating point operations.
-For problems of such a small scale, when our features are based on a polynomial approximation Franke Function, it can be important to remember that we don't necessarily have to turn to these fancier solutions as the simple Newton method will do the job perfectly.
+For problems of such a small scale, when our features are based on a polynomial approximation of the Franke Function, it can be important to remember that we don't necessarily have to turn to these fancier solutions as the simple Newton method will do the job perfectly very well for little computational cost.
 Keeping this lesson in mind, we wish to further investigate the various scheduler methods in order to gain a better understanding of how well they perform, which will become useful for larger problems.
 
-From FIGURE MOMENTUM v CONSTANT we can observe that adding momentum reduces the amount of epochs necessary to reach a lower MSE score.
-Already after EPOCHS epochs, the momentum scheduler has achieved MOMENTUM MSE AT EPOCHS, a notable improvement over CONSTANT MSE AT EPOCHS achieved by a constant learning rate.
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/constant_v_momentum.png}}
+\caption{Constant vs momentum MSE for test data plotted for 200 epochs using full gradient descent}
+\label{constant_v_momentum}
+\end{figure}
 
-From FIGURE MOMENTUM  v CONSTANT, we can observe that adding momentum does not improve upon the solutions found by the constant scheduler when using full gradient descent, reaching once again an MSE of 0.21 for the test data.
+From  Fig.~\ref{constant_v_momentum}, we can observe that adding momentum does not improve upon the solutions found by the constant scheduler when using full gradient descent, reaching once again an MSE of 0.21 for the test data.
 This implies that when both schedulers are gridsearched for optimal hyperparameters, the optimal value for the momentum hyperparameter is 0.
@@
 
-BATCH SIZE FIG (should mention in text that 320 is the size of the training data. Should mention base batch size for grid search was 18)
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/SGD_batch_momentum_constant.png}}
+\caption{Constant vs momentum test MSE for different batch sizes, plotted for 200 epochs using full gradient descent. A batch size of 320 is full GD, whilst a batch size of 1 is fully stochastic GD.}
+\label{batch_constant_v_momentum}
+\end{figure}
 
-By introducing mini-batch gradient descent, we can observe that
-
-By introducing mini-batch gradient descent of different batch sizes, we can observe from BATCH SIZE FIG that the scheduler methods perform better when optimized for different batch sizes over 100 epochs.
-Both the constant and the momentum scheduler seem to vary equally over batch sizes, with the optimal batch size being equal to one.
-This means that our schedulers work the best for a fully stochastic gradient descent (SGD).
-The MSE with full gradient descent, as can be seen in BATCH SIZE FIG when batch size equals 320, provides an MSE of 4.211 of the test data, whilst SGD provides an MSE of 0.28 for the test data in an equal amount of time.
-Because we first gridsearched the optimal values of eta and lambda first with a predefined base batch size, we must acknowledge that BATCH SIZE FIG will have some bias towards the batch size used when found the optimal hyperparameters.
+By introducing mini-batch gradient descent, we can observe that stochastic gradient descent with momentum starts outperforming the constant scheduler consistantly, requiring fewer steps to reach a low MSE.
+Mini-batch gradient descent clearly yields a lower MSE for the test data than full gradient descent for the same amount of epochs.
+The best result for both constant and momentum were reached for a batch size equal to 2, close to a fully stochastic gradient descent.
+The test data MSE reached 0.044, an impressive improvement over constant's previous result of 0.198 MSE with full gradient descent.
+Because we first gridsearched the optimal values of eta and lambda first with a predefined base batch size, we must acknowledge that Fig.~\ref{batch_constant_v_momentum} will have some bias towards the batch size used when found the optimal hyperparameters.
 However, given that we have gridsearched with different base batch sizes accross our testing, we have found the results to be similar accross different base batch sizes, and that this bias which we induced to save computational time was insignificant.
-A full comparison of MSE over batch sizes for all the different schedulers can be viewed in APPENDIX in the appendix.
+A full comparison of MSE over batch sizes for all the different schedulers can be viewed in APPENDIX in the appendix, but all schedulers yield the lowest MSE for stochastic gradient descent.
 
@@
 Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs.
-We can also observe that the different schedulers all reach an MSE of 0.02 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.2 of the test data.
+We can also observe that the different schedulers all reach an MSE of 0.200 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.200 of the test data.
 Looking at TIME TABLE FOR ALL SCHEDULERS, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds.
@@
 
-FIGURE NUM (heatmap)
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/100e_gridsearch.png}}
+\caption{MSE score gridsearch for hyperparameters eta, lambda for different schedulers}
+\label{heatmap}
+\end{figure}
 
-An interesting insight from FIGURE NUM is that all the schedulers prefer high learning rates and regularization.
+An interesting insight from Fig.~\ref{heatmap} is that all the schedulers prefer high learning rates and regularization.
 As this plot was generated for 100 epochs due to the computational complexity of gridsearching the six schedulers, the bias towards higher learning rates is certainly due to the high learning rates being able to reach a low MSE in the short training time.
@@
 
-Another observation from FIG NUM is that the schedulers final error for 100 epochs are more or less the same.
+Another observation from Fig.~\ref{heatmap} is that the schedulers final error for 100 epochs are more or less the same.
 As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased.
-However, as can be seen in HEATMAP FIG, Adagrad offers a 12\% \footnote{$\frac{constant MSE}{adagrad MSE} * 100 = \frac{0.018}{0.015} * 100 = 12\%$} decrease in MSE compared to the constant scheduler.
+However, as can be seen in Fig.~\ref{heatmap}, Adagrad offers a 12\% \footnote{$\frac{constant MSE}{adagrad MSE} * 100 = \frac{0.018}{0.015} * 100 = 12\%$} decrease in MSE compared to the constant scheduler.
 It is clear from the results thus far that Adagrad is the best scheduler for approximating the Franke function when using linear regression with gradient descent.
@@
 
-5000e_adam_v_adagrad
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/5000e_adam_v_adagrad.png}}
+\caption{MSE test error over 5000 epochs for Adam and Adagrad schedulers}
+\label{adam_v_adagrad}
+\end{figure}
 
-To test this, we compared Adagrad with Adam over 5000 epochs, cross validated for 5 folds.
+To test this, we compared Adagrad with Adam over 5000 epochs, cross validated for 5 folds, as can be seen in Fig.~\ref{heatmap}.
 Both schedulers perform almost equally, with Adagrad reaching a slightly lower MSE of 0.011 for the test data when compared to Adam's 0.012.
@@
 Our MSE might approach the analytical MSE, but this might take an unreasonable amount of epochs considering the computational runtime.
+Still, we have obtained an MSE lower than the Newton method's result of 0.020, even though it was at a much greater computation.
 Next, we investigate the use of linear regression in a neural network with hidden nodes, such that we can free ourselves of the limitations of the polynomial approximation.

### 2022-11-16 src/task_b.py
Working task_b.py

@@ -10,7 +10,7 @@ np.random.seed(1337)
     X,
     X_train,
     X_test,
-    z,
+    t,
     z_train,
     z_test,
     centering,
@@ -32,8 +32,8 @@ params = [eta, rho, rho2]
 #dims = (2, 266, 133, 1) # second best
 #dims = (2, 133, 133, 1)
 #dims = (2, 320, 160, 1) # second best 
-dims = (2, 133, 133, 133, 1) # best so far 
-train_epochs = 100
+dims = (2, 66, 66, 66, 1) # best so far 
+train_epochs = 200
 
 
 # ------------------------- FFNN -------------------------
@@ -66,17 +66,19 @@ optimal_batch = neural.optimize_batch(
     epochs=30,
 )
 """
-scores = neural.fit(
-    X_train[:, 1:3],
-    z_train,
+scores = neural.cross_val(
+    5,
+    X[:, 1:3],
+    t.reshape(400,1),
     sched,
     *params,
-    batches=X_train.shape[0],
+    batches=X.shape[0],
     epochs=train_epochs,
-    lam=0.0001,
-    X_test=X_test[:, 1:3],
-    t_test=z_test,
+    lam=0.0,
+    use_best_weights=True,
 )
+
+print(f"Final error: {scores['final_test_error']}")
 
 # ------------------------- MLPRegressor -------------------------
 mlp = MLPRegressor(

### 2022-11-16 doc/paper.tex
Snaper time

--- 2022-11-16 doc/paper.tex
@@
 However, we also found that simple logistic regression performed similarly, with a much lower computational cost.
-For regression applied to the Franke Function, we found that our neural net performed worse than a simple polynomial fit using linear regression, attaining a MSE of $0.007$, more than twice as high as the one achieved using linear regression.
-
-%we found our neural net lacking, not being able to beat our polynomial fit from our previous project, and only attaining an MSE of 1 BILLION, which is 2000 BILLION lower than our best polynomial fit of the Franke Function.
-
-% \tableofcontents
-% \newpage
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+For regression applied to the Franke Function, we found that our neural net performed worse than a simple polynomial fit using linear regression, attaining a MSE of $0.004781$, 24\% higher than our best polynomial fit MSE of $0.0038302$.
 

### 2022-11-16 doc/paper.tex

@@
 
-The Franke Function is a simple weighted sum of polynomials, taking two inputs and returning a real value. It is commonly used to test machine learning algorithms for regression problems.
+The Franke Function is a simple weighted sum of polynomials, taking two inputs and returning a real value.
+It is commonly used to test machine learning algorithms for regression problems.
+For our linear regression problems, our training data will be based on a 9th degree polynomial approximation to the Franke Function, with 56 features.
+For the feed forward neural network, we will use 2 of the features (x, y) of the Franke Function.
 
@@
 \hline
+Franke Function & synthetic & 56 & regression \\
 Franke Function & synthetic & 2 & regression \\
@@
 
--hessian vs constant?
--constant vs momentum
--adaptive methods
--batch sizes
+In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function with 55 features.
+We started off by comparing Newtons method for gradient descent to a constant learning rate using full batch gradient descent.
 
-In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function.
-IF HESSIAN: Using Newtons method, the computational cost is much greater than simply replacing the Hessian matrix inversion with a constant learning rate as can be seen by TIME COMPARISON.
-Furthermore, we can observe from FIGURE COMPARING HESSIAN TO CONSTANT that the constant scheduler does a good job at reducing the MSE, given that the final MSE score is CONSTANT SCORE compared to the Newton method's HESSIAN SCORE.
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/constant_v_hessian.png}}
+\caption{Constant scheduler with gridsearched hyperparameters vs hessian (Newton's method) MSE for test data plotted for 200 epochs}
+\label{constant_v_hessian}
+\end{figure}
 
-From FIGURE MOMENTUM v CONSTANT we can observe that adding momentum reduces the amount of epochs necessary to reach a lower MSE score.
-Already after EPOCHS epochs, the momentum scheduler has achieved MOMENTUM MSE AT EPOCHS, a notable improvement over CONSTANT MSE AT EPOCHS achieved by a constant learning rate.
+From observing Fig.~\ref{constant_v_hessian} we see that Newton's method using inversion of the  Hessian matrix greatly outperforms the constant scheduler by reaching an MSE of 0.020 for the testing data within the first epoch.
+Newton's method quickly converges, but does so with some error SULI P121, meaning it will not improve past this result.
+The analytical solution provides an MSE of approximately 0.003 REBER ET HANSSEN, a whole order of magnitude lower than the MSE Newton's method achieves.
+By comparison, the constant scheduler given 200 epochs reaches an MSE of 0.198 for the testing data.
+Given this result, it would seem indefensible to claim that Newtons method requires improvements, especially considering the time it took the Newton method 0.1512 seconds to compute 200 epochs, again beating the result from the constant scheduler which took 0.4991 seconds \footnote{this result should be taken with a grain of salt, given that the constant scheduler is run as a part of a larger architecture, whilst the Hessian calculation is done in a seperate script.
+Still, this provides an intuition that Newton's method does not take much time to compute this specific problem}.
+This result is due to the Hessian matrix only being a mere 56x56 matrix for this problem, such that the Hessian matrix inversion will not be computationally complex.
+However, if we were to increase the complexity of the problem, Newtons method would rapidly increase in computational complexity as it is a large matrix inversion with many floating point operations.
+For problems of such a small scale, when our features are based on a polynomial approximation of the Franke Function, it can be important to remember that we don't necessarily have to turn to these fancier solutions as the simple Newton method will do the job perfectly very well for little computational cost.
+Keeping this lesson in mind, we wish to further investigate the various scheduler methods in order to gain a better understanding of how well they perform, which will become useful for larger problems.
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/constant_v_momentum.png}}
+\caption{Constant vs momentum MSE for test data plotted for 200 epochs using full gradient descent}
+\label{constant_v_momentum}
+\end{figure}
+
+From  Fig.~\ref{constant_v_momentum}, we can observe that adding momentum does not improve upon the solutions found by the constant scheduler when using full gradient descent, reaching once again an MSE of 0.21 for the test data.
+This implies that when both schedulers are gridsearched for optimal hyperparameters, the optimal value for the momentum hyperparameter is 0.
+This result makes sense, as the momentum hyperparameter is used to reduce noise from derivatives calculated from smaller batch sizes, so-called mini batches.
+When utilizing a full gradient descent, where the batch size equals the size of the training data, momentum will not improve our results.
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/SGD_batch_momentum_constant.png}}
+\caption{Constant vs momentum test MSE for different batch sizes, plotted for 200 epochs using full gradient descent. A batch size of 320 is full GD, whilst a batch size of 1 is fully stochastic GD.}
+\label{batch_constant_v_momentum}
+\end{figure}
+
+By introducing mini-batch gradient descent, we can observe that stochastic gradient descent with momentum starts outperforming the constant scheduler consistantly, requiring fewer steps to reach a low MSE.
+Mini-batch gradient descent clearly yields a lower MSE for the test data than full gradient descent for the same amount of epochs.
+The best result for both constant and momentum were reached for a batch size equal to 2, close to a fully stochastic gradient descent.
+The test data MSE reached 0.044, an impressive improvement over constant's previous result of 0.198 MSE with full gradient descent.
+Because we first gridsearched the optimal values of eta and lambda first with a predefined base batch size, we must acknowledge that Fig.~\ref{batch_constant_v_momentum} will have some bias towards the batch size used when found the optimal hyperparameters.
+However, given that we have gridsearched with different base batch sizes accross our testing, we have found the results to be similar accross different base batch sizes, and that this bias which we induced to save computational time was insignificant.
+A full comparison of MSE over batch sizes for all the different schedulers can be viewed in APPENDIX in the appendix, but all schedulers yield the lowest MSE for stochastic gradient descent.
+
+FIG ALL
+TIME TABLE FOR ALL SCHEDULERS
+
 Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs.
-We can also observe that the different schedulers that they all reach REASONABLE MSE within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data.
-Looking at TIME TABLE FOR ALL SCHEDULERS, BEST took the least time, beating out our baseline constant scheduler by TIME DIFF.
-Surprisingly, RMS prop performed more poorly than expected.
+We can also observe that the different schedulers all reach an MSE of 0.200 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.200 of the test data.
+Looking at TIME TABLE FOR ALL SCHEDULERS, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds.
+Based on FIG ALL, Adam seems to reach an MSE of 0.2 the fastest when strictly looking at the number of epochs required, however, whilst Adagrad requires more epochs the computations are computed faster in our implementation.
+Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.
 
-Before continuing, it is interesting to note that all the schedulers work best with a combination of similar eta and lambda values when gridsearched for achieving the minimal MSE.
-It is also interesting that they in the long term reach more or less the same error as can be observed in HEATMAP FIG.
-This could be because given more epochs, the other schedulers are given more time to catch up as the best performing schedulers, such as Adam, only decrease their MSE slightly.
-However, as can be seen in HEATMAP FIG, Adam offers a $\frac{0.018}{0.016} \dot 100 = 11.11\%$ decrease in MSE compared to the constant schedulers.
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/100e_gridsearch.png}}
+\caption{MSE score gridsearch for hyperparameters eta, lambda for different schedulers}
+\label{heatmap}
+\end{figure}
 
-However, when trained for 500 epochs, Adam with optimal parameters does not yield any improvement in MSE.
-Scikits Adam scheduler does perform better, further decreasing the MSE slightly.
-Even though we have cross validated both our own Adam implementation and Scikit's with 5 folds, these differences in results might be due to a more favorable weight initiation in Scikits model.
-Nonetheless, neither models are able to beat the analytical solution provided by linear regression methods using polynomial approximations, with our best result of ADAM FINAL MSE and Scikits final MSE of SCIKIT FINAL MSE both falling short of the analytical MSE of 0.003.
-In theory, the feed forward neural network should be able to better approximate the Franke function given that it is not limited to polynomial solutions.
-This result might be due to 500 epochs not being enough training time.
+An interesting insight from Fig.~\ref{heatmap} is that all the schedulers prefer high learning rates and regularization.
+As this plot was generated for 100 epochs due to the computational complexity of gridsearching the six schedulers, the bias towards higher learning rates is certainly due to the high learning rates being able to reach a low MSE in the short training time.
+With the lambda = 0 column representing the ordinary least squares solutions, these are clearly consistantly outperformed by Ridge regression as the best MSE is achieved with regularization.
+For this specific plot, momentum does not seem to make out a noticeable difference for the best solutions, as the optimal solution equals their momentumless counterparts.
+That being said, in FIG ALL it is clear that the added momentum aids in a quicker convergence when compared to the constant scheduler.
+
+Another observation from Fig.~\ref{heatmap} is that the schedulers final error for 100 epochs are more or less the same.
+As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased.
+However, as can be seen in Fig.~\ref{heatmap}, Adagrad offers a 12\% \footnote{$\frac{constant MSE}{adagrad MSE} * 100 = \frac{0.018}{0.015} * 100 = 12\%$} decrease in MSE compared to the constant scheduler.
+It is clear from the results thus far that Adagrad is the best scheduler for approximating the Franke function when using linear regression with gradient descent.
+We believe that the reason Adagrad wins out is that these result are based on training for 100 epochs.
+The downfall of Adagrad in the literature seems to be that the aggressive decaying of the learning rate hinders a quick convergence deeper into training SOURCE https://towardsdatascience.com/learning-parameters-part-5-65a2f3583f7d.
+We are thus cautionary to claim that Adagrad will do the best job over a much larger amount of epochs when compared to adaptive schedulers that claim to improve this aspect such as Adam.
+
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/5000e_adam_v_adagrad.png}}
+\caption{MSE test error over 5000 epochs for Adam and Adagrad schedulers}
+\label{adam_v_adagrad}
+\end{figure}
+
+To test this, we compared Adagrad with Adam over 5000 epochs, cross validated for 5 folds, as can be seen in Fig.~\ref{heatmap}.
+Both schedulers perform almost equally, with Adagrad reaching a slightly lower MSE of 0.011 for the test data when compared to Adam's 0.012.
+It could be argued that 5000 epochs is not enough for the improvements of Adam to be noticeable, as the MSE could still be pushed much lower.
+However, for these computationally expensive runs it is clear that Adagrad is the scheduler of choice, especially considering our previous conclusion that Adagrad used less time to train.
+It should be noted that scikit learn's implementation outperforms both our implementation of Adagrad and Adam, with an MSE of 0.004.
+This could be due to a more favorable wight initialization in scikits model, along with additional optimizations implemented in their library.
+That we don't beat the analytical solution is no surprise, as we have limited ourself to a polynomial approximation of the Franke Function with 56 features.
+Our MSE might approach the analytical MSE, but this might take an unreasonable amount of epochs considering the computational runtime.
+Still, we have obtained an MSE lower than the Newton method's result of 0.020, even though it was at a much greater computation.
+Next, we investigate the use of linear regression in a neural network with hidden nodes, such that we can free ourselves of the limitations of the polynomial approximation.
 

### 2022-11-17 doc/paper.tex
Add appendix table

@@
 \section{Introduction}
+
 Many problems in machine learning cannot be solved analytically.
 Thus, we turn to numerical methods, but this change implicitly comes with computationally heavy caveats as they require large matrix calculations over many iterations.
-Finding a solution becomes a waiting game, finding the many solutions of varying hyperparameters in order to make an informed model selection becomes virtually impossible.
+Finding a solution becomes a waiting game, and finding the many solutions of varying hyperparameters in order to make an informed model selection becomes virtually impossible.
 The problem of model selection is at the heart machine learning, as optimal model selection allows us to make predictions with the greatest possible accuracy.
@@
 
-In this paper we will investigate implantations of the numerical optimization technique of gradient descent by researching stochastic gradient descent, along with constant, adaptive and momentum-enhanced schedulers with the aim of decreasing the computational cost of our calculations.
-We will study different supervised learning models which utilize gradient descent such as linear regression, logistic regression and a feed forward neural network (FFNN) in order to gain an understanding of model selection across regression, binary- and multi-class classification problems.
+In this paper we will investigate the numerical optimization technique of gradient descent and various techniques used to boost its performance.
+Thus, we will compare full batch versus stochastic gradient descent, different schedulers such as constant, adaptive and momentum-encanced schedulers, with the aim of decreasing the computational cost of our calculations.
+We will then study different supervised learning models which utilize gradient descent such as linear regression, logistic regression and a feed forward neural network (FFNN) in order to gain an understanding of model selection across regression, binary- and multi-class classification problems.
+Furthermore, we will expand our research to include comparisons between different activation functions and model architectures, gaining insights of the different trade-offs each decision entails.
+Through this, we will attempt to answer the following questions: How complex does our model have to be to achieve satisfying results?
+Which tradeoffs have to be made to reach these results while balancing the computational effort required to achieve them?
+
+% In this paper we are going to study supervised learning algorithms used for classification and regression problems, which utilize the numerical optimization technique of gradient descent as a means of finding an optimal solution.
+We will start by investigating different methods used to boost the performance of our gradient descent, and then proceed by diving deeper into the problem of model selection and the various trade offs which must be made to reach satisfying results while balancing the computational effort required to achieve them.
 
@@
 
+% \subsubsection{Linear Regression}
+
+% TODO
+
 \subsubsection{Logistic Regression}
 
-As part of the research conducted in this project, we introduce a new type of regression model, known as Logistic regression.
-This is a powerful, yet simple algorithm used in machine learning for the purpose of approximating a probability associated with the occurrence of a specific event given measured data.
-The fundamental idea behind logistic regression is based on the exact approach used in Linear Regression, meaning that we wish to describe a relationship between dependent and explanatory variables in terms of a set of estimated parameters.
+Logistic regression is a powerful, yet simple algorithm used in machine learning for the purpose of approximating a probability associated with the occurrence of a specific event given measured data.
+The fundamental idea behind logistic regression is based on the exact approach used in linear regression, meaning that we wish to describe a relationship between dependent and explanatory variables in terms of a set of estimated parameters.
 However, in this case we are not using the weighted sum of our basis functions directly, but feeding them into an activation function.
@@
 Rounding of said probability leads to labeling of each input as belonging to one of two classes, meaning that each object fed into the model will be classified as being of class 0 or 1 (in the case of binary classification).
-
-% To begin with however, the probability output by the activation function will be incorrect for majority of inputs, due to the random nature of the initialized parameters.
-While this may seem to be an issue, the weights can be adjusted to give better classification results through the application of the gradient descent coupled with an adjustable variable called learning rate.
-As we also shall see, the implementation of an activation function at the output gives a significant advantage over the classic Linear regression in terms of classification power.
+However, this method is not limited to binary classification, as the substitution of the activation function for SoftMax can allow for multiclass classification.
 
@@
 
-In our previous project, we have primarily dealt with problems that have an analytical solution.
+In linear regression problems, the optimal solution can be found analytically.
 This means that once a model has been decided on (polynomial degree and value for lambda for example), finding the optimal solution is a deterministic process which does not require further tuning.
 However, for the methods we use in this project - such as Logistic Regression - no analytical solutions exist, thus we will have to turn to numerical techniques to find the 'ideal' solution.
-Our method of choice in this case is gradient descent, an optimization technique which reduces the residuals by utilizing the idea of moving downhill a convex function until a globally minimal scalar value is reached.
+Our method of choice in this case is gradient descent, an optimization technique which reduces the residuals by utilizing the idea of moving downhill a function representing our cost until a local optimum is reached.
 The gradient itself is nothing more than a first order derivative of a given function, which points in the direction of the greatest change.
 
-The application of gradient descent to our problems as means of finding the ideal solution implies that the parameters be initialized randomly.
-The most common approach is to set each parameter equal to a normal distribution with mean zero and variance equal to one, and iteratively update the parameters by calculating the gradient at each given point in the residual space \cite[preface p.
+The application of gradient descent to our problems as a means of finding the ideal solution implies that the parameters used in our prediction be initialized randomly.
+The most common approach is to set the parameters according to a normal distribution with mean zero and variance equal to one, and iteratively update the parameters by calculating the gradient at each given point in the residual space \cite[preface p.
 x]{dds}.
@@
 \begin{equation}
-\frac{\partial{\cal C}}{\partial w_{jk}}  = \frac{\partial C}{\partial a_{j}}\frac{\partial a_j}{\partial w_{jk}}
+\frac{\partial{\cal C}}{\partial w_{jk}}  = \frac{\partial C}{\partial a_{j}}\frac{\partial a_j}{\partial w_{jk}}.
 \end{equation}
 
-Although gradient decent is an excellent algorithm for finding optimal solutions, it can also become quite computationally expensive for problems where large numbers of datapoints are available.
-However, this issue may be effectively countered by applying a variant of standard gradient approach known as Stochastic Gradient Decent or simply SGD.
+Although gradient descent is an excellent algorithm for finding optimal solutions, it can also become quite computationally expensive for problems where large numbers of datapoints are available.
+However, this issue may be effectively countered by applying a variant of the standard gradient descent known as Stochastic Gradient Decent or simply SGD.
 The SGD is based on randomly choosing one or a small number of instances from the dataset to update the parameters, hence the name stochastic.
-While it may seem like a poor idea to approximate the gradient instead of using the true gradient, the SGD has proven to give a large computational advantage and will often allow the model to converge into a global minima faster than GD.
+While it may seem like a poor idea to approximate the gradient instead of using the true gradient, SGD has proven to give a large computational advantage over full batch gradient descent.
 Furthermore, the stochastic nature of this method can allow SGD to avoid getting stuck in local minima \cite[46]{sr}.
@@
 These methods work by adaptively adjusting the learning rate, thus allowing us to avoid local minima and converge even faster.
-One of the simpler methods involves adding a fraction of the previous change to the current value used to update the weights, thus keeping a sort of memory of the previous iterations.
+One of the simpler methods involves adding a fraction of the previous change to the current value used to update the weights, hence keeping a sort of memory of the previous iterations.
 This is called momentum, due to the physical analogue of a moving particle having momentum.
@@
 Examples of schedulers using the second moment are RMS-prop, Adam and Adagrad.
-
-% Adaptive optimization methods don't always have good generalization.
-Recent studies have shown that adaptive methods such as ADAM, RMSPorp, and AdaGrad tend to have poor generalization compared to SGD or SGD with momentum, particularly in the high-dimensional limit (i.e.
-the number of parameters exceeds the number of data points).
-Although it is not clear at this stage why these methods perform so well in training deep neural networks, simpler procedures like properly-tuned SGD may work as well or better in these applications.
+The mathematical formulas for these scheduling algorithms can be found in Hjorth-Jensens lecture \cite{mortensched}.
 
@@
 
-For some problems, like approximating a simple function like the Franke Function in project 1, it is sufficient to choose a set of basis functions and find their optimal linear combination.
+For some problems, it is sufficient to choose a set of basis functions and find their optimal linear combination.
 Given that our problem can be reasonably approximated by this basis, we can achieve good results.
-However, by limiting ourselves to linear functions, we are possibly limiting our model, as our dataset may well be better approximated by a different basis, our perhaps not even by a linear combination of functions at all.
+However, by limiting ourselves to linear functions, we are possibly limiting our model, as our dataset may well be better approximated by a different basis, or perhaps not even by a linear combination of functions at all.
 Because of this, linear regression is not able to understand the interaction between two arbitrary input variables \cite[165]{gbc}.
@@
 
-% For an FFNN with an arbitrary amount of hidden layers, the output layer is thus defined as
-%
-% \begin{equation}
-% a^{l+1}_i = f^{l+1}\left[\!\sum_{j=1}^{N_l} w_{ij}^3 f^l\left(\sum_{k=1}^{N_{l-1}}w_{jk}^{l-1}\left(\dots f^1\left(\sum_{n=1}^{N_0} w_{mn}^1 x_n+ b_m^1\right)\dots\right)+b_k^2\right)+b_1^3\right].
-% \end{equation}
-%
-% \begin{equation}
-% 	a^{L}_i = f^{L}\left[\!\sum_{j=1}^{N_{L-1}} w_{ij}^3 f^{L-1}\left(\sum_{k=1}^{N_{L-2}}w_{jk}^{L-2}\left(\dots f^1\left(\sum_{n=1}^{N_0} w_{mn}^1 x_n+ b_m^1\right)\dots\right)+b_k^2\right)+b_1^3\right].
-% \end{equation}
-
 \subsubsection{Activation functions}
@@
 \begin{equation}
-y = \frac{1}{1 + exp(x)}
+y = \frac{1}{1 + exp(x)}.
 \end{equation}
@@
 w_{jk}^l\leftarrow  = w_{jk}^l- \eta \delta_j^la_k^{l-1} \\
-b_j^l \leftarrow b_j^l-\eta \frac{\partial {\cal C}}{\partial b_j^l}=b_j^l-\eta \delta_j^l
+b_j^l \leftarrow b_j^l-\eta \frac{\partial {\cal C}}{\partial b_j^l}=b_j^l-\eta \delta_j^l.
 \end{gather}
@@
 
-The Franke Function is a simple weighted sum of polynomials, taking two inputs and returning a real value.
+The Franke Function is a simple weighted sum of exponentials, taking two inputs and returning a real value.
 It is commonly used to test machine learning algorithms for regression problems.
 For our linear regression problems, our training data will be based on a 9th degree polynomial approximation to the Franke Function, with 56 features.
-For the feed forward neural network, we will use 2 of the features (x, y) of the Franke Function.
+For the feed forward neural network, we will use the x and y values directly.
+The Franke Function is defined as follows.
+
+\begin{multline}
+f(x,y) = \frac{3}{4}e^{\left(-\frac{(9x-2)^2}{4}-\frac{(9y-2)^2}{4}\right)} +  \frac{3}{4}e^{\left(-\frac{(9x+1)^2}{49}-\frac{(9y+1)}{10}\right)} +
+\frac{1}{2}e^{\left(-\frac{(9x-7)^2}{4}-\frac{(9y-3)^2}{4}\right)} - 
+\frac{1}{5}e^{\left(-(9x-4)^2-(9y-7)^2\right)}.
+\end{multline}
+
+In practice, we add a normally distributed noise $N(0,0.05)$ to this function.
 
@@
 \hline
-Franke Function & synthetic & 56 & regression \\
-Franke Function & synthetic & 2 & regression \\
+Franke Function (raw data) & 400 & 2 & regression \\
+Franke Function (polynomial features) & 400 & 56 & regression \\
 Wisconsin Breast Cancer & 569 & 30 & binary classification \\
@@
 
-In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function with 55 features.
-We started off by comparing Newtons method for gradient descent to a constant learning rate using full batch gradient descent.
+In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function with 56 features.
+To gain an intuition, we compare momentum and
 
-\begin{figure}[h]
-\centerline{\includegraphics[width=5in]{figure/constant_v_hessian.png}}
-\caption{Constant scheduler with gridsearched hyperparameters vs hessian (Newton's method) MSE for test data plotted for 200 epochs}
-\label{constant_v_hessian}
-\end{figure}
+% \begin{figure}[h]
+% \centerline{\includegraphics[width=5in]{figure/constant_v_hessian.png}}
+% \caption{Constant scheduler with gridsearched hyperparameters vs hessian (Newton's method) MSE for test data plotted for 200 epochs}
+% \label{constant_v_hessian}
+% \end{figure}
 
-From observing Fig.~\ref{constant_v_hessian} we see that Newton's method using inversion of the  Hessian matrix greatly outperforms the constant scheduler by reaching an MSE of 0.020 for the testing data within the first epoch.
-Newton's method quickly converges, but does so with some error SULI P121, meaning it will not improve past this result.
-The analytical solution provides an MSE of approximately 0.003 REBER ET HANSSEN, a whole order of magnitude lower than the MSE Newton's method achieves.
+% From observing Fig.~\ref{constant_v_hessian} we see that Newton's method using inversion of the  Hessian matrix greatly outperforms the constant scheduler by reaching an MSE of 0.020 for the testing data within the first epoch.
+Newton's method quickly converges, but does so with some error \cite[121]{num}, meaning it will not improve past this result.
+The analytical solution provides an MSE of approximately 0.004 \cite{ericjona}, a whole order of magnitude lower than the MSE Newton's method achieves.
 By comparison, the constant scheduler given 200 epochs reaches an MSE of 0.198 for the testing data.
-Given this result, it would seem indefensible to claim that Newtons method requires improvements, especially considering the time it took the Newton method 0.1512 seconds to compute 200 epochs, again beating the result from the constant scheduler which took 0.4991 seconds \footnote{this result should be taken with a grain of salt, given that the constant scheduler is run as a part of a larger architecture, whilst the Hessian calculation is done in a seperate script.
+Given this result, it would seem indefensible to claim that Newtons method requires improvements, especially considering the time it took.
+Newton's method took 0.1512 seconds to compute 200 epochs, again beating the result from the constant scheduler which took 0.4991 seconds \footnote{this result should be taken with a grain of salt, given that the constant scheduler is run as a part of a larger architecture, whilst the Hessian calculation is done in a seperate script.
 Still, this provides an intuition that Newton's method does not take much time to compute this specific problem}.
@@
 \centerline{\includegraphics[width=5in]{figure/constant_v_momentum.png}}
-\caption{Constant vs momentum MSE for test data plotted for 200 epochs using full gradient descent}
+\caption{Constant vs momentum MSE for test data plotted for 200 epochs using full batch gradient descent}
 \label{constant_v_momentum}
@@
 
-From  Fig.~\ref{constant_v_momentum}, we can observe that adding momentum does not improve upon the solutions found by the constant scheduler when using full gradient descent, reaching once again an MSE of 0.21 for the test data.
-This implies that when both schedulers are gridsearched for optimal hyperparameters, the optimal value for the momentum hyperparameter is 0.
+From  Fig.~\ref{constant_v_momentum}, we can observe that adding momentum improves upon the solutions found by the constant scheduler when using full batch gradient descent, reaching an MSE of NUMBER for the test data.
+By adding a memory of the previous gradients, we gain speed in the direction where most gradients point, speeding up our descent.
+This allows us to obtain a lower MSE scores in fever epochs, when compared to the constant scheduler's MSE of 0.198.
+
+% This implies that when both schedulers are gridsearched for optimal hyperparameters, the optimal value for the momentum hyperparameter is 0.
 This result makes sense, as the momentum hyperparameter is used to reduce noise from derivatives calculated from smaller batch sizes, so-called mini batches.
-When utilizing a full gradient descent, where the batch size equals the size of the training data, momentum will not improve our results.
+When utilizing a full batch gradient descent, where the batch size equals the size of the training data, momentum will not improve our results.
 
@@
 \centerline{\includegraphics[width=5in]{figure/SGD_batch_momentum_constant.png}}
-\caption{Constant vs momentum test MSE for different batch sizes, plotted for 200 epochs using full gradient descent. A batch size of 320 is full GD, whilst a batch size of 1 is fully stochastic GD.}
+\caption{Constant vs momentum test MSE for different batch sizes, plotted for 200 epochs using full batch gradient descent.
+A batch size of 320 is full GD, whilst a batch size of 1 is fully stochastic GD.}
 \label{batch_constant_v_momentum}
@@
 
-By introducing mini-batch gradient descent, we can observe that stochastic gradient descent with momentum starts outperforming the constant scheduler consistantly, requiring fewer steps to reach a low MSE.
-Mini-batch gradient descent clearly yields a lower MSE for the test data than full gradient descent for the same amount of epochs.
+By introducing mini-batch gradient descent, we can observe that stochastic gradient descent with momentum keeps outperforming the constant scheduler.
+Mini-batch gradient descent clearly yields a lower MSE for the test data than full batch gradient descent for the same amount of epochs.
 The best result for both constant and momentum were reached for a batch size equal to 2, close to a fully stochastic gradient descent.
-The test data MSE reached 0.044, an impressive improvement over constant's previous result of 0.198 MSE with full gradient descent.
+The test data MSE reached 0.044, an impressive improvement over constant's previous result of 0.198 MSE with full batch gradient descent.
 Because we first gridsearched the optimal values of eta and lambda first with a predefined base batch size, we must acknowledge that Fig.~\ref{batch_constant_v_momentum} will have some bias towards the batch size used when found the optimal hyperparameters.
@@
 \ref{msebymodelfranke} & \texttt{python3 bycomplexityfranke.py}\\
+\ref{constant_v_momentum} & \texttt{python3 constant\_v\_momentum.py}\\
+\ref{batch_constant_v_momentum} & \texttt{python3 constant\_v\_momentum.py}\\
+\ref{heatmap} & \texttt{python3 task\_a.py}\\
+\ref{adam_v_adagrad} & \texttt{python3 unknown.py}\\
+\ref{predterrain} & \texttt{python3 task\_b.py}\\
 \hline

### 2022-11-17 src/FFNN.py
Add more comments to FFNN

@@ -46,6 +46,7 @@ class FFNN:
         self.checkpoint_file = checkpoint_file
         self.seed = seed
 
+        # weight initialization
         for i in range(len(self.dimensions) - 1):
             if self.seed is not None:
                 np.random.seed(self.seed)
@@ -61,7 +62,7 @@ class FFNN:
         X: np.ndarray,
         t: np.ndarray,
         scheduler_class,
-        *args,  # arguments for the scheduler (sans batchsize)
+        *args,  # arguments for the scheduler
         batches: int = 1,
         epochs: int = 1000,
         lam: float = 0,
@@ -82,7 +83,7 @@ class FFNN:
         :param use_best_weights: save the best weights and only use them. This runs slower because
         it has to check every epoch if it is better than the best. Only use with X_test set
 
-        :return: error over epochs for training and test runs for plotting
+        :return: scores dictionary containing test and train error amongst other things
         """
 
         # --------- setup ---------
@@ -98,6 +99,7 @@ class FFNN:
         if X_test is not None and t_test is not None:
             test_set = True
 
+        # --- arrays of scores over epochs ----
         train_errors = np.empty(epochs)
         train_errors.fill(
             np.nan
@@ -118,9 +120,7 @@ class FFNN:
         if self.seed is not None:
             np.random.seed(self.seed)
 
-        # if X_test is not None:
-        #     test_preds = np.zeros((X_test.shape[0], epochs))
-
+        # logic for getting the best weights
         best_weights = deepcopy(self.weights)
         best_test_error = 10e20
         best_test_acc = 0
@@ -137,6 +137,7 @@ class FFNN:
         if test_set:
             cost_function_test = self.cost_func(t_test)
 
+        # create schedulers for each weight matrix
         for i in range(len(self.weights)):
             self.schedulers_weight.append(scheduler_class(*args))
             self.schedulers_bias.append(scheduler_class(*args))
@@ -152,14 +153,13 @@ class FFNN:
                         X_batch = X[i * batch_size :, :]
                         t_batch = t[i * batch_size :, :]
                     else:
-                        # print(f"{t=}")
                         X_batch = X[i * batch_size : (i + 1) * batch_size, :]
                         t_batch = t[i * batch_size : (i + 1) * batch_size, :]
 
                     self._feedforward(X_batch)
                     self._backpropagate(X_batch, t_batch, lam)
 
-                # reset schedulers every epoch
+                # reset schedulers every epoch (some schedulers pass in this call)
[…]

### 2022-11-17 doc/paper.tex
Added some stuff

@@
 \usepackage{xurl}
-
-\definecolor{codegreen}{rgb}{0,0.6,0}
-\definecolor{codegray}{rgb}{0.5,0.5,0.5}
-\definecolor{codepurple}{rgb}{0.58,0,0.82}
-\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
-
-% \expandafter\def\expandafter\UrlBreaks\expandafter{\UrlBreaks%  save the current one
-%   \do\a\do\b\do\c\do\d\do\e\do\f\do\g\do\h\do\i\do\j%
-%   \do\k\do\l\do\m\do\n\do\o\do\p\do\q\do\r\do\s\do\t%
-%   \do\u\do\v\do\w\do\x\do\y\do\z\do\A\do\B\do\C\do\D%
-%   \do\E\do\F\do\G\do\H\do\I\do\J\do\K\do\L\do\M\do\N%
-%   \do\O\do\P\do\Q\do\R\do\S\do\T\do\U\do\V\do\W\do\X%
-%   \do\Y\do\Z}
-
-\lstdefinestyle{mystyle}{
-    backgroundcolor=\color{backcolour},
-    commentstyle=\color{codegreen},
-    keywordstyle=\color{magenta},
-    numberstyle=\tiny\color{codegray},
-    stringstyle=\color{codepurple},
-    basicstyle=\ttfamily\footnotesize,
-    breakatwhitespace=false,
-    breaklines=true,
-    captionpos=b,
-    keepspaces=true,
-    numbers=left,
-    numbersep=5pt,
-    showspaces=false,
-    showstringspaces=false,
-    showtabs=false,
-    tabsize=2
-}
-
-\lstset{breaklines=true, style=mystyle}
-%% The class has several options
-%  onecolumn/twocolumn - format for one or two columns per page
-%  10pt/11pt/12pt - use 10, 11, or 12 point font
-%  oneside/twoside - format for oneside/twosided printing
-%  final/draft - format for final/draft copy
-%  cleanfoot - take out copyright info in footer leave page number
-%  cleanhead - take out the conference banner on the title page
-%  titlepage/notitlepage - put in titlepage or leave out titlepage
-%  
-%% The default is oneside, onecolumn, 10pt, final
 
@@
 To optimize our network, we explored the use of different schedulers in our gradient descent method, and compared their performance against eachother.
-Amongst our schedulers, we found that Adam and Adagrad performed the best.
-Overall, we found that our FFNN performed well in both classification tasks we tested, attaining a test accuracy of $98.76$ on the Wisconsin Breast Cancer Dataset and a test accuracy of $99.07$ on the MNIST 8x8 dataset.
-However, we also found that simple logistic regression performed similarly, with a much lower computational cost.
+Amongst our schedulers, we found that Adam and Adagrad performed best achieving a METRIC THAT IS BETTER.
+In our applications of our neural network, we found that it performed well in both classification tasks we tested, attaining a test accuracy of $98.59\%$ on the Wisconsin Breast Cancer Dataset and a test accuracy of $99.07\%$ on the MNIST 8x8 dataset.
+However, we also found that simple logistic regression performed similarly in terms of accuracy, achieving a test accuracy of $98.76\%$ and $99.54\%$ on the cancer and MNIST dataset, respectively.
+Moreover, these results came at a much lower computational cost.
 For regression applied to the Franke Function, we found that our neural net performed worse than a simple polynomial fit using linear regression, attaining a MSE of $0.004781$, 24\% higher than our best polynomial fit MSE of $0.0038302$.
+We conclude that allthough neural networks are powerful tools which can be used in many contexts, there are many cases in which simpler and less computationally intensive methods can be used to gain similar or even superior performance.
 
@@
 
+Our implementation can be found at \texttt{github.com/EricEReber/FFNN\_optimization.git}.
+
 We implement the FFNN as a class, using NumPy's many matrix functions.
@@
 Because of this, no explicit implementation of linear and logistic regression has been made, which means that almost all relevant code is contained in \texttt{src/FFNN.py} and \texttt{src/Schedulers.py}.
+This unity in implementation ensures that we can directly compare metrics obtained from the different models we are comparing, e.g neural network versus logistic regression.
 When instatiating the model, we set the weights to normal distributions with mean zero and variance one.
@@
 It is commonly used to test machine learning algorithms for regression problems.
-For our linear regression problems, our training data will be based on a 9th degree polynomial approximation to the Franke Function, with 56 features.
+For our linear regression problems, our training data will be based on a 9th degree polynomial fit of the Franke Function, with 56 features.
 For the feed forward neural network, we will use the x and y values directly.
@@
 
-In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function with 56 features.
+In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function with 56 polynomial features.
 To gain an intuition, we compare momentum and
@@
 \centerline{\includegraphics[width=5in]{figure/constant_v_momentum.png}}
-\caption{Constant vs momentum MSE for test data plotted for 200 epochs using full batch gradient descent}
+\caption{Constant vs momentum MSE for test data plotted for 200 epochs using full batch gradient descent. The final MSE for Constant was 0.198, ten percent higher than Momentum's 0.180.}
 \label{constant_v_momentum}
@@
 
-From  Fig.~\ref{constant_v_momentum}, we can observe that adding momentum improves upon the solutions found by the constant scheduler when using full batch gradient descent, reaching an MSE of NUMBER for the test data.
+From  Fig.~\ref{constant_v_momentum}, we can observe that adding momentum improves upon the solutions found by the constant scheduler when using full batch gradient descent, reaching an MSE of 0.1799 for the test data.
 By adding a memory of the previous gradients, we gain speed in the direction where most gradients point, speeding up our descent.
@@
 \caption{Constant vs momentum test MSE for different batch sizes, plotted for 200 epochs using full batch gradient descent.
-A batch size of 320 is full GD, whilst a batch size of 1 is fully stochastic GD.}
+A batch size of 320 is full batch GD, whilst a batch size of 1 is fully stochastic GD.}
 \label{batch_constant_v_momentum}
@@
 
-FIG ALL
-TIME TABLE FOR ALL SCHEDULERS
+\begin{figure}[h]
+\centerline{\includegraphics[width=5in]{figure/200e_schedulers.png}}
+\caption{Test MSE over epochs for all schedulers}
+\label{figall}
+\end{figure}
 
-Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from FIGURE ALL that we can reach even greater results in fewer epochs.
+Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from Fig.~\ref{figall} that we can reach even greater results in fewer epochs.
 We can also observe that the different schedulers all reach an MSE of 0.200 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.200 of the test data.
-Looking at TIME TABLE FOR ALL SCHEDULERS, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds.
-Based on FIG ALL, Adam seems to reach an MSE of 0.2 the fastest when strictly looking at the number of epochs required, however, whilst Adagrad requires more epochs the computations are computed faster in our implementation.
+Looking at Table.~\ref{timetable}, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds.
+Based on Fig.~\ref{figall}, Adam seems to reach an MSE of 0.2 the fastest when strictly looking at the number of epochs required, however, whilst Adagrad requires more epochs the computations are computed faster in our implementation.
 Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.
+
+\begin{table}[h]
+\caption{Table over time take for each scheduler}
+\begin{center}
+\label{timetable}
+\begin{tabular}{| c | l |}
+\hline
+Scheduler & Time to reach 0.2 test MSE \\
+\hline
+Constant & 0.601 \\
+Momentum & 0.513 \\
+RMS\_prop & 0.330 \\
+AdaGrad & 0.013 \\
+AdaGrad with Momentum & 0.014 \\
+Adam & 0.041 \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 For this specific plot, momentum does not seem to make out a noticeable difference for the best solutions, as the optimal solution equals their momentumless counterparts.
-That being said, in FIG ALL it is clear that the added momentum aids in a quicker convergence when compared to the constant scheduler.
+That being said, in Fig.~\ref{figall} it is clear that the added momentum aids in a quicker convergence when compared to the constant scheduler.
 
 Another observation from Fig.~\ref{heatmap} is that the schedulers final error for 100 epochs are more or less the same.
-As the largest change in MSE occurs in the first 50 epochs as seen in ALL FIGS, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased.
+As the largest change in MSE occurs in the first 50 epochs as seen in Fig.~\ref{figall}, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased.
 However, as can be seen in Fig.~\ref{heatmap}, Adagrad offers a 12\% \footnote{$\frac{constant MSE}{adagrad MSE} * 100 = \frac{0.018}{0.015} * 100 = 12\%$} decrease in MSE compared to the constant scheduler.
@@
 We believe that the reason Adagrad wins out is that these result are based on training for 100 epochs.
-The downfall of Adagrad in the literature seems to be that the aggressive decaying of the learning rate hinders a quick convergence deeper into training SOURCE https://towardsdatascience.com/learning-parameters-part-5-65a2f3583f7d.
+The downfall of Adagrad in the literature seems to be that the aggressive decaying of the learning rate hinders a quick convergence deeper into training \cite{datasci}.
 We are thus cautionary to claim that Adagrad will do the best job over a much larger amount of epochs when compared to adaptive schedulers that claim to improve this aspect such as Adam.
@@
 
-Having gained an understanding of how the different schedulers function, and having created the necessary infrastructure to study our models performance based on a number of different factors, we use these methods to optimize our neural network for linear regression.
+Having gained an understanding of how the different schedulers function, and having created the necessary infrastructure to study our models' performance based on a number of different factors, we use these methods to optimize our neural network for linear regression.
 More specifically, we attempt to fit the Franke function.
@@
 
-Using these parameters along with Adam and LRELU, we acheived and MSE of VERYNICE, which is better than the results gained using gradient descent on our polynomial fit.
-Furthermore, we beat our gradient descended polynomial fit after a very short amount of epochs, achieving MSEs lower than 0.011 (the best gradient descended polynomial fit) after only EPOCH epochs.
-We can see that by not forcing our model to try and find a polynomial fit, we are achieving better results more quickly.
+Using these parameters along with Adam and LRELU, we acheived an MSE of $0.004781$, which is better than the results gained using gradient descent on our polynomial fit.
+We can see that by not forcing our model to try and find a polynomial fit, we are achieving better results.
 However, we are still not able to beat the analytical solution to the best polynomial fit at polynomial degree 9, which is somewhat disappointing.
@@
 Looking at our predicted terrain (Fig.~\ref{predterrain}), we can see that our prediction is somewhat reasonable, but not entirely convincing.
+Furthermore, we can see that the polynomial fit seems to match the actual terrain more closely, which it achieves in a fraction of the runtime.
 
 \begin{figure}[h]
-\centerline{\includegraphics[width=6in]{figure/Franke_fit_66x3_lam0.png}}
-\caption{Comparison of our prediction of }
+\centerline{\includegraphics[width=7in]{figure/Franke_fit_66x3_lam0.png}}
+\caption{Surface plot of the actual terrain created by the Franke function, the best fit achieved by our Neural Network, and the best polynomial fit for polynomial degree 9}
 \label{predterrain}
@@
 But to get the best results, there are a large amount of parameters that need to be decided on.
-We must find the best architecture, the best activation functions, best scheduler and the best parameters for the scheduler.
+We must find the best architecture, the best activation functions, the best scheduler and the best parameters for said scheduler.
 Finding the absolute best combination of all these choices is impossible, so compromises must be made.
@@
 Using our gridsearch method on this model, we found that the best scheduler was Adam, with an eta of $0.001$ and a lambda of $0.0001$.
-With this combination of architecture and scheduler, we were able to achieve a crossvalidated test accuracy of $98.76\%$, which is a great result.
-This outperforms SciKit's MLP classifier, which achieved a crossvalidated test accuracy of around 95\%.
+With this combination of architecture and scheduler, we were able to achieve a crossvalidated test accuracy of $98.59\%$, which is a great result.
+This outperforms SciKit's MLP classifier, which achieved a crossvalidated test accuracy of around $95\%$.
 Looking at the confusion matrix (Fig.~\ref{cancerconf}), we see that we are achieving good predictions for both true positives and true negatives.
@@
 \begin{figure}[h]
-\centerline{\includegraphics[width=5in]{figure/cancerconf.png}}
+\centerline{\includegraphics[width=5in]{figure/conf_mat_NN_cancer.png}}
 \caption{Confusion matrix for the cancer dataset using a neural net with a single hidden layer of 60 nodes}
@@
 Using logistic regression on the cancer dataset, we found through our gridsearch that Adam with an eta $0.1$ gave the best results.
-Here we found that logistic regression performed essentially equivalent or even better, at a much lower computational cost.
+Here we found that logistic regression performed essentially equivalent to our neural net, achieving a test accuracy of $98.76\%$.
 Looking at Fig.~\ref{cancerlogneur}, we see that logistic regression keeps up and even surpasses our neural network in accuracy over epochs.
@@
 \begin{figure}[h]
-\centerline{\includegraphics[width=5in]{figure/cancerlogneur.png}}
+\centerline{\includegraphics[width=5in]{figure/comp_NN_logreg_cancer.png}}
 \caption{Comparison between the test accuracy of logistic regression and a neural network on classifying the cancer dataset}
@@
 \begin{figure}[h]
-\centerline{\includegraphics[width=5in]{figure/cancerconflog.png}}
+\centerline{\includegraphics[width=5in]{figure/conf_mat_logreg_cancer.png}}
 \caption{Confusion matrix for the cancer dataset using logistic regression}
@@
 
-In this paper we have explored the use of a Feed Forward Neural Network for both regression and classification. We have studied how different schedulers, affect the gradient descent
+In this paper we have studied supervised learning algorithms used for classification and regression problems, using the numerical optimization technique of gradient descent as a means of finding an optimal solution.
+We have observed how different schedulers improve the performance of our gradient descent, and found that amongst all the schedulers we tested, Adam and AdaGrad exhibited the best performance overall, achieving a METRIC THAT IS BETTER.
+These schedulers were then used for numerical linear regression, logistic regression and a feed forward neural network.
+Solving both classification and regression problems with these models, we found that our neural network
+
+FUTURE WORK AN ANALYSIS
+
+We then implemented these schedulers in a Feed Forward Neural Network, and applied said network to both regression and classification problems.
+
+Using these schedulers, we applied a Feed Forward Neural network to both classification and regression problems.
+
+Through the application of a Feed Forward Neural Network to both regression and classification problems, we have found that for the problems we have studied, simpler and less computationally intensive methods proved to be superior.
+
+In this paper we have explored the use of a Feed Forward Neural Network, as well as numerical  for both regression and classification alon. Furthermore, we have applied
 
@@
 \hline
-\ref{verification} & \texttt{python3 testing/verification.py}\\
-\ref{cancerconflog} & \texttt{python3 task\_d.py}\\
-\ref{cancerconf} & \texttt{python3 task\_d.py}\\
-\ref{cancerlogneur} & \texttt{python3 task\_d.py}\\
-\ref{cancerarchcomp} & \texttt{python3 bycomplexitycancer.py}\\
-\ref{cancerarchcompsigmoid} & \texttt{python3 bycomplexitycancer.py}\\
-\ref{msebymodelfranke} & \texttt{python3 bycomplexityfranke.py}\\
+\ref{verification} & \texttt{python3 verification.py}\\
 \ref{constant_v_momentum} & \texttt{python3 constant\_v\_momentum.py}\\
@@
 \ref{adam_v_adagrad} & \texttt{python3 unknown.py}\\
+\ref{msebymodelfranke} & \texttt{python3 bycomplexityfranke.py}\\
 \ref{predterrain} & \texttt{python3 task\_b.py}\\
+\ref{cancerarchcompsigmoid} & \texttt{python3 bycomplexitycancer.py}\\
+\ref{cancerarchcomp} & \texttt{python3 bycomplexitycancer.py}\\
+\ref{cancerconf} & \texttt{python3 task\_d.py}\\
+\ref{cancerlogneur} & \texttt{python3 task\_d.py}\\
+\ref{cancerconflog} & \texttt{python3 task\_d.py}\\
 \hline

### 2022-11-17 doc/paper.tex
rewrite abstract

@@
 We conclude that allthough neural networks are powerful tools which can be used in many contexts, there are many cases in which simpler and less computationally intensive methods can be used to gain similar or even superior performance.
+
+To rigidly test different models for regression and classification against eachother, we developed a flexible Feed Forward Neural Network that can be specified for various architectures.
+We used this network to test different schedulers, activation functions and model complexities in order to achieve a balance of satisfying results and low computational cost accross different datasets.
+Pitting the schedulers against eachother to fit a polynomial approximation of the Franke Function using linear regression, Adam and Adagrad came out on top, with Adagrad using less time to complete the fit.
+Furthermore, Adagrad's cross validated test MSE of 0.011 being slightly lower than Adam's 0.012 when ran for 5000 epochs.
+For neural networks problems, Adam proved to be the superior scheduler, ---
 

### 2022-11-17 src/bycomplexityfranke.py
Comments added to scripts

@@ -12,7 +12,14 @@ from sklearn.model_selection import train_test_split, GridSearchCV, KFold
 
 np.random.seed(1337)
 
-# read in data
+"""
+When run, the following code generates a plot the accuracy achieved by a model
+as function of its complexity, i.e the number og hidden layers and nodes 
+in each of said layers. 
+"""
+
+# ------------------------ Loading data ------------------------
+
 (
     betas_to_plot,
     N,
@@ -32,7 +39,7 @@ z_test = z_test.reshape(z_test.shape[0], 1)
 z = z.ravel()
 z = z.reshape(z.shape[0], 1)
 
-# epochs to run for
+# ----------------------- Setting params -----------------------
 epochs = 10
 folds = 5
 scheduler = Adam
@@ -40,6 +47,7 @@ args = [0.001, 0.9, 0.999]
 
 funcs = [sigmoid, lambda x: x, CostOLS]
 
+# ------------------------ Plots ------------------------
 results = plot_arch(
     FFNN,
     400,

### 2022-11-17 src/constant_v_hessian.py

@@ -1,12 +1,22 @@
+# Our own library
 from utils import *
 from Schedulers import *
 from FFNN import FFNN
+# Other libraries
 import timeit
 from tabulate import tabulate
 
+
+"""
+The following script first performs a gridsearch for optimal parameters lambda and eta for use 
+with regular full-batch gradient descent. It the proceeds with the measurement of the runtime 
+of our FFNN and Newton-Raphon's method, before it finally prints said runtime and test error 
+acieved by each technique for the sake of comparison of both methods against eachother.
+"""
+
 np.random.seed(1337)
 
-# read in data
+# ---------------------- Loading data ---------------------- 
 (
     betas_to_plot,
     N,
@@ -24,7 +34,7 @@ np.random.seed(1337)
 z_train = z_train.reshape(z_train.shape[0], 1)
 z_test = z_test.reshape(z_test.shape[0], 1)
 
-# define params
+# --------------------- Setting params --------------------- 
 epochs = 200
 folds = 5
 
@@ -35,11 +45,11 @@ lam = np.logspace(-5, -1, 5)
 constant_scheduler = Constant
 constant_params = []
 
-# linear regression NN
+# ------------------ Linear Regression NN ------------------
 dims = (X.shape[1], 1)
 neural = FFNN(dims, seed=1337)
 
-# optimize constant learning rate, regularization
+# Gridsearching parameters learning rate and lambda for "Constant" scheduler, i.e full-batch gradient descent
 optimal_params, optimal_lambda, _ = neural.optimize_scheduler(
     X_train,
     z_train,
@@ -52,7 +62,7 @@ optimal_params, optimal_lambda, _ = neural.optimize_scheduler(
     folds=folds,
 )
 
-# take average runtime constant
+# Measurement of average runtime of our FFNN model
 time = np.zeros(2)
 start = timeit.default_timer()
 constant_scores = neural.cross_val(
@@ -70,7 +80,7 @@ time[0] = (stop - start) / folds
 print(f"Constant time to train {epochs} epochs: {time[0]}")
 constant_test_errors = constant_scores["test_errors"]
 
-# take average runtime Newton's
+# Measurement of average runtime of Newton-Raphson's method 
 start = timeit.default_timer()
 hessian_scores, _ = hessian_cv(folds, X_train, z_train, epochs=epochs)
 stop = timeit.default_timer()
@@ -78,7 +88,7 @@ time[1] = (stop - start) / folds
[…]

### 2022-11-17 src/constant_v_momentum.py

@@ -3,7 +3,15 @@ from Schedulers import *
 from FFNN import FFNN
 import timeit
 
-# read in data
+
+
+"""
+The following script first performs a gridsearch for optimal parameters lambda and eta for use 
+with regular full-batch gradient descent and momentum gradient descent. It the proceeds with the 
+measurement of the runtime of each of said methods, before it finally prints said runtime and 
+test error acieved by each technique for the sake of comparison of both schedulers against eachother.
+"""
+# ---------------------- Loading data ---------------------- 
 (
     betas_to_plot,
     N,
@@ -21,7 +29,7 @@ import timeit
 z_train = z_train.reshape(z_train.shape[0], 1)
 z_test = z_test.reshape(z_test.shape[0], 1)
 
-# define params
+# --------------------- Setting params --------------------- 
 epochs = 200
 folds = 5
 
@@ -34,7 +42,7 @@ constant_params = []
 momentum_params = np.linspace(0, 0.1, 5)
 params = [momentum_params, constant_params]
 
-# linear regression NN
+# ------------------ Linear Regression NN ------------------
 dims = (X.shape[1], 1)
 neural = FFNN(dims, seed=1337)
 
@@ -43,7 +51,7 @@ batches_list = np.logspace(
 )
 
 sns.set(font_scale=3)
-# optimize constant, momentum hyperparameters
+# Gridsearching parameters learning rate and lambda for "Constant" and Momentum scheduler
 for i in range(len(schedulers)):
     optimal_params, optimal_lambda, _ = neural.optimize_scheduler(
         X_train,
@@ -57,7 +65,7 @@ for i in range(len(schedulers)):
         folds=folds,
     )
 
-    # take average runtime constant
+    # Measurement of average runtime of our FFNN model
     start = timeit.default_timer()
     constant_scores = neural.cross_val(
         folds,

### 2022-11-17 src/task_c.py
Comments in task_c.py

@@ -1,12 +1,22 @@
+# Imports
 import matplotlib.pyplot as plt
 import numpy as np
 import seaborn as sns
-
+# Imports from our own library
 from utils import *
 from Schedulers import *
 from FFNN import FFNN
 from sklearn.neural_network import MLPRegressor
 
+"""
+This script is used for comparision of results achieved when fitting our FFNN model 
+with functions sigmoid, RELU and LRELU as activation functions in the hidden layers. 
+THe program first gridsearches optimal eta, lambda and batch size for each activation 
+function, before finally plotting the mse achieved by the model when using each of the 
+functions mentioned. 
+"""
+
+# -------------------------- Loading data --------------------------
 np.random.seed(42069)
 (
     betas_to_plot,
@@ -26,7 +36,7 @@ np.random.seed(42069)
 z_train = z_train.reshape(z_train.shape[0], 1)
 z_test = z_test.reshape(z_test.shape[0], 1)
 
-# ------------------------- Params -------------------------
+# ------------------------- Setting params -------------------------
 eta = 0.00005
 momentum = 0.5
 rho = 0.9
@@ -49,7 +59,8 @@ hidden_func_list = [sigmoid, RELU, LRELU]
 
 for i in range(len(hidden_func_list)):
     neural = FFNN(dimensions=dims, hidden_func=hidden_func_list[i], seed=42069)
-
+    
+    # Gridsearching optimal parameters for each of the hidden activation functions
     optimal_params, optimal_lambda, _ = neural.optimize_scheduler(
         X_train[:, 1:3],
         z_train,
@@ -64,7 +75,8 @@ for i in range(len(hidden_func_list)):
     )
 
     params = [optimal_params[0], rho, rho2]
-
+    
+    # Gridsearching optimal batchsize for each hidden activation functions using optimal eta and lambda 
     optimal_batch = neural.optimize_batch(
         X_train[:, 1:3],
         z_train,
@@ -76,7 +88,8 @@ for i in range(len(hidden_func_list)):
         batches_list=batch_sizes,
         epochs=20,
     )
-
+    
+    # Fitting the model after achieving optimal hyperparameters eta, lambda and batchsize
     scores = neural.fit(
         X_train[:, 1:3],
         z_train,
@@ -92,6 +105,8 @@ for i in range(len(hidden_func_list)):
     test_error = scores["train_error"]
[…]

### 2022-11-17 doc/paper.tex
Final

--- 2022-11-17 doc/paper.tex
@@
 
-We developed a Feed Forward Neural Network (FFNN) and applied it to both regression and classification problems.
+% We developed a Feed Forward Neural Network (FFNN) and applied it to both regression and classification problems.
 Furthermore, we compared our results to those attained by linear and logistic regression.
@@
 Amongst our schedulers, we found that Adam and Adagrad performed best achieving a METRIC THAT IS BETTER.
-In our applications of our neural network, we found that it performed well in both classification tasks we tested, attaining a test accuracy of $98.59\%$ on the Wisconsin Breast Cancer Dataset and a test accuracy of $99.07\%$ on the MNIST 8x8 dataset.
+
+To rigidly test different models for regression and classification against eachother, we developed a flexible Feed Forward Neural Network that can be specified for various architectures.
+We used this network to test different schedulers, activation functions and model complexities in order to achieve a balance of satisfying results and low computational cost accross different datasets.
+Pitting the schedulers against eachother to fit a polynomial approximation of the Franke Function using linear regression, Adam and Adagrad came out on top, achieving MSE's that were $32\%$ lower than the other schedulers after 100 epochs\footnote{Adam, Adagrad and Adagrad with momentum had an average test MSE of $0.017$, while RMS prop, Constant and Momentum had an average test MSE of $0.025$}.
+%Of these methods, Adagrad used the least amount of time to reach a given MSE.
+Furthermore, Adagrad's cross validated test MSE of 0.011 being slightly lower than Adam's 0.012 when ran for 5000 epochs.
+For neural networks problems, Adam proved to be the superior scheduler for both classification and regression.
+Using this scheduler, we attained a test accuracy of $98.59\%$ on the Wisconsin Breast Cancer Dataset and a test accuracy of $99.07\%$ on the MNIST 8x8 dataset.
 However, we also found that simple logistic regression performed similarly in terms of accuracy, achieving a test accuracy of $98.76\%$ and $99.54\%$ on the cancer and MNIST dataset, respectively.
 Moreover, these results came at a much lower computational cost.
-For regression applied to the Franke Function, we found that our neural net performed worse than a simple polynomial fit using linear regression, attaining a MSE of $0.004781$, 24\% higher than our best polynomial fit MSE of $0.0038302$.
+For regression applied to the Franke Function, we found that our neural net performed worse than a simple polynomial fit using linear regression, attaining a MSE of $0.004781$, 24\% higher than the MSE gained by doing simple polynomial fit using analytical linear regression, which was $0.0038302$.
 We conclude that allthough neural networks are powerful tools which can be used in many contexts, there are many cases in which simpler and less computationally intensive methods can be used to gain similar or even superior performance.
@@
 In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function with 56 polynomial features.
-To gain an intuition, we compare momentum and
+This simple problem will allow us to gain an intuition of how the different schedulers compare against eachother.
 
-% \begin{figure}[h]
-% \centerline{\includegraphics[width=5in]{figure/constant_v_hessian.png}}
-% \caption{Constant scheduler with gridsearched hyperparameters vs hessian (Newton's method) MSE for test data plotted for 200 epochs}
-% \label{constant_v_hessian}
-% \end{figure}
+From  Fig.~\ref{constant_v_momentum}, we can observe that adding momentum improves upon the solutions found by the constant scheduler when using full batch gradient descent, reaching an MSE of 0.180 for the test data.
+By adding a memory of the previous gradients, we gain speed in the direction where most gradients point, speeding up our descent.
+This allows us to obtain a lower MSE scores in fever epochs, when compared to the constant scheduler's MSE of 0.198.
 
-% From observing Fig.~\ref{constant_v_hessian} we see that Newton's method using inversion of the  Hessian matrix greatly outperforms the constant scheduler by reaching an MSE of 0.020 for the testing data within the first epoch.
-Newton's method quickly converges, but does so with some error \cite[121]{num}, meaning it will not improve past this result.
-The analytical solution provides an MSE of approximately 0.004 \cite{ericjona}, a whole order of magnitude lower than the MSE Newton's method achieves.
-By comparison, the constant scheduler given 200 epochs reaches an MSE of 0.198 for the testing data.
-Given this result, it would seem indefensible to claim that Newtons method requires improvements, especially considering the time it took.
-Newton's method took 0.1512 seconds to compute 200 epochs, again beating the result from the constant scheduler which took 0.4991 seconds \footnote{this result should be taken with a grain of salt, given that the constant scheduler is run as a part of a larger architecture, whilst the Hessian calculation is done in a seperate script.
-Still, this provides an intuition that Newton's method does not take much time to compute this specific problem}.
-This result is due to the Hessian matrix only being a mere 56x56 matrix for this problem, such that the Hessian matrix inversion will not be computationally complex.
-However, if we were to increase the complexity of the problem, Newtons method would rapidly increase in computational complexity as it is a large matrix inversion with many floating point operations.
-For problems of such a small scale, when our features are based on a polynomial approximation of the Franke Function, it can be important to remember that we don't necessarily have to turn to these fancier solutions as the simple Newton method will do the job perfectly very well for little computational cost.
-Keeping this lesson in mind, we wish to further investigate the various scheduler methods in order to gain a better understanding of how well they perform, which will become useful for larger problems.
-
-\begin{figure}[h]
+\begin{figure}[H]
 \centerline{\includegraphics[width=5in]{figure/constant_v_momentum.png}}
@@
 
-From  Fig.~\ref{constant_v_momentum}, we can observe that adding momentum improves upon the solutions found by the constant scheduler when using full batch gradient descent, reaching an MSE of 0.1799 for the test data.
-By adding a memory of the previous gradients, we gain speed in the direction where most gradients point, speeding up our descent.
-This allows us to obtain a lower MSE scores in fever epochs, when compared to the constant scheduler's MSE of 0.198.
+By introducing mini-batch gradient descent, we can observe how decreasing the batch size improves our performance further (Fig.~\ref{batch_constant_v_momentum}).
+Mini-batch gradient descent clearly yields a lower MSE for the test data than full batch gradient descent for the same amount of epochs.
+The best result for both constant and momentum were reached for a batch size equal to 2, close to a fully stochastic gradient descent.
+The test data MSE reached $0.044$, an impressive improvement over constant's previous result of $0.198$ MSE with full batch gradient descent.
+Because we first gridsearched the optimal values of eta and lambda first with a predefined base batch size, we must acknowledge that these results will have some bias towards the batch size used when searching for the optimal hyperparameters.
+However, given that we have gridsearched with different base batch sizes accross our testing, we have found the results to be similar accross different base batch sizes, and that this bias which we induced to save computational time was insignificant.
+A full comparison of MSE over batch sizes for all the different schedulers can be viewed in Figure ~\ref{sgdbatch} in the appendix, but all schedulers yield increasingly better results as the batch size decreases.
 
-% This implies that when both schedulers are gridsearched for optimal hyperparameters, the optimal value for the momentum hyperparameter is 0.
-This result makes sense, as the momentum hyperparameter is used to reduce noise from derivatives calculated from smaller batch sizes, so-called mini batches.
-When utilizing a full batch gradient descent, where the batch size equals the size of the training data, momentum will not improve our results.
-
-\begin{figure}[h]
-\centerline{\includegraphics[width=5in]{figure/SGD_batch_momentum_constant.png}}
+\begin{figure}[H]
+\centerline{\includegraphics[width=6in]{figure/SGD_batch_momentum_constant.png}}
 \caption{Constant vs momentum test MSE for different batch sizes, plotted for 200 epochs using full batch gradient descent.
@@
 
-By introducing mini-batch gradient descent, we can observe that stochastic gradient descent with momentum keeps outperforming the constant scheduler.
-Mini-batch gradient descent clearly yields a lower MSE for the test data than full batch gradient descent for the same amount of epochs.
-The best result for both constant and momentum were reached for a batch size equal to 2, close to a fully stochastic gradient descent.
-The test data MSE reached 0.044, an impressive improvement over constant's previous result of 0.198 MSE with full batch gradient descent.
-Because we first gridsearched the optimal values of eta and lambda first with a predefined base batch size, we must acknowledge that Fig.~\ref{batch_constant_v_momentum} will have some bias towards the batch size used when found the optimal hyperparameters.
-However, given that we have gridsearched with different base batch sizes accross our testing, we have found the results to be similar accross different base batch sizes, and that this bias which we induced to save computational time was insignificant.
-A full comparison of MSE over batch sizes for all the different schedulers can be viewed in APPENDIX in the appendix, but all schedulers yield the lowest MSE for stochastic gradient descent.
+Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, Adagrad (with and without momentum) and adam we can observe from Fig.~\ref{figall} that we can reach even greater results in fewer epochs.
+In order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.200.
+Looking at Table.~\ref{timetable}, Adagrad took the least time, reaching the target MSE almost 50 times faster our constant scheduler.
+Based on Fig.~\ref{figall}, Adam seems to reach an MSE of 0.2 the fastest when strictly looking at the number of epochs required.
+However, whilst Adagrad requires more epochs the computations are computed faster in our implementation.
+Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.
 
-\begin{figure}[h]
-\centerline{\includegraphics[width=5in]{figure/200e_schedulers.png}}
+\begin{figure}[H]
+\centerline{\includegraphics[width=7in]{figure/new_adam_v_adagrad.png}}
 \caption{Test MSE over epochs for all schedulers}
@@
 
-Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, adagrad, adagrad with momentum and adam we can observe from Fig.~\ref{figall} that we can reach even greater results in fewer epochs.
-We can also observe that the different schedulers all reach an MSE of 0.200 within a couple of epochs, but in order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.200 of the test data.
-Looking at Table.~\ref{timetable}, Adagrad took the least time, beating out our baseline constant scheduler by 0,222106 seconds.
-Based on Fig.~\ref{figall}, Adam seems to reach an MSE of 0.2 the fastest when strictly looking at the number of epochs required, however, whilst Adagrad requires more epochs the computations are computed faster in our implementation.
-Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.
-
-\begin{table}[h]
+\begin{table}[H]
 \caption{Table over time take for each scheduler}
@@
 
-\begin{figure}[h]
-\centerline{\includegraphics[width=5in]{figure/100e_gridsearch.png}}
+An interesting insight from Fig.~\ref{heatmap} is that all the schedulers prefer high learning rates and regularization.
+As this plot was generated for 100 epochs due to the computational complexity of gridsearching the six schedulers, the bias towards higher learning rates is certainly due to the high learning rates being able to reach a low MSE in the short training time.
+With the lambda = 0 column representing the ordinary least squares solutions, these are clearly consistantly outperformed by Ridge regression as the best MSE is achieved with regularization.
+We can see that the addition of a momentum term makes a small difference for both the constant scheduler, as well as with Adagrad.
+However, we can see that adaptively changing the learning rate, like Adagrad, RMS prop and Adam does, has a much bigger impact on the final MSE than the momentum term.
+
+\begin{figure}[H]
+\centerline{\includegraphics[width=8in]{figure/new_100e_gridsearch.png}}
 \caption{MSE score gridsearch for hyperparameters eta, lambda for different schedulers}
@@
 
-An interesting insight from Fig.~\ref{heatmap} is that all the schedulers prefer high learning rates and regularization.
-As this plot was generated for 100 epochs due to the computational complexity of gridsearching the six schedulers, the bias towards higher learning rates is certainly due to the high learning rates being able to reach a low MSE in the short training time.
-With the lambda = 0 column representing the ordinary least squares solutions, these are clearly consistantly outperformed by Ridge regression as the best MSE is achieved with regularization.
-For this specific plot, momentum does not seem to make out a noticeable difference for the best solutions, as the optimal solution equals their momentumless counterparts.
-That being said, in Fig.~\ref{figall} it is clear that the added momentum aids in a quicker convergence when compared to the constant scheduler.
-
-Another observation from Fig.~\ref{heatmap} is that the schedulers final error for 100 epochs are more or less the same.
+%100 epochs are more or less the same.
 As the largest change in MSE occurs in the first 50 epochs as seen in Fig.~\ref{figall}, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased.
@@
 
-\begin{figure}[h]
-\centerline{\includegraphics[width=5in]{figure/5000e_adam_v_adagrad.png}}
-\caption{MSE test error over 5000 epochs for Adam and Adagrad schedulers}
-\label{adam_v_adagrad}
-\end{figure}
-
-To test this, we compared Adagrad with Adam over 5000 epochs, cross validated for 5 folds, as can be seen in Fig.~\ref{heatmap}.
+% To test this, we compared Adagrad with Adam over 5000 epochs, cross validated for 5 folds, as can be seen in Fig.~\ref{adam_v_adagrad}.
 Both schedulers perform almost equally, with Adagrad reaching a slightly lower MSE of 0.011 for the test data when compared to Adam's 0.012.
@@
 Next, we investigate the use of linear regression in a neural network with hidden nodes, such that we can free ourselves of the limitations of the polynomial approximation.
+
+% \begin{figure}[H]
+% \centerline{\includegraphics[width=5in]{figure/5000e_adam_v_adagrad.png}}
+% \caption{MSE test error over 5000 epochs for Adam and Adagrad schedulers}
+% \label{adam_v_adagrad}
+% \end{figure}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \centerline{\includegraphics[width=6in]{figure/msebymodelfranke.png}}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \centerline{\includegraphics[width=7in]{figure/Franke_fit_66x3_lam0.png}}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \centerline{\includegraphics[width=6in]{figure/cancerarchcompsigmoid.png}}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \centerline{\includegraphics[width=6in]{figure/cancerarchcomp.png}}
@@
 
-\begin{figure}[h]
+\begin{figure}[H]
 \centerline{\includegraphics[width=5in]{figure/conf_mat_NN_cancer.png}}
@@
 In this case, the logistic regression was about 65\% faster\footnote{The time taken to complete a crossvalidation with 5 folds up to 200 epochs was about 52 seconds for our neural network, but only 18 seconds for logistic regression}, with results surpassing our neural network (see Fig.~\ref{cancerconflog}).
+In the MNIST dataset, logistic regression also performed well, achieving an accuracy of $99.54\%$.
 This implies that our classification datasets were perhaps too simple, and that using Neural Networks was not needed to achieve good results.
 
-\begin{figure}[h]
+\begin{figure}[H]
 \centerline{\includegraphics[width=5in]{figure/comp_NN_logreg_cancer.png}}
@@
 \end{figure}
-
-\begin{figure}[h]
+\begin{figure}[H]
 \centerline{\includegraphics[width=5in]{figure/conf_mat_logreg_cancer.png}}
@@
 \end{figure}
-
 \section{Conclusion}
@@
 In this paper we have studied supervised learning algorithms used for classification and regression problems, using the numerical optimization technique of gradient descent as a means of finding an optimal solution.
-We have observed how different schedulers improve the performance of our gradient descent, and found that amongst all the schedulers we tested, Adam and AdaGrad exhibited the best performance overall, achieving a METRIC THAT IS BETTER.
-These schedulers were then used for numerical linear regression, logistic regression and a feed forward neural network.
-Solving both classification and regression problems with these models, we found that our neural network
-
-FUTURE WORK AN ANALYSIS
-
-We then implemented these schedulers in a Feed Forward Neural Network, and applied said network to both regression and classification problems.
-
-Using these schedulers, we applied a Feed Forward Neural network to both classification and regression problems.
-
-Through the application of a Feed Forward Neural Network to both regression and classification problems, we have found that for the problems we have studied, simpler and less computationally intensive methods proved to be superior.
-
-In this paper we have explored the use of a Feed Forward Neural Network, as well as numerical  for both regression and classification alon. Furthermore, we have applied
+We have observed how different schedulers improve the performance of our gradient descent, and found that amongst all the schedulers we tested, Adam and AdaGrad exhibited the best performance overall, achieving MSE's that were $32\%$ lower than the other schedulers in our testing.
+Using these schedulers, we solved regression and classification problems with our neural network implementation, but found that it underperformed compared to simpler models.
+For regression, our network achieved an MSE of $0.004781$, $24\%$ higher than a regular polynomial fit using the normal equations.
+In classification, our network achieved accuracies above $98\%$ for both the Wisconsin Breast Cancer and MNIST datasets, but regular logistic regression also achieved accuracies above $98\%$, in a much shorter time.
+To build upon the resarch conducted in this paper, one avenue to explore is to use larger and more complicated datasets, or to use different forms of neural networks such as convolutional neural networks.
 
@@
 
-\begin{table*}
+\begin{table*}[h]
 \caption{This table shows the file which created every plot in the report}
@@
 \ref{cancerconflog} & \texttt{python3 task\_d.py}\\
+\ref{sgdbatch} & \texttt{python3 task\_a.py}\\
 \hline
@@
 
+\begin{figure}[H]
+\centerline{\includegraphics[width=5in]{figure/SGD_batch_size_100e.png}}
+\caption{MSE over epochs for different batch sizes, for all different schedulers}
+\label{sgdbatch}
+\end{figure}
+
 \end{document}

### 2022-11-17 doc/paper.tex

@@
 However, given that we have gridsearched with different base batch sizes accross our testing, we have found the results to be similar accross different base batch sizes, and that this bias which we induced to save computational time was insignificant.
-A full comparison of MSE over batch sizes for all the different schedulers can be viewed in Figure ~\ref{sgdbatch} in the appendix, but all schedulers yield increasingly better results as the batch size decreases.
+A full comparison of MSE over batch sizes for all the different schedulers can be viewed in Figure ~\ref{sgdbatch} in the appendix, where we can see that all schedulers yield increasingly better results as the batch size decreases.
 
 \begin{figure}[H]
-\centerline{\includegraphics[width=6in]{figure/SGD_batch_momentum_constant.png}}
+\centerline{\includegraphics[width=6in]{figure/new_SGD_batch_momentum_constant.png}}
 \caption{Constant vs momentum test MSE for different batch sizes, plotted for 200 epochs using full batch gradient descent.
@@
 
+Based on these findings, we decided to push Adagrad and Adam further, allowing them to run for much longer.
+After 5000 epochs, we reached an MSE of $0.012$ for Adam and $0.011$ for Adagrad.
+The plot for the MSE over epochs can be found in the appendix (Fig.~\ref{adam_v_adagrad}).
+
 %100 epochs are more or less the same.
@@
 Next, we investigate the use of linear regression in a neural network with hidden nodes, such that we can free ourselves of the limitations of the polynomial approximation.
-
-% \begin{figure}[H]
-% \centerline{\includegraphics[width=5in]{figure/5000e_adam_v_adagrad.png}}
-% \caption{MSE test error over 5000 epochs for Adam and Adagrad schedulers}
-% \label{adam_v_adagrad}
-% \end{figure}
 
@@
 
+\begin{figure}[H]
+\centerline{\includegraphics[width=5in]{figure/5000e_adam_v_adagrad.png}}
+\caption{MSE test error over 5000 epochs for Adam and Adagrad schedulers}
+\label{adam_v_adagrad}
+\end{figure}
+
 \end{document}

### 2022-11-17 src/constant_v_hessian.py
Remove plot

@@ -1,99 +1,0 @@
-# Our own library
-from utils import *
-from Schedulers import *
-from FFNN import FFNN
-# Other libraries
-import timeit
-from tabulate import tabulate
-
-
-"""
-The following script first performs a gridsearch for optimal parameters lambda and eta for use 
-with regular full-batch gradient descent. It the proceeds with the measurement of the runtime 
-of our FFNN and Newton-Raphon's method, before it finally prints said runtime and test error 
-acieved by each technique for the sake of comparison of both methods against eachother.
-"""
-
-np.random.seed(1337)
-
-# ---------------------- Loading data ---------------------- 
-(
-    betas_to_plot,
-    N,
-    X,
-    X_train,
-    X_test,
-    z,
-    z_train,
-    z_test,
-    centering,
-    x,
-    y,
-    z,
-) = read_from_cmdline()
-z_train = z_train.reshape(z_train.shape[0], 1)
-z_test = z_test.reshape(z_test.shape[0], 1)
-
-# --------------------- Setting params --------------------- 
-epochs = 200
-folds = 5
-
-full_batch = 1
-eta = np.logspace(-5, -1, 5)
-lam = np.logspace(-5, -1, 5)
-
-constant_scheduler = Constant
-constant_params = []
-
-# ------------------ Linear Regression NN ------------------
-dims = (X.shape[1], 1)
-neural = FFNN(dims, seed=1337)
-
-# Gridsearching parameters learning rate and lambda for "Constant" scheduler, i.e full-batch gradient descent
-optimal_params, optimal_lambda, _ = neural.optimize_scheduler(
-    X_train,
-    z_train,
-    constant_scheduler,
-    eta,
-    lam,
-    constant_params,
-    batches=full_batch,
-    epochs=epochs,
-    folds=folds,
-)
-
-# Measurement of average runtime of our FFNN model
-time = np.zeros(2)
-start = timeit.default_timer()
-constant_scores = neural.cross_val(
-    folds,
-    X_train,
-    z_train,
-    constant_scheduler,
-    *optimal_params,
-    batches=full_batch,
-    epochs=epochs,
-    lam=optimal_lambda,
-)
-stop = timeit.default_timer()
-time[0] = (stop - start) / folds
-print(f"Constant time to train {epochs} epochs: {time[0]}")
-constant_test_errors = constant_scores["test_errors"]
-
[…]

### 2022-11-18 doc/paper.tex
Ad

@@
 
-%100 epochs are more or less the same.
-As the largest change in MSE occurs in the first 50 epochs as seen in Fig.~\ref{figall}, perhaps the other schedulers are given enough time to catch up as the best performing schedulers, such as Adagrad, starts slowing down the rate at which the MSE is decreased.
-However, as can be seen in Fig.~\ref{heatmap}, Adagrad offers a 12\% \footnote{$\frac{constant MSE}{adagrad MSE} * 100 = \frac{0.018}{0.015} * 100 = 12\%$} decrease in MSE compared to the constant scheduler.
-It is clear from the results thus far that Adagrad is the best scheduler for approximating the Franke function when using linear regression with gradient descent.
-We believe that the reason Adagrad wins out is that these result are based on training for 100 epochs.
-The downfall of Adagrad in the literature seems to be that the aggressive decaying of the learning rate hinders a quick convergence deeper into training \cite{datasci}.
-We are thus cautionary to claim that Adagrad will do the best job over a much larger amount of epochs when compared to adaptive schedulers that claim to improve this aspect such as Adam.
-
-% To test this, we compared Adagrad with Adam over 5000 epochs, cross validated for 5 folds, as can be seen in Fig.~\ref{adam_v_adagrad}.
-Both schedulers perform almost equally, with Adagrad reaching a slightly lower MSE of 0.011 for the test data when compared to Adam's 0.012.
-It could be argued that 5000 epochs is not enough for the improvements of Adam to be noticeable, as the MSE could still be pushed much lower.
-However, for these computationally expensive runs it is clear that Adagrad is the scheduler of choice, especially considering our previous conclusion that Adagrad used less time to train.
-It should be noted that scikit learn's implementation outperforms both our implementation of Adagrad and Adam, with an MSE of 0.004.
-This could be due to a more favorable wight initialization in scikits model, along with additional optimizations implemented in their library.
-That we don't beat the analytical solution is no surprise, as we have limited ourself to a polynomial approximation of the Franke Function with 56 features.
-Our MSE might approach the analytical MSE, but this might take an unreasonable amount of epochs considering the computational runtime.
-Still, we have obtained an MSE lower than the Newton method's result of 0.020, even though it was at a much greater computation.
-Next, we investigate the use of linear regression in a neural network with hidden nodes, such that we can free ourselves of the limitations of the polynomial approximation.
-
 \subsection{Neural Network for regression}
@@
 \ref{heatmap} & \texttt{python3 task\_a.py}\\
-\ref{adam_v_adagrad} & \texttt{python3 unknown.py}\\
 \ref{msebymodelfranke} & \texttt{python3 bycomplexityfranke.py}\\
@@
 \ref{sgdbatch} & \texttt{python3 task\_a.py}\\
+\ref{adam_v_adagrad} & \texttt{python3 linreg\_franke.py}\\
 \hline
@@
 \begin{figure}[H]
-\centerline{\includegraphics[width=5in]{figure/SGD_batch_size_100e.png}}
+\centerline{\includegraphics[width=7in]{figure/SGD_batch_size_100e.png}}
 \caption{MSE over epochs for different batch sizes, for all different schedulers}
@@
 \begin{figure}[H]
-\centerline{\includegraphics[width=5in]{figure/5000e_adam_v_adagrad.png}}
+\centerline{\includegraphics[width=7in]{figure/5000e_adam_v_adagrad.png}}
 \caption{MSE test error over 5000 epochs for Adam and Adagrad schedulers}

### 2022-11-18 src/FFNN.py
clean up and reformat code

@@ -173,15 +173,14 @@ class FFNN:
                 if train_error > 10e20:
                     # if this happens, we have a problem
                     length = 10
-                    train_error=None
-                    test_error=None
-                    train_acc=None
-                    test_acc=None
+                    train_error = None
+                    test_error = None
+                    train_acc = None
+                    test_acc = None
                     break
                 if test_set:
                     prediction_test = self.predict(X_test, raw=True)
                     test_error = cost_function_test(prediction_test)
-                    # test_preds[:, e] = prediction_test.ravel()
 
                     if use_best_weights:
                         if test_error < best_test_error:
@@ -194,7 +193,6 @@ class FFNN:
 
                 else:
                     test_errors = np.nan  # a
-
 
                 train_acc = None
                 test_acc = None
@@ -251,7 +249,6 @@ class FFNN:
         )
         print()
 
-
         # update weights if specified
         if use_best_weights:
             self.weights = best_weights
@@ -290,7 +287,7 @@ class FFNN:
     ):
         """Crossvalidate our fitting
         Takes the same parameters as fit (except for test set)
-        
+
         returns a scores dictionary, but this time averaged over the number of folds
         """
         if self.seed:
@@ -360,9 +357,6 @@ class FFNN:
             This vector is thresholded if we are dealing with classification and raw is not True
         """
 
-        # if self.output_func.__name__ == "sigmoid":
-        #   return np.where(self.feedforward(X) > 0.5, 1, 0)
-        # else:
         predict = self._feedforward(X)
         if raw:
             return predict
@@ -380,7 +374,6 @@ class FFNN:
         Parameters:
             path (str): The path to the file to be written to
         """
-        # print(f'Writing weights to file "{path}"')
         np.set_printoptions(threshold=np.inf)
         with open(path, "w") as file:
             text = str(self.dimensions) + "\n"
@@ -527,7 +520,6 @@ class FFNN:
             gradient_weights = np.mean(gradient_weights_matrix, axis=0)
             delta_accumulated = np.mean(delta_matrix, axis=0)
 
-            # gradient_weights = self.a_matrices[i][:, 1:].T @ delta_matrix
             gradient_weights += self.weights[i][1:, :] * lam
 
             update_matrix = np.vstack(

### 2022-11-18 src/Franke_opt.py

@@ -34,7 +34,7 @@ z_test = z_test.reshape(z_test.shape[0], 1)
 rho = 0.9
 rho2 = 0.99
 
-dims = (2, 66,66,66, 1)
+dims = (2, 66, 66, 66, 1)
 train_epochs = 1000
 
 eta = np.logspace(-4, -1, 4)
@@ -45,10 +45,9 @@ schedulers = [Adagrad, Adam]
 adagrad_params = []
 adam_params = [rho, rho2]
 
-params_list = [ 
-        adagrad_params. 
-        adam_params,
-        ]
+params_list = [
+    adagrad_params.adam_params,
+]
 
 optimal_params_list = []
 optimal_eta = np.zeros(len(schedulers))
@@ -60,10 +59,10 @@ neural = FFNN(dims, hidden_func=LRELU, seed=1337)
 for i in range(len(schedulers)):
     plt.subplot(321 + i)
     plt.suptitle("Test loss for eta, lambda grid", fontsize=22)
-    
+
     optimal_params, optimal_lambda, loss_heatmap = neural.optimize_scheduler(
         X[:, 1:3],
-        z.reshape(400,1),
+        z.reshape(400, 1),
         schedulers[i],
         eta,
         lams,
@@ -81,7 +80,10 @@ for i in range(len(schedulers)):
     ax = sns.heatmap(loss_heatmap, xticklabels=lams, yticklabels=eta, annot=True)
     ax.add_patch(
         Rectangle(
-            (np.where(lams == optimal_lambda)[0], np.where(eta == optimal_params[0])[0]),
+            (
+                np.where(lams == optimal_lambda)[0],
+                np.where(eta == optimal_params[0])[0],
+            ),
             width=1,
             height=1,
             fill=False,

### 2022-11-18 src/task_a.py

@@ -1,3 +1,6 @@
+"""
+This script plots a heatmap displaying a gridsearch over our hyperparameters eta, lambda for different schedulers. The gridsearch also searches for the optimal momentum hyperparameter for momentum-enhanced schedulers. 
+"""
 # Our own library of functions
 import matplotlib.pyplot as plt
 import numpy as np
@@ -29,8 +32,8 @@ z_train = z_train.reshape(z_train.shape[0], 1)
 z_test = z_test.reshape(z_test.shape[0], 1)
 
 # epochs to run for
-epochs = 10
-folds = 2
+epochs = 100
+folds = 5
 
 # no hidden layers, no activation function
 dims = (X.shape[1], 1)
@@ -39,7 +42,7 @@ neural = FFNN(dims, seed=1337)
 # parameters to test for
 eta = np.logspace(-5, -1, 5)
 lam = np.logspace(-5, -1, 5)
-momentums = np.linspace(0, 0.1, 5)
+momentums = np.linspace(0, 1, 5)
 lam[0] = 0
 rho = 0.9
 rho2 = 0.999
@@ -76,7 +79,7 @@ optimal_eta = np.zeros(len(schedulers))
 optimal_lambdas = np.zeros(len(schedulers))
 optimal_batches = np.zeros(len(schedulers), dtype=int)
 
-sns.set(font_scale=2)
+sns.set(font_scale=1.4)
 # gridsearch eta, lambda
 for i in range(len(schedulers)):
     plt.subplot(321 + i)
@@ -88,7 +91,7 @@ for i in range(len(schedulers)):
         eta,
         lam,
         params_list[i],
-        batches=batches_list[-1],
+        batches=batches_list[3],
         epochs=epochs,
         folds=folds,
     )
@@ -118,11 +121,10 @@ for i in range(len(schedulers)):
 plt.show()
 
 # search batch size
-optimal_batches = np.zeros(len(schedulers))
 for i in range(len(schedulers)):
     plt.subplot(321 + i)
     plt.suptitle("MSE over epochs for different \n batch sizes", fontsize=42)
-    optimal_batch, batches_list_search = neural.optimize_batch(
+    _, batches_list_search = neural.optimize_batch(
         X_train,
         z_train,
         X_test,
@@ -134,11 +136,11 @@ for i in range(len(schedulers)):
         epochs=epochs,
     )
 
-    optimal_batches[i] = optimal_batch
     for j in range(len(batches_list)):
         plt.plot(
             batches_list_search[j, :],
[…]

### 2022-11-18 src/task_c.py

@@ -42,10 +42,8 @@ momentum = 0.5
 rho = 0.9
 rho2 = 0.99
 sched = Adam
-# sched = Momentum
 params = [eta, rho, rho2]
 opt_params = [rho, rho2]
-# params = [eta, momentum]
 
 dims = (2, 224, 112, 1)
 train_epochs = 10

### 2022-11-18 src/task_d.py

@@ -1,7 +1,3 @@
-"""
-Performs classification using a neural net and logistic regression, using
-params found by gridsearch performed in src/cancer_opt.py
-"""
 # Our own library of functions
 import matplotlib.pyplot as plt
 import numpy as np
@@ -15,6 +11,11 @@ from sklearn.model_selection import train_test_split, GridSearchCV, KFold
 import time
 
 np.random.seed(1337)
+
+"""
+Performs classification using a neural net and logistic regression, using
+params found by gridsearch performed in src/cancer_opt.py
+"""
 
 # ------------------ Read in data ------------------
 cancer = load_breast_cancer()
@@ -36,7 +37,6 @@ optimal_lambda = 0.0001
 # ------------------------- FFNN -------------------------
 X_train, X_test, t_train, t_test = train_test_split(X, z)
 
-# dims = (30, 60, 1)
 dims = (30, 100, 1)
 neural = FFNN(
     dims, hidden_func=LRELU, output_func=sigmoid, cost_func=CostLogReg, seed=1337

### 2022-11-18 src/time_plot.py
add new optimal params

@@ -42,21 +42,21 @@ schedulers = [Constant, Momentum, Adagrad, AdagradMomentum, RMS_prop, Adam]
 
 # optimal parameters from task_a.py
 optimal_params = [
-    [0.01],
-    [0.01, 0.1],
     [0.1],
-    [0.1, 0.075],
-    [0.001, 0.9],
+    [0.1, 0.25],
+    [0.01],
+    [0.01, 0.75],
+    [0.01, 0.9],
     [0.01, 0.9, 0.999],
 ]
-optimal_lambdas = [0.1, 0.1, 0.001, 0.001, 0.001, 0.01]
+optimal_lambdas = [0.01, 0.01, 0.001, 0.0001, 0.001, 0.0001]
 
 # no hidden layers, no activation function
 dims = (X.shape[1], 1)
 neural = FFNN(dims, seed=1337)
 
 time = np.zeros(len(schedulers))
-MSE_to_reach = 0.2
+MSE_to_reach = 0.03
 
 # runtime comparison
 for i in range(len(schedulers)):

### 2022-11-18 doc/paper.tex
Update some images

@@
 We used this network to test different schedulers, activation functions and model complexities in order to achieve a balance of satisfying results and low computational cost accross different datasets.
-Pitting the schedulers against eachother to fit a polynomial approximation of the Franke Function using linear regression, Adam and Adagrad came out on top, achieving MSE's that were $32\%$ lower than the other schedulers after 100 epochs\footnote{Adam, Adagrad and Adagrad with momentum had an average test MSE of $0.017$, while RMS prop, Constant and Momentum had an average test MSE of $0.025$}.
+Pitting the schedulers against eachother to fit a polynomial approximation of the Franke Function using linear regression, Adam and Adagrad came out on top, achieving MSE's that were $33\%$ lower than the other schedulers after 100 epochs\footnote{Adam, Adagrad and Adagrad with momentum had an average test MSE of $0.0166$, while RMS prop, Constant and Momentum had an average test MSE of $0.0250$}.
 %Of these methods, Adagrad used the least amount of time to reach a given MSE.
@@
 Moreover, these results came at a much lower computational cost.
-For regression applied to the Franke Function, we found that our neural net performed worse than a simple polynomial fit using linear regression, attaining a MSE of $0.004781$, 24\% higher than the MSE gained by doing simple polynomial fit using analytical linear regression, which was $0.0038302$.
+For regression applied to the Franke Function, we found that our neural net performed worse than a simple polynomial fit using linear regression, attaining a MSE of $0.004781$, which is 24\% higher than the MSE gained by doing simple polynomial fit using analytical linear regression, which was $0.0038302$.
 We conclude that allthough neural networks are powerful tools which can be used in many contexts, there are many cases in which simpler and less computationally intensive methods can be used to gain similar or even superior performance.
@@
 Many problems in machine learning cannot be solved analytically.
-Thus, we turn to numerical methods, but this change implicitly comes with computationally heavy caveats as they require large matrix calculations over many iterations.
+Thus, in our search for the optimal solution, we turn to numerical methods.
+This transition, however, implicitly comes with computationally heavy caveats due to numerical optimization often requiring calcluations of large matrices over many iterations.
 Finding a solution becomes a waiting game, and finding the many solutions of varying hyperparameters in order to make an informed model selection becomes virtually impossible.
-The problem of model selection is at the heart machine learning, as optimal model selection allows us to make predictions with the greatest possible accuracy.
-Finding ways around the liminations of numerical methods becomes a must for proper usage of these powerful machine learning tools.
+This is a problem that lies at the heart of machine learning, as the optimal model selection allows us to make predictions with the greatest possible accuracy.
+Finding ways around the limitations of numerical methods becomes a must for proper usage of these powerful machine learning tools.
 
@@
 
-% \subsubsection{Linear Regression}
+\subsubsection{Linear Regression}
+
+Linear regression aims to find a relationship between a set of dependent and explanatory variables.
+Assume we have a dataset, which contains a series of inputs $\{x_i\}$, and corresponding target values $\{y_i\}$.
+Our assumption is that there is some relationship $y_i = f(x_i) + \epsilon_i$ where $\epsilon_i \sim N(0,\sigma^2)$, and we wish to create a model which approximates this relationship.
+More specifically, we wish to find a vector $\beta$ which gives us the "best" approximation $\bm{\tilde{y}}$ of $\bm{y}$, by using the parameters in the following equation:
+
+\begin{equation}
+X \bm{\beta} = \bm{\tilde{y}}
+\label{first}
+\end{equation}
+
+Here, $X$ is our design matrix, containing our explanatory variables, our features, while $\bm{\tilde{y}}$ is our predicted vector.
+By defining a cost function, such as Mean Squared Error, we are able to find the parameter $\beta$ which minimizes the difference between our prediction and the target values.
+By using the normal equations, we can solve this problem analytically.
 
@@
 This means that once a model has been decided on (polynomial degree and value for lambda for example), finding the optimal solution is a deterministic process which does not require further tuning.
-However, for the methods we use in this project - such as Logistic Regression - no analytical solutions exist, thus we will have to turn to numerical techniques to find the 'ideal' solution.
+However, for the methods we use in this project - such as Logistic Regression - no analytical solutions exist, thus we will have to rely on numerical techniques to find the 'ideal' solution.
 Our method of choice in this case is gradient descent, an optimization technique which reduces the residuals by utilizing the idea of moving downhill a function representing our cost until a local optimum is reached.
-The gradient itself is nothing more than a first order derivative of a given function, which points in the direction of the greatest change.
+The gradient itself is nothing more than a first order derivative of a given function with respect to all its parameters, which points in the direction of the greatest change.
 
@@
 These methods work by adaptively adjusting the learning rate, thus allowing us to avoid local minima and converge even faster.
-One of the simpler methods involves adding a fraction of the previous change to the current value used to update the weights, hence keeping a sort of memory of the previous iterations.
-This is called momentum, due to the physical analogue of a moving particle having momentum.
+One of the simpler methods involves adding a fraction of the previous change to the current value used to update the weights, meaning it keeps a sort of memory of the previous iterations.
+This hyperparameter is called momentum, due to the physical analogue of a moving particle having momentum.
 With this addition, we will keep moving in parameter space even if our current gradient is small, because we have gained momentum in a certain direction based on our previous iterations.
@@
 
-Like logistic regression, the optimal weights for each layer in an FFNN can not be found analytically, and we have to turn to gradient descent.
+As in the case of logistic regression, the optimal weights for each layer in an FFNN can not be found analytically, and we have to turn to gradient descent.
 By repeated use of the chain rule, we are able to calculate the gradient for each layer iteratively, using the gradient of the layer after it and starting at the output.
@@
 
-Training a model is implemented as repeated calls to feedforward and backpropagation, once per batch per epoch.
+Training of a model is implemented as repeated calls to feedforward and backpropagation, once per batch per epoch.
 During training, one is able to specify which scheduler to use.
@@
 
-In practice, we add a normally distributed noise $N(0,0.05)$ to this function.
+In practice, to better test our models ability to generalize and to mimic real terrain data, we add a normally distributed noise $N(0,0.05)$ to this function.
 
@@
 
-This dataset contains 30 features describing important characteristics of tumors.
-The features are real numbers.
+This dataset contains 30 features describing important characteristics of tumors, where each feature is a real numbers.
 Each instance has an associated class, indicating whether the tumor was benign or not.
@@
 
-In order to study the attributes of the different scheduler methods we performed numerical linear regression using gradient descent on the Franke Function with 56 polynomial features.
+In order to study the attributes of the different scheduler methods, we performed numerical linear regression using gradient descent on the Franke Function with 56 polynomial features.
 This simple problem will allow us to gain an intuition of how the different schedulers compare against eachother.
@@
 The best result for both constant and momentum were reached for a batch size equal to 2, close to a fully stochastic gradient descent.
-The test data MSE reached $0.044$, an impressive improvement over constant's previous result of $0.198$ MSE with full batch gradient descent.
-Because we first gridsearched the optimal values of eta and lambda first with a predefined base batch size, we must acknowledge that these results will have some bias towards the batch size used when searching for the optimal hyperparameters.
+The test data MSE reached $0.044$, an impressive improvement over constant's previous result of $0.198$ MSE when using full batch gradient descent.
+Because we first gridsearched the optimal values of eta and lambda with a predefined base batch size, we must acknowledge that these results will have some bias towards the batch size used when searching for the optimal hyperparameters.
 However, given that we have gridsearched with different base batch sizes accross our testing, we have found the results to be similar accross different base batch sizes, and that this bias which we induced to save computational time was insignificant.
@@
 
-Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, Adagrad (with and without momentum) and adam we can observe from Fig.~\ref{figall} that we can reach even greater results in fewer epochs.
-In order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of 0.200.
-Looking at Table.~\ref{timetable}, Adagrad took the least time, reaching the target MSE almost 50 times faster our constant scheduler.
-Based on Fig.~\ref{figall}, Adam seems to reach an MSE of 0.2 the fastest when strictly looking at the number of epochs required.
-However, whilst Adagrad requires more epochs the computations are computed faster in our implementation.
-Surprisingly, RMS prop performed more poorly than expected compared to other adaptive schedulers.
+Furthermore, with the introduction of various adaptive scheduler methods such as RMS prop, Adagrad (with and without momentum) and Adam, we can observe from Fig.~\ref{figall} that we can reach even greater results in fewer epochs.
+In order to get a better understanding of the computational cost we have also compared the time it took to fit the data to an MSE of $0.03$.
+Looking at Table.~\ref{timetable}, RMS prop took the least time, reaching the target MSE over 5 times faster than our constant scheduler.
+If we take a look at figure ~\ref{figall}, we can see that RMS\_prop has a fast descent in the first iterations, however, it struggles in the later parts, reaching MSE values that are much higher than Adam and Adagrad.
 
 \begin{figure}[H]
-\centerline{\includegraphics[width=7in]{figure/new_adam_v_adagrad.png}}
+\centerline{\includegraphics[width=7in]{figure/fin_all_schedulers.jpg}}
 \caption{Test MSE over epochs for all schedulers}
@@
 \hline
-Constant & 0.601 \\
-Momentum & 0.513 \\
-RMS\_prop & 0.330 \\
-AdaGrad & 0.013 \\
-AdaGrad with Momentum & 0.014 \\
-Adam & 0.041 \\
+Constant & 0.489 \\
+Momentum & 0.379 \\
+RMS\_prop & 0.089 \\
+AdaGrad & 0.205 \\
+AdaGrad with Momentum & 0.135 \\
+Adam & 0.209 \\
 \hline
@@
 
-An interesting insight from Fig.~\ref{heatmap} is that all the schedulers prefer high learning rates and regularization.
-As this plot was generated for 100 epochs due to the computational complexity of gridsearching the six schedulers, the bias towards higher learning rates is certainly due to the high learning rates being able to reach a low MSE in the short training time.
-With the lambda = 0 column representing the ordinary least squares solutions, these are clearly consistantly outperformed by Ridge regression as the best MSE is achieved with regularization.
-We can see that the addition of a momentum term makes a small difference for both the constant scheduler, as well as with Adagrad.
-However, we can see that adaptively changing the learning rate, like Adagrad, RMS prop and Adam does, has a much bigger impact on the final MSE than the momentum term.
+Moving over to Fig.~\ref{heatmap}, we gain interesting insight into the behavior of our schedulers, with all of them preferring rather high learning rates and mid to low values of regularization.
+This plot was generated for 100 epochs due to the computational complexity of gridsearching for these parameters for all six schedulers.
+The observable bias towards higher values of learning rates is certainly a side effect of the limiting the number of epochs, as this short interval favours parameters which allow the schedulers to reach low MSEs fast, even if it might lead to instabilities in later iterations.
+These high learning rates coupled with values of lambda higher than zero, indicate that Ridge regression systematically outperforms ordinary least squares.
+We can see that the addition of a momentum term makes a small difference for both the constant scheduler and Adagrad.
+On the other hand, we see that adaptively changing the learning rate, like Adagrad, RMS prop and Adam does, has a much bigger impact on the final MSE than the momentum term.
 
 \begin{figure}[H]
-\centerline{\includegraphics[width=8in]{figure/new_100e_gridsearch.png}}
-\caption{MSE score gridsearch for hyperparameters eta, lambda for different schedulers}
+\centerline{\includegraphics[width=8in]{figure/fin_heatmap.png}}
+\caption{Heatmap for each scheduler, showing the optimal lambda and eta for each method, as well as the best MSE reached (highlighted in red)}
 \label{heatmap}
@@
 The plot for the MSE over epochs can be found in the appendix (Fig.~\ref{adam_v_adagrad}).
+Our models have also been compared to SciKit's implementations (Fig.~\ref{scikitcompare}).
 
@@
 There are many reasons that this may be the case.
-As we have stated before, once we start using numerical methods, computational effort becomes a large factor in any test one does, and many decisions have to made regarding how much time can be spent on each test.
-Due to these computational constraints, we have not been able to grid search for the best parameters for each model complexity, nor have we been able to train for more than 200 epochs for each model.
-These shortcomings will greatly affect how our models perform, as it is unlikely that each model uses the same learning rate and scheduler, or that more complex models use the same amount of epochs to train their weights.
-Based on this, we can assume that the higher complexity models have not been trained enough, and that this explains why both their train and test MSE are higher than the less complex models.
+As we have stated before, once we start using numerical methods, computational effort becomes a large factor in any test, and many decisions have to made regarding how much time can be spent on each test.
+Due to these computational constraints, we have not been able to grid search for the best parameters over each model complexity, nor have we been able to train for more than 200 epochs for each model.
+These shortcomings greatly affect how our models perform, as it is unlikely that each model uses the same learning rate and scheduler, or that more complex models use the same amount of epochs to train their weights.
+Based on this, we assume that the higher complexity models have not been trained enough, and that this explains why both their train and test MSE are higher than the less complex models.
 Regardless, it seems that for up to about 200 total nodes, three hidden layers does the best, and as such we chose an architecture of three hidden layers with 66 nodes each (which is where we achieved our lowest MSE).
@@
 Using this architecture, we performed a gridsearch.
-Due to time constraints, we have felt it necessary to limit our gridsearch to only two schedulers.
-Based on our testing with linear regression with gradient descent, we have seen that Adam and Adagrad seems to be the best performers.
-Based on Goodfellows recommendations, we started with the RELU activation function \cite[188]{gbc}.
+Again, due to computational constraints, we have felt it necessary to limit our gridsearch to only two schedulers.
+Based on our testing of linear regression with gradient descent, we have seen that Adam and Adagrad seems to be the best performers in terms of MSE\footnote{Adagrad with momentum was discarded even though it performed well, due to the gridsearch for this method involving the momentum parameter as well, making the gridsearch take much longer than the other methods}.
+Following Goodfellows recommendations, we started with the RELU activation function \cite[188]{gbc}.
 With this, we found Adam to be the best scheduler, with an eta of 0.01 and no regularization.
@@
 We can see that by not forcing our model to try and find a polynomial fit, we are achieving better results.
-However, we are still not able to beat the analytical solution to the best polynomial fit at polynomial degree 9, which is somewhat disappointing.
+Unfortunately, we are still unable to beat the analytical solution to the best polynomial fit at polynomial degree 9, which is somewhat disappointing.
 A possible explanation for this is that the Franke function can be represented as a weighted sum of polynomials, and as such a polynomial fit is especially suited to approximate this function.
@@
 To test our neural networks ability to solve classification problems, we first applied it to the Wisconsin Breast Cancer dataset.
-But to get the best results, there are a large amount of parameters that need to be decided on.
-We must find the best architecture, the best activation functions, the best scheduler and the best parameters for said scheduler.
-Finding the absolute best combination of all these choices is impossible, so compromises must be made.
-
+As with regression, in order to achieve satisfying results, the hyperparameters of our model need to be tuned, which requires us to consider the tradeoff between performance and computational complexity.
 We start by trying to find a good model complexity, i.e the best number of hidden layers, and the amount of nodes in them.
@@
 With such a dataset, it is possible to gain a very good fit using more and more complex models, although the computational effort obviously increases as well.
-Moreover, it is not given that instances outside this dataset are as homogenous as the relatively small (569) sample we have here, so one should be careful to use an overcomplex model even if it does not hurt the test accuracy for this particular dataset.
+Moreover, it is not given that instances outside this dataset are as homogenous as the relatively small (569) sample we have here, meaning one should be careful to use an overcomplex model even if it does not hurt the test accuracy for this particular dataset.
 Using the RELU activation function, the results are quite different, as can be seen in Fig.~\ref{cancerarchcomp}.
@@
 Clearly, we are again running into the "dying RELU problem".
-Here we can clearly see some of the benefits of the sigmoid, as even if the gradients become very small, they will never be zero, as with RELU.
+Here we can clearly see some of the benefits of the sigmoid, as even when the gradients become very small, they will never be zero, as with RELU.
 Regardless, we see similar performance for RELU using only one hidden layer.
@@
 \ref{adam_v_adagrad} & \texttt{python3 linreg\_franke.py}\\
+\ref{scikitcompare} & \texttt{python3 scikit\_compare.py}\\
 \hline
@@
 \begin{figure}[H]
-\centerline{\includegraphics[width=7in]{figure/SGD_batch_size_100e.png}}
+\centerline{\includegraphics[width=7in]{figure/fin_sgd_batch_size_all.png}}
 \caption{MSE over epochs for different batch sizes, for all different schedulers}
@@
 \begin{figure}[H]
-\centerline{\includegraphics[width=7in]{figure/5000e_adam_v_adagrad.png}}
+\centerline{\includegraphics[width=8in]{figure/5000e_adam_v_adagrad.png}}
 \caption{MSE test error over 5000 epochs for Adam and Adagrad schedulers}
@@
 
+\begin{figure}[H]
+\centerline{\includegraphics[width=7in]{figure/scikit_compare.png}}
+\caption{Comparison between our own implementation of Adam and Adagrad with SciKit's implementation}
+\label{scikitcompare}
+\end{figure}
+
 \end{document}

### 2022-11-18 src/constant_v_momentum.py
final commit

@@ -39,7 +39,7 @@ lam = np.logspace(-5, -1, 5)
 
 schedulers = [Momentum, Constant]
 constant_params = []
-momentum_params = np.linspace(0, 0.1, 5)
+momentum_params = np.linspace(0, 1, 5)
 params = [momentum_params, constant_params]
 
 # ------------------ Linear Regression NN ------------------

### 2022-11-18 src/task_d_digits.py

@@ -11,6 +11,9 @@ from sklearn.datasets import load_digits
 from sklearn.model_selection import train_test_split, GridSearchCV, KFold
 
 np.random.seed(1337)
+"""
+This script performs multinomial classification using neural networks and logistic regression on the MNIST dataset. Prints scores.
+"""
 
 # read in data
 digits = load_digits()

### 2022-11-18 doc/paper.tex

@@
 \centerline{\includegraphics[width=8in]{figure/fin_heatmap.png}}
-\caption{Heatmap for each scheduler, showing the optimal lambda and eta for each method, as well as the best MSE reached (highlighted in red)}
+\caption{MSE score gridsearch for hyperparameters eta, lambda for different schedulers}
 \label{heatmap}
@@
 \begin{figure}[H]
-\centerline{\includegraphics[width=7in]{figure/fin_sgd_batch_size_all.png}}
+\centerline{\includegraphics[width=7in]{figure/SGD_batch_size_100e.png}}
 \caption{MSE over epochs for different batch sizes, for all different schedulers}
@@
 \begin{figure}[H]
-\centerline{\includegraphics[width=8in]{figure/5000e_adam_v_adagrad.png}}
+\centerline{\includegraphics[width=7in]{figure/5000e_adam_v_adagrad.png}}
 \caption{MSE test error over 5000 epochs for Adam and Adagrad schedulers}

### 2022-11-18 doc/paper.tex

--- 2022-11-18 doc/paper.tex
@@
 This simple problem will allow us to gain an intuition of how the different schedulers compare against eachother.
-
-From  Fig.~\ref{constant_v_momentum}, we can observe that adding momentum improves upon the solutions found by the constant scheduler when using full batch gradient descent, reaching an MSE of 0.180 for the test data.
+From  Fig.~\ref{constant_v_momentum}, we can observe that adding momentum improves upon the solutions found by the constant scheduler considerably, reaching an MSE of $0.09$ for the test data after 200 epochs.
 By adding a memory of the previous gradients, we gain speed in the direction where most gradients point, speeding up our descent.
-This allows us to obtain a lower MSE scores in fever epochs, when compared to the constant scheduler's MSE of 0.198.
+This allows us to obtain a lower MSE scores in fever epochs, when compared to the constant scheduler's MSE of $0.21$.
 
@@
 \centerline{\includegraphics[width=5in]{figure/constant_v_momentum.png}}
-\caption{Constant vs momentum MSE for test data plotted for 200 epochs using full batch gradient descent. The final MSE for Constant was 0.198, ten percent higher than Momentum's 0.180.}
+\caption{Constant vs momentum MSE for test data plotted for 200 epochs using full batch gradient descent. The final MSE for Constant was $0.21$, considerably higher than Momentum's $0.09$.}
 \label{constant_v_momentum}
